### 2010-01-05 manuscript/PNAS/PNAS_submission.tex
initial commit

%% PNAStmpl.tex
%% Template file to use for PNAS articles prepared in LaTeX
%% Version: Apr 14, 2008

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BASIC CLASS FILE 
%% PNAStwo for two column articles is called by default.
%% Uncomment PNASone for single column articles. One column class
%% and style files are available upon request from pnas@nas.edu.
%% (uncomment means get rid of the '%' in front of the command)

%\documentclass{pnasone}
\documentclass{pnastwo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Changing position of text on physical page:
%% Since not all printers position
%% the printed page in the same place on the physical page,
%% you can change the position yourself here, if you need to:

% \advance\voffset -.5in % Minus dimension will raise the printed page on the 
                         %  physical page; positive dimension will lower it.

%% You may set the dimension to the size that you need.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL GRAPHICS STYLE FILE

%% Requires graphics style file (graphicx.sty), used for inserting
%% .eps files into LaTeX articles.
%% Note that inclusion of .eps files is for your reference only;
%% when submitting to PNAS please submit figures separately.

%% Type into the square brackets the name of the driver program 
%% that you are using. If you don't know, try dvips, which is the
%% most common PC driver, or textures for the Mac. These are the options:

% [dvips], [xdvi], [dvipdf], [dvipdfm], [dvipdfmx], [pdftex], [dvipsone],
% [dviwindo], [emtex], [dviwin], [pctexps], [pctexwin], [pctexhp], [pctex32],
% [truetex], [tcidvi], [vtex], [oztex], [textures], [xetex]

%\usepackage[dvips]{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL POSTSCRIPT FONT FILES

%% PostScript font files: You may need to edit the PNASoneF.sty
%% or PNAStwoF.sty file to make the font names match those on your system. 
%% Alternatively, you can leave the font style file commands commented out
%% and typeset your article using the default Computer Modern 
%% fonts (recommended). If accepted, your article will be typeset
%% at PNAS using PostScript fonts.

% Choose PNASoneF for one column; PNAStwoF for two column:
%\usepackage{PNASoneF}
%\usepackage{PNAStwoF}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% ADDITIONAL OPTIONAL STYLE FILES

%% The AMS math files are commonly used to gain access to useful features
%% like extended math fonts and math commands.

\usepackage{amssymb,amsfonts,amsmath}

\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% OPTIONAL MACRO FILES
%% Insert self-defined macros here.
%% \newcommand definitions are recommended; \def definitions are supported

%\newcommand{\mfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
%\def\s{\sigma}
\newcommand{\mB}{\mathcal{B}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mM}{\mathcal{M}}
% \newcommand{\mN}{\mathcal{N}}

\newcommand{\hL}{\widehat{L}}
\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_F \mB}
\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{thex}{Thought experiment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Don't type in anything in the following section:
%%%%%%%%%%%%
%% For PNAS Only:
\contributor{Submitted to Proceedings
of the National Academy of Sciences of the United States of America}
\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
\copyrightyear{2008}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% For titles, only capitalize the first letter
%% \title{Almost sharp fronts for the surface quasi-geostrophic equation}

\title{Are mental properties supervenient on brain properties?}

%% Enter authors via the \author command.  
%% Use \affil to define affiliations.
%% (Leave no spaces between author name and \affil command)

%% Note that the \thanks{} command has been disabled in favor of
%% a generic, reserved space for PNAS publication footnotes.

%% \author{<author name>
%% \affil{<number>}{<Institution>}} One number for each institution.
%% The same number should be used for authors that
%% are affiliated with the same institution, after the first time
%% only the number is needed, ie, \affil{number}{text}, \affil{number}{}
%% Then, before last author ...
%% \and
%% \author{<author name>
%% \affil{<number>}{}}

%% For example, assuming Garcia and Sonnery are both affiliated with
%% Universidad de Murcia:
%% \author{Roberta Graff\affil{1}{University of Cambridge, Cambridge,
%% United Kingdom},
%% Javier de Ruiz Garcia\affil{2}{Universidad de Murcia, Bioquimica y Biologia
%% Molecular, Murcia, Spain}, \and Franklin Sonnery\affil{2}{}}

\author{Joshua T. Vogelstein\affil{1}{Johns Hopkins University, Baltimore, MD}, 
R. Jacob Vogelstein\affil{1}{}, 
Carey E. Priebe\affil{1}{}}

\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

%% The \maketitle command is necessary to build the title page.
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{article}

\begin{abstract}

The ``mind-brain supervenience'' theorem suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain. The validity of this theorem has been argued in philosphical terms for over 2,500 years, but it has not previously been approachable through experimental investigation. To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis. To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience. The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two. In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.

\end{abstract}

%% When adding keywords, separate each term with a straight line: |
\keywords{graph theory | universal consistency | neural circuit}

%% Optional for entering abbreviations, separate the abbreviation from
%% its definition with a comma, separate each pair with a semicolon:
%% for example:
%% \abbreviations{SAM, self-assembled monolayer; OTS,
%% octadecyltrichlorosilane}

% \abbreviations{}

%% The first letter of the article should be drop cap: \dropcap{}
%\dropcap{I}n this article we study the evolution of ''almost-sharp'' fronts

%% Enter the text of your article beginning here and ending before
%% \begin{acknowledgements}
%% Section head commands for your reference:
%% \section{}
%% \subsection{}
%% \subsubsection{}

Questioning the relationship between the mind (thoughts, beliefs, preferences, emotions, intelligence, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}. Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}. Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience theorem, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}. Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation. Here we define new versions of supervenience that formulate the theorem in rigorous mathematical terms and that can be experimentally tested.

Let $\mB$ be the observation space for some physical property, such as brain connectivity structure (i.e., connectome; see \cite{SpornsKotter05,LichtmanSanes08,Seung09}). Let $\mM$ be the (finite) indicator space for some mental property, such as knowing calculus. Thus, for $b \in \mB$ and $m \in \mM$, the pair $(b,m)$ represents a brain property/mind property pair.

%Let $\{(B_i,M_i)\}_{i=1}^n \overset{iid}{\sim} F_{BM}$
Let $(B,M), (B_1,M_1),\cdots,(B_n,M_n)$ be random observation pairs taking their values in $\mB \times \mM$, independently and identically distributed according to some joint probability distribution $F=F_{BM}$. Abusing notation to conceptually identify the properties with their spaces, the statistical supervenience relation $\mM \overset{S}{\sim}_F \mB$ says that $M_i \neq M_j \implies B_i \neq B_j$ (almost surely; where $\implies$ does not suggest causation). That is, observing $B=b$ can allow us to assign $m$ to $M$. While previously proposed notions of mind-brain-supervenience claim that all mental properties supervene on physical properties \cite{Kim95}, here we consider empirically investigating only whether a particular mental property $\mM$ statistically supervenes on a particular physical property $\mB$.

Let $g:\mB \rightarrow \mM$ be a classifier, which takes as input an observed brain connectivity structure $b$ and produces a classification $\widehat{m}=g(b)$ for the unobserved mental property $m$. The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, where $L_{F}(g) = P_{F}[g(B) \neq M]$ denotes the probability of misclassification for classifier $g$ under joint distribution $F=F_{BM}$. We can therefore rigorously define \textit{statistical supervenience}:

\begin{defi}
\label{def1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{F}(g^*) = 0$.
\end{defi}

Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$. Therefore, we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:

\begin{defi}
\label{def2}
Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
\end{defi}

Now, generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \rightarrow \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;(b_1,m_1),\cdots,(b_n,m_n))$. Let $L_{F}(g_n) = E[P_{F}[g_n(B;(B_1,M_1),\cdots,(B_n,M_n)) \neq M|(B_1,M_1),\cdots,(B_n,M_n)]]$.

Consider the problem of testing for $\varepsilon$-supervenience. Let the null hypothesis be given by $H_0: L_{F}(g_n) \geq \varepsilon$ so that if we reject at level $\alpha >0$ in favor of the alternative hypothesis $H_A: L_{F}(g_n) < \varepsilon$ then we can conclude, with $100(1-\alpha$)\% confidence, that $\MeB$. Letting $\hL^{n'}_{F}(g_n)$ denote the hold-out estimate of misclassification performance based on $n'$ test observations, we note that $\hL^{n'}_{F}(g_n)$ is distributed $Binomial(n',L_{F}(g_n))$. The test rejects for small $\hL^{n'}_{F}(g_n)$. The level $\alpha$ critical value $c_{\alpha}(n',\varepsilon)$ is available under the least favorable distribution $Binomial(n',\varepsilon)$. Furthermore, $\MeB$ implies $L_{F}(g^*) < \varepsilon$, and thus if $g_n$ is a {\it consistent} classifier for $F=F_{BM}$ --- that is, if $\lim_n L_{F}(g_n) = L_{F}(g^*)$ --- then the power of this test (the probability of rejecting when in fact the alternative is true) goes to unity as $n,n' \rightarrow \infty$. Thus we have an inference procedure: \begin{thm} \label{thm1} Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$. Furthermore, given a consistent classifier the power of the test converges to unity. \end{thm}

Since the joint distribution $F=F_{BM}$ is unknown, the utility of Theorem \ref{thm1} requires that $g_n$ be a {\it universally consistent} classifier --- that is, $\lim_n L_{F}(g_n) = L_{F}(g^*)$ for all distributions $F=F_{BM}$. Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{BM}$ \cite{DGL96}. Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}. For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation. Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not satisfy Popper's {\it falsifiability} requirement \cite{Popper}. Given these caveats, consider the following thought experiment:

\begin{thex}

Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts). Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable. Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators. A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 1 for proof). Therefore, Theorem \ref{thm1} applies and the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds. This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.

\end{thex}

While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality. Assuming we have confirmed $\MeB$ for a particular mental/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.

Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities. Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment). Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.

%% == end of paper:

%% Optional Materials and Methods Section
%% The Materials and Methods section header will be added automatically.

%% Enter any subheads and the Materials and Methods text below.
%\begin{materials}
% Materials text
%\end{materials}

%% Optional Appendix or Appendices
%% \appendix Appendix text...
%% or, for appendix with title, use square brackets:
%% \appendix[Appendix Title]

\section*{Appendix 1: $k_n$-nearest neighbor universal consistency for graphs}
\label{proof}

Assume first that all graphs are simple, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another. Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$. If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$. Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$. Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.

In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability. Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $P[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{F}(g_n) \rightarrow c \leq L_{F}(g^*) + \delta$, yielding universal consistency.

If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm. Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$. Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}. Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
%\tr{Trivially, this argument also applies to graphs with (self-) loops, and
%directed, colored, or weighted edges, and hypergraphs.}

Several points of note: Isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.

%\appendix
\section*{Appendix 2: Simulation} \label{simulation}

As an example of a feasible experiment, one may consider a species whose nervous system consists of the same (small) number of labeled neurons for each organism. {\it Caenorhabditis elegans} is believed to be such a species \cite{Durbin87}. The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons. While the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}. Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}. Thus, one may design an experiment by describing the joint distribution $F_{BM}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $P[M=m_1]$. Here the mental property corresponds to the C.~elegans exhibiting or not exhibiting a particular behavior (e.g., response to an odor).

Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $F_{BM}$ (Figure \ref{fig1}). To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$. The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ k_{ij})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $k_{ij}$ uniformly sampled from $[-5,5]$. We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm. The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency. (Better classifiers can be constructed for the joint distribution $F_{BM}$ used here; however, we demand universal consistency.)

Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations. 3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day. Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.

\begin{acknowledgments}
The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.
\end{acknowledgments}

%% PNAS does not support submission of supporting .tex files such as BibTeX.
%% Instead all references must be included in the article .tex document. 
%% If you currently use BibTeX, your bibliography is formed because the 
%% command \verb+\bibliography{}+ brings the <filename>.bbl file into your
%% .tex document. To conform to PNAS requirements, copy the reference listings
%% from your .bbl file and add them to the article .tex file, using the
%% bibliography environment described above.

%%  Contact pnas@nas.edu if you need assistance with your
%%  bibliography.

% Sample bibliography item in PNAS format:
%% \bibitem{in-text reference} comma-separated author names up to 5,
%% for more than 5 authors use first author last name et al. (year published)
%% article title  {\it Journal Name} volume #: start page-end page.
%% ie,
% \bibitem{Neuhaus} Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) 
% A short C-terminal sequence is necessary and sufficient for the
% targeting of chitinases to the plant vacuole. 
% {\it Proc Natl Acad Sci USA} 88:10362-10366.

%% Enter the largest bibliography number in the facing curly brackets
%% following \begin{thebibliography}

\begin{thebibliography}{26}

\bibitem{Plato97}
Plato.
\newblock {\em Plato: complete works}.
\newblock Hackett Pub Co,  (1997).

\bibitem{Descartes1641}
Descartes, R.
\newblock {\em Meditationes de prima philosophia}.
\newblock  (1641).

\bibitem{Davidson70}
Davidson, D.
\newblock {\em Experience and Theory}, chapter Mental Events.
\newblock Duckworth (1970).

\bibitem{SpornsKotter05}
Sporns, O., Tononi, G., and Kotter, R.
\newblock {\em PLoS Computational Biology}{ \bf 1}(4), e42 (2005).

\bibitem{LichtmanSanes08}
Lichtman, J.~W., Livet, J., and Sanes, J.~R.
\newblock {\em Nat Rev Neurosci}{ \bf 9}(6), 417--422 Jun  (2008).

\bibitem{Seung09}
Seung, H.
\newblock {\em Neuron}{ \bf 62}(1), 17--29 (2009).

\bibitem{Kim95}
Kim, J.
\newblock {\em Phiilosophy of Mind}.
\newblock Westview Press, second edition,  (2005).

\bibitem{DGL96}
Devroye, L., Gy{\"o}rfi, L., and Lugosi, G.
\newblock {\em {A Probabilistic Theory of Pattern Recognition}}.
\newblock Springer,  (1996).

\bibitem{Devroye83}
Devroye, L.
\newblock {\em Probability Theory and Related Fields}{ \bf 62}(4), 475--483
  (1983).

\bibitem{Popper}
Popper, K.
\newblock { \bf } (1959).

\bibitem{WhiteBrenner86}
White, J., Southgate, E., Thomson, J.~N., and Brenner, S.
\newblock {\em Philosophical Transactions of Royal Society London. Series B,
  Biological Sciences}{ \bf 314}(1165), 1--340 (1986).

\bibitem{DenkHorstmann04}
W.~Denk, W. and Horstmann, H.
\newblock {\em PLOS Biol.}{ \bf 2}, e329 (2004).

\bibitem{BriggmanDenk06}
Briggman, K. and Denk, W.
\newblock {\em Current opinion in neurobiology}{ \bf 16}(5), 562--570 (2006).

\bibitem{MackeBorst08}
Macke, J.~H., Maack, N., Gupta, R., Denk, W., Schölkopf, B., and Borst, A.
\newblock {\em J Neurosci Methods}{ \bf 167}(2), 349--357 Jan  (2008).

\bibitem{Mishchenko09}
Mishchenko, Y.
\newblock {\em J Neurosci Methods}{ \bf 176}(2), 276--289 Jan  (2009).

\bibitem{LuLichtman09}
Lu, J., Fiala, J.~C., and Lichtman, J.~W.
\newblock {\em PLoS ONE}{ \bf 4}(5), e5655 05  (2009).

\bibitem{HornJohnson90}
Horn, R. and Johnson, C.
\newblock {\em {Matrix analysis}}.
\newblock Cambridge Univ Pr,  (1990).

\bibitem{ConroyLouis97}
Conroy, J.~M., Kratzer, S.~G., and Podrazik, L.~J.
\newblock In {\em Society for Industrial and Applied Mathematics},  (1997).

\bibitem{ZaslavskiyVert08}
Zaslavskiy, M., Bach, F., and Vert, J.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}{
  \bf 31}, 2227--2242 (2008).

\bibitem{Durbin87}
Durbin, R.~M.
\newblock {\em Studies on the Development and Organisation of the Nervous
  System of Caenorhabditis elegans}.
\newblock PhD thesis, University of Cambridge,  (1987).

\bibitem{deBonoMaricq05}
de~Bono, M. and Maricq, A.~V.
\newblock {\em Annu Rev Neurosci}{ \bf 28}, 451--501 (2005).

\bibitem{VarshneyChklovskii09}
Varshney, L., Chen, B., Paniagua, E., Hall, D., and Chklovskii, D.
\newblock {\em ArXiv}{ \bf } (2009).

\bibitem{ChalasaniBargmann07}
Chalasani, S.~H., Chronis, N., Tsunozaki, M., Gray, J.~M., Ramot, D., Goodman,
  M.~B., and Bargmann, C.~I.
\newblock {\em Nature}{ \bf 450}(7166), 63--70 Nov  (2007).

\bibitem{VaziriShank08}
Vaziri, A., Tang, J., Shroff, H., and Shank, C.~V.
\newblock {\em Proc Natl Acad Sci U S A}{ \bf 105}(51), 20221--20226 Dec
  (2008).

\bibitem{HelmstaedterDenk08}
Helmstaedter, M., Briggman, K.~L., and Denk, W.
\newblock {\em Curr Opin Neurobiol}{ \bf 18}(6), 633--641 Dec  (2008).

\bibitem{BuckinghamSattelle08}
Buckingham, S.~D. and Sattelle, D.~B.
\newblock {\em Invert Neurosci}{ \bf 8}(3), 121--131 Sep  (2008).

\end{thebibliography}

\end{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Adding Figure and Table References
%% Be sure to add figures and tables after \end{article}
%% and before \end{document}

%% For figures, put the caption below the illustration.
%%
%% \begin{figure}
%% \caption{Almost Sharp Front}\label{afoto}
%% \end{figure}

\begin{figure}[h!]
% \centering 
\includegraphics[width=.5\linewidth]{Lhatplot}
\caption{C.~elegans graph classification simulation results. $\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$. Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$. E.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01. We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$. $L_{F}(g^*) \approx 0$ for this simulation.
}
\label{fig1}
\end{figure}

%% For Tables, put caption above table
%%
%% Table caption should start with a capital letter, continue with lower case
%% and not have a period at the end
%% Using @{\vrule height ?? depth ?? width0pt} in the tabular preamble will
%% keep that much space between every line in the table.

%% \begin{table}
%% \caption{Repeat length of longer allele by age of onset class}
%% \begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}lrcccc}
%% table text
%% \end{tabular}
%% \end{table}

%% For two column figures and tables, use the following:

%% \begin{figure*}
%% \caption{Almost Sharp Front}\label{afoto}
%% \end{figure*}

%% \begin{table*}
%% \caption{Repeat length of longer allele by age of onset class}
%% \begin{tabular}{ccc}
%% table text
%% \end{tabular}
%% \end{table*}

\end{document}

### 2010-01-05 manuscript/supervenience.tex
initial commit

--- 2010-01-05 manuscript/PNAS/PNAS_submission.tex
@@
-%% PNAStmpl.tex
-%% Template file to use for PNAS articles prepared in LaTeX
-%% Version: Apr 14, 2008
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BASIC CLASS FILE 
-%% PNAStwo for two column articles is called by default.
-%% Uncomment PNASone for single column articles. One column class
-%% and style files are available upon request from pnas@nas.edu.
-%% (uncomment means get rid of the '%' in front of the command)
-
-%\documentclass{pnasone}
-\documentclass{pnastwo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Changing position of text on physical page:
-%% Since not all printers position
-%% the printed page in the same place on the physical page,
-%% you can change the position yourself here, if you need to:
-
-% \advance\voffset -.5in % Minus dimension will raise the printed page on the 
-                         %  physical page; positive dimension will lower it.
-
-%% You may set the dimension to the size that you need.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL GRAPHICS STYLE FILE
-
-%% Requires graphics style file (graphicx.sty), used for inserting
-%% .eps files into LaTeX articles.
-%% Note that inclusion of .eps files is for your reference only;
-%% when submitting to PNAS please submit figures separately.
-
-%% Type into the square brackets the name of the driver program 
-%% that you are using. If you don't know, try dvips, which is the
-%% most common PC driver, or textures for the Mac. These are the options:
-
-% [dvips], [xdvi], [dvipdf], [dvipdfm], [dvipdfmx], [pdftex], [dvipsone],
-% [dviwindo], [emtex], [dviwin], [pctexps], [pctexwin], [pctexhp], [pctex32],
-% [truetex], [tcidvi], [vtex], [oztex], [textures], [xetex]
-
-%\usepackage[dvips]{graphicx}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL POSTSCRIPT FONT FILES
-
-%% PostScript font files: You may need to edit the PNASoneF.sty
-%% or PNAStwoF.sty file to make the font names match those on your system. 
-%% Alternatively, you can leave the font style file commands commented out
-%% and typeset your article using the default Computer Modern 
-%% fonts (recommended). If accepted, your article will be typeset
-%% at PNAS using PostScript fonts.
-
-% Choose PNASoneF for one column; PNAStwoF for two column:
-%\usepackage{PNASoneF}
-%\usepackage{PNAStwoF}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-
-%% The AMS math files are commonly used to gain access to useful features
-%% like extended math fonts and math commands.
-
-\usepackage{amssymb,amsfonts,amsmath}
-
+%\input{/Users/joshyv/Research/misc/latex_paper.tex}
+\documentclass{article}
+\usepackage{fancyhdr}
+\usepackage{hyperref}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{amsthm}
@@
 \usepackage{verbatim}
+% \usepackage{ulem}
+% \usepackage{color}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO FILES
-%% Insert self-defined macros here.
-%% \newcommand definitions are recommended; \def definitions are supported
+% \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
+% \newcommand{\conv}{\rightarrow}
+% \newcommand{\Real}{\mathbb{R}}
+% \providecommand{\tr}[1]{\textcolor{red}{#1}}
 
-%\newcommand{\mfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2}}
-%\def\s{\sigma}
 \newcommand{\mB}{\mathcal{B}}
@@
 \newtheorem{thex}{Thought experiment}
+\lhead{Vogelstein JT, et al}
+\rhead{Neurocognitive Graph Theory}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2008}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+\title{Are mental properties supervenient on brain properties?}
+
+\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^{1,2}$, Carey E. Priebe$^1$\\
+$^1$Department of Applied Mathematics and Sciences, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% For titles, only capitalize the first letter
-%% \title{Almost sharp fronts for the surface quasi-geostrophic equation}
-
-\title{Are mental properties supervenient on brain properties?}
-
-%% Enter authors via the \author command.  
-%% Use \affil to define affiliations.
-%% (Leave no spaces between author name and \affil command)
-
-%% Note that the \thanks{} command has been disabled in favor of
-%% a generic, reserved space for PNAS publication footnotes.
-
-%% \author{<author name>
-%% \affil{<number>}{<Institution>}} One number for each institution.
-%% The same number should be used for authors that
-%% are affiliated with the same institution, after the first time
-%% only the number is needed, ie, \affil{number}{text}, \affil{number}{}
-%% Then, before last author ...
-%% \and
-%% \author{<author name>
-%% \affil{<number>}{}}
-
-%% For example, assuming Garcia and Sonnery are both affiliated with
-%% Universidad de Murcia:
-%% \author{Roberta Graff\affil{1}{University of Cambridge, Cambridge,
-%% United Kingdom},
-%% Javier de Ruiz Garcia\affil{2}{Universidad de Murcia, Bioquimica y Biologia
-%% Molecular, Murcia, Spain}, \and Franklin Sonnery\affil{2}{}}
-
-\author{Joshua T. Vogelstein\affil{1}{Johns Hopkins University, Baltimore, MD}, 
-R. Jacob Vogelstein\affil{1}{}, 
-Carey E. Priebe\affil{1}{}}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
-
-%% The \maketitle command is necessary to build the title page.
 \maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{article}
-
+%\tableofcontents
 \begin{abstract}
@@
 
-%% When adding keywords, separate each term with a straight line: |
-\keywords{graph theory | universal consistency | neural circuit}
-
-%% Optional for entering abbreviations, separate the abbreviation from
-%% its definition with a comma, separate each pair with a semicolon:
-%% for example:
-%% \abbreviations{SAM, self-assembled monolayer; OTS,
-%% octadecyltrichlorosilane}
-
-% \abbreviations{}
-
-%% The first letter of the article should be drop cap: \dropcap{}
-%\dropcap{I}n this article we study the evolution of ''almost-sharp'' fronts
-
-%% Enter the text of your article beginning here and ending before
-%% \begin{acknowledgements}
-%% Section head commands for your reference:
-%% \section{}
-%% \subsection{}
-%% \subsubsection{}
+\vspace*{0.5 in}
 
@@
 
-\begin{acknowledgments}
-The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.
-\end{acknowledgments}
+\paragraph{Acknowledgments}
+The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
+% \end{acknowledgments}
 
-%% PNAS does not support submission of supporting .tex files such as BibTeX.
-%% Instead all references must be included in the article .tex document. 
-%% If you currently use BibTeX, your bibliography is formed because the 
-%% command \verb+\bibliography{}+ brings the <filename>.bbl file into your
-%% .tex document. To conform to PNAS requirements, copy the reference listings
-%% from your .bbl file and add them to the article .tex file, using the
-%% bibliography environment described above.
-
-%%  Contact pnas@nas.edu if you need assistance with your
-%%  bibliography.
-
-% Sample bibliography item in PNAS format:
-%% \bibitem{in-text reference} comma-separated author names up to 5,
-%% for more than 5 authors use first author last name et al. (year published)
-%% article title  {\it Journal Name} volume #: start page-end page.
-%% ie,
-% \bibitem{Neuhaus} Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) 
-% A short C-terminal sequence is necessary and sufficient for the
-% targeting of chitinases to the plant vacuole. 
-% {\it Proc Natl Acad Sci USA} 88:10362-10366.
-
-%% Enter the largest bibliography number in the facing curly brackets
-%% following \begin{thebibliography}
-
-\begin{thebibliography}{26}
-
-\bibitem{Plato97}
-Plato.
-\newblock {\em Plato: complete works}.
-\newblock Hackett Pub Co,  (1997).
-
-\bibitem{Descartes1641}
-Descartes, R.
-\newblock {\em Meditationes de prima philosophia}.
-\newblock  (1641).
-
-\bibitem{Davidson70}
-Davidson, D.
-\newblock {\em Experience and Theory}, chapter Mental Events.
-\newblock Duckworth (1970).
-
-\bibitem{SpornsKotter05}
-Sporns, O., Tononi, G., and Kotter, R.
-\newblock {\em PLoS Computational Biology}{ \bf 1}(4), e42 (2005).
-
-\bibitem{LichtmanSanes08}
-Lichtman, J.~W., Livet, J., and Sanes, J.~R.
-\newblock {\em Nat Rev Neurosci}{ \bf 9}(6), 417--422 Jun  (2008).
-
-\bibitem{Seung09}
-Seung, H.
-\newblock {\em Neuron}{ \bf 62}(1), 17--29 (2009).
-
-\bibitem{Kim95}
-Kim, J.
-\newblock {\em Phiilosophy of Mind}.
-\newblock Westview Press, second edition,  (2005).
-
-\bibitem{DGL96}
-Devroye, L., Gy{\"o}rfi, L., and Lugosi, G.
-\newblock {\em {A Probabilistic Theory of Pattern Recognition}}.
-\newblock Springer,  (1996).
-
-\bibitem{Devroye83}
-Devroye, L.
-\newblock {\em Probability Theory and Related Fields}{ \bf 62}(4), 475--483
-  (1983).
-
-\bibitem{Popper}
-Popper, K.
-\newblock { \bf } (1959).
-
-\bibitem{WhiteBrenner86}
-White, J., Southgate, E., Thomson, J.~N., and Brenner, S.
-\newblock {\em Philosophical Transactions of Royal Society London. Series B,
-  Biological Sciences}{ \bf 314}(1165), 1--340 (1986).
-
-\bibitem{DenkHorstmann04}
-W.~Denk, W. and Horstmann, H.
-\newblock {\em PLOS Biol.}{ \bf 2}, e329 (2004).
-
-\bibitem{BriggmanDenk06}
-Briggman, K. and Denk, W.
-\newblock {\em Current opinion in neurobiology}{ \bf 16}(5), 562--570 (2006).
-
-\bibitem{MackeBorst08}
-Macke, J.~H., Maack, N., Gupta, R., Denk, W., Schölkopf, B., and Borst, A.
-\newblock {\em J Neurosci Methods}{ \bf 167}(2), 349--357 Jan  (2008).
-
-\bibitem{Mishchenko09}
-Mishchenko, Y.
-\newblock {\em J Neurosci Methods}{ \bf 176}(2), 276--289 Jan  (2009).
-
-\bibitem{LuLichtman09}
-Lu, J., Fiala, J.~C., and Lichtman, J.~W.
-\newblock {\em PLoS ONE}{ \bf 4}(5), e5655 05  (2009).
-
-\bibitem{HornJohnson90}
-Horn, R. and Johnson, C.
-\newblock {\em {Matrix analysis}}.
-\newblock Cambridge Univ Pr,  (1990).
-
-\bibitem{ConroyLouis97}
-Conroy, J.~M., Kratzer, S.~G., and Podrazik, L.~J.
-\newblock In {\em Society for Industrial and Applied Mathematics},  (1997).
-
-\bibitem{ZaslavskiyVert08}
-Zaslavskiy, M., Bach, F., and Vert, J.
-\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}{
-  \bf 31}, 2227--2242 (2008).
-
-\bibitem{Durbin87}
-Durbin, R.~M.
-\newblock {\em Studies on the Development and Organisation of the Nervous
-  System of Caenorhabditis elegans}.
-\newblock PhD thesis, University of Cambridge,  (1987).
-
-\bibitem{deBonoMaricq05}
-de~Bono, M. and Maricq, A.~V.
-\newblock {\em Annu Rev Neurosci}{ \bf 28}, 451--501 (2005).
-
-\bibitem{VarshneyChklovskii09}
-Varshney, L., Chen, B., Paniagua, E., Hall, D., and Chklovskii, D.
-\newblock {\em ArXiv}{ \bf } (2009).
-
-\bibitem{ChalasaniBargmann07}
-Chalasani, S.~H., Chronis, N., Tsunozaki, M., Gray, J.~M., Ramot, D., Goodman,
-  M.~B., and Bargmann, C.~I.
-\newblock {\em Nature}{ \bf 450}(7166), 63--70 Nov  (2007).
-
-\bibitem{VaziriShank08}
-Vaziri, A., Tang, J., Shroff, H., and Shank, C.~V.
-\newblock {\em Proc Natl Acad Sci U S A}{ \bf 105}(51), 20221--20226 Dec
-  (2008).
-
-\bibitem{HelmstaedterDenk08}
-Helmstaedter, M., Briggman, K.~L., and Denk, W.
-\newblock {\em Curr Opin Neurobiol}{ \bf 18}(6), 633--641 Dec  (2008).
-
-\bibitem{BuckinghamSattelle08}
-Buckingham, S.~D. and Sattelle, D.~B.
-\newblock {\em Invert Neurosci}{ \bf 8}(3), 121--131 Sep  (2008).
-
-\end{thebibliography}
-
-\end{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Adding Figure and Table References
-%% Be sure to add figures and tables after \end{article}
-%% and before \end{document}
-
-%% For figures, put the caption below the illustration.
-%%
-%% \begin{figure}
-%% \caption{Almost Sharp Front}\label{afoto}
-%% \end{figure}
+%\subparagraph*{References}
+\bibliography{biblist}
+\addcontentsline{toc}{section}{References}
+%\bibliography{Science}
+\bibliographystyle{unsrt}
+% \bibliographystyle{nature}
 
 \begin{figure}[h!]
-% \centering 
-\includegraphics[width=.5\linewidth]{Lhatplot}
+\centering \includegraphics[width=.9\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.
@@
 
-%% For Tables, put caption above table
-%%
-%% Table caption should start with a capital letter, continue with lower case
-%% and not have a period at the end
-%% Using @{\vrule height ?? depth ?? width0pt} in the tabular preamble will
-%% keep that much space between every line in the table.
-
-%% \begin{table}
-%% \caption{Repeat length of longer allele by age of onset class}
-%% \begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}lrcccc}
-%% table text
-%% \end{tabular}
-%% \end{table}
-
-%% For two column figures and tables, use the following:
-
-%% \begin{figure*}
-%% \caption{Almost Sharp Front}\label{afoto}
-%% \end{figure*}
-
-%% \begin{table*}
-%% \caption{Repeat length of longer allele by age of onset class}
-%% \begin{tabular}{ccc}
-%% table text
-%% \end{tabular}
-%% \end{table*}
-
 \end{document}

### 2010-01-05 code/Untitled 2.R

@@ -1,0 +1,1 @@
+ass=1:10

### 2010-01-05 code/Untitled2.R

@@ -1,0 +1,8 @@
+twosam <- function(y1, y2) { 
+n1 <- length(y1); n2 <- length(y2) 
+yb1 <- mean(y1); yb2 <- mean(y2) 
+s1 <- var(y1); s2 <- var(y2) 
+s <- ((n1-1)*s1 + (n2-1)*s2)/(n1+n2-2) 
+tst <- (yb1 - yb2)/sqrt(s*(1/n1 + 1/n2)) 
+tst 
+} 

### 2010-01-05 code/sim.R

@@ -1,0 +1,211 @@
+
+library(MASS)
+library(lpSolve)
+
+getAk = function(n,p,m,q,labeled=T)
+ ### generates adjacency matrix A for K(n,p,m,q)
+ {
+ A=matrix(sample(0:1,n^2,T,c(1-p,p)),nrow=n,ncol=n)
+ if(labeled)  thism = 1:m
+ else         thism = sample(1:n,m,F)
+ A[thism,thism]=sample(0:1,m^2,T,c(1-q,q))
+ A=A*upper.tri(A)
+ A=A+t(A)
+ return(A)
+ }
+
+Lstar = function( n=20,p=0.5,m=8,q=0.9 , q0vec ,pi0=0.5)
+ ### class-conditional distributions are K(n,p,m,q0) and K(n,p,m,q)
+ ### pi0 = 0.5  => class priors pi0=pi1=1/2
+ ### Bayes optimal error rates L* ... assuming my math & coding are correct ?!?
+ {
+ pi1 = 1-pi0
+ nc2 = choose(n,2)
+ mc2 = choose(m,2)
+ q1=q
+ LstarL = LstarU = rep(0,length(q0vec))
+ for(i in 1:length(q0vec))
+  {
+  q0=q0vec[i]
+  for(e in 0:nc2)
+   {
+   for(eA in 0:min(mc2,e))
+    {
+    x = choose(mc2,eA)*choose(nc2-mc2,e-eA)
+    thing0 = pi0*q0^eA*(1-q0)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
+    thing1 = pi1*q1^eA*(1-q1)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
+    minj = thing0 < thing1
+    LstarL[i] = LstarL[i] + x*thing0*minj + x*(1-minj)*thing1
+    }
+   ttt0=ttt1=0
+   for(eA in 0:min(mc2,e))
+    {
+    x = choose(mc2,eA)*choose(nc2-mc2,e-eA)
+    ttt0 = ttt0 + x*pi0*q0^eA*(1-q0)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
+    ttt1 = ttt1 + x*pi1*q1^eA*(1-q1)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
+    }
+   minj = ttt0 < ttt1
[…]

### 2010-01-05 code/twosam.R

@@ -1,0 +1,10 @@
+## ass
+
+twosam <- function(y1, y2) { 
+n1 <- length(y1); n2 <- length(y2) 
+yb1 <- mean(y1); yb2 <- mean(y2) 
+s1 <- var(y1); s2 <- var(y2) 
+s <- ((n1-1)*s1 + (n2-1)*s2)/(n1+n2-2) 
+tst <- (yb1 - yb2)/sqrt(s*(1/n1 + 1/n2)) 
+tst 
+} 

### 2010-03-13 manuscript/edelman/cover_letter.tex
modified draft somewhat

% request for removal from telemarketing lists

\documentclass[10pt]{letter}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=5.9in
\topmargin=-.5in
\textheight=9in

\name{	Joshua T.\ Vogelstein \\ 
		Johns Hopkins University \\
		Ph.D.\ Candidate in the Dept.\ of Neuro. \\
		% Postdoc.\ at Nat'l Center of Applied Neuro. \\
		% and the Department of Applied Mathematics and Statistics \\
		3400 N.\ Charles St.\ \\
		Krieger Hall, Rm.\ 338 \\
		Baltimore, MD 21205 \\
      	(443) 858-9911}

\date{\today}

\begin{document}

\begin{letter}{	Prof.\ Gerald Edelman, M.D.\, Ph.D.\ \\
               	Chairman,The Scripps Research Institute \\
				10550 N.\ Torrey Pines Rd.\ \\
               	La Jolla, CA  92037-1000}

\opening{Dear Professor Edelman:}

On behalf of myself and my coauthors, I submit for your review a copy of our
manuscript entitled, ``Are mental properties supervenient on brain properties?''
We hope that you will consider ``communicating'' this manuscript to the
prestigious Proceedings of the National Academy of Sciences (PNAS).

At the core of this paper is a thought experiment in which we mathematically
prove that if the graph corresponding to brain connectivity (i.e., the
\emph{connectome}) is statistically related to a particular mental property
(e.g., propensity for mathematics), one could build a classifier to determine
whether any individual's brain exhibits that property with an arbitrarily small
misclassification rate.  The proof follows directly from the probabilistic
theory of pattern recognition and is made possible by our novel exposition of
universal consistency for a $k_n$-nearest neighbor classifer on graphs.  In
addition to these theoretical results, the manuscript describes a simulation
that suggests how one could actually build such a classifier for the ``brain''
of \emph{Caenorhabditis elegans} using today's technology.

Our results have many implications, some philosophical, some statistical, and
some pragmatic.  First, the philosophical implications:  The relationship
between the mind and brain has been investigated for millennia, using tools
ranging from abstract arguments to microstimulation of individual neurons.
However, until now, no one (to our knowledge) has proposed a framework that
allows for empirical investigation of this relationship at the scale of the
connectome.  Such a framework will undoubtedly be of value as the multiple
ongoing large-scale efforts to collect a connectome (cf.\ NIH Human Connectome
Project) begin paying dividends.

%Rather, it seems the typical approach is to first \emph{assume} such
%a relationship exists, and then find the \emph{location} of that relationship
%(to make a crude analogy, much of contemporary neuroscience seems to have taken
%a lesson from Decartes, who, assuming a relationship between an immaterial soul,
%and the body, found that the pineal gland was the location).

In addition to the philosophical implications, our work has significant
statistical implications.  Whereas several groups have recently proposed
algorithms for classifying graphs, there exists very little theoretical work on
the limitations of these tools.  In contrast, our work is founded on an
existance proof of a universally consistent classifier on graphs, the
limitations of which are both known and well-understood.  Although we only apply
this classifier to simulated brain-derived graphs, the methods are applicable to
many other kinds of graphical data. The recent emergence of networks and graphs 
as the preferred representations of structured relationships in multiple
scientific disciplines suggests that the development of statistical tools for
rigorous analysis of network data is a pressing need for the scientific
community at large.  
%Furthermore, we show using the arbitrarily slow convergence
%theorems, that only a one-sided result is possible, i.e., while we can be
%$100(1-\alpha)\%$ confident of the relationship between the properties under
%investigation, we can \emph{never} be confident in its negation.  This result
%is widely applicable outside the investigation of mind-brain supervenience.

Finally, the results described in this manuscript have pragmatic implications.
Studies relating brain structure to brain function are being published with
increasing frequency due to the increasing availability of diffusion-weighted
magnetic resonance imaging and other connectivity-imaging technologies.  However,
without sophisticated statistical tools for the analysis of such data, to-date
the findings have (mostly) been limited to linear correlations between scalar
measures of connectivity and the brain function under investigation.  The
graph-theoretic approach proposed here allows for utilizing the entirety of the
information obtained from such measures, rather than using simple descriptive
statistics.

In summary, we believe that the work described in this manuscript makes
significant contributions in several contemporary fields of inquiry, and is
therefore suited for a general-audience peer-reviewed journal such as PNAS.  Your
interest in the philosophical underpinnings of neuroscience, as well as in
inferring causal relationships from neural data and generating large-scale
simulations of neural populations suggests to us that this manuscript would of
interest to you.  Please do not hesitate to contact us with any questions or
comments.

\closing{Respectfully,}

\end{letter}

\end{document}

### 2010-03-13 manuscript/supervenience.tex
modified draft somewhat

@@
 \usepackage{verbatim}
+\usepackage{color}
 % \usepackage{ulem}
@@
 \newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}
+\providecommand{\tr}[1]{\textcolor{red}{#1}}
 
@@
 The ``mind-brain supervenience'' theorem suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
-The validity of this theorem has been argued in philosphical terms for over 2,500 years, but it has not previously been approachable through experimental investigation.
+The validity of this theorem has been argued in philosophical terms for over 2,500 years.
+\tr{The alternative hypotheses, including various non-physical causes of mental properties, seems rather implausible to many.
+That said, testing these hypotheses } has not previously been approachable through experimental investigation.
 To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
@@
 Here we define new versions of supervenience that formulate the theorem in rigorous mathematical terms and that can be experimentally tested.
+
+\tr{The primary contributions of this work are as follows.
+First, a notion of supervenience emendable to empirical investigation is formally introduced.
+This renders the mind-brain dualism debate a hypothesis, rather than an assumption.
+Second, the limits of hypothesis testing to unravel supervenience hypotheses are provided.
+Third, we posit a very general model of brains and their associated mental properties that admits a universally consistent classifier; that is, we prove that assuming the brain can be accurately represented by this model, and there is sufficient data, this algorithm is guaranteed to obtain the best possible results.
+Fourth, simulations show this algorithm converges relatively rapidly.}
 
@@
 
-Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$.
-Therefore, we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
+\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$.
+Therefore,} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
 
@@
 
-Now, generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \rightarrow \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;(b_1,m_1),\cdots,(b_n,m_n))$.
-Let $L_{F}(g_n) = E[P_{F}[g_n(B;(B_1,M_1),\cdots,(B_n,M_n)) \neq M|(B_1,M_1),\cdots,(B_n,M_n)]]$.
+\tr{Unfortunately, in general $F$ is unknown, but can be estimated from the data.
+Therefore,} generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \rightarrow \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $\tr{\vec{d}_n=}(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;\tr{\vec{d}_n})$.
+Let $L_{F}(g_n) = E[P_{F}[g_n(B;\tr{\vec{D}_n}) \neq M|\tr{\vec{D}_n}]]$.
 
@@
 For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation.
-Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-Given these caveats, consider the following thought experiment:
+Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.\footnote{\tr{The strictly must be there because it is possible for the null to be rejected, given the data has a particular distribution, it is simply not necessarily rejectable.
+This is a subtle point that was lost on me until very recently.}} Given these caveats, consider the following thought experiment:
 
@@
 
-Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts).
+Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts).
 Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable.
@@
 Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.
+\tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that ``hope'' of empirically investigating supervenience is possible.}
 
@@
 
-Assume first that all graphs are simple, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another.
+Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another.
 Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$.
@@
 
-Several points of note: Isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.
+Several points of note: isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.
 
@@
 To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$.
-The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ k_{ij})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $k_{ij}$ uniformly sampled from $[-5,5]$.
+The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $\tr{z_{ij}}$ uniformly sampled from $[-5,5]$.
 We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm.
@@
 %\bibliography{Science}
-\bibliographystyle{unsrt}
+\bibliographystyle{ieeetr}
 % \bibliographystyle{nature}
@@
 $\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$.
-Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$.
-E.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
+Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
 We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.

### 2010-03-13 code/Untitled 2.R

@@ -1,1 +1,0 @@
-ass=1:10

### 2010-03-13 code/Untitled2.R

@@ -1,8 +1,0 @@
-twosam <- function(y1, y2) { 
-n1 <- length(y1); n2 <- length(y2) 
-yb1 <- mean(y1); yb2 <- mean(y2) 
-s1 <- var(y1); s2 <- var(y2) 
-s <- ((n1-1)*s1 + (n2-1)*s2)/(n1+n2-2) 
-tst <- (yb1 - yb2)/sqrt(s*(1/n1 + 1/n2)) 
-tst 
-} 

### 2010-03-13 code/sim.R

@@ -1,211 +1,0 @@
-
-library(MASS)
-library(lpSolve)
-
-getAk = function(n,p,m,q,labeled=T)
- ### generates adjacency matrix A for K(n,p,m,q)
- {
- A=matrix(sample(0:1,n^2,T,c(1-p,p)),nrow=n,ncol=n)
- if(labeled)  thism = 1:m
- else         thism = sample(1:n,m,F)
- A[thism,thism]=sample(0:1,m^2,T,c(1-q,q))
- A=A*upper.tri(A)
- A=A+t(A)
- return(A)
- }
-
-Lstar = function( n=20,p=0.5,m=8,q=0.9 , q0vec ,pi0=0.5)
- ### class-conditional distributions are K(n,p,m,q0) and K(n,p,m,q)
- ### pi0 = 0.5  => class priors pi0=pi1=1/2
- ### Bayes optimal error rates L* ... assuming my math & coding are correct ?!?
- {
- pi1 = 1-pi0
- nc2 = choose(n,2)
- mc2 = choose(m,2)
- q1=q
- LstarL = LstarU = rep(0,length(q0vec))
- for(i in 1:length(q0vec))
-  {
-  q0=q0vec[i]
-  for(e in 0:nc2)
-   {
-   for(eA in 0:min(mc2,e))
-    {
-    x = choose(mc2,eA)*choose(nc2-mc2,e-eA)
-    thing0 = pi0*q0^eA*(1-q0)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
-    thing1 = pi1*q1^eA*(1-q1)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
-    minj = thing0 < thing1
-    LstarL[i] = LstarL[i] + x*thing0*minj + x*(1-minj)*thing1
-    }
-   ttt0=ttt1=0
-   for(eA in 0:min(mc2,e))
-    {
-    x = choose(mc2,eA)*choose(nc2-mc2,e-eA)
-    ttt0 = ttt0 + x*pi0*q0^eA*(1-q0)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
-    ttt1 = ttt1 + x*pi1*q1^eA*(1-q1)^(mc2-eA)*p^(e-eA)*(1-p)^(nc2-mc2-(e-eA))
-    }
-   minj = ttt0 < ttt1
[…]

### 2010-03-13 code/twosam.R

@@ -1,10 +1,0 @@
-## ass
-
-twosam <- function(y1, y2) { 
-n1 <- length(y1); n2 <- length(y2) 
-yb1 <- mean(y1); yb2 <- mean(y2) 
-s1 <- var(y1); s2 <- var(y2) 
-s <- ((n1-1)*s1 + (n2-1)*s2)/(n1+n2-2) 
-tst <- (yb1 - yb2)/sqrt(s*(1/n1 + 1/n2)) 
-tst 
-} 

### 2010-03-13 manuscript/supervenience.tex
modified draft somewhat

@@
 
-The ``mind-brain supervenience'' theorem suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
-The validity of this theorem has been argued in philosophical terms for over 2,500 years.
-\tr{The alternative hypotheses, including various non-physical causes of mental properties, seems rather implausible to many.
-That said, testing these hypotheses } has not previously been approachable through experimental investigation.
-To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
+The ``mind-brain supervenience'' \tr{conjecture} suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
+The validity of this \tr{conjecture} has been argued in philosophical terms for over 2,500 years.
+\tr{The alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many.
+Proving or disproving these conjectures, however, has remained elusive.}\footnote{\tr{i took out this stuff:} has not previously been approachable through experimental investigation.} To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
@@
 Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
-Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience theorem, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}.
+Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience \tr{conjecture}, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}.
 Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation.
-Here we define new versions of supervenience that formulate the theorem in rigorous mathematical terms and that can be experimentally tested.
+Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms and that can be experimentally tested \tr{as a hypothesis}.
 

### 2010-03-24 manuscript/supervenience.tex
incorporated jacobs comments

@@
 This renders the mind-brain dualism debate a hypothesis, rather than an assumption.
-Second, the limits of hypothesis testing to unravel supervenience hypotheses are provided.
-Third, we posit a very general model of brains and their associated mental properties that admits a universally consistent classifier; that is, we prove that assuming the brain can be accurately represented by this model, and there is sufficient data, this algorithm is guaranteed to obtain the best possible results.
-Fourth, simulations show this algorithm converges relatively rapidly.}
+Second, in addition to expanding the space of questions amenable to hypothesis testing, we also demonstrate the the limits of hypothesis testing.
+Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.
+Fourth, we prove that this formulation admits a universally consistent classifier (a result of general interest to graph classification, independent of the subject of the graphs).
+Fifth we demonstrate through simulation that the proposed universally consistent classifer has reasonable convergence properties on simulated data.}
 
@@
 
-\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$.
-Therefore,} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
+\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{\tr{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$.
+Therefore,}} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
 
@@
 Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.
-\tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that ``hope'' of empirically investigating supervenience is possible.}
+\tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that empirically investigating supervenience is at least theoretically possible.}
 

### 2010-03-29 manuscript/supervenience.tex
finished jacobs edits, now to cep

@@
 The validity of this \tr{conjecture} has been argued in philosophical terms for over 2,500 years.
-\tr{The alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many.
-Proving or disproving these conjectures, however, has remained elusive.}\footnote{\tr{i took out this stuff:} has not previously been approachable through experimental investigation.} To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
+\tr{Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.}\footnote{\tr{i took out this stuff:} has not previously been approachable through experimental investigation.} To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
@@
 \tr{The primary contributions of this work are as follows.
-First, a notion of supervenience emendable to empirical investigation is formally introduced.
+First, a notion of supervenience amenable to empirical investigation is formally introduced.
 This renders the mind-brain dualism debate a hypothesis, rather than an assumption.
@@
 Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.
-Fourth, we prove that this formulation admits a universally consistent classifier (a result of general interest to graph classification, independent of the subject of the graphs).
-Fifth we demonstrate through simulation that the proposed universally consistent classifer has reasonable convergence properties on simulated data.}
+Fourth, we prove that this formulation admits a universally consistent classifier that is guaranteed to find the relationship between minds and brains, if one exists.
+Fifth we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.}
 

### 2010-08-17 PLoS_ONE/resubmission/supervenience.tex
got reviews back from plos one, rearranged a bit and started making edits

%\input{/Users/joshyv/Research/misc/latex_paper.tex}
\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{times}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{color}
\pagestyle{fancy}

\oddsidemargin=0.0in %%this makes the odd side margin go to the default of 1inch
\evensidemargin=0.0in
\textwidth=6.5in
\headwidth=6.5in
\textheight=9in %%sets the textwidth to 6.5, which leaves 1 for the remaining right margin with 8 1/2X11inch paper
\headheight=12pt
\topmargin=-0.25in
%\headheight=0in
%\headsep=0in
%\pagestyle{headings}

\usepackage{hyperref}
% \usepackage{ulem}
% \usepackage{color}

% \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
\newcommand{\conv}{\rightarrow}
% \newcommand{\Real}{\mathbb{R}}
% \providecommand{\tr}[1]{\textcolor{red}{#1}}

\newcommand{\mB}{\mathcal{B}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mM}{\mathcal{M}}
\newcommand{\PP}{\mathbb{P}}           % probability
\newcommand{\EE}{\mathbb{E}}           % expected value
\newcommand{\II}{\mathbb{I}}           % expected value
\newcommand{\Real}{\mathbb{R}}           % expected value

\newcommand{\del}{\delta}
\newcommand{\sig}{\sigma}
\newcommand{\lam}{\lambda}
\newcommand{\gam}{\gamma}
\newcommand{\eps}{\varepsilon}

\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\mb}[1]{\boldsymbol{#1}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\mv}[1]{\vec{#1}}
\providecommand{\mh}[1]{\widehat{#1}}
\providecommand{\mt}[1]{\widetilde{#1}}
\providecommand{\mhc}[1]{\hat{\mathcal{#1}}}
\providecommand{\mhb}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\mvb}[1]{\vec{\boldsymbol{#1}}}
\providecommand{\mtb}[1]{\widetilde{\boldsymbol{#1}}}

\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

% \newcommand{\mN}{\mathcal{N}}

\newcommand{\hL}{\widehat{L}}
\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_F \mB}
\newcommand{\MsB}{\mM \overset{S}{\sim}_F \mB}
\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}
\providecommand{\tr}[1]{\textcolor{black}{#1}}
\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{thex}{Thought experiment}
\lhead{Vogelstein JT, et al}
\rhead{Neurocognitive Graph Theory}

\title{Are mental properties supervenient on brain properties?}

\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^{1,2}$, Carey E. Priebe$^1$\\
$^1$Department of Applied Mathematics and Sciences, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}

\begin{document}

\maketitle
%\tableofcontents
\begin{abstract}

	The ``mind-brain supervenience'' \tr{conjecture} suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain. The validity of this \tr{conjecture} has been argued in philosophical terms for over 2,500 years. \tr{Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.}

	To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis. To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience. The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two. In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.

	The potential significance of this work can be divided into distinct disciplines.  To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.  To the statistician, herein lies the first proof to our knowledge of the existence of a universally consistent classifier on graphs, and a constructivist one at that.  To the neuroscientist, a theoretically possible experiment is proposed to garnish support for a hypothesis that is widely believed: that mental properties supervene on brain properties.

\end{abstract}

\vspace*{0.5 in}

\section{Introduction}

Questioning the relationship between the mind (thoughts, beliefs, preferences, emotions, intelligence, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}. Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}. Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience \tr{conjecture}, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}. Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation. Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms and that can be experimentally tested \tr{as a hypothesis}.

\tr{The primary contributions of this work are as follows.  First, a notion of supervenience amenable to empirical investigation is formally introduced.  This renders the mind-brain dualism debate a hypothesis, rather than an assumption.  Second, in addition to expanding the space of questions amenable to hypothesis testing, we also demonstrate the limits of hypothesis testing.  Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.  Fourth, we prove that this formulation admits a universally consistent classifier that is guaranteed to find the relationship between minds and brains, if one exists.  Fifth we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.}

\section{Preliminaries} % (fold)
\label{sec:preliminaries}

The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.  We therefore first define the objects of interest, that is, minds and brains.

Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.  The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.  In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.  Thus, each different $b \in \mB$ corresponds to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.

Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.  The set of possible minds $\mM$ is also unrestricted. In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences. Thus, each different $m \in \mM$ corresponds to a difference in at least one thought, belief, or preference.

The mind-brain supervenience conjecture, is a relation between these two sets, the set of mental states and the set of brain states.  Donald Davidson canonized this conjecture in 1970 with the following quote: \cite{Davidson70}
\begin{quotation}
\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect,
\end{quotation}
can be concisely and formally stated: $m \neq m' \implies b \neq b' \, \forall (b,m), (b',m') \in \mB \times \mM$.  While mind-brain supervenience is a relatively strong claim, importantly, one can imagine far stronger relations, such as the following.

First, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.  An injective relation is any relation that preserves distinctness.  Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (b,m), (b',m') \in \mB \times \mM$ (note that the directionality of the implication has been switched relative to supervenience). For instance, it might be the case that a brain could change without the mind changing.  Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.  It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.  In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective. This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.  Minds supervening on brains does not imply that brains supervene on minds.

% In this case, it would be said that both $M$ supervenes on $B$ and $B$ supervenes on $M$. But it is not necessarily the case that if supervenience holds in one direction that it does or does not hold in the other direction.

Second, it may be the case that minds supervene on brains, but that brains do not cause minds.  For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.  Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other. This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.

% Supervenience is also not \emph{causality}.  Consider the coin flip example.  Nowhere did the notion of causality even arise.  It could be the case that $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.  Thus, supervenience does not imply causality.

Third, supervenience does not imply \emph{identity}.  Consider, for example, acceleration and velocity.  Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).  Similarly, velocity supervenes on position, as velocity cannot change without position changing.  Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.  Rather, position can change with constant velocity, meaning without acceleration changing.

What supervenience does imply, however, is the following.   Imagine finding two different minds.  If $\MeB$, then the brains subvening under those two minds must be different.  In other words, there cannot be two different minds, either of which could supervene on a single brain.  Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
	\caption{Possible relations between minds and brains.  (A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.  (B) Minds supervene on brains, and it so happens that there is a surjective relation from brains to minds. (C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
	\label{fig:rel}
\end{figure}

All of the above relations are \emph{logical} relations, not probabilistic relations. To facilitate both statistical analysis and empirical investigation, we project this supervenience notion into a statistical setting.  To proceed, we first define a \emph{model}, $\PP[B,M]=F_{BM}$, which specifies the probability of any element $(b,m)$ occurring from the space of all possible elements $\mB \times \mM$ (also called the \emph{sample space}).  The model could in theory come from any possible joint distribution defined on the sample space, $F_{BM} \in \mc{F}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
\begin{enumerate}
	\item $0 \leq \PP[B=b,M=m] \leq 1$ for all $(b,m) \in \mM \times \mB$
	\item $\PP[\Omega]=1$ and $\PP[\emptyset]=0$, where $\Omega$ is the whole sample sample, that is, all elements, $(b,m) \in \mB \times \mM$
	\item any countable sequence of pairwise disjoint elements, $(b_1,m_1), (b_2,m_2), \ldots$ satisfies \\
	\mbox{$\PP[(b_1,m_1) \cup (b_2,m_2) \cup \ldots] = \sum_i \PP[(b_i,m_i)]$}.
\end{enumerate}
Given a model, one can then calculate any marginal or conditional distributions as a function of the model.  For instance, $\PP[B]=\int_{m \in \mM} \PP[B,M]\text{d}m$, or $\PP[M|B]=\PP[B,M]/\PP[B]$.  Given the definition of a model, statistical supervenience can be defined as follows:
\begin{defi}
\label{def1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0 \, \forall (b,m), (b',m') \in \mB \times \mM$. Alternately, $\MeB$ can be written as $\PP[m = m' | b = b']=1  \, \forall (b,m), (b',m') \in \mB \times \mM$.
\end{defi}
Statistical supervenience is therefore a probabilistic relation on sets.  Note that statistical supervenience is distinct from statistical correlation.  \emph{Statistical correlation} between brain states and mental states is defined as $\rho_{BM}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.  If $\rho_{BM}=1$, then both $\MeB$ and $\mB \overset{S}{\sim}_F \mM$. Thus, correlation implies supervenience, but supervenience does not imply correlation.

 % Consider the following analogy: one desires to understand the relationship between height and weight.  In such a scenario, height, $h$, could be assumed to be an element from the set of all possible heights, $\mc{H}=(0,\infty)$ cm; and weight, $w$, could be assumed to be an element from the set of all possible weights, $\mc{W}=(0,\infty)$ kgs.   Then, it may be assumed that an agent's height and weight is sampled identically and independently from some \emph{model}, that is $(h,w)\overset{iid}{\sim}\PP[H,W]=F_{HW}$. To gain insight into $F_{HW}$, one might then state hypotheses, collect data, and perform analyses.  Implicit in the exposition of a model is that both objects, height and weight here, are \emph{random objects}.  Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.

% section preliminaries (end)

\section{Results}

\subsection{Theoretical results} % (fold)
\label{sub:theoretical_results}

% subsection theoretical_results (end)
If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.  This means that there exists a unique mapping from each brain to a single mind.  In other words, one can in principle construct a function $g(b): \mB \mapsto \mM$, that is a deterministic mapping from brains to minds.  It may be the case that subsets of brains from equivalence classes, such that any brain in that subset is mapped to the same mind (see, for example, $b_1$ and $b_2$ in Figure \ref{fig:rel}(A)).  Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).  Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.  Define misclassification rate as:
\begin{align}
	L_F(g) = P_{F}[g(B) \neq M] = \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} \text{d}b \text{d}m
\end{align}
where $\II\{\cdot\}$ indicates the indicator function, taking value one if its argument is true, and zero otherwise.  $L_F(g)$ therefore effectively counts the fraction of time $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is
\begin{align}
	g^* = \argmin_{g \in \mc{G}} L_F(g)
\end{align}
where $\mc{G}$ is the set of all possible classifiers.  Thus, the \emph{Bayes error}, or Bayes risk, $L_F(g^*)$ is the minimum possible misclassification rate. The primary result of casting supervenience is a statistical framework is the following theorem: 
\begin{thm}
\label{thm1} 
$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{F}(g^*) = 0$.
\end{thm}
% \begin{thm} 
% 	\label{thm1} 
% 	Statistical supervenience implies that there exists a classifier that is \emph{always correct}, that is, $g(b)=m$, for all $(b,m) \in \mB \times \mM$.  
% \end{thm}

\noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.  This means that one could draw a decision boundary between all equivalence classes of brains, each class corresponding to a different mind, and no mind will reside within two different equivalence classes.  Thus, the optimal classifier would correctly find these decision boundaries, and therefore have no opportunity to err. $\square$

This relationship between statistical supervenience and Bayes error can therefore be described concisely: \mbox{$\MsB \Leftrightarrow L_F(g^*)=0$}.  Thus, the above arguments shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent. Further, statistical supervenience can be thought of as a constraint on the possible models.  Specifically, let $\mc{F}_s \subset \mc{F}$ be subset of models for which supervenience holds.  Then, $\mc{F}_s = \{F | L_F(g^*)=0\} \subset \mc{F}$.

\subsection{Hypothesis testing} % (fold)
\label{sub:hypothesis_testing}

While the above theorem is of potential theoretical interest, because the arguments rest on knowing $F_{BM}$ and $g^*$, which are typically unknown, they are pragmatically useless.  However, both $F_{BM}$ and $g^*$ could be estimated from data.  Let $(b_1,m_1), (b_2,m_2), \ldots, (b_n,m_n)$ be random samples taking their values in $\mB \times \mM$, independently and identically distributed according to model $F_{BM}$.  Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mB \times \mM)^n \mapsto \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $\tr{\mc{T}_n=}\{(b_1,m_1),\cdots,(b_n,m_n)\}$ and produces a classification $g_n(b;\tr{\mc{T}_n})=\mh{m}$.  Misclassification rate for this classifier will therefore be a random variable, because the training data $\mc{T}_n$ are random samples.  Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
\begin{align}
\EE[L_{F}(g_n)] = \EE[P_{F}[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP_F[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
\end{align}
Unfortunately, in practice, computing $\EE[L_F(g_n)]$, requires integrating over all possible training data corpuses, and by definition, we only have access to a single training data corpus.  We therefore define ``hold out'' misclassification performance:
\begin{align}
\EE[L_F(g_n)] %= \int \PP_F[g_n(B)=M | \mc{T}_{n'-n}] \PP[\mc{T}_{n'-n}] d\mc{T}_{n'n} \nonumber \\ 
&\approx  \hL^{n'}_{F}(g_n) = \sum_{\mc{T}_{n-n'}} \PP_F[g_n(B)=M | \mc{T}_{n-n'}] \PP[\mc{T}_{n-n'}],
\end{align}
where $n'<n$ is the number of held-out training samples, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.   $n' \hL^{n'}_{F}(g_n)$ is the expected number of misclassified minds.  Simplifying further, assuming no sum, rather, just one ``hold-out'' set, then $n' \hL^{n'}_{F}(g_n)$ and has a binomial distribution, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect, thus $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{F}(g_n))$.

Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
\begin{defi}
\label{def2}
Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
\end{defi}

Given this relaxation, consider the problem of testing for $\eps-$supervenience.  First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.  Because we hope to reject the null, in favor of the alternative, let the null hypothesis be $H_0$: $L_{F}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{F}(g_n) < \eps$.  We reject for low values of the test-statistic, $n' \hL^{n'}_{F}(g_n)$.  Specifically, if $n' \hL^{n'}_{F}(g_n)$ is less than the critical value,  $c_{\alpha}(n',\varepsilon)$, then we reject.  The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.  Thus, if $n' \hL^{n'}_{F}(g_n) < c_{\alpha}(n',\varepsilon)$, we can conclude with $100(1-\alpha$)\% confidence that $\MeB$.  The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$. Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{F}_\eps = \{F | L_F(g^*)< \eps\}$.  One could then sort  $\eps$-supervenience subsets,  $\mc{F}_s \subseteq \mc{F}_{\eps} \subseteq \mc{F}_{\eps'} \subseteq \mc{F}$, for any $\eps < \eps'$.

% subsection hypothesis_testing (end)

\subsection{Power and consistency} % (fold)
\label{ssub:power_and_consistency}

Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis (in other words, the probability that it will not make a Type II error).  Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.  A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.  A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_F(g_n)] \conv L_F(g^*)$ as $n\conv \infty$. As the notation suggests, consistency of a classifier is a function of the true model, $F$.  Without any prior knowledge of what the model might be, one desires a \emph{universally consistent} classifier, that is a classifier that is consistent for all $F \in \mc{F}$.

Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{BM}$ \cite{DGL96}. Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}. For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation.  This means that we can never be confident that $\MeB$ does \emph{not} hold; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data. Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.  Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}. Given these limitations on even universal consistency, it is still the best one can hope for.  Therefore, we hope to obtain a universally consistent classifier to test for $\eps$-supervenience.

In 1977, Stone proved that a certain collection of $k_n$ nearest neighbor algorithms is universally consistent, assuming $b$ was a random $d$-dimensional vector $\mc{B} \subseteq \Real^d$ \cite{Stone77} (see Appendix \ref{app:knn} for an explanation of $k_n$ nearest neighbor algorithms, and constraints to ensure universal consistency) (since then, other algorithms have also been shown to be universally consistent, most notably neural networks \cite{Cybenko89}).  This beautiful result would be directly applicable to mind-brain supervenience, if we were satisfied representing brains by random vectors.  However, brains are rich with structure, and therefore, other, more sophisticated data structures can better capture aspects of the brain that many of us believe to be true. Therefore, in the next section, we introduce the concept of a \emph{brain-graph}, which can have a much richer structure than a vector.  We then generalize the $k_n$ nearest neighbor algorithm to operate in this domain, and extend the universal consistency proof as well.

% subsubsection power_and_consistency (end)

\subsection{Brain-graphs} % (fold)
\label{sub:brain_graphs}

In 1891, H Waldeyer-Hartz first formally proposed the ``neuron doctrine'' \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex \emph{network} of ``neurons'' (a term invented in the above review), largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.  This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.  In parallel with the development and refinement of concepts of how collections of neurons may collectively operate to subserve various brain functions, a simultaneous development has taken place in the statistical analysis of networks, or graphs \cite{Bollobas, otherbook}.  We therefore propose to represent the brain as a random graph, where nodes represent neurons and edges represent synapses.  Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $\{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neurons, and arcs (or edges) $\{A_{ij}\}$, where $A_{ij}$ represents the synapse from neuron $V_j$ to $V_i$. For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that (i) there are no self loops (no autapses, or self-connections), (ii) all edges are binary, meaning that $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric, that is, $a_{ij}=a_{ji}$.  Further assuming that the vertices are \emph{labeled}, meaning that there is a known one-to-one mapping

from vertex $V_i$ in

Formally, assume that each brains, $b \in \mc{B}$

% subsection brain_graphs (end)

\section{Old results} % (fold)
\label{sec:old_results}

% (fold)
% Let $\mB$ be the observation space for some physical property, such as brain connectivity structure (i.e., connectome; see \cite{SpornsKotter05,LichtmanSanes08,Seung09}). Let $\mM$ be the (finite) indicator space for some mental property, such as knowing calculus. Thus, for $b \in \mB$ and $m \in \mM$, the pair $(b,m)$ represents a brain property/mind property pair.
% 
% %Let $\{(B_i,M_i)\}_{i=1}^n \overset{iid}{\sim} F_{BM}$
% Let $(B,M), (B_1,M_1),\cdots,(B_n,M_n)$ be random observation pairs taking their values in $\mB \times \mM$, independently and identically distributed according to some joint probability distribution $F=F_{BM}$. Abusing notation to conceptually identify the properties with their spaces, the statistical supervenience relation $\mM \overset{S}{\sim}_F \mB$ says that $M_i \neq M_j \implies B_i \neq B_j$ (almost surely; where $\implies$ does not suggest causation). That is, observing $B=b$ can allow us to assign $m$ to $M$. While previously proposed notions of mind-brain-supervenience claim that all mental properties supervene on physical properties \cite{Kim95}, here we consider empirically investigating only whether a particular mental property $\mM$ statistically supervenes on a particular physical property $\mB$.
% 
% Let $g:\mB \mapsto \mM$ be a classifier, which takes as input an observed brain connectivity structure $b$ and produces a classification $\widehat{m}=g(b)$ for the unobserved mental property $m$. The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, where $L_{F}(g) = P_{F}[g(B) \neq M]$ denotes the probability of misclassification for classifier $g$ under joint distribution $F=F_{BM}$. We can therefore rigorously define \textit{statistical supervenience}:
% 
% \begin{defi}
% \label{def1} 
% $\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{F}(g^*) = 0$.
% \end{defi}
% 
% \tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,}  we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
% 
% \begin{defi}
% \label{def2}
% Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
% \end{defi}
% 
% \tr{Unfortunately, in general $F$ is unknown, but can be estimated from the data. Therefore,} generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \mapsto \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $\tr{\vec{d}_n=}(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;\tr{\vec{d}_n})$. Let $L_{F}(g_n) = E[P_{F}[g_n(B;\tr{\vec{D}_n}) \neq M|\tr{\vec{D}_n}]]$.
% 
% Consider the problem of testing for $\varepsilon$-supervenience. Let the null hypothesis be given by $H_0: L_{F}(g_n) \geq \varepsilon$ so that if we reject at level $\alpha >0$ in favor of the alternative hypothesis $H_A: L_{F}(g_n) < \varepsilon$ then we can conclude, with $100(1-\alpha$)\% confidence, that $\MeB$. Letting $\hL^{n'}_{F}(g_n)$ denote the hold-out estimate of misclassification performance based on $n'$ test observations, we note that $\hL^{n'}_{F}(g_n)$ is distributed $\text{Binomial}(n',L_{F}(g_n))$.
% 
% % section old_results
% 
% The test rejects for small $\hL^{n'}_{F}(g_n)$. The level $\alpha$ critical value $c_{\alpha}(n',\varepsilon)$ is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$. Furthermore, $\MeB$ implies $L_{F}(g^*) < \varepsilon$, and thus if $g_n$ is a {\it consistent} classifier for $F=F_{BM}$ --- that is, if $\lim_n L_{F}(g_n) = L_{F}(g^*)$ --- then the power of this test (the probability of rejecting when in fact the alternative is true) goes to unity as $n,n' \rightarrow \infty$. Thus we have an inference procedure: 
% 
% \begin{thm} 
% 	\label{thm1} 
% 	Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$. Furthermore, given a consistent classifier the power of the test converges to unity. 
% \end{thm}
% 
% Since the joint distribution $F=F_{BM}$ is unknown, the utility of Theorem \ref{thm1} requires that $g_n$ be a {\it universally consistent} classifier --- that is, $\lim_n L_{F}(g_n) = L_{F}(g^*)$ for all distributions $F=F_{BM}$. Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{BM}$ \cite{DGL96}. Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}. For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation. Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}. Given these caveats, consider the following thought experiment:
% (end)

\begin{thex}
Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts). Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable. Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators. A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 1 for proof). Therefore, Theorem \ref{thm1} applies and the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds. This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes. Furthermore, Appendix 1 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.  
\end{thex}

\section{Discussion}

While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality. Assuming we have confirmed $\MeB$ for a particular mental/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.

Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm (or other norm; see Appendix 1 for details) may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities. Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment). Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure. \tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that empirically investigating supervenience is at least theoretically possible.}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
% \section{Materials and Methods}

\appendix
\section{$k_n$ nearest neighbor algorithm} % (fold)
\label{app:knn}

Assume that $b$ is a real $d$-dimensional vector, $b \in \mc{B} \subseteq \Real^d$, and $m$ is a binary indicator, $m \in \mc{M} = \{0,1\}$.  Then, further assume that we have observed a collection of training data, $\mc{T}_n =\{(b_i,m_i)\}_{i=1}^n$, each sampled identically and independently from some unknown joint distribution, $(b_i,m_i)\overset{iid}{\sim} F_{BM}$.  A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.  The $1$-nearest neighbor classifier works as follows.  Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.  Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, where parenthetical indices, $(i)$, indicate rank order.  One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.  The $1$ nearest neighbor algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $m=m_{(1)}$.  The $k_n$ nearest neighbor is a straightforward generalization of this approach.  It says that the test mind is in the same class as which ever class is the majority class of the $k_n$ nearest neighbors, $m=\II\{\sum_{i=1}^{k_n} m_{(i)} > k_n/2\}$.  Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.

Unfortunately, no such algorithm is universally consistent.  Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training points.  Then, a collection of such algorithms, $\{g_n\}$,  with $k_n$ increasing with $n$, can be universally consistent under certain constraints. In particular, as $n$ increases, $k_n$ must also increase, but not quite as quickly.  Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$. In Stone' original proof, the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.  Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.

This algorithm also readily generalizes to the $C$-way classification problem, where there $C \geq 2$ classes.  Instead of finding the majority class of the $k_n$ nearest neighbors, one instead finds the plurality class, that is, the class with the most training samples within the $k_n$ nearest neighbors.  This algorithm generalizes even further, when $\mc{M}$ is no longer discrete, but rather continuous, for instance, $\mc{M}=\Real$.  In such a scenario, this approach is called $k_n$ nearest neighbor \emph{regression}.

% section _k_n_nearest_neighbor_algorithm (end)

\section{$k_n$-nearest neighbor universal consistency for graphs}
\label{proof}

Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another. Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$. If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$. Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$. Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.

In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability. Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $\PP[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{F}(g_n) \rightarrow c \leq L_{F}(g^*) + \delta$, yielding universal consistency.

If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm. Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$. Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}. Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
%\tr{Trivially, this argument also applies to graphs with (self-) loops, and
%directed, colored, or weighted edges, and hypergraphs.}

Several points of note:  isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.

Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg. 455 for proof when $b_i$'s are vectors).  Furthermore, we can relax the constraint that mental properties are finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).  The proof for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.

%\appendix
\section{Simulation} \label{simulation}

As an example of a feasible experiment, one may consider a species whose nervous system consists of the same (small) number of labeled neurons for each organism. {\it Caenorhabditis elegans} is believed to be such a species \cite{Durbin87}. The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons. While the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}. Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}. Thus, one may design an experiment by describing the joint distribution $F_{BM}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$. Here the mental property corresponds to the C.~elegans exhibiting or not exhibiting a particular behavior (e.g., response to an odor).

Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $F_{BM}$ (Figure \ref{fig1}). To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$. The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $\tr{z_{ij}}$ uniformly sampled from $[-5,5]$. We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm. The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency. (Better classifiers can be constructed for the joint distribution $F_{BM}$ used here; however, we demand universal consistency.)

\begin{figure}[!ht]
\centering \includegraphics[width=.5\linewidth]{Lhatplot}
\caption{C.~elegans graph classification simulation results. $\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$. Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01. We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$. $L_{F}(g^*) \approx 0$ for this simulation.
}
\label{fig1}
\end{figure}

Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations. 3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day. Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.

% Do NOT remove this, even if you are not including acknowledgments
\section{Acknowledgments}

The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.

% \section{References}
% The bibtex filename
\bibliography{biblist}
\bibliographystyle{ieeetr}

% \clearpage
% \section{Figure Legends}
% 
% Figure 1: C.~elegans graph classification simulation results. $\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$. Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01. We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$. $L_{F}(g^*) \approx 0$ for this simulation.
%

\end{document}

### 2010-08-17 PLoS_ONE/resubmission/supervenience_old.tex
got reviews back from plos one, rearranged a bit and started making edits

--- 2010-08-17 PLoS_ONE/resubmission/supervenience.tex
@@
 where $n'<n$ is the number of held-out training samples, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.
-$n' \hL^{n'}_{F}(g_n)$ is the expected number of misclassified minds.
-Simplifying further, assuming no sum, rather, just one ``hold-out'' set, then $n' \hL^{n'}_{F}(g_n)$ and has a binomial distribution, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect, thus $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{F}(g_n))$.
+$n' \hL^{n'}_{F}(g_n)$ is the expected number of misclassified minds, and has a binomial distribution, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect, thus $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{F}(g_n))$.
 
@@
 
-Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis (in other words, the probability that it will not make a Type II error).
+Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value, and its power, the probability that the test will reject a false null hypothesis (in other words, the probability that it will not make a Type II error).
 Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
@@
 
-In 1977, Stone proved that a certain collection of $k_n$ nearest neighbor algorithms is universally consistent, assuming $b$ was a random $d$-dimensional vector $\mc{B} \subseteq \Real^d$ \cite{Stone77} (see Appendix \ref{app:knn} for an explanation of $k_n$ nearest neighbor algorithms, and constraints to ensure universal consistency) (since then, other algorithms have also been shown to be universally consistent, most notably neural networks \cite{Cybenko89}).
+In 1977, Stone proved that a certain collection of $k_n$ nearest neighbor algorithms is universally consistent, assuming $b$ was a random $d$-dimensional vector $\mc{B} \subseteq \Real^d$ \cite{Stone77} (see Appendix \ref{app:knn} for an explanation of $k_n$ nearest neighbor algorithms, and constraints to ensure universal consistency).
 This beautiful result would be directly applicable to mind-brain supervenience, if we were satisfied representing brains by random vectors.
@@
 In 1891, H Waldeyer-Hartz first formally proposed the ``neuron doctrine'' \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex \emph{network} of ``neurons'' (a term invented in the above review), largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
-This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
-In parallel with the development and refinement of concepts of how collections of neurons may collectively operate to subserve various brain functions, a simultaneous development has taken place in the statistical analysis of networks, or graphs \cite{Bollobas, otherbook}.
-We therefore propose to represent the brain as a random graph, where nodes represent neurons and edges represent synapses.
-Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $\{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neurons, and arcs (or edges) $\{A_{ij}\}$, where $A_{ij}$ represents the synapse from neuron $V_j$ to $V_i$.
-For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that (i) there are no self loops (no autapses, or self-connections), (ii) all edges are binary, meaning that $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric, that is, $a_{ij}=a_{ji}$.
-Further assuming that the vertices are \emph{labeled}, meaning that there is a known one-to-one mapping
-
-from vertex $V_i$ in
-
-Formally, assume that each brains, $b \in \mc{B}$
+This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years.
 

### 2010-08-17 PNAS/PNAS_submission.tex
got reviews back from plos one, rearranged a bit and started making edits

--- 2010-01-05 manuscript/PNAS/PNAS_submission.tex
@@
 \author{Joshua T. Vogelstein\affil{1}{Johns Hopkins University, Baltimore, MD}, 
-R. Jacob Vogelstein\affil{1}{}, 
+R. Jacob Vogelstein\affil{1}{}\affil{2}{Johns Hopkins University Applied Physics Laboratory, Laurel, MD}, 
 Carey E. Priebe\affil{1}{}}
@@
 \begin{acknowledgments}
-The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.
+The authors would like to acknowledge helpful discussions with J Lande, A Vaziri, and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
 \end{acknowledgments}
@@
 
-\begin{thebibliography}{26}
+\begin{thebibliography}{10}
 
 \bibitem{Plato97}
-Plato.
-\newblock {\em Plato: complete works}.
-\newblock Hackett Pub Co,  (1997).
+Plato, {\em Plato: complete works}.
+\newblock Hackett Pub Co, 1997.
 
 \bibitem{Descartes1641}
-Descartes, R.
-\newblock {\em Meditationes de prima philosophia}.
-\newblock  (1641).
+R.~Descartes, {\em Meditationes de prima philosophia}.
+\newblock 1641.
 
 \bibitem{Davidson70}
-Davidson, D.
-\newblock {\em Experience and Theory}, chapter Mental Events.
-\newblock Duckworth (1970).
+D.~Davidson, {\em Experience and Theory}, ch.~Mental Events.
+\newblock Duckworth, 1970.
 
 \bibitem{SpornsKotter05}
-Sporns, O., Tononi, G., and Kotter, R.
-\newblock {\em PLoS Computational Biology}{ \bf 1}(4), e42 (2005).
+O.~Sporns, G.~Tononi, and R.~Kotter, ``The human connectome: A structural
+  description of the human brain,'' {\em PLoS Computational Biology}, vol.~1,
+  no.~4, p.~e42, 2005.
 
 \bibitem{LichtmanSanes08}
-Lichtman, J.~W., Livet, J., and Sanes, J.~R.
-\newblock {\em Nat Rev Neurosci}{ \bf 9}(6), 417--422 Jun  (2008).
+J.~W. Lichtman, J.~Livet, and J.~R. Sanes, ``A technicolour approach to the
+  connectome.,'' {\em Nat Rev Neurosci}, vol.~9, pp.~417--422, Jun 2008.
 
 \bibitem{Seung09}
-Seung, H.
-\newblock {\em Neuron}{ \bf 62}(1), 17--29 (2009).
+H.~Seung, ``{Reading the Book of Memory: Sparse Sampling versus Dense Mapping
+  of Connectomes},'' {\em Neuron}, vol.~62, no.~1, pp.~17--29, 2009.
 
 \bibitem{Kim95}
-Kim, J.
-\newblock {\em Phiilosophy of Mind}.
-\newblock Westview Press, second edition,  (2005).
+J.~Kim, {\em Phiilosophy of Mind}.
+\newblock Westview Press, second~ed., 2005.
 
 \bibitem{DGL96}
-Devroye, L., Gy{\"o}rfi, L., and Lugosi, G.
-\newblock {\em {A Probabilistic Theory of Pattern Recognition}}.
-\newblock Springer,  (1996).
+L.~Devroye, L.~Gy{\"o}rfi, and G.~Lugosi, {\em {A Probabilistic Theory of
+  Pattern Recognition}}.
+\newblock Springer, 1996.
 
 \bibitem{Devroye83}
-Devroye, L.
-\newblock {\em Probability Theory and Related Fields}{ \bf 62}(4), 475--483
-  (1983).
+L.~Devroye, ``{On arbitrarily slow rates of global convergence in density
+  estimation},'' {\em Probability Theory and Related Fields}, vol.~62, no.~4,
+  pp.~475--483, 1983.
 
 \bibitem{Popper}
-Popper, K.
-\newblock { \bf } (1959).
+K.~Popper, ``{The logic of scientific discovery},'' 1959.
 
 \bibitem{WhiteBrenner86}
-White, J., Southgate, E., Thomson, J.~N., and Brenner, S.
-\newblock {\em Philosophical Transactions of Royal Society London. Series B,
-  Biological Sciences}{ \bf 314}(1165), 1--340 (1986).
+J.~White, E.~Southgate, J.~N. Thomson, and S.~Brenner, ``The structure of the
+  nervous system of the nematode caenorhabditis elegans.,'' {\em Philosophical
+  Transactions of Royal Society London. Series B, Biological Sciences},
+  vol.~314, no.~1165, pp.~1--340, 1986.
 
 \bibitem{DenkHorstmann04}
-W.~Denk, W. and Horstmann, H.
-\newblock {\em PLOS Biol.}{ \bf 2}, e329 (2004).
+W.~W.~Denk and H.~Horstmann, ``Serial block-face scanning electron microscopy
+  to reconstruct three-dimensional tissue nanostructure,'' {\em PLOS Biol.},
+  vol.~2, p.~e329, 2004.
 
 \bibitem{BriggmanDenk06}
-Briggman, K. and Denk, W.
-\newblock {\em Current opinion in neurobiology}{ \bf 16}(5), 562--570 (2006).
+K.~Briggman and W.~Denk, ``{Towards neural circuit reconstruction with volume
+  electron microscopy techniques},'' {\em Current opinion in neurobiology},
+  vol.~16, no.~5, pp.~562--570, 2006.
 
 \bibitem{MackeBorst08}
-Macke, J.~H., Maack, N., Gupta, R., Denk, W., Schölkopf, B., and Borst, A.
-\newblock {\em J Neurosci Methods}{ \bf 167}(2), 349--357 Jan  (2008).
+J.~H. Macke, N.~Maack, R.~Gupta, W.~Denk, B.~Schölkopf, and A.~Borst,
+  ``Contour-propagation algorithms for semi-automated reconstruction of neural
+  processes.,'' {\em J Neurosci Methods}, vol.~167, pp.~349--357, Jan 2008.
 
 \bibitem{Mishchenko09}
-Mishchenko, Y.
-\newblock {\em J Neurosci Methods}{ \bf 176}(2), 276--289 Jan  (2009).
+Y.~Mishchenko, ``Automation of 3d reconstruction of neural tissue from large
+  volume of conventional serial section transmission electron micrographs.,''
+  {\em J Neurosci Methods}, vol.~176, pp.~276--289, Jan 2009.
 
 \bibitem{LuLichtman09}
-Lu, J., Fiala, J.~C., and Lichtman, J.~W.
-\newblock {\em PLoS ONE}{ \bf 4}(5), e5655 05  (2009).
+J.~Lu, J.~C. Fiala, and J.~W. Lichtman, ``Semi-automated reconstruction of
+  neural processes from large numbers of fluorescence images,'' {\em PLoS ONE},
+  vol.~4, p.~e5655, 05 2009.
 
 \bibitem{HornJohnson90}
-Horn, R. and Johnson, C.
-\newblock {\em {Matrix analysis}}.
-\newblock Cambridge Univ Pr,  (1990).
+R.~Horn and C.~Johnson, {\em {Matrix analysis}}.
+\newblock Cambridge Univ Pr, 1990.
 
 \bibitem{ConroyLouis97}
-Conroy, J.~M., Kratzer, S.~G., and Podrazik, L.~J.
-\newblock In {\em Society for Industrial and Applied Mathematics},  (1997).
+J.~M. Conroy, S.~G. Kratzer, and L.~J. Podrazik, ``A continuous method for the
+  quadratic assignment problem,'' in {\em Society for Industrial and Applied
+  Mathematics}, 1997.
 
 \bibitem{ZaslavskiyVert08}
-Zaslavskiy, M., Bach, F., and Vert, J.
-\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}{
-  \bf 31}, 2227--2242 (2008).
+M.~Zaslavskiy, F.~Bach, and J.~Vert, ``{A path following algorithm for the
+  graph matching problem},'' {\em IEEE Transactions on Pattern Analysis and
+  Machine Intelligence}, vol.~31, pp.~2227--2242, 2008.
 
 \bibitem{Durbin87}
-Durbin, R.~M.
-\newblock {\em Studies on the Development and Organisation of the Nervous
+R.~M. Durbin, {\em Studies on the Development and Organisation of the Nervous
   System of Caenorhabditis elegans}.
-\newblock PhD thesis, University of Cambridge,  (1987).
+\newblock PhD thesis, University of Cambridge, 1987.
 
 \bibitem{deBonoMaricq05}
-de~Bono, M. and Maricq, A.~V.
-\newblock {\em Annu Rev Neurosci}{ \bf 28}, 451--501 (2005).
+M.~de~Bono and A.~V. Maricq, ``Neuronal substrates of complex behaviors in c.
+  elegans.,'' {\em Annu Rev Neurosci}, vol.~28, pp.~451--501, 2005.
 
 \bibitem{VarshneyChklovskii09}
-Varshney, L., Chen, B., Paniagua, E., Hall, D., and Chklovskii, D.
-\newblock {\em ArXiv}{ \bf } (2009).
+L.~Varshney, B.~Chen, E.~Paniagua, D.~Hall, and D.~Chklovskii, ``{Structural
+  Properties of the Caenorhabditis elegans Neuronal Network},'' {\em ArXiv},
+  2009.
 
 \bibitem{ChalasaniBargmann07}
-Chalasani, S.~H., Chronis, N., Tsunozaki, M., Gray, J.~M., Ramot, D., Goodman,
-  M.~B., and Bargmann, C.~I.
-\newblock {\em Nature}{ \bf 450}(7166), 63--70 Nov  (2007).
+S.~H. Chalasani, N.~Chronis, M.~Tsunozaki, J.~M. Gray, D.~Ramot, M.~B. Goodman,
+  and C.~I. Bargmann, ``Dissecting a circuit for olfactory behaviour in
+  caenorhabditis elegans.,'' {\em Nature}, vol.~450, pp.~63--70, Nov 2007.
 
 \bibitem{VaziriShank08}
-Vaziri, A., Tang, J., Shroff, H., and Shank, C.~V.
-\newblock {\em Proc Natl Acad Sci U S A}{ \bf 105}(51), 20221--20226 Dec
-  (2008).
+A.~Vaziri, J.~Tang, H.~Shroff, and C.~V. Shank, ``Multilayer three-dimensional
+  super resolution imaging of thick biological samples.,'' {\em Proc Natl Acad
+  Sci U S A}, vol.~105, pp.~20221--20226, Dec 2008.
 
 \bibitem{HelmstaedterDenk08}
-Helmstaedter, M., Briggman, K.~L., and Denk, W.
-\newblock {\em Curr Opin Neurobiol}{ \bf 18}(6), 633--641 Dec  (2008).
+M.~Helmstaedter, K.~L. Briggman, and W.~Denk, ``3d structural imaging of the
+  brain with photons and electrons.,'' {\em Curr Opin Neurobiol}, vol.~18,
+  pp.~633--641, Dec 2008.
 
 \bibitem{BuckinghamSattelle08}
-Buckingham, S.~D. and Sattelle, D.~B.
-\newblock {\em Invert Neurosci}{ \bf 8}(3), 121--131 Sep  (2008).
+S.~D. Buckingham and D.~B. Sattelle, ``Strategies for automated analysis of c.
+  elegans locomotion.,'' {\em Invert Neurosci}, vol.~8, pp.~121--131, Sep 2008.
 
@@
 % \centering 
-\includegraphics[width=.5\linewidth]{Lhatplot}
+% \includegraphics[width=.5\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.
@@
 We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{F}(g^*) \approx 0$ for this simulation.
-}
+$L_{F}(g^*) \approx 0$ for this simulation.}
 \label{fig1}

### 2010-08-17 PNAS/cover_letter.tex
got reviews back from plos one, rearranged a bit and started making edits

--- 2010-03-13 manuscript/edelman/cover_letter.tex
@@
 \evensidemargin=.2in
-\textwidth=5.9in
-\topmargin=-.5in
-\textheight=9in
+\textwidth=6in
+\topmargin=-1.2in
+\textheight=10in
 
-\name{	Joshua T.\ Vogelstein \\ 
+\name{	Joshua T.\ Vogelstein, PhD. \\ 
 		Johns Hopkins University \\
-		Ph.D.\ Candidate in the Dept.\ of Neuro. \\
-		% Postdoc.\ at Nat'l Center of Applied Neuro. \\
-		% and the Department of Applied Mathematics and Statistics \\
+		% Ph.D.\ Candidate in the Dept.\ of Neuro. \\
+		% Postdoc., 
+		Dept. Applied Math.\ \& Stat's \\
 		3400 N.\ Charles St.\ \\
-		Krieger Hall, Rm.\ 338 \\
-		Baltimore, MD 21205 \\
+		Barton 320 \\
+		Baltimore, MD 21210 \\
       	(443) 858-9911}
@@
 
-\begin{letter}{	Prof.\ Gerald Edelman, M.D.\, Ph.D.\ \\
-               	Chairman,The Scripps Research Institute \\
-				10550 N.\ Torrey Pines Rd.\ \\
-               	La Jolla, CA  92037-1000}
+\begin{letter}{	%Prof.\ Gerald Edelman, M.D.\, Ph.D.\ \\
+               	% Chairman,The Scripps Research Institute \\
+               	% 10550 N.\ Torrey Pines Rd.\ \\
+               	% La Jolla, CA  92037-1000
+}
 
-\opening{Dear Professor Edelman:}
+\opening{Dear Editorial Board:}
 
-On behalf of myself and my coauthors, I submit for your review a copy of our
-manuscript entitled, ``Are mental properties supervenient on brain properties?''
-We hope that you will consider ``communicating'' this manuscript to the
-prestigious Proceedings of the National Academy of Sciences (PNAS).
+On behalf of myself and my coauthors, I submit for your review a copy of our manuscript entitled, ``Are mental properties supervenient on brain properties?'' We hope that you will consider publishing this manuscript to the prestigious Proceedings of the National Academy of Sciences (PNAS).
 
-At the core of this paper is a thought experiment in which we mathematically
-prove that if the graph corresponding to brain connectivity (i.e., the
-\emph{connectome}) is statistically related to a particular mental property
-(e.g., propensity for mathematics), one could build a classifier to determine
-whether any individual's brain exhibits that property with an arbitrarily small
-misclassification rate.  The proof follows directly from the probabilistic
-theory of pattern recognition and is made possible by our novel exposition of
-universal consistency for a $k_n$-nearest neighbor classifer on graphs.  In
-addition to these theoretical results, the manuscript describes a simulation
-that suggests how one could actually build such a classifier for the ``brain''
-of \emph{Caenorhabditis elegans} using today's technology.
+At the core of this paper is a thought experiment in which we mathematically prove that if the graph corresponding to brain connectivity (i.e., the \emph{connectome}) is statistically related to a particular mental property (e.g., propensity for mathematics), one could build a classifier to determine whether any individual's brain exhibits that property with an arbitrarily small misclassification rate.
+The proof follows directly from the probabilistic theory of pattern recognition and is made possible by our novel exposition of universal consistency for a $k_n$-nearest neighbor classifer on graphs.
+In addition to these theoretical results, the manuscript describes a simulation that suggests how one could actually build such a classifier for the ``brain'' of \emph{Caenorhabditis elegans} using today's technology.
 
-Our results have many implications, some philosophical, some statistical, and
-some pragmatic.  First, the philosophical implications:  The relationship
-between the mind and brain has been investigated for millennia, using tools
-ranging from abstract arguments to microstimulation of individual neurons.
-However, until now, no one (to our knowledge) has proposed a framework that
-allows for empirical investigation of this relationship at the scale of the
-connectome.  Such a framework will undoubtedly be of value as the multiple
-ongoing large-scale efforts to collect a connectome (cf.\ NIH Human Connectome
-Project) begin paying dividends.
+Our results have many implications, some philosophical, some statistical, and some pragmatic.
+First, the philosophical implications: The relationship between the mind and brain has been investigated for millennia, using tools ranging from abstract arguments to microstimulation of individual neurons.
+However, until now, no one (to our knowledge) has proposed a framework that allows for empirical investigation of this relationship at the scale of the connectome.
+Such a framework will undoubtedly be of value as the multiple ongoing large-scale efforts to collect a connectome (cf.\ NIH Human Connectome Project) begin paying dividends.
 
-%Rather, it seems the typical approach is to first \emph{assume} such
-%a relationship exists, and then find the \emph{location} of that relationship
-%(to make a crude analogy, much of contemporary neuroscience seems to have taken
-%a lesson from Decartes, who, assuming a relationship between an immaterial soul,
-%and the body, found that the pineal gland was the location).
-
-In addition to the philosophical implications, our work has significant
-statistical implications.  Whereas several groups have recently proposed
-algorithms for classifying graphs, there exists very little theoretical work on
-the limitations of these tools.  In contrast, our work is founded on an
-existance proof of a universally consistent classifier on graphs, the
-limitations of which are both known and well-understood.  Although we only apply
-this classifier to simulated brain-derived graphs, the methods are applicable to
-many other kinds of graphical data. The recent emergence of networks and graphs 
-as the preferred representations of structured relationships in multiple
-scientific disciplines suggests that the development of statistical tools for
-rigorous analysis of network data is a pressing need for the scientific
-community at large.  
-%Furthermore, we show using the arbitrarily slow convergence
-%theorems, that only a one-sided result is possible, i.e., while we can be
-%$100(1-\alpha)\%$ confident of the relationship between the properties under
-%investigation, we can \emph{never} be confident in its negation.  This result
-%is widely applicable outside the investigation of mind-brain supervenience.
+In addition to the philosophical implications, our work has significant statistical implications.
+Whereas several groups have recently proposed algorithms for classifying graphs, there exists very little theoretical work on the limitations of these tools.
+In contrast, our work is founded on an existence proof of a universally consistent classifier on graphs, the limitations of which are both known and well-understood.
+Although we only apply this classifier to simulated brain-derived graphs, the methods are applicable to many other kinds of graphical data.
+The recent emergence of networks and graphs as the preferred representations of structured relationships in multiple scientific disciplines suggests that the development of statistical tools for rigorous analysis of network data is a pressing need for the scientific community at large.
 
 Finally, the results described in this manuscript have pragmatic implications.
-Studies relating brain structure to brain function are being published with
-increasing frequency due to the increasing availability of diffusion-weighted
-magnetic resonance imaging and other connectivity-imaging technologies.  However,
-without sophisticated statistical tools for the analysis of such data, to-date
-the findings have (mostly) been limited to linear correlations between scalar
-measures of connectivity and the brain function under investigation.  The
-graph-theoretic approach proposed here allows for utilizing the entirety of the
-information obtained from such measures, rather than using simple descriptive
-statistics.
+Studies relating brain structure to brain function are being published with increasing frequency due to the increasing availability of diffusion-weighted magnetic resonance imaging and other connectivity-imaging technologies.
+However, without sophisticated statistical tools for the analysis of such data, to-date the findings have (mostly) been limited to linear correlations between scalar measures of connectivity and the brain function under investigation.
+The graph-theoretic approach proposed here allows for utilizing the entirety of the information obtained from such measures, rather than using simple descriptive statistics.
 
-In summary, we believe that the work described in this manuscript makes
-significant contributions in several contemporary fields of inquiry, and is
-therefore suited for a general-audience peer-reviewed journal such as PNAS.  Your
-interest in the philosophical underpinnings of neuroscience, as well as in
-inferring causal relationships from neural data and generating large-scale
-simulations of neural populations suggests to us that this manuscript would of
-interest to you.  Please do not hesitate to contact us with any questions or
-comments.
+In summary, we believe that the work described in this manuscript makes significant contributions in several contemporary fields of inquiry, and is therefore suited for a general-audience peer-reviewed journal such as PNAS.
+%Your interest in the philosophical underpinnings of neuroscience, as well as in inferring causal relationships from neural data and generating large-scale simulations of neural populations suggests to us that this manuscript would of interest to you.
+Please do not hesitate to contact us with any questions or comments.
 

### 2010-08-17 other/discussion.tex
got reviews back from plos one, rearranged a bit and started making edits

\paragraph{Multiple Realizability}

$B \in \mB_i$

\paragraph{delta varepsilon}

$M \overset{\delta,\varepsilon}{\sim} \mB \Leftrightarrow \exists f(\cdot), \; B \in \mB$ s.t. $P[d(f^\ast(B)\neq M)< \delta]< \varepsilon$

\paragraph{superdupervenience}

from stanford encyclopedia:

Supervenience claims, by themselves, do nothing more than state that certain patterns of property (or fact) variation hold. They are silent about why those patterns hold, and about the precise nature of the dependency involved (see Kim 1993, 167; 1998, 9-15; Blackburn 1984, 186; Schiffer 1987, 153-154; and McGinn 1993, 57). But supervenience theses are not plausibly brute, that is, unexplainable. It is natural to look further, and to try to explain why A-properties supervene on B-properties. When such supervenience is explainable, there is ÔsuperdupervenienceÕ (a term coined by William Lycan; see also Schiffer 1987; Horgan 1993; and Wilson 1999).

\newpage
\paragraph{related work}

\paragraph{math}

graph classification papers: \cite{KashimaInokuchi02, KudoMatsumoto04, Emmert-StreibKilian05, PredictingStructuredData07, BunkeRiesen08, SaigoTsuda09, TrentinIorio09, ZaslavskiyVert08, BunkeRiesen08, RiesenBunke09}

T. Gaertner, A survey of kernels for structured data, SIGKDD Exploration Newsletter 5 (1) (2003) 49Ð58.

M. Gori, G. Monfardini, F. Scarselli, A new model for learning in graph domains, in: Proceedings of IJCNN-05, August 2005. in 1990, and his Ph.D. in Automation and Information

A. Sperduti, A. Starita, Supervised neural networks for the classification of structures, IEEE Transactions on Neural Networks 8 (3) (1997) 714-735.

P.A. Flach, N. Lachiche, Naive Bayesian classiﬁcation of structured data, Machine Learning 57 (3) (2004) 233–269.

D.J. Hand, K. Yu, Idiot’s bayes—not so stupid after all?, International Statistical Review 69 (2001) 385–398.

B.J. Jain, P. Geibel, F. Wysotzski.  SVM learning with the Schur-Hadamard inner product for graphs Neurocomputing (2005) 64:93-105

\paragraph{DTI}

book on DTI: \cite{DTIbook09}

higher res technology, DSI: \cite{AlexanderPark01}

q-ball imaging (between DTI and DSI): \cite{Tuch03}

\paragraph{biology}

human connectome: \cite{SpornsKotter05, Seung09}

neuron level methods papers: \cite{DenkHorstmann04, BriggmanDenk06, LuLichtman09}

neuroscience finding papers will depend somewhat on the specific covariates we get, as there are many possible papers here.  the DTI book has a few in it.

anatomy relationship to mental properties: \cite{HutchinsonSchlaug03}

\paragraph{c elegans}

got the graph: \cite{WhiteBrenner86}
looked at ``network statistics'' (like in-degree and out-degree distributions), might be useful for simulations: \cite{ReiglChklovski04}

### 2010-08-17 supervenience.tex
got reviews back from plos one, rearranged a bit and started making edits

--- 2010-03-29 manuscript/supervenience.tex
@@
 \newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}
-\providecommand{\tr}[1]{\textcolor{red}{#1}}
+\providecommand{\tr}[1]{\textcolor{black}{#1}}
+\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
+\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose
 
@@
 The validity of this \tr{conjecture} has been argued in philosophical terms for over 2,500 years.
-\tr{Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.}\footnote{\tr{i took out this stuff:} has not previously been approachable through experimental investigation.} To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
+\tr{Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.}
+
+To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
@@
 In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.
+
+The potential significance of this work can be divided into distinct disciplines.
+To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
+To the statistician, herein lies the first proof to our knowledge of the existence of a universally consistent classifier on graphs, and a constructivist one at that.
+To the neuroscientist, a theoretically possible experiment is proposed to garnish support for a hypothesis that is widely believed: that mental properties supervene on brain properties.
 
@@
 
-\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,} \footnote{\tr{The following was stricken from the record, as I believed it to be a non-sequitor, but i partially replace it below: Unfortunately, it is in general impossible to determine whether $L_F(g^*)=0$ without knowing $F$.
-Therefore,}} we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
+\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,}  we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
 
@@
 For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation.
-Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.\footnote{\tr{The strictly must be there because it is possible for the null to be rejected, given the data has a particular distribution, it is simply not necessarily rejectable.
-This is a subtle point that was lost on me until very recently.}} Given these caveats, consider the following thought experiment:
+Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+Given these caveats, consider the following thought experiment:
 
 \begin{thex}
-
 Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts).
@@
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
-
+Furthermore, Appendix 1 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
 \end{thex}
@@
 
-Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities.
+Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm (or other norm; see Appendix 1 for details) may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities.
 Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment).
@@
 
+Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg.
+455 for proof when $b_i$'s are vectors).
+Furthermore, we can relax the constraint that mental properties finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).
+The proofs for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.
+
 %\appendix
@@
 
-Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day.
-Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
-
-\paragraph{Acknowledgments}
-The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
-% \end{acknowledgments}
-
-%\subparagraph*{References}
-\bibliography{biblist}
-\addcontentsline{toc}{section}{References}
-%\bibliography{Science}
-\bibliographystyle{ieeetr}
-% \bibliographystyle{nature}
-
 \begin{figure}[h!]
@@
 
+Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
+3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day.
+Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
+
+\paragraph{Acknowledgments}
+The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
+% \end{acknowledgments}
+
+%\subparagraph*{References}
+\bibliography{biblist}
+% \addcontentsline{toc}{section}{References}
+%\bibliography{Science}
+% \bibliographystyle{plain}
+\bibliographystyle{nature}
+
 \end{document}

### 2010-08-17 code/getLhat.R

@@ -1,87 +1,0 @@
-
-#  see <http://mit.edu/lrv/www/elegans/>
-#  for labeled graphs of c. elegans (in matlab).
-#  ---cep
-
-# We consider k(n) nrst nbr classification
-# of labeled multigraphs on 279 vertices,
-# under Frobenius norm.
-# Class-conditional distributions F_{G|Y=j}
-# are specified via Poisson parameters for each edge.
-# Class-conditional distribution F_{G|Y=0}
-# is given by "celegans + noise".
-# (noise parameter "eps".)
-# Class-conditional distribution F_{G|Y=1}
-# is given by "celegans + noise + egg".
-# (noise parameter "eps", egg parameters "m" & "sig".)
-# NB: a minor but deterministic class-conditional difference
-# is imposed, so that L^{\star} = 0.
-
-getLhat = function()
-{
-load("celegansGraph.Rd")  ## NB: dim(Ac) => 279 279
-set.seed(12345)
-tstn = 500  ## NB: L=0.1 & 900 test observations => stdev(Lhat^(R)) = 0.01
-maxn = 200  ## maximum number of training observations per class
-## NB: we've conditioned on cc ss n_0=n_1 for both training and testing
-n = 279
-m = 14  ## egg consists of 5% of the vertices ... just because.
-thesem = sample(n,m)  ## NB: egg vertices; same for all mc replicates!
-eps = 0.05
-sig = 5
-E0 = Ac + matrix(runif(n^2,0,eps),nrow=n,ncol=n)
-E1=E0
-egg = matrix(runif(m^2,-sig/2,sig/2),nrow=m,ncol=m)
-E1[thesem,thesem] = E1[thesem,thesem] + egg
-for(i in thesem) for(j in thesem) if(E1[i,j]<0) E1[i,j]=0
-## NB: Frobenius norm => we can use vector representation for graphs!
-Amat = matrix(0,nrow=(2*maxn),ncol=(n*n))  ## training data
-Bmat = matrix(0,nrow=(2*tstn),ncol=(n*n))  ## test data
-for(i in 1:n) for(j in 1:n)
- {
- indexij = i+(j-1)*n
[…]

### 2010-08-17 code/igraph.R

@@ -1,86 +1,0 @@
-
-
-require(igraph)
-
-load("celegansGraph.Rd")
-gg <- graph.adjacency(Ag,mode="undirected")
-gg <- set.graph.attribute(gg,"animal","C. Elegans")
-gg <- set.graph.attribute(gg,"type","Gap Junction")
-
-gg <- set.vertex.attribute(gg,"category",value=c("M","I","S")[vcols])
-gg <- set.vertex.attribute(gg,"type",value=types)
-gg <- set.vertex.attribute(gg,"name",value=neurons)
-
-ggu <- graph.adjacency(Ag>0,mode="undirected")
-ggu <- set.graph.attribute(ggu,"animal","C. Elegans")
-ggu <- set.graph.attribute(ggu,"type","Gap Junction")
-
-ggu <- set.vertex.attribute(ggu,"category",value=c("M","I","S")[vcols])
-ggu <- set.vertex.attribute(ggu,"type",value=types)
-ggu <- set.vertex.attribute(ggu,"name",value=neurons)
-
-gc <- graph.adjacency(Ac>0,mode="directed")
-gc <- set.graph.attribute(gc,"animal","C. Elegans")
-gc <- set.graph.attribute(gc,"type","Chemical Synapse")
-
-gc <- set.vertex.attribute(gc,"category",value=c("M","I","S")[vcols])
-gc <- set.vertex.attribute(gc,"type",value=types)
-gc <- set.vertex.attribute(gc,"name",value=neurons)
-
-gc <- set.edge.attribute(gc,"weight",value=Ac[get.edgelist(gc,names=F)+1])
-
-spg <- shortest.paths(gg)
-spg[is.infinite(spg)] <- 2+max(spg[!is.infinite(spg)])
-spc <- shortest.paths(gc)
-spc[is.infinite(spc)] <- 2+max(spc[!is.infinite(spc)])
-spg <- spg/max(spg)
-spc <- spc/max(spc)
-M <- matrix(0,nrow=2*nrow(Ag),ncol=2*ncol(Ag))
-M[1:279,1:279] <- spg
-M[-(1:279),-(1:279)] <- spc
-A <- (spg+spc)/2
[…]

### 2010-08-17 code/sim3b.R

@@ -1,86 +1,0 @@
-
-# see <http://mit.edu/lrv/www/elegans/> for labeled graphs of c. elegans.
-
-# We consider k(n)-nearest neighbor classification
-# of labeled multigraphs (directed, with loops)
-# on 279 vertices, under Frobenius norm.
-# Class-conditional distributions F_{G|Y=j}
-# are specified via Poisson parameters for each edge.
-# Class-conditional distribution F_{G|Y=0}
-# is given by "celegans + noise".
-# (noise parameter "eps".)
-# Class-conditional distribution F_{G|Y=1}
-# is given by "celegans + noise + egg".
-# (noise parameter "eps", egg parameter "sig".)
-# egg vertices:
-#  based on Chalasni et al.,
-#  the following neurons are involved in odor behaviors:
-#  AWC, AIA, AIB, AIZ, AIY.
-#  each of the above actually has a left neuron and a right neuron,
-#  meaning 10 neurons total.
-#  so, in our sim, the egg is defined by this set of m=10 neurons.
-
-load("celegansGraph.Rd")  ## NB: dim(Ac) => 279 279
-set.seed(12345)
-tstn = 500  ## NB: L=0.1 & 1000 test observations => stdev(Lhat^(R)) < 0.01
-maxn = 200  ## maximum number of training observations per class
-## NB: we condition on class-cond'l sample size n_0=n_1 for both training and testing
-n = 279
-##  m=10 and
-##  AWCL=69 AIBL=80 AWCR=82 AIRB=94 AIAL=110 AIAR=127 AIZR=129 AIZL=133 AIYR=134 AIYL=138
-m=10
-thesem = c(69,80,82,94,110,127,129,133,134,138)
-eps = 0.05  # noise parameter "eps"
-sig = 5     # egg parameter "sig"
-E0 = Ac + matrix(runif(n^2,0,eps),nrow=n,ncol=n)
-E1=E0
-egg = matrix(runif(m^2,-sig/2,sig/2),nrow=m,ncol=m)
-E1[thesem,thesem] = E1[thesem,thesem] + egg
-for(i in thesem) for(j in thesem) if(E1[i,j]<0) E1[i,j]=0
[…]

### 2010-11-17 supervenience.tex
tried to make readable for others

@@
 \documentclass{article}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{amsthm}
+%\usepackage{cite}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{times}
 \usepackage{fancyhdr}
-\usepackage{hyperref}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amsthm}
 \usepackage{graphicx}
@@
 \usepackage{color}
+\pagestyle{fancy}
+
+\oddsidemargin=0.0in %%this makes the odd side margin go to the default of 1inch
+\evensidemargin=0.0in
+\textwidth=6.5in
+\headwidth=6.5in
+\textheight=9in %%sets the textwidth to 6.5, which leaves 1 for the remaining right margin with 8 1/2X11inch paper
+\headheight=12pt
+\topmargin=-0.25in
+%\headheight=0in
+%\headsep=0in
+%\pagestyle{headings}
+
+\usepackage{hyperref}
 % \usepackage{ulem}
@@
 % \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
-% \newcommand{\conv}{\rightarrow}
+\newcommand{\conv}{\rightarrow}
 % \newcommand{\Real}{\mathbb{R}}
@@
 \newcommand{\mM}{\mathcal{M}}
+\newcommand{\PP}{\mathbb{P}}           % probability
+\newcommand{\EE}{\mathbb{E}}           % expected value
+\newcommand{\II}{\mathbb{I}}           % expected value
+\newcommand{\Real}{\mathbb{R}}           % expected value
+
+\newcommand{\del}{\delta}
+\newcommand{\sig}{\sigma}
+\newcommand{\lam}{\lambda}
+\newcommand{\gam}{\gamma}
+\newcommand{\eps}{\varepsilon}
+
+\providecommand{\mc}[1]{\mathcal{#1}}
+\providecommand{\mb}[1]{\boldsymbol{#1}}
+\providecommand{\mbb}[1]{\mathbb{#1}}
+\providecommand{\mv}[1]{\vec{#1}}
+\providecommand{\mh}[1]{\widehat{#1}}
+\providecommand{\mt}[1]{\widetilde{#1}}
+\providecommand{\mhc}[1]{\hat{\mathcal{#1}}}
+\providecommand{\mhb}[1]{\hat{\boldsymbol{#1}}}
+\providecommand{\mvb}[1]{\vec{\boldsymbol{#1}}}
+\providecommand{\mtb}[1]{\widetilde{\boldsymbol{#1}}}
+
+\newcommand{\argmax}{\operatornamewithlimits{argmax}}
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
+
 % \newcommand{\mN}{\mathcal{N}}
@@
 \newcommand{\hL}{\widehat{L}}
-\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_F \mB}
-\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_F \mB}
+\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_{\PP} \mB}
+\newcommand{\MsB}{\mM \overset{S}{\sim}_F \mB}
+\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{\PP} \mB}
 \providecommand{\tr}[1]{\textcolor{black}{#1}}
@@
 \lhead{Vogelstein JT, et al}
-\rhead{Neurocognitive Graph Theory}
+\rhead{Statistical Supervenience}
 
@@
 
-\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^{1,2}$, Carey E. Priebe$^1$\\
-$^1$Department of Applied Mathematics and Sciences, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
+$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723
+}
 
@@
 
-To address the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to scientific methods and statistical analysis.
+To enable addressing the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
 The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two.
-In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.
+%In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.
 
@@
 To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
-To the statistician, herein lies the first proof to our knowledge of the existence of a universally consistent classifier on graphs, and a constructivist one at that.
-To the neuroscientist, a theoretically possible experiment is proposed to garnish support for a hypothesis that is widely believed: that mental properties supervene on brain properties.
+To the statistician, this work points out the limitations of hypothesis testing in a novel domain.
+To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be considered supervenience investigations, further motivating brain-graph related research paradigms.
 
@@
 \vspace*{0.5 in}
+
+\section{Introduction}
 
@@
 
-\tr{The primary contributions of this work are as follows.
-First, a notion of supervenience amenable to empirical investigation is formally introduced.
-This renders the mind-brain dualism debate a hypothesis, rather than an assumption.
-Second, in addition to expanding the space of questions amenable to hypothesis testing, we also demonstrate the the limits of hypothesis testing.
-Third we posit a very general model of brains and their associated mental properties that admits statistical analysis in a graph theoretical and statistical framework.
-Fourth, we prove that this formulation admits a universally consistent classifier that is guaranteed to find the relationship between minds and brains, if one exists.
-Fifth we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.}
+The primary contributions of this work are as follows.
+A notion of supervenience amenable to empirical investigation is formally introduced.
+This renders the mind-brain dualism debate a hypothesis, rather than an assumption, both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
+Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.
+Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists.
+Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.
+This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, suggesting avenues for future research.
 
-Let $\mB$ be the observation space for some physical property, such as brain connectivity structure (i.e., connectome; see \cite{SpornsKotter05,LichtmanSanes08,Seung09}).
-Let $\mM$ be the (finite) indicator space for some mental property, such as knowing calculus.
-Thus, for $b \in \mB$ and $m \in \mM$, the pair $(b,m)$ represents a brain property/mind property pair.
+ % we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.
+Finally, we discuss both contemporary and future applications of this framework to neurobiological studies in humans.
 
-%Let $\{(B_i,M_i)\}_{i=1}^n \overset{iid}{\sim} F_{BM}$
-Let $(B,M), (B_1,M_1),\cdots,(B_n,M_n)$ be random observation pairs taking their values in $\mB \times \mM$, independently and identically distributed according to some joint probability distribution $F=F_{BM}$.
-Abusing notation to conceptually identify the properties with their spaces, the statistical supervenience relation $\mM \overset{S}{\sim}_F \mB$ says that $M_i \neq M_j \implies B_i \neq B_j$ (almost surely; where $\implies$ does not suggest causation).
-That is, observing $B=b$ can allow us to assign $m$ to $M$.
-While previously proposed notions of mind-brain-supervenience claim that all mental properties supervene on physical properties \cite{Kim95}, here we consider empirically investigating only whether a particular mental property $\mM$ statistically supervenes on a particular physical property $\mB$.
+\section{Preliminaries} % (fold)
+\label{sec:preliminaries}
 
-Let $g:\mB \rightarrow \mM$ be a classifier, which takes as input an observed brain connectivity structure $b$ and produces a classification $\widehat{m}=g(b)$ for the unobserved mental property $m$.
-The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, where $L_{F}(g) = P_{F}[g(B) \neq M]$ denotes the probability of misclassification for classifier $g$ under joint distribution $F=F_{BM}$.
-We can therefore rigorously define \textit{statistical supervenience}:
+\subsection{Supervenience} % (fold)
+\label{sub:supervenience}
 
+% subsection supervenience (end)
+The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
+Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}.
+Therefore, here we define minds, brains, and supervenience.
+
+Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
+The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
+In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.
+Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.
+
+Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
+The set of possible minds $\mM$ is also unrestricted.
+In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences.
+Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
+
+The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
+Donald Davidson canonized this conjecture in 1970 with the following quote: \cite{Davidson70}
+\begin{quotation}
+\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
+\end{quotation}
+This conjecture may be concisely and formally stated: $m \neq m' \implies b \neq b' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$, where $(m,b)$ is a mind-brain pair, and $(m',b')$ is a different mind-brain pair.
+While mind-brain supervenience is a relatively strong claim, importantly, one can imagine far stronger relations.
+
+For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.
+An injective relation is any relation that preserves distinctness.
+Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$ (note that the directionality of the implication has been switched relative to supervenience).
+For instance, it might be the case that a brain could change without the mind changing.
+Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
+It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
+In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
+This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.
+Minds supervening on brains does not imply that brains supervene on minds.
+
+% In this case, it would be said that both $M$ supervenes on $B$ and $B$ supervenes on $M$.
+But it is not necessarily the case that if supervenience holds in one direction that it does or does not hold in the other direction.
+
+Second, it may be the case that minds supervene on brains, but that brains do not cause minds.
+For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.
+Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other.
+This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
+
+% Supervenience is also not \emph{causality}.
+Consider the coin flip example.
+Nowhere did the notion of causality even arise.
+It could be the case that $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
+Thus, supervenience does not imply causality.
+
+Third, supervenience does not imply \emph{identity}.
+Consider, for example, acceleration and velocity.
+Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).
+Similarly, velocity supervenes on position, as velocity cannot change without position changing.
+Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
+Rather, position can change with constant velocity, meaning without acceleration changing.
+
+What supervenience does imply, however, is the following.
+Imagine finding two different minds.
+If $\MeB$, then the brains subvening under those two minds must be different.
+In other words, there cannot be two different minds, either of which could supervene on a single brain.
+Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
+	\caption{Possible relations between minds and brains.
+(A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
+(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds.
+(C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
+	\label{fig:rel}
+\end{figure}
+
+All of the above relations are \emph{logical} relations, not probabilistic relations.
+To facilitate both statistical analysis and empirical investigation, we project this supervenience notion into a statistical setting.
+To proceed, we first define a \emph{model}, $\mc{P}=\{\PP[M,B]\}$, the set of all possible joint distributions, $\PP[M,B]$, that could describe the probabilistic relationship between minds and brains.
+Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$ (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
+% \begin{enumerate}
+% 	\item $0 \leq \PP[B=b,M=m] \leq 1$ for all $(m,b) \in \mM \times \mB$
+% 	\item $\PP[\Omega]=1$ and $\PP[\emptyset]=0$, where $\Omega$ is the whole sample sample, that is, all elements, $(m,b) \in \mc{M} \times \mc{B}$
+% 	\item any countable sequence of pairwise disjoint elements, $(m_1,b_1), (m_2,b_2), \ldots$ satisfies \\
+% 	\mbox{$\PP[(m_1,b_1) \cup (m_2,b_2) \cup \ldots] = \sum_i \PP[(m_i,b_i)]$}.
+% \end{enumerate}
+Given a distribution, one can then calculate any marginal or conditional distributions.
+For instance, $\PP[B]=\int_{m \in \mM} \PP[M,B]dm$, or $\PP[M|B]=\PP[M,B]/\PP[B]$.
+Having specified a model, statistical supervenience can be defined as follows:
 \begin{defi}
 \label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{F}(g^*) = 0$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0 \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$.
+Alternately, $\MeB$ can be written as $\PP[m = m' | b = b']=1  \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$.
+\end{defi}
+Statistical supervenience is therefore a probabilistic relation on sets.
+Note that statistical supervenience is distinct from statistical correlation.
+\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
+If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_F \mM$.
+Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+
+ % Consider the following analogy: one desires to understand the relationship between height and weight.
+In such a scenario, height, $h$, could be assumed to be an element from the set of all possible heights, $\mc{H}=(0,\infty)$ cm; and weight, $w$, could be assumed to be an element from the set of all possible weights, $\mc{W}=(0,\infty)$ kgs.
+Then, it may be assumed that an agent's height and weight is sampled identically and independently from some \emph{model}, that is $(h,w)\overset{iid}{\sim}\PP[H,W]=F_{HW}$.
+To gain insight into $F_{HW}$, one might then state hypotheses, collect data, and perform analyses.
+Implicit in the exposition of a model is that both objects, height and weight here, are \emph{random objects}.
+Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.
+
+\subsection{Brain-graphs} % (fold)
+\label{sub:brain_graphs}
+
+Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.
+For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.
+It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine'' \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex \emph{network} of ``neurons'' (a term he invented), largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
+This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
+Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.
+Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.
+Below, we formalize this concept to enable statistical testing in the graph domain.
+
+% subsection brain_graphs (end)
+
+% section preliminaries (end)
+
+\section{Results}
+
+\subsection{Theoretical results} % (fold)
+\label{sub:theoretical_results}
+
+% subsection theoretical_results (end)
+If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.
+This means that there exists a unique mapping from each brain to a single mind.
+In other words, one can in principle construct a function $g(b): \mB \mapsto \mM$, that is a deterministic mapping from brains to minds.
+It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind (see, for example, $b_1$ and $b_2$ in Figure \ref{fig:rel}(A)).
+Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
+Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
+Define misclassification rate as:
+\begin{align}
+	L_{\PP}(g) = \PP[g(B) \neq M] = \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
+\end{align}
+where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.
+$L_{\PP}(g)$ therefore effectively counts the fraction of time $g$ misclassifies $b$.
+The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is
+\begin{align}
+	g^* = \argmin_{g \in \mc{G}} L_{\PP}(g)
+\end{align}
+where $\mc{G}$ is the set of all possible classifiers.
+Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
+The primary result of casting supervenience is a statistical framework is the following theorem:
+\begin{thm}
+\label{thm1} 
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{\PP}(g^*) = 0$.
+\end{thm}
+% \begin{thm} 
+% 	\label{thm1} 
+% 	Statistical supervenience implies that there exists a classifier that is \emph{always correct}, that is, $g(b)=m$, for all $(m,b) \in \mc{M} \times \mc{B}$.  
+% \end{thm}
+
+\noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.
+This means that one could draw a decision boundary between all equivalence classes of brains, each class corresponding to a different mind, and no mind will reside within two different equivalence classes.
+Thus, the optimal classifier would correctly find these decision boundaries, and therefore have no opportunity to err.
+$\square$
+
+This relationship between statistical supervenience and Bayes error can therefore be described concisely: \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.
+Thus, the above arguments shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
+Further, statistical supervenience can be thought of as a constraint on the possible models.
+Specifically, let $\mc{P}_s \subseteq \mc{P}$ be subset of models for which supervenience holds.
+Then, $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
+
+\subsection{Hypothesis testing} % (fold)
+\label{sub:hypothesis_testing}
+
+While the above theorem is of potential theoretical interest, because the arguments rest on knowing $\PP[M,B]$ and $g^*$, which are typically unknown, they are pragmatically useless.
+However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
+Let $(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)$ be random samples taking their values in $\mc{M} \times \mc{B}$, independently and identically distributed according to model $\PP[M,B]$.
+Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and $n$ training pairs $\tr{\mc{T}_n=}\{(m_1,b_1),\cdots,(m_n,b_n)\}$, and produces a classification $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
+Misclassification rate for this classifier will therefore be a random variable, because the training data $\mc{T}_n$ are random samples.
+Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
+\begin{align}
+\EE[L_{\PP}(g_n)] = \EE[\PP_F[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
+\end{align}
+Unfortunately, in practice, computing $\EE[L_{\PP}(g_n)]$, requires integrating over all possible training data corpuses of size $n$, and by definition, we only have access to a single training data corpus.
+We therefore define ``hold out'' misclassification performance:
+\begin{align}
+\EE[L_{\PP}(g_n)] %= \int \PP[g_n(B)=M | \mc{T}_{n'-n}] \PP[\mc{T}_{n'-n}] d\mc{T}_{n'n} \nonumber \\ 
+&\approx  \hL^{n'}_{F}(g_n) = \sum_{\mc{T}_{n-n'}} \PP[g_n(B)=M | \mc{T}_{n-n'}] \PP[\mc{T}_{n-n'}],
+\end{align}
+where $n'<n$ is the number of held-out training samples, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.
+$n' \hL^{n'}_{F}(g_n)$ is the expected number of misclassified minds.
+Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{F}(g_n)$ and has a binomial distribution, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect, thus $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{\PP}(g_n))$.
+
+Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
+\begin{defi}
+\label{def2}
+Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\MeB$, if and only if $L_{\PP}(g^*) < \varepsilon$.
 \end{defi}
 
-\tr{(Note that this definition does not imply a one-to-one mapping.)}  \tr{To allow for the possibility of only \emph{partial} supervenience,}  we relax the above statistical supervenience to define $\varepsilon$-\textit{supervenience}:
+Given this relaxation, consider the problem of testing for $\eps-$supervenience.
+First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
+Because we hope to reject the null, in favor of the alternative, let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
+We reject for low values of the test-statistic, $n' \hL^{n'}_{F}(g_n)$.
+Specifically, if $n' \hL^{n'}_{F}(g_n)$ is less than the critical value,  $c_{\alpha}(n',\varepsilon)$, then we reject.
+The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
+Thus, if $n' \hL^{n'}_{F}(g_n) < c_{\alpha}(n',\varepsilon)$, we can conclude with $100(1-\alpha$)\% confidence that $\MeB$.
+The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$.
+Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{P}_\eps = \{\PP | L_{\PP}(g^*)< \eps\}$.
+One could then sort  $\eps$-supervenience subsets,  $\mc{P}_s \subseteq \mc{P}_{\eps} \subseteq \mc{P}_{\eps'} \subseteq \mc{P}$, for any $\eps < \eps'$.
 
-\begin{defi}
-\label{def2}
-Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{BM}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
-\end{defi}
+% subsection hypothesis_testing (end)
 
-\tr{Unfortunately, in general $F$ is unknown, but can be estimated from the data.
-Therefore,} generalizing the concept of a classifier $g$ to allow consideration of training data, consider $g_n:\mB \times (\mB \times \mM)^n \rightarrow \mM$ which takes as input an observed brain connectivity structure $b$ and $n$ training pairs $\tr{\vec{d}_n=}(b_1,m_1),\cdots,(b_n,m_n)$ and produces a classification $\widehat{m}=g_n(b;\tr{\vec{d}_n})$.
-Let $L_{F}(g_n) = E[P_{F}[g_n(B;\tr{\vec{D}_n}) \neq M|\tr{\vec{D}_n}]]$.
+\subsection{Power and consistency} % (fold)
+\label{ssub:power_and_consistency}
 
-Consider the problem of testing for $\varepsilon$-supervenience.
-Let the null hypothesis be given by $H_0: L_{F}(g_n) \geq \varepsilon$ so that if we reject at level $\alpha >0$ in favor of the alternative hypothesis $H_A: L_{F}(g_n) < \varepsilon$ then we can conclude, with $100(1-\alpha$)\% confidence, that $\MeB$.
-Letting $\hL^{n'}_{F}(g_n)$ denote the hold-out estimate of misclassification performance based on $n'$ test observations, we note that $\hL^{n'}_{F}(g_n)$ is distributed $Binomial(n',L_{F}(g_n))$.
-The test rejects for small $\hL^{n'}_{F}(g_n)$.
-The level $\alpha$ critical value $c_{\alpha}(n',\varepsilon)$ is available under the least favorable distribution $Binomial(n',\varepsilon)$.
-Furthermore, $\MeB$ implies $L_{F}(g^*) < \varepsilon$, and thus if $g_n$ is a {\it consistent} classifier for $F=F_{BM}$ --- that is, if $\lim_n L_{F}(g_n) = L_{F}(g^*)$ --- then the power of this test (the probability of rejecting when in fact the alternative is true) goes to unity as $n,n' \rightarrow \infty$.
-Thus we have an inference procedure: \begin{thm} \label{thm1} Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$.
-Furthermore, given a consistent classifier the power of the test converges to unity.
-\end{thm}
+Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis (in other words, the probability that it will not make a Type II error).
+Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
+A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.
+A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{\PP}(g_n)] \conv L_{\PP}(g^*)$ as $n\conv \infty$.
+As the notation suggests, consistency of a classifier is a function of the true model, $F$.
+Without any prior knowledge of what the model might be, one desires a \emph{universally consistent} classifier, that is a classifier that is consistent for all $\PP \in \mc{P}$.
 
-Since the joint distribution $F=F_{BM}$ is unknown, the utility of Theorem \ref{thm1} requires that $g_n$ be a {\it universally consistent} classifier --- that is, $\lim_n L_{F}(g_n) = L_{F}(g^*)$ for all distributions $F=F_{BM}$.
-Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{BM}$ \cite{DGL96}.
-Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
+Unfortunately, the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ depends on the (unknown) distribution $\PP=\PP[M,B]$ \cite{DGL96}.
+Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
 For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation.
-Thus, without restrictions on $F_{BM}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-Given these caveats, consider the following thought experiment:
+This means that we can never be confident that $\MeB$ does \emph{not} hold; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
+Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.
+Thus, without restrictions on $\PP[M,B]$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+Given these limitations on even universal consistency, it is still the best one can hope for.
+Therefore, we hope to obtain a universally consistent classifier to test for $\eps$-supervenience.
 
-\begin{thex}
-Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing connections between neurons (or white matter tracts).
-Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable.
-Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 1 for proof).
-Therefore, Theorem \ref{thm1} applies and the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
-Furthermore, Appendix 1 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
-\end{thex}
+% subsubsection power_and_consistency (end)
 
-While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality.
+\subsection{Universal consistency for brain-graphs} % (fold)
+\label{sub:uc}
+
+% In 1977, Stone proved that a certain collection of $k_n$ nearest neighbor algorithms is universally consistent, assuming $b$ was a random $d$-dimensional vector $\mc{B} \subseteq \Real^d$ \cite{Stone77}  (since then, other algorithms have also been shown to be universally consistent, most notably neural networks \cite{Cybenko89}).
+This beautiful result would be directly applicable to mind-brain supervenience, if we were satisfied representing brains by random vectors.
+However, brains are rich with structure, and therefore, other, more sophisticated data structures can better capture aspects of the brain that many of us believe to be true.
+We therefore introduce the concept of a \emph{brain-graph}, which can have a much richer structure than a vector.
+
+Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $V = \{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neural units, and arcs (or edges) $A=\{A_{ij}\}$, where $A_{ij}$ represents the connectivity from neural unit $V_j$ to $V_i$.
+For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that for all $i,j \in [n]$: (i) there are no self loops, $a_{ii}=0$, (ii) all edges are binary, $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric,  $a_{ij}=a_{ji}$.
+Further assume that the vertices are \emph{labeled}, meaning that there is a known one-to-one mapping from each vertex in any brain to a vertex in any other brain, and that each brain-graph has the same number of vertices.
+Thus, each brain-graph is characterized entirely by its adjacency matrix, $a \in \mc{A}$, where $\mc{A} \subseteq \{0,1\}^{n \times n}$.
+Further, let the mental property under investigation be binary, so $m \in \mc{M} = \{0,1\}$.
+The Bayes classifier is:
+\begin{align}
+	\mh{m} = \argmax_m \PP[b,m] = \argmax_m \PP[b|m] \PP[m] 
+\end{align}
+where the likelihood term, $\PP[B|M]$, specifies the probability of observing any particular adjacency matrix for any class, and the prior term, $\PP[M]$, specifies the prior probability of either class.
+
+Under the above model assumptions, the number of possible brain-graph/mental-property pairs is finite, that is, $|\mc{M} \times \mc{B}| = d < \infty$.
+Therefore, one can simply enumerate all possible brain-graphs for each class, such that each class is represented by a multinomial distribution.
+The maximum likelihood estimator (MLE) of the multinomial parameter for each class as guaranteed to converge to the true (but unknown) parameter.
+Formally, one can state $\mh{\PP}_{MLE}[B|M] \conv \PP[B|M]$ as $n \conv \infty$, almost surely.
+Furthermore, the MLE  estimator of the prior also converges to the true (but unknown) prior distribution, $\mh{\PP}_{MLE}[M] \conv \PP[M]$, as $n \conv \infty$, almost surely.
+
+The Bayes plugin classifier is defined by:
+\begin{align}
+	\mh{m} = \argmax_m  \mh{\PP}[b|m] \mh{\PP}[m] 
+\end{align}
+where $\mh{\PP}[B|M]$ and $\mh{\PP}[M]$ are plugin estimators.
+Since the MLE estimators defined above are plugin estimators, and consistent, one obtains a consistent classifier by plugging in the MLE parameters.
+This logic holds for any model in which the cardinality of the sample space is finite, that is $|\mc{M}\times \mc{B}| < \infty$.
+In particular, one can relax both the assumed symmetry and hollowness of the adjacency matrix, and this result still holds.
+Furthermore, one can allow for multi-graphs, in which there are a finite number of different kinds of edges (different colored edges), such that $a_{ij} \in \{0,1,\ldots, K\}$, with $K < \infty$, and the set of possible brain-graphs remains finite.
+Finally, one can allow $\mc{M}$ to be any finite space, not just binary.
+Note that in all these cases, the process one uses to enumerate and sort the brain-graph collections is essentially irrelevant, assuming the same process is used for each class.
+Further note that more robust estimators, such as certain M-estimators and the \emph{maximum a posteriori} estimators, also admit universally consistent classifiers, with perhaps faster convergence properties.
+
+The above approach, however, is insufficient in several important cases.
+First, when the edges live in a continuous space, such as $a_{ij} \in \Real$, corresponding to weighted graphs, the MLE plugin is not well defined.
+In such scenarios, the $k_n$ nearest neighbor ($k$-NN) algorithm can be used (see Appendix 1 for a description of the $k$-NN algorithm) \cite{Stone77}.
+Although originally defined to operate on finite dimensional Euclidean space, the same algorithm can be applied here, once one has first embedded the brain-graphs into such a space.
+One possible embedding is to stack the columns of the adjacency matrix, yielding an $n^2$ element vector, instead of an $n \times n$ element matrix.
+Given such an embedding, the $k$-NN classifier may be used without modification, assuming one has defined a suitable distance operator (such as $L_2$ or $L_1$).
+This algorithm may also be used in the above cases where the support of the sample space is finite, and, convergence of the $k$-NN algorithm for those problems will be faster than the MLE plugins, whenever the distance used is ``appropriate'' (where appropriate is assessed by convergence rates under the true (but unknown) model).
+Furthermore, one can allow for multiple kinds of edges, as before, yielding adjacency tensors, with $a_{ij} \in \Real^d$.
+Or, one can allow the mental property to be multidimensional as well, $m \in \Real^l$.
+The $k$-NN algorithm converges in both these generalizations.
+Finally, one could allow the number of vertices to differ across brain samples.
+In such a scenario, one can ``pad'' the adjacency matrices of the smaller brain-graphs, and the result still holds.
+
+The above universal consistency results hold for unlabeled graphs as well, with a minor modification.
+Assuming the brain-graphs are unlabeled, and we have a test brain-graph $b$, one could first align each training brain-graph with the test brain-graph, and then use the above algorithms.
+To align any pair of graphs, one can solve the graph isomorphism problem, $\mh{Q}_i=\argmin_{Q \in \mc{Q}} \norm{Q b_i Q\T - b}$, where $\mc{Q}$ is the set of all permutation matrices (that is, zero matrices with a single unit value in each row and column).
+Given $\mh{Q}_i$, one can compute $\mt{b}_i=\mh{Q}_i b \mh{Q}_i\T$, and then proceed as one would normally.
+Unfortunately, the graph isomorphism problem is known to be NP-incomplete (that is, it is not known to believe in either P or NP-complete) \cite{GareyJohnson79}, and therefore no polynomial time algorithm is available, making this approach quite time-consuming.
+
+% Thus, while the above Bayes plugin estimators are sufficient in some scenarios, they are not a panacea.
+In particular, the Bayes plugin classifier suffers by having slow convergence properties.
+More specifically, it lacks the ability to generalize, in that any graph not previously observed will be classified according to the prior, no matter how ``close'' it is to other previously observed graphs.
+A different algorithm called the $k_n$ nearest neighbor (knn) algorithm, can significantly expedite convergence by defining a dissimilarity metric, $d(\cdot,\cdot): \mc{B} \times \mc{B} \mapsto \Real$,  which effectively assumes a notion of proximity.
+Although originally proven to be universally consistent when observations are vector-valued random graphs, by embedding the graphs into a vector space, the result holds true for graph data as well (see the appendix for details).
+Using $k$-NN algorithm therefore can classify previously unseen graphs based on their proximity to graphs in the training set.
+
+% subsection brain_graphs (end)
+
+\subsection{Theoretical implications} % (fold)
+\label{sub:implications}
+
+Returning to the question of supervenience, consider the following thought experiment.
+Let the physical property under consideration be brain-graphs, and the mental property under investigation be binary (such as, knows calculus or not).
+Now, imagine collecting very large amounts of very accurate exchangeable brain-graph data and the associated mental property values.
+A $k_n$-nearest neighbor algorithm using an isomorphism-matching Frobenius norm is universally consistent.
+Therefore, %Theorem \ref{thm1} applies and
+the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
+This logic holds for directed graphs or multigraphs or hypergraphs with edge weights, with or without labeled vertices.
+
+% subsection implications (end)
+
+% subsection practical_applications (end)
+
+\section{Discussion}
+
+\subsection{Summary} % (fold)
+\label{par:summary}
+
+We have introduced the notion of $\eps$-supervenience, which states that the misclassification rate for any mental/brain property pair is less than $\eps$.
+Furthermore, we have shown that certain algorithms are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mental/brain pair properties.
+Furthermore, this is a one-sided test, so although power converges to unity, one can never determine whether more data is necessary to get a lower $p$-value, or the particular $\eps$-supervenience does not hold.
+
+\subsection{Practical issues} % (fold)
+\label{par:practical_issues}
+
+Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans when the vertices represent individual neurons is practically possible for any particular mental property at this time.
+Rather, the claim is that \emph{if} one had sufficient data (a very large number of exchangeable mind/brain property pairs), then \emph{in theory}, some level $\eps$-supervenience hypothesis test could be performed.
+Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$), for any particular dataset.
+Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under an $L_p$ norm may have a rate of convergence so slow and a computational demand so high as to be impractical.
+Whichever classifier is used, it is likely to benefit from a large amount of domain specific knowledge, which the proposed classifier completely neglects.
+
+% Further, collecting enough sufficiently accurate, exchangeable brain-graph data and the associated mental property values may be beyond current technological capabilities.
+
+ % (see Appendix 2 for a simulated example of a feasible experiment).
+Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.
+
+% Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2)
+
+% paragraph practical_issues (end)
+
+%A simulation of a realistically performable experiment in C.
+elegans suggests that convergence rates are not so poor as to make these results completely impractical.
+Several points merit further discussion, in our humble opinion.
+
+% paragraph summary (end)
+
+\subsection{A unified quest} % (fold)
+\label{sub:practical_applications}
+
+A central (perhaps \emph{the} central) quest in much of neuroscience, psychology, and cognitive science is to discover the brain properties that subvene under various mental properties, although questions are rarely cast in the supervenience formalism.
+Moreover, the particular brain properties that are often believed to subvene under these mental properties are neural circuits, or brain-subgraphs.
+To this end, many investigations in these fields include schematic diagrams showing a particular brain-subgraph subvening under a particular mental phenotype.
+This practice transcends the evolutionary hierarchy of neuroscientific research.
+For instance, in the invertebrate literature, vertices correspond to particular labeled neurons, and edges correspond to synapse \cite{NorthGreenspan07}.
+In the vertebrate literature, vertices tend to correspond to types of neurons in particular regions, and edges correspond to tendencies of connections \cite{Shepherd04}.
+For primates \cite{Felleman_VanEssen91} and humans \cite{Mori05} vertices tend to represent functionally distinct neuroanatomical regions, and edges represent regional interconnectivity.
+Furthermore, this practice also transcends analytical background, including anatomists \cite{Abeles91}, philosophers \cite{Koch_Davis94}, statisticians \cite{Rao_Lewicki02}, and physicists \cite{Chow_Dalibard03}.
+The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (perhaps causality, not supervenience, is the true desideratum).
+
+% \subsection{Causality} % (fold)
+% \label{par:causality}
+% 
+% While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality.
 Assuming we have confirmed $\MeB$ for a particular mental/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.
+% 
+% % paragraph causality (end)
 
-Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2) a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under our Frobenius norm (or other norm; see Appendix 1 for details) may have a rate of convergence so slow and a computational demand so high as to be impractical; and (3) collecting enough sufficiently accurate independent and identically distributed brain-graph data and the associated mental property indicators may be beyond current technological capabilities.
-Regardless, related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggesting feasibility of such an experiment in the near future (see Appendix 2 for a simulated example of a feasible experiment).
-Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.
-\tr{This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first \emph{proof} (to our knowledge) that empirically investigating supervenience is at least theoretically possible.}
+\subsection{Human applications} % (fold)
+\label{par:human_applications}
 
-%% == end of paper:
+In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has seen an explosion of studies investigating the human connectome, and relating connectomes to cognitive properties \cite{Jones10}.
+These studies can all be thought of as testing various $\eps$-supervenience hypotheses.
+For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenics and controls \cite{ArdekaniSzeszko10}.
+As the resolution and signal-to-noise ratio of magnetic resonance imaging continues to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
+Furthermore, the utilization of other imaging technologies, such as polarized light imaging \cite{PalmAmunts10} and high-throughput electron microscopy \cite{DenkHorstmann04,HayworthLichtman06}, will continue to improve the effective resolution of these inferred connectomes from human brains.
+While determining whether a particular individual ``knows calculus'' might be quite distant, many other cognitive and psychological supervenience hypothesis have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be vanishing.
 
-%% Optional Materials and Methods Section
-%% The Materials and Methods section header will be added automatically.
+% Related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggests feasibility of such an experiment in the future.
 
-%% Enter any subheads and the Materials and Methods text below.
-%\begin{materials}
-% Materials text
-%\end{materials}
+% paragraph relationship_to_previous_work (end)
 
-%% Optional Appendix or Appendices
-%% \appendix Appendix text...
-%% or, for appendix with title, use square brackets:
-%% \appendix[Appendix Title]
+\subsection{Quantum networks} % (fold)
+\label{par:quantum_networks}
 
-\section*{Appendix 1: $k_n$-nearest neighbor universal consistency for graphs}
-\label{proof}
+Recently, several authors have suggested the possibility that brains are better characterized as quantum networks, instead of classical networks.
+The above results hold regardless of whether computations in the brain are quantum or classical, as quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
+This means, if the above analysis failed to reject the null at level $\alpha$, the interpretation does not change if one assumes quantum versus classical computations.
 
-Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another.
-Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$.
-If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$.
-Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$.
-Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.
+% paragraph quantum_networks (end)
 
-In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability.
-Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $P[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{F}(g_n) \rightarrow c \leq L_{F}(g^*) + \delta$, yielding universal consistency.
+\subsection{Stochastic Supervenience} % (fold)
+\label{par:stochastic_supervenience}
 
-If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm.
-Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$.
-Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}.
-Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
-%\tr{Trivially, this argument also applies to graphs with (self-) loops, and
-%directed, colored, or weighted edges, and hypergraphs.}
+Possible explanations of how it might be the case that $\eps>0$ include stochastic supervenience \cite{Craver09}, and supernatural causal effects.
+Thus, the above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
 
-Several points of note:  isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.
+% paragraph stochastic_supervenience (end)
 
-Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg.
-455 for proof when $b_i$'s are vectors).
-Furthermore, we can relax the constraint that mental properties finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).
-The proofs for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.
+\subsection{Dynamics vs. statics} % (fold)
+\label{par:dynamics_vs_statics}
 
-%\appendix
-\section*{Appendix 2: Simulation} \label{simulation}
+The above \emph{gedanken} experiment did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
+Similarly, for the question of mind-brain supervenience in humans, one need not every observe any activity of the brain, one must merely observe the model, which determines the activity (in a potentially stochastic process).
+Thus, this approach to understanding the relationship between mind and brain could be considered a rather drastic departure from most of systems neuroscience, in which the goal is typically to understanding the neural activity ``code.''  In contrast, if mind-brain supervenience is true, this motivates a search for the neural connectivity ``code,'' an \emph{engram} for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
+
+% paragraph dynamics_vs_statics (end)
+
+\subsection{Concluding thoughts} % (fold)
+\label{par:concluding_thoughts}
+
+% paragraph concluding_thoughts (end)
+
+This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
+The above discussion suggests that many previously conducted investigations either assume supervenience, or test it.
+Further, new technologies facilitate testing supervenience of mental properties on brain-graphs more easily.
+
+\appendix
+\section{$k_n$ nearest neighbor algorithm} % (fold)
+\label{app:knn}
+
+Assume that $b$ is a real $d$-dimensional vector, $b \in \mc{B} \subseteq \Real^d$, and $m$ is a binary indicator, $m \in \mc{M} = \{0,1\}$.
+Then, further assume that we have observed a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled identically and independently from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$.
+A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
+It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim \PP[M,B]$, and $m$ is simply unobserved.
+
+% This algorithm also readily generalizes to the $C$-way classification problem, where there $C \geq 2$ classes.
+Instead of finding the majority class of the $k_n$ nearest neighbors, one instead finds the plurality class, that is, the class with the most training samples within the $k_n$ nearest neighbors.
+This algorithm generalizes even further, when $\mc{M}$ is no longer discrete, but rather continuous, for instance, $\mc{M}=\Real$.
+In such a scenario, this approach is called $k_n$ nearest neighbor \emph{regression}.
+
+% Another algorithm, proven to be universally consistent, is the $k_n$ nearest neighbor algorithm \cite{Stone77}, can rectify these limitations of the Bayes plugin algorithms.
+
+The $1$-nearest neighbor ($1$-NN) classifier works as follows.
+Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
+Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, where parenthetical indices, $(i)$, indicate rank order.
+One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
+The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $m=m_{(1)}$.
+The $k_n$ nearest neighbor is a straightforward generalization of this approach.
+It says that the test mind is in the same class as which ever class is the majority class of the $k_n$ nearest neighbors, $m=\II\{\sum_{i=1}^{k_n} m_{(i)} > k_n/2\}$.
+Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
+
+Unfortunately, no such algorithm is universally consistent.
+Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training points.
+Then, a collection of such algorithms, $\{g_n\}$,  with $k_n$ increasing with $n$, can be universally consistent under certain constraints.
+In particular, as $n$ increases, $k_n$ must also increase, but not quite as quickly.
+Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
+In Stone' original proof, the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
+% section _k_n_nearest_neighbor_algorithm (end)
+
+\section{Simulation} % (fold)
+\label{sub:simulation}
 
@@
 Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}.
-Thus, one may design an experiment by describing the joint distribution $F_{BM}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $P[M=m_1]$.
+Thus, one may design an experiment by describing the joint distribution $\PP[M,B]$ via class-conditional distributions $\PP[B|M=m_j]$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
 Here the mental property corresponds to the C.~elegans exhibiting or not exhibiting a particular behavior (e.g., response to an odor).
 
-Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $F_{BM}$ (Figure \ref{fig1}).
+Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
 To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$.
@@
 The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $F_{BM}$ used here; however, we demand universal consistency.)
+(Better classifiers can be constructed for the joint distribution $\PP[M,B]$ used here; however, we demand universal consistency.)
 
-\begin{figure}[h!]
-\centering \includegraphics[width=.9\linewidth]{Lhatplot}
+\begin{figure}[!ht]
+\centering \includegraphics[width=.5\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.
-$\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$.
-Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
-We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{F}(g^*) \approx 0$ for this simulation.
+$\hL^{1000}_{\PP}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $\PP[M,B]$.
+Each dot depicts an estimate for $L_{\PP}(g_n)$; standard errors are $(L_{\PP}(g_n)(1-L_{\PP}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
+We reject $H_0: L_{\PP}(g^*) \geq 0.10$ at $\alpha=0.01$.
+$L_{\PP}(g^*) \approx 0$ for this simulation.
 }
@@
 
-\paragraph{Acknowledgments}
-The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
-% \end{acknowledgments}
+\clearpage
 
-%\subparagraph*{References}
-\bibliography{biblist}
-% \addcontentsline{toc}{section}{References}
-%\bibliography{Science}
-% \bibliographystyle{plain}
-\bibliographystyle{nature}
+% subsection simulation (end)
+
+% \section{$k_n$-nearest neighbor universal consistency for graphs}
+% \label{proof}
+% 
+% Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another.
+Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$.
+If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$.
+Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$.
+Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.
+% 
+% In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability.
+Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $\PP[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{\PP}(g_n) \rightarrow c \leq L_{\PP}(g^*) + \delta$, yielding universal consistency.
+% 
+% If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm.
+Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$.
+Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}.
+Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
+% %\tr{Trivially, this argument also applies to graphs with (self-) loops, and
+% %directed, colored, or weighted edges, and hypergraphs.}
+% 
+% Several points of note:  isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.
+% 
+% Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg.
+455 for proof when $b_i$'s are vectors).
+Furthermore, we can relax the constraint that mental properties are finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).
+The proof for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.
+
+\section*{Acknowledgments}
+
+The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, and S.~Seung.
+
+% \section{References}
+% \bibliography{biblist}
+\bibliography{/Users/jovo/Research/papers/meta/biblist}
+\bibliographystyle{ieeetr}
 

### 2010-11-21 supervenience.tex
reduced draft to good bits

@@
 \newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_{\PP} \mB}
-\newcommand{\MsB}{\mM \overset{S}{\sim}_F \mB}
+\newcommand{\MsB}{\mM \overset{S}{\sim}_{\PP} \mB}
 \newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{\PP} \mB}
@@
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
-The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two.
+The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two.
 %In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.
@@
 The potential significance of this work can be divided into distinct disciplines.
-To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
-To the statistician, this work points out the limitations of hypothesis testing in a novel domain.
+To the philosopher, this work demonstrates an example of morphing philosophical conjectures into statistical hypotheses that are amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
+To the statistician, this work points out the limitations of hypothesis testing in a novel domain, and is suggestive of more general limitations previously (perhaps) under-appreciated.
 To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be considered supervenience investigations, further motivating brain-graph related research paradigms.
@@
 
-\vspace*{0.5 in}
+% \vspace*{0.5 in}
 
@@
 
-Questioning the relationship between the mind (thoughts, beliefs, preferences, emotions, intelligence, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
+Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
 Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
@@
 Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation.
-Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms and that can be experimentally tested \tr{as a hypothesis}.
+This work is an attempt to bridge the gap between these philosophical conjectures and experimentally testable hypotheses.
+% Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms, and that can be experimentally tested \tr{as a hypothesis}.
 
@@
 Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.
-Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists.
+Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists, given sufficient data.
 Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.
@@
 
-\section{Preliminaries} % (fold)
+\section{Statistical supervenience} % (fold)
 \label{sec:preliminaries}
 
-\subsection{Supervenience} % (fold)
-\label{sub:supervenience}
+% \subsection{Mind Brain Supervenience} % (fold)
+% \label{sub:supervenience}
 
 % subsection supervenience (end)
-The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
+% The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
 Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}.
 Therefore, here we define minds, brains, and supervenience.
+% The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
 
-Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
-The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
-In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.
-Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.
-
-Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-The set of possible minds $\mM$ is also unrestricted.
-In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences.
-Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-
-The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
-Donald Davidson canonized this conjecture in 1970 with the following quote: \cite{Davidson70}
+Donald Davidson canonized the mind-brain supervenience relation in 1970 with the following quote: \cite{Davidson70}
 \begin{quotation}
@@
 \end{quotation}
-This conjecture may be concisely and formally stated: $m \neq m' \implies b \neq b' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$, where $(m,b)$ is a mind-brain pair, and $(m',b')$ is a different mind-brain pair.
-While mind-brain supervenience is a relatively strong claim, importantly, one can imagine far stronger relations.
+This conjecture may be concisely and formally stated.
+Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
+% The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
+In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.
+%Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.
+Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
+%The set of possible minds $\mM$ is also unrestricted.
+In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences.
+%Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
+Thus%, the above mind brain supervenience conjecture may be formally stated
+: $m \neq m' \implies b \neq b' \, \forall (m,b) \neq (m',b') \in \mc{M} \times \mc{B}$, where $(m,b)$ and $(m',b')$ are mind-brain pairs.
+This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details).
 
-For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.
-An injective relation is any relation that preserves distinctness.
-Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$ (note that the directionality of the implication has been switched relative to supervenience).
-For instance, it might be the case that a brain could change without the mind changing.
-Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
-It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
-In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
-This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.
-Minds supervening on brains does not imply that brains supervene on minds.
-
-% In this case, it would be said that both $M$ supervenes on $B$ and $B$ supervenes on $M$.
-But it is not necessarily the case that if supervenience holds in one direction that it does or does not hold in the other direction.
-
-Second, it may be the case that minds supervene on brains, but that brains do not cause minds.
-For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.
-Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other.
-This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
-
-% Supervenience is also not \emph{causality}.
-Consider the coin flip example.
-Nowhere did the notion of causality even arise.
-It could be the case that $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
-Thus, supervenience does not imply causality.
-
-Third, supervenience does not imply \emph{identity}.
-Consider, for example, acceleration and velocity.
-Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).
-Similarly, velocity supervenes on position, as velocity cannot change without position changing.
-Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
-Rather, position can change with constant velocity, meaning without acceleration changing.
-
-What supervenience does imply, however, is the following.
-Imagine finding two different minds.
-If $\MeB$, then the brains subvening under those two minds must be different.
-In other words, there cannot be two different minds, either of which could supervene on a single brain.
-Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
-
-\begin{figure}[htbp]
-	\centering
-		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
-	\caption{Possible relations between minds and brains.
-(A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
-(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds.
-(C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
-	\label{fig:rel}
-\end{figure}
-
-All of the above relations are \emph{logical} relations, not probabilistic relations.
-To facilitate both statistical analysis and empirical investigation, we project this supervenience notion into a statistical setting.
-To proceed, we first define a \emph{model}, $\mc{P}=\{\PP[M,B]\}$, the set of all possible joint distributions, $\PP[M,B]$, that could describe the probabilistic relationship between minds and brains.
-Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$ (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
+To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
+Let $\PP[M,B]$ indicate a joint distribution of minds and brains.
+%Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$.
+% (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
 % \begin{enumerate}
@@
 % \end{enumerate}
-Given a distribution, one can then calculate any marginal or conditional distributions.
+%Given a distribution, one can then calculate any marginal or conditional distributions.
 For instance, $\PP[B]=\int_{m \in \mM} \PP[M,B]dm$, or $\PP[M|B]=\PP[M,B]/\PP[B]$.
@@
 \label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0 \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$.
-Alternately, $\MeB$ can be written as $\PP[m = m' | b = b']=1  \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$, for all $(m,b) \neq (m',b') \in \mc{M} \times \mc{B}$.
 \end{defi}
-Statistical supervenience is therefore a probabilistic relation on sets.
-Note that statistical supervenience is distinct from statistical correlation.
-\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
-If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_F \mM$.
-Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+\noindent Statistical supervenience is therefore a probabilistic relation on sets (related to, but distinct from correlation; see Appendix \ref{sec:relations} for details).
 
@@
 Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.
-
-\subsection{Brain-graphs} % (fold)
-\label{sub:brain_graphs}
-
-Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.
-For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.
-It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine'' \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex \emph{network} of ``neurons'' (a term he invented), largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
-This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
-Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.
-Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.
-Below, we formalize this concept to enable statistical testing in the graph domain.
+% 
+% \subsection{Brain-graphs} % (fold)
+% \label{sub:brain_graphs}
 
@@
 
-\section{Results}
+%\section{Results}
 
-\subsection{Theoretical results} % (fold)
+\section{Statistical supervenience is equivalent to perfect classification accuracy} % (fold)
 \label{sub:theoretical_results}
@@
 In other words, one can in principle construct a function $g(b): \mB \mapsto \mM$, that is a deterministic mapping from brains to minds.
-It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind (see, for example, $b_1$ and $b_2$ in Figure \ref{fig:rel}(A)).
+It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind (see Appendix \ref{sec:relations} for examples).
 Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
@@
 Define misclassification rate as:
-\begin{align}
-	L_{\PP}(g) = \PP[g(B) \neq M] = \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
-\end{align}
-where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.
-$L_{\PP}(g)$ therefore effectively counts the fraction of time $g$ misclassifies $b$.
-The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is
-\begin{align}
-	g^* = \argmin_{g \in \mc{G}} L_{\PP}(g)
-\end{align}
-where $\mc{G}$ is the set of all possible classifiers.
+% \begin{align}
+	$L_{\PP}(g) = \PP[g(B) \neq M]$, %= \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
+% \end{align}
+% where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.  
+% $L_{\PP}(g)$ therefore effectively 
+which counts the fraction of time $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
+% \begin{align}
+	$g^* = \argmin_{g} L_{\PP}(g)$.
+% \end{align}
+% where $\mc{G}$ is the set of all possible classifiers.  
 Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
@@
 \label{thm1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $L_{\PP}(g^*) = 0$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_{\PP} \mB$, if and only if $L_{\PP}(g^*) = 0$.
+Formally, \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.
 \end{thm}
@@
 
-This relationship between statistical supervenience and Bayes error can therefore be described concisely: \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.
 Thus, the above arguments shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
 Further, statistical supervenience can be thought of as a constraint on the possible models.
-Specifically, let $\mc{P}_s \subseteq \mc{P}$ be subset of models for which supervenience holds.
+Specifically, let $\mc{P}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{P}_s$ be subset of models for which supervenience holds.
 Then, $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
 
-\subsection{Hypothesis testing} % (fold)
+\section{A hypothesis test for supervenience} % (fold)
 \label{sub:hypothesis_testing}
@@
 However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
-Let $(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)$ be random samples taking their values in $\mc{M} \times \mc{B}$, independently and identically distributed according to model $\PP[M,B]$.
-Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and $n$ training pairs $\tr{\mc{T}_n=}\{(m_1,b_1),\cdots,(m_n,b_n)\}$, and produces a classification $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
-Misclassification rate for this classifier will therefore be a random variable, because the training data $\mc{T}_n$ are random samples.
-Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
-\begin{align}
-\EE[L_{\PP}(g_n)] = \EE[\PP_F[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
-\end{align}
-Unfortunately, in practice, computing $\EE[L_{\PP}(g_n)]$, requires integrating over all possible training data corpuses of size $n$, and by definition, we only have access to a single training data corpus.
+Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to model $\PP[M,B]$.
+Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
+Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
+The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
+% Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
+% \begin{align}
+% \EE[L_{\PP}(g_n)] = \EE[\PP_F[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
+% \end{align}
+% Unfortunately, in practice, computing $\EE[L_{\PP}(g_n)]$, requires integrating over all possible training data corpuses of size $n$, and by definition, we only have access to a single training data corpus.
 We therefore define ``hold out'' misclassification performance:
-\begin{align}
-\EE[L_{\PP}(g_n)] %= \int \PP[g_n(B)=M | \mc{T}_{n'-n}] \PP[\mc{T}_{n'-n}] d\mc{T}_{n'n} \nonumber \\ 
-&\approx  \hL^{n'}_{F}(g_n) = \sum_{\mc{T}_{n-n'}} \PP[g_n(B)=M | \mc{T}_{n-n'}] \PP[\mc{T}_{n-n'}],
-\end{align}
-where $n'<n$ is the number of held-out training samples, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.
-$n' \hL^{n'}_{F}(g_n)$ is the expected number of misclassified minds.
-Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{F}(g_n)$ and has a binomial distribution, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect, thus $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{\PP}(g_n))$.
+% \begin{align}
+% \EE[L_{\PP}(g_n)] 
+% &\approx  \hL^{n'}_{F}(g_n) = \sum_{\mc{T}_{n-n'}} \PP[g_n(B)=M | \mc{T}_{n-n'}] \PP[\mc{T}_{n-n'}],
+% \end{align}
+$\hL^{n'}_{F}(g_n) = \PP[g_n(B)=M | \mc{T}_{n-n'}]$,
+where $n'<n$ is the number of held-out training samples (samples not used to obtain $g_n(\cdot)$).
+%, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.  
+ % $n' \hL^{n'}_{F}(g_n)$ is t
+The approximate expected number of misclassified minds
+%. Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{F}(g_n)$ 
+therefore
+% and 
+has a binomial distribution:  $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{\PP}(g_n))$. %, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect.
 
@@
 Given this relaxation, consider the problem of testing for $\eps-$supervenience.
-First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
-Because we hope to reject the null, in favor of the alternative, let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-We reject for low values of the test-statistic, $n' \hL^{n'}_{F}(g_n)$.
-Specifically, if $n' \hL^{n'}_{F}(g_n)$ is less than the critical value,  $c_{\alpha}(n',\varepsilon)$, then we reject.
+%First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
+% Because we hope to reject the null, in favor of the alternative, l
+Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
+We reject for  values of the test-statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{F}(g_n) <c_{\alpha}(n',\varepsilon)$.
 The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
-Thus, if $n' \hL^{n'}_{F}(g_n) < c_{\alpha}(n',\varepsilon)$, we can conclude with $100(1-\alpha$)\% confidence that $\MeB$.
+Thus, rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
 The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$.
-Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{P}_\eps = \{\PP | L_{\PP}(g^*)< \eps\}$.
+%Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{P}_\eps = \{\PP | L_{\PP}(g^*)< \eps\}$.
 One could then sort  $\eps$-supervenience subsets,  $\mc{P}_s \subseteq \mc{P}_{\eps} \subseteq \mc{P}_{\eps'} \subseteq \mc{P}$, for any $\eps < \eps'$.
@@
 
-\subsection{Power and consistency} % (fold)
-\label{ssub:power_and_consistency}
+% \section{Power and consistency} % (fold)
+% \label{ssub:power_and_consistency}
 
-Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis (in other words, the probability that it will not make a Type II error).
+Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis% (in other words, the probability that it will not make a Type II error)
+.
 Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
@@
 A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{\PP}(g_n)] \conv L_{\PP}(g^*)$ as $n\conv \infty$.
-As the notation suggests, consistency of a classifier is a function of the true model, $F$.
+%As the notation suggests, consistency of a classifier is a function of the true model, $\PP$.
 Without any prior knowledge of what the model might be, one desires a \emph{universally consistent} classifier, that is a classifier that is consistent for all $\PP \in \mc{P}$.
+% The above described hypothesis tests suggests the following theorem:
+% 
+% \begin{thm} \label{thm1} 
+% Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$.
+Furthermore, given a consistent classifier the power of the test converges to unity.
+% \end{thm}
 
@@
 Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation.
-This means that we can never be confident that $\MeB$ does \emph{not} hold; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.
+For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, %.
+This means that we can never be confident that $\MeB$ does \emph{not} hold; rather,
+it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
+%Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.
 Thus, without restrictions on $\PP[M,B]$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-Given these limitations on even universal consistency, it is still the best one can hope for.
+%Given these limitations on even universal consistency, it is still the best one can hope for.
 Therefore, we hope to obtain a universally consistent classifier to test for $\eps$-supervenience.
@@
 
-\subsection{Universal consistency for brain-graphs} % (fold)
+\section{A thought experiment to guarantee unity power} % (fold)
 \label{sub:uc}
 
-% In 1977, Stone proved that a certain collection of $k_n$ nearest neighbor algorithms is universally consistent, assuming $b$ was a random $d$-dimensional vector $\mc{B} \subseteq \Real^d$ \cite{Stone77}  (since then, other algorithms have also been shown to be universally consistent, most notably neural networks \cite{Cybenko89}).
-This beautiful result would be directly applicable to mind-brain supervenience, if we were satisfied representing brains by random vectors.
-However, brains are rich with structure, and therefore, other, more sophisticated data structures can better capture aspects of the brain that many of us believe to be true.
-We therefore introduce the concept of a \emph{brain-graph}, which can have a much richer structure than a vector.
+% Below, we will assume that brains are characterized by graphs, referred to hereafter as ``brain-graphs.''  This assumption follows naturally from the neuron-doctrine
+% % Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.
+For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.
+It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine''
+% \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex {network} of neurons, largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
+This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
+%Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.
+Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.
+Below, we formalize this concept to enable statistical testing in the graph domain.
+% Below, we will assume that $m$ is binary, that is, $m \in \{0,1\}$ (for example, $m$ might indicate whether the person knows calculus or not).
 
-Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $V = \{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neural units, and arcs (or edges) $A=\{A_{ij}\}$, where $A_{ij}$ represents the connectivity from neural unit $V_j$ to $V_i$.
+To ensure consistency and therefore unity power, the classifier $g_n(\cdot)$ must be able to converge to the truth, regardless of the true model, $\PP$.
+We therefore make explicit a model for brain, and show that under this very general model, universally consistent classifiers are available.
+
+\begin{thex}
+Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a brain-graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
+Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable.
+Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
+A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix \ref{sec:knn} for proof).
+Therefore, %Theorem \ref{thm1} applies and
+the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
+This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
+Furthermore, Appendix \ref{sec:knn} also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
+\end{thex}
+
+%(fold)
+% Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $V = \{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neural units, and arcs (or edges) $A=\{A_{ij}\}$, where $A_{ij}$ represents the connectivity from neural unit $V_j$ to $V_i$.
 For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that for all $i,j \in [n]$: (i) there are no self loops, $a_{ii}=0$, (ii) all edges are binary, $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric,  $a_{ij}=a_{ji}$.
@@
 The Bayes classifier is:
-\begin{align}
-	\mh{m} = \argmax_m \PP[b,m] = \argmax_m \PP[b|m] \PP[m] 
-\end{align}
-where the likelihood term, $\PP[B|M]$, specifies the probability of observing any particular adjacency matrix for any class, and the prior term, $\PP[M]$, specifies the prior probability of either class.
-
-Under the above model assumptions, the number of possible brain-graph/mental-property pairs is finite, that is, $|\mc{M} \times \mc{B}| = d < \infty$.
+% \begin{align}
+% 	\mh{m} = \argmax_m \PP[b,m] = \argmax_m \PP[b|m] \PP[m] 
+% \end{align}
+% where the likelihood term, $\PP[B|M]$, specifies the probability of observing any particular adjacency matrix for any class, and the prior term, $\PP[M]$, specifies the prior probability of either class.
+% 
+% Under the above model assumptions, the number of possible brain-graph/mental-property pairs is finite, that is, $|\mc{M} \times \mc{B}| = d < \infty$.
 Therefore, one can simply enumerate all possible brain-graphs for each class, such that each class is represented by a multinomial distribution.
@@
 Furthermore, the MLE  estimator of the prior also converges to the true (but unknown) prior distribution, $\mh{\PP}_{MLE}[M] \conv \PP[M]$, as $n \conv \infty$, almost surely.
-
-The Bayes plugin classifier is defined by:
-\begin{align}
-	\mh{m} = \argmax_m  \mh{\PP}[b|m] \mh{\PP}[m] 
-\end{align}
-where $\mh{\PP}[B|M]$ and $\mh{\PP}[M]$ are plugin estimators.
+% 
+% The Bayes plugin classifier is defined by:
+% \begin{align}
+% 	\mh{m} = \argmax_m  \mh{\PP}[b|m] \mh{\PP}[m] 
+% \end{align}
+% where $\mh{\PP}[B|M]$ and $\mh{\PP}[M]$ are plugin estimators.
 Since the MLE estimators defined above are plugin estimators, and consistent, one obtains a consistent classifier by plugging in the MLE parameters.
@@
 Further note that more robust estimators, such as certain M-estimators and the \emph{maximum a posteriori} estimators, also admit universally consistent classifiers, with perhaps faster convergence properties.
-
-The above approach, however, is insufficient in several important cases.
+% 
+% The above approach, however, is insufficient in several important cases.
 First, when the edges live in a continuous space, such as $a_{ij} \in \Real$, corresponding to weighted graphs, the MLE plugin is not well defined.
@@
 In such a scenario, one can ``pad'' the adjacency matrices of the smaller brain-graphs, and the result still holds.
-
-The above universal consistency results hold for unlabeled graphs as well, with a minor modification.
+% 
+% The above universal consistency results hold for unlabeled graphs as well, with a minor modification.
 Assuming the brain-graphs are unlabeled, and we have a test brain-graph $b$, one could first align each training brain-graph with the test brain-graph, and then use the above algorithms.
@@
 Unfortunately, the graph isomorphism problem is known to be NP-incomplete (that is, it is not known to believe in either P or NP-complete) \cite{GareyJohnson79}, and therefore no polynomial time algorithm is available, making this approach quite time-consuming.
-
-% Thus, while the above Bayes plugin estimators are sufficient in some scenarios, they are not a panacea.
-In particular, the Bayes plugin classifier suffers by having slow convergence properties.
-More specifically, it lacks the ability to generalize, in that any graph not previously observed will be classified according to the prior, no matter how ``close'' it is to other previously observed graphs.
-A different algorithm called the $k_n$ nearest neighbor (knn) algorithm, can significantly expedite convergence by defining a dissimilarity metric, $d(\cdot,\cdot): \mc{B} \times \mc{B} \mapsto \Real$,  which effectively assumes a notion of proximity.
-Although originally proven to be universally consistent when observations are vector-valued random graphs, by embedding the graphs into a vector space, the result holds true for graph data as well (see the appendix for details).
-Using $k$-NN algorithm therefore can classify previously unseen graphs based on their proximity to graphs in the training set.
-
-% subsection brain_graphs (end)
-
-\subsection{Theoretical implications} % (fold)
-\label{sub:implications}
-
-Returning to the question of supervenience, consider the following thought experiment.
+% 
+% 
+% % subsection brain_graphs 
+% 
+% \section{Theoretical implications} % (fold)
+% \label{sub:implications}
+% 
+% Returning to the question of supervenience, consider the following thought experiment.
 Let the physical property under consideration be brain-graphs, and the mental property under investigation be binary (such as, knows calculus or not).
@@
 Therefore, %Theorem \ref{thm1} applies and
-the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
+% the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
 This logic holds for directed graphs or multigraphs or hypergraphs with edge weights, with or without labeled vertices.
-
-% subsection implications (end)
-
-% subsection practical_applications (end)
+% 
+% 
+% % subsection implications (end)
+% 
+% 
+% % subsection practical_applications (end)
+%(end)
 
@@
 We have introduced the notion of $\eps$-supervenience, which states that the misclassification rate for any mental/brain property pair is less than $\eps$.
-Furthermore, we have shown that certain algorithms are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mental/brain pair properties.
+Furthermore, we have shown that $k$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mental/brain pair properties.
 Furthermore, this is a one-sided test, so although power converges to unity, one can never determine whether more data is necessary to get a lower $p$-value, or the particular $\eps$-supervenience does not hold.
@@
 Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$), for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under an $L_p$ norm may have a rate of convergence so slow and a computational demand so high as to be impractical.
+Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under an $L_p$ norm may have a rate of convergence so slow and a computational demand so high as to be impractical (though, see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
 Whichever classifier is used, it is likely to benefit from a large amount of domain specific knowledge, which the proposed classifier completely neglects.
 
-% Further, collecting enough sufficiently accurate, exchangeable brain-graph data and the associated mental property values may be beyond current technological capabilities.
-
- % (see Appendix 2 for a simulated example of a feasible experiment).
-Nevertheless, our thought experiment suggests that we can hope to determine that a given mental property under consideration $\varepsilon$-supervenes on a brain's connectivity structure.
-
-% Practical issues regarding actually conducting the above thought experiment include: (1) as stated, we must consider the space $\mB$ to be the quotient space of graphs mod graph isomorphism, unless the vertices are {\it labeled}; (2)
-
 % paragraph practical_issues (end)
-
-%A simulation of a realistically performable experiment in C.
-elegans suggests that convergence rates are not so poor as to make these results completely impractical.
-Several points merit further discussion, in our humble opinion.
 
@@
 Furthermore, the utilization of other imaging technologies, such as polarized light imaging \cite{PalmAmunts10} and high-throughput electron microscopy \cite{DenkHorstmann04,HayworthLichtman06}, will continue to improve the effective resolution of these inferred connectomes from human brains.
-While determining whether a particular individual ``knows calculus'' might be quite distant, many other cognitive and psychological supervenience hypothesis have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be vanishing.
-
-% Related experimental work includes collecting various types of brain graph data \cite{WhiteBrenner86, DenkHorstmann04, BriggmanDenk06} and various approaches to inference on brain graphs \cite{MackeBorst08, Mishchenko09, LuLichtman09}, suggests feasibility of such an experiment in the future.
+While determining whether a particular individual knows calculus might be quite distant, many other cognitive and psychological supervenience hypothesis have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be vanishing.
 
@@
 
-Recently, several authors have suggested the possibility that brains are better characterized as quantum networks, instead of classical networks.
+Recently, several authors have suggested the possibility that brains are better characterized as quantum networks, instead of classical networks \cite{Penrose,check_book}.
 The above results hold regardless of whether computations in the brain are quantum or classical, as quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
@@
 The above \emph{gedanken} experiment did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
-Similarly, for the question of mind-brain supervenience in humans, one need not every observe any activity of the brain, one must merely observe the model, which determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain could be considered a rather drastic departure from most of systems neuroscience, in which the goal is typically to understanding the neural activity ``code.''  In contrast, if mind-brain supervenience is true, this motivates a search for the neural connectivity ``code,'' an \emph{engram} for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
+Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model, which determines the activity (in a potentially stochastic process).
+Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understanding the neural activity ``code.''  In contrast, if mind-brain supervenience is true, this motivates a search for the neural connectivity ``code,'' an \emph{engram} for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
 
@@
 
+\clearpage
 \appendix
+
+\section{Relations between sets} % (fold)
+\label{sec:relations}
+
+In this appendix we describe several kinds of relations that are not implied by supervenience.
+
+%For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.
+
+First, a supervenient relation does not imply an injective relation.
+An injective relation is any relation that preserves distinctness.
+Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$ (note that the directionality of the implication has been switched relative to supervenience).
+For instance, it might be the case that a brain could change without the mind changing.
+Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
+It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
+In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
+This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.
+Minds supervening on brains does not imply that brains supervene on minds.
+
+Second,  supervenience does not imply causality. %it may be the case that minds supervene on brains, but that brains do not cause minds.  
+For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.
+Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other.
+This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
+
+Third, supervenience does not imply identity.
+Consider, for example, acceleration and velocity.
+Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).
+Similarly, velocity supervenes on position, as velocity cannot change without position changing.
+Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
+Rather, position can change with constant velocity, meaning without acceleration changing.
+
+What supervenience does imply, however, is the following.
+Imagine finding two different minds.
+If $\MeB$, then the brains subvening under those two minds must be different.
+In other words, there cannot be two different minds, either of which could supervene on a single brain.
+Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
+
+\begin{figure}[htbp]
+	\centering
+		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
+	\caption{Possible relations between minds and brains.
+(A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
+(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds.
+(C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
+	\label{fig:rel}
+\end{figure}
+% section relations_between_sets (end)
+
+Note that statistical supervenience is distinct from statistical correlation.
+\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
+If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_F \mM$.
+Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+
 \section{$k_n$ nearest neighbor algorithm} % (fold)
-\label{app:knn}
+\label{sec:knn}
 
-Assume that $b$ is a real $d$-dimensional vector, $b \in \mc{B} \subseteq \Real^d$, and $m$ is a binary indicator, $m \in \mc{M} = \{0,1\}$.
+% Assume that $b$ is a real $d$-dimensional vector, $b \in \mc{B} \subseteq \Real^d$, and $m$ is a binary indicator, $m \in \mc{M} = \{0,1\}$.
 Then, further assume that we have observed a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled identically and independently from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$.
@@
 
-% This algorithm also readily generalizes to the $C$-way classification problem, where there $C \geq 2$ classes.
-Instead of finding the majority class of the $k_n$ nearest neighbors, one instead finds the plurality class, that is, the class with the most training samples within the $k_n$ nearest neighbors.
-This algorithm generalizes even further, when $\mc{M}$ is no longer discrete, but rather continuous, for instance, $\mc{M}=\Real$.
-In such a scenario, this approach is called $k_n$ nearest neighbor \emph{regression}.
-
-% Another algorithm, proven to be universally consistent, is the $k_n$ nearest neighbor algorithm \cite{Stone77}, can rectify these limitations of the Bayes plugin algorithms.
-
 The $1$-nearest neighbor ($1$-NN) classifier works as follows.
 Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
-Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, where parenthetical indices, $(i)$, indicate rank order.
-One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
+Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and their corresponding mental properties, , $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order.
+%One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
 The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $m=m_{(1)}$.
@@
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
+When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into finite Euclidean space, in which case Stone's theorem applies.
 % section _k_n_nearest_neighbor_algorithm (end)
@@
 \section{Simulation} % (fold)
-\label{sub:simulation}
+\label{sec:sim}
 
@@
 Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
-To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$.
+To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$, and $E_{ij}$ is the number of synapses between neuron $i$ and $j$ for a particular sample.
 The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $\tr{z_{ij}}$ uniformly sampled from $[-5,5]$.
@@
 
-% \section{$k_n$-nearest neighbor universal consistency for graphs}
-% \label{proof}
-% 
-% Assume first that all graphs are simple \tr{(meaning undirected with no loops and binary edges)}, on the same set of vertices, and that the graphs are labeled so that we know which vertex in one graph corresponds to which vertex in another.
-Then the Frobenius distance function $d(b_1,b_2)$ can be written in terms of the associated adjacency matrices $A_1$ and $A_2$: $d(b_1,b_2) = ||A_1-A_2||_F$.
-If the graphs are identical, then $d(b_1,b_2) = 0$, and if the graphs are different, then $d(b_1,b_2) \geq 1$.
-Since the space $\mB$ is finite, $n$ large enough guarantees that with probability approaching unity at least $k_n$ training samples coincide with each atom, so long as $k_n/n \rightarrow 0$.
-Then $k_n \rightarrow \infty$ guarantees that the nearest neighbor vote-winner for each atom will eventually coincide with Bayes' choice, yielding universal consistency.
-% 
-% In the foregoing argument, there exists a smallest non-zero atomic probability $p_{min}$, and ``$n$ large enough'' is driven by this probability.
-Generalizing to countable $\mB$ with discrete weights, we see that given $\delta > 0$, there is a finite set $S$ with $\PP[S]>1-\delta$ and smallest atomic probability $p_{min}$, so that $L_{\PP}(g_n) \rightarrow c \leq L_{\PP}(g^*) + \delta$, yielding universal consistency.
-% 
-% If the graphs may have different numbers of vertices, and are unlabeled, we consider the isomorphism-matching Frobenius norm.
-Assume without loss of generality that $b_1$ has at least as many vertices as $b_2$, and write $A_2^P$ for the adjacency matrix associated with $b_2$ ``padded'' to include extra isolated vertices so that $A_2^P$ is the same size as $A_1$.
-Then $d(b_1,b_2) = \min_Q ||Q A_1 Q^T - A_2^P||_F$ where the minimum is taken over all permutation matrices \cite{HornJohnson90}.
-Under the equivalence relation induced by this isomorphism-matching, the foregoing universal consistency argument holds.
-% %\tr{Trivially, this argument also applies to graphs with (self-) loops, and
-% %directed, colored, or weighted edges, and hypergraphs.}
-% 
-% Several points of note:  isolated vertices are ignored in our equivalence relation; the class-conditional signal is entirely encompassed by the connectivity structure; the graph isomorphism problem is computationally hard \cite{ConroyLouis97,ZaslavskiyVert08}; and the argument employed here does not capture the concept of ``nearness implies likelihood of similar class''---we simply rely on atomic behavior.
-% 
-% Finally, the above proof can be straightforwardly generalized to utilize any matrix norm, $\norm{b_1-b_2}_A=\norm{A(b_1-b_2)}$, assuming that $A A\T$ is positive definite (see \cite{DGL96} pg.
-455 for proof when $b_i$'s are vectors).
-Furthermore, we can relax the constraint that mental properties are finite, that is, we can allow any $|\mM| = \infty$ (see \cite{Stone77} for proof when $b_i$'s are vectors).
-The proof for the case when $b$ is an adjacency matrix follows immediately upon first defining a bijective embedding of the matrix into a vector space, and then embedding each adjacency matrix in that space.
-
 \section*{Acknowledgments}

### 2010-12-31 supervenience.tex
new version, incorporating reviewer comments
finished cep mods

@@
 
-To enable addressing the question of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
+To enable addressing the question, through empirical means, of whether the mind supervenes on the brain, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
@@
 The potential significance of this work can be divided into distinct disciplines.
-To the philosopher, this work demonstrates an example of morphing philosophical conjectures into statistical hypotheses that are amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
+To the philosopher, this work demonstrates an example of morphing philosophical conjectures into statistical hypotheses that are amenable to experimental investigations, allowing the philosopher to add empirical support to her rational arguments.
 To the statistician, this work points out the limitations of hypothesis testing in a novel domain, and is suggestive of more general limitations previously (perhaps) under-appreciated.
@@
 %Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-Thus%, the above mind brain supervenience conjecture may be formally stated
-: $m \neq m' \implies b \neq b' \, \forall (m,b) \neq (m',b') \in \mc{M} \times \mc{B}$, where $(m,b)$ and $(m',b')$ are mind-brain pairs.
+Thus, the above supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
+: $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
 This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details).
@@
 \label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$, for all $(m,b) \neq (m',b') \in \mc{M} \times \mc{B}$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
+%, $\forall$ $(m,b), (m',b') \in \mc{M} \times \mc{B}$.
 \end{defi}
@@
 Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as:
+Define misclassification rate as
 % \begin{align}
@@
 % $L_{\PP}(g)$ therefore effectively 
-which counts the fraction of time $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
+which denotes the probability that $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
 % \begin{align}
@@
 Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
-The primary result of casting supervenience is a statistical framework is the following theorem:
+The primary result of casting supervenience as a statistical framework is the following theorem:
 \begin{thm}
@@
 
-Thus, the above arguments shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
+Thus, the above argument shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
 Further, statistical supervenience can be thought of as a constraint on the possible models.
@@
 % \EE[L_{\PP}(g_n)] 
-% &\approx  \hL^{n'}_{F}(g_n) = \sum_{\mc{T}_{n-n'}} \PP[g_n(B)=M | \mc{T}_{n-n'}] \PP[\mc{T}_{n-n'}],
+% &\approx  \hL^{n'}_{\PP}(g_n) = \sum_{\mc{T}_{\mt{n}}} \PP[g_n(B)=M | \mc{T}_{\mt{n}}] \PP[\mc{T}_{\mt{n}}],
 % \end{align}
-$\hL^{n'}_{F}(g_n) = \PP[g_n(B)=M | \mc{T}_{n-n'}]$,
-where $n'<n$ is the number of held-out training samples (samples not used to obtain $g_n(\cdot)$).
-%, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{n-n'}$, such that $\hL^{n'}_F(g_n)$ converges.  
- % $n' \hL^{n'}_{F}(g_n)$ is t
-The approximate expected number of misclassified minds
-%. Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{F}(g_n)$ 
+$\hL^{n'}_{\PP}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
+where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
+%, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{\mt{n}}$, such that $\hL^{n'}_{\PP}(g_n)$ converges.  
+ % $n' \hL^{n'}_{\PP}(g_n)$ is t
+The approximate  number of misclassified minds
+%. Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{\PP}(g_n)$ 
 therefore
 % and 
-has a binomial distribution:  $n' \hL^{n'}_{F}(g_n) \sim \text{Binomial}(n',L_{\PP}(g_n))$. %, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect.
+has a binomial distribution:  $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{\PP}(g_{\mt{n}}))$.
+%, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect.
 
@@
 Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-We reject for  values of the test-statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{F}(g_n) <c_{\alpha}(n',\varepsilon)$.
+We reject for  values of the test-statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
 The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
@@
 
-\section{A thought experiment to guarantee unity power} % (fold)
+\section{A thought experiment demonstrating unity power} % (fold)
 \label{sub:uc}
@@
 Therefore, %Theorem \ref{thm1} applies and
-the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
+the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
@@
 Therefore, %Theorem \ref{thm1} applies and
-% the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mental/brain property pair, if indeed $\varepsilon$-supervenience holds.
+% the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
 This logic holds for directed graphs or multigraphs or hypergraphs with edge weights, with or without labeled vertices.
@@
 
-We have introduced the notion of $\eps$-supervenience, which states that the misclassification rate for any mental/brain property pair is less than $\eps$.
-Furthermore, we have shown that $k$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mental/brain pair properties.
-Furthermore, this is a one-sided test, so although power converges to unity, one can never determine whether more data is necessary to get a lower $p$-value, or the particular $\eps$-supervenience does not hold.
+We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind/brain property pair is less than $\eps$.
+Furthermore, when we restrict to the setting of Thought Experiment 1, we have shown that $k$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
+Alas, this is a one-sided test, so although power converges to unity, one can never determine whether more data is necessary to get a lower p-value, or the particular $\eps$-supervenience does not hold.
 
@@
 Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$), for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under an $L_p$ norm may have a rate of convergence so slow and a computational demand so high as to be impractical (though, see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
+Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (though, see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
 Whichever classifier is used, it is likely to benefit from a large amount of domain specific knowledge, which the proposed classifier completely neglects.
@@
 
-A central (perhaps \emph{the} central) quest in much of neuroscience, psychology, and cognitive science is to discover the brain properties that subvene under various mental properties, although questions are rarely cast in the supervenience formalism.
+A central (perhaps \emph{the} central) quest in much of neuroscience, psychology, and cognitive science is to discover the brain properties that subvene under various mental properties, although questions are rarely cast within a  supervenience formalism.
 Moreover, the particular brain properties that are often believed to subvene under these mental properties are neural circuits, or brain-subgraphs.
@@
 This practice transcends the evolutionary hierarchy of neuroscientific research.
-For instance, in the invertebrate literature, vertices correspond to particular labeled neurons, and edges correspond to synapse \cite{NorthGreenspan07}.
-In the vertebrate literature, vertices tend to correspond to types of neurons in particular regions, and edges correspond to tendencies of connections \cite{Shepherd04}.
-For primates \cite{Felleman_VanEssen91} and humans \cite{Mori05} vertices tend to represent functionally distinct neuroanatomical regions, and edges represent regional interconnectivity.
+For instance, in the invertebrate literature, vertices correspond to particular labeled neurons, and edges correspond to synapses \cite{NorthGreenspan07}.
+In the vertebrate literature, vertices often correspond to types of neurons in particular regions, and edges correspond to tendencies of connections \cite{Shepherd04}.
+For primates \cite{Felleman_VanEssen91} and humans \cite{Mori05} vertices frequently represent functionally distinct neuroanatomical regions, and edges represent regional interconnectivity.
 Furthermore, this practice also transcends analytical background, including anatomists \cite{Abeles91}, philosophers \cite{Koch_Davis94}, statisticians \cite{Rao_Lewicki02}, and physicists \cite{Chow_Dalibard03}.
-The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (perhaps causality, not supervenience, is the true desideratum).
+The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (although perhaps causality, not supervenience, is the true desideratum).
+Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neuropsychological investigations.
 
@@
 % While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality.
-Assuming we have confirmed $\MeB$ for a particular mental/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.
+Assuming we have confirmed $\MeB$ for a particular mind/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.
 % 
@@
 
-In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has seen an explosion of studies investigating the human connectome, and relating connectomes to cognitive properties \cite{Jones10}.
+In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has seen an explosion of studies investigating the human connectome, and relating connectomes to cognitive properties.
+% \cite{Jones10}.
 These studies can all be thought of as testing various $\eps$-supervenience hypotheses.
 For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenics and controls \cite{ArdekaniSzeszko10}.
-As the resolution and signal-to-noise ratio of magnetic resonance imaging continues to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
+As the resolution and signal-to-noise ratio of magnetic resonance imaging continue to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
 Furthermore, the utilization of other imaging technologies, such as polarized light imaging \cite{PalmAmunts10} and high-throughput electron microscopy \cite{DenkHorstmann04,HayworthLichtman06}, will continue to improve the effective resolution of these inferred connectomes from human brains.
-While determining whether a particular individual knows calculus might be quite distant, many other cognitive and psychological supervenience hypothesis have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be vanishing.
+While determining from a brain scan whether a particular individual knows calculus might be quite distant, many other cognitive and psychological supervenience hypotheses have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be diminishing.
 
@@
 
-Recently, several authors have suggested the possibility that brains are better characterized as quantum networks, instead of classical networks \cite{Penrose,check_book}.
-The above results hold regardless of whether computations in the brain are quantum or classical, as quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
+Recently, several authors have suggested the possibility that brains are better characterized as quantum computers, instead of classical computers \cite{Penrose99,Satinover02}.
+Assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
+This follows because quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
 This means, if the above analysis failed to reject the null at level $\alpha$, the interpretation does not change if one assumes quantum versus classical computations.
@@
 
-In this appendix we describe several kinds of relations that are not implied by supervenience.
+In this appendix we aim to provide more intuition regarding supervenience, by discussing the limitations and extent of its implications.
 
@@
 An injective relation is any relation that preserves distinctness.
-Thus if minds are injective on brains, then $b\neq b' \implies m \neq m' \, \forall (m,b), (m',b') \in \mc{M} \times \mc{B}$ (note that the directionality of the implication has been switched relative to supervenience).
+Thus if minds are injective on brains, then $b\neq b' \implies m \neq m'$ (note that the directionality of the implication has been switched relative to supervenience).
 For instance, it might be the case that a brain could change without the mind changing.
 Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
-It is possible that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
+It is possible (likely?) that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
 In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
@@
 
-What supervenience does imply, however, is the following.
-Imagine finding two different minds.
-If $\MeB$, then the brains subvening under those two minds must be different.
-In other words, there cannot be two different minds, either of which could supervene on a single brain.
-Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
-
 \begin{figure}[htbp]
@@
 
+What supervenience does imply, however, is the following.
+Imagine finding two different minds.
+If $\MeB$, then the brains subvening under those two minds must be different.
+In other words, there cannot be two different minds, either of which could supervene on a single brain.
+Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
+
 Note that statistical supervenience is distinct from statistical correlation.
 \emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
-If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_F \mM$.
+If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_{\PP} \mM$.
 Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
@@
 
-% Assume that $b$ is a real $d$-dimensional vector, $b \in \mc{B} \subseteq \Real^d$, and $m$ is a binary indicator, $m \in \mc{M} = \{0,1\}$.
-Then, further assume that we have observed a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled identically and independently from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$.
+Consider the following problem setup.
+We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangably from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
 A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim \PP[M,B]$, and $m$ is simply unobserved.
+It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim \PP[M,B]$, and $m$ is unobserved.
+Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$.
 
@@
 The $k_n$ nearest neighbor is a straightforward generalization of this approach.
-It says that the test mind is in the same class as which ever class is the majority class of the $k_n$ nearest neighbors, $m=\II\{\sum_{i=1}^{k_n} m_{(i)} > k_n/2\}$.
+It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $m=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
 Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
@@
 Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone' original proof, the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone' original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
 When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into finite Euclidean space, in which case Stone's theorem applies.
+Stone's original proof applied to the scenario when $|\mc{M}|$ was infinite, resulting in a regression algorithm as well.
+
 % section _k_n_nearest_neighbor_algorithm (end)
@@
 Thus, one may design an experiment by describing the joint distribution $\PP[M,B]$ via class-conditional distributions $\PP[B|M=m_j]$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
-Here the mental property corresponds to the C.~elegans exhibiting or not exhibiting a particular behavior (e.g., response to an odor).
+Here the mental property corresponds to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
 
-Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$ and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
-To generate the data, we let the class-conditional random variable $E_{ij} | M=m_0$ be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$, and $E_{ij}$ is the number of synapses between neuron $i$ and $j$ for a particular sample.
-The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$ for neurons $i,j \in \mD$, where $\mD$ is the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}, with signal parameter $\tr{z_{ij}}$ uniformly sampled from $[-5,5]$.
+Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
+To generate the data, let $E_{ij}$ be an integer-valued random variable whose value indicates the number of synapses (edges) between neurons (vertices) $i$ and $j$.
+Let the class-conditional random variable $E_{ij} | M=m_0$  be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$.
+Let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
+Therefore, the distribution of these edges must differ between the two classes.
+The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$, where the signal parameter $z_{ij}=\eta$ for all edges not in $\mc{E}$, and $z_{ij}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{E}$.
+
 We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm.
@@
 \centering \includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.
-$\hL^{1000}_{\PP}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $\PP[M,B]$.
-Each dot depicts an estimate for $L_{\PP}(g_n)$; standard errors are $(L_{\PP}(g_n)(1-L_{\PP}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
+\caption{C.~elegans graph classification simulation results.  $\hL$  (the misclassification rate estimated upon with $1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $\PP[M,B]$.
+Each dot depicts an estimate for $L_{\PP}(g_{\mt{n}})$; standard errors are $(L_{\PP}(g_{\mt{n}})(1-L_{\PP}(g_{\mt{n}}))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$; $k_n = 53$; $\hL^{1000}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
 We reject $H_0: L_{\PP}(g^*) \geq 0.10$ at $\alpha=0.01$.
@@
 Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day.
+3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a C.
+elegans brain-graph within a day.
 Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
@@
 % \bibliography{biblist}
-\bibliography{/Users/jovo/Research/papers/meta/biblist}
+\bibliography{/Users/jovo/Research/latex/biblist}
 \bibliographystyle{ieeetr}

### 2011-02-17 PLoS_ONE/2010_07_15/cover_letter.tex
submitted to plos one, revised version
Signed-off-by: jovo <joshuav@jhu.edu>

--- 2010-08-17 PNAS/cover_letter.tex
@@
 \topmargin=-1.2in
-\textheight=10in
+\textheight=10.2in
 
@@
 
-On behalf of myself and my coauthors, I submit for your review a copy of our manuscript entitled, ``Are mental properties supervenient on brain properties?'' We hope that you will consider publishing this manuscript to the prestigious Proceedings of the National Academy of Sciences (PNAS).
+On behalf of myself and my coauthors, I submit for your review a copy of our manuscript entitled, ``Are mental properties supervenient on brain properties?'' We hope that you will consider publishing this manuscript to the prestigious Public Library of Science (PLoS) ONE journal.
 
@@
 Whereas several groups have recently proposed algorithms for classifying graphs, there exists very little theoretical work on the limitations of these tools.
-In contrast, our work is founded on an existence proof of a universally consistent classifier on graphs, the limitations of which are both known and well-understood.
-Although we only apply this classifier to simulated brain-derived graphs, the methods are applicable to many other kinds of graphical data.
+In contrast, our work is founded on a constructive proof of a universally consistent classifier on graphs, the limitations of which are both known and well-understood.
+Furthermore, the proof is shown to be only one-sided, meaning that we can never gain confidence in the negation of the null hypothesis.
+Although we only apply this classifier to simulated brain-derived graphs, the methods are applicable to many other kinds of graph data.
 The recent emergence of networks and graphs as the preferred representations of structured relationships in multiple scientific disciplines suggests that the development of statistical tools for rigorous analysis of network data is a pressing need for the scientific community at large.
@@
 However, without sophisticated statistical tools for the analysis of such data, to-date the findings have (mostly) been limited to linear correlations between scalar measures of connectivity and the brain function under investigation.
-The graph-theoretic approach proposed here allows for utilizing the entirety of the information obtained from such measures, rather than using simple descriptive statistics.
+The graph-theoretic approach proposed here allows investigators to utilize the entirety of the information obtained from such measures, rather than using simple descriptive statistics.
 
-In summary, we believe that the work described in this manuscript makes significant contributions in several contemporary fields of inquiry, and is therefore suited for a general-audience peer-reviewed journal such as PNAS.
+In summary, we believe that the work described in this manuscript makes significant contributions in several contemporary fields of inquiry, and is therefore suited for a general-audience journal such as PLoS ONE.
 %Your interest in the philosophical underpinnings of neuroscience, as well as in inferring causal relationships from neural data and generating large-scale simulations of neural populations suggests to us that this manuscript would of interest to you.

### 2011-02-17 PLoS_ONE/2010_07_15/plos_supervenience.tex
submitted to plos one, revised version
Signed-off-by: jovo <joshuav@jhu.edu>

--- 2010-03-29 manuscript/supervenience.tex
@@
-%\input{/Users/joshyv/Research/misc/latex_paper.tex}
-\documentclass{article}
+% Template for PLoS
+% Version 1.0 January 2009
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+
+\documentclass[10pt]{article}
+
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
+
+% graphicx package, useful for including eps and pdf graphics
+% include graphics with the command \includegraphics
+\usepackage{graphicx}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+\usepackage{color}
+
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace} 
+%\doublespacing
+
 \usepackage{fancyhdr}
@@
 \newtheorem{thex}{Thought experiment}
-\lhead{Vogelstein JT, et al}
-\rhead{Neurocognitive Graph Theory}
 
-\title{Are mental properties supervenient on brain properties?}
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
 
-\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^{1,2}$, Carey E. Priebe$^1$\\
-$^1$Department of Applied Mathematics and Sciences, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% Use the PLoS provided bibtex style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
+
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
+
+%% END MACROS SECTION
 
@@
 
-\maketitle
-%\tableofcontents
-\begin{abstract}
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf{Are mental properties supervenient on brain properties?}
+}
+% Insert Author names, affiliations and corresponding author email.
+\\
+Joshua T.~Vogelstein$^{1,\ast}$, 
+R.~Jacob Vogelstein$^{1,2}$, 
+Carey E.~Priebe$^{1}$
+\\
+\bf{1} Department of Applied Mathematics and Sciences, Johns Hopkins University, Baltimore, MD, USA
+\\
+\bf{2} National Security Technology Department, Johns Hopkins University Applied Physics Laboratory, Baltimore, MD, USA
+\\
+$\ast$ E-mail: joshuav@jhu.edu
+\end{flushleft}
+
+% Please keep the abstract between 250 and 300 words
+\section*{Abstract}
 
@@
 
-\end{abstract}
+The potential significance of this work can be divided into distinct disciplines.
+To the philosopher, this work demonstrates that philosophical conjectures can be morphed into statistical hypotheses, amenable to experimental investigations, allowing the philosopher to add empirical support to their rational arguments.
+To the statistician, herein lies the first proof to our knowledge of the existence of a universally consistent classifier on graphs, and a constructivist one at that.
+To the neuroscientist, a theoretically possible experiment is proposed to garnish support for a hypothesis that is widely believed: that mental properties supervene on brain properties.
 
-\vspace*{0.5 in}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLoS ONE authors please skip this step. 
+% Author Summary not valid for PLoS ONE submissions.   
+% \section*{Author Summary}
+
+\section*{Introduction}
 
@@
 Fifth we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.}
+
+% Results and Discussion can be combined.
+\section*{Results}
 
@@
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
+Furthermore, Appendix 1 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
+\end{thex}
 
-\end{thex}
+\section*{Discussion}
 
@@
 
-%% == end of paper:
-
-%% Optional Materials and Methods Section
-%% The Materials and Methods section header will be added automatically.
-
-%% Enter any subheads and the Materials and Methods text below.
-%\begin{materials}
-% Materials text
-%\end{materials}
-
-%% Optional Appendix or Appendices
-%% \appendix Appendix text...
-%% or, for appendix with title, use square brackets:
-%% \appendix[Appendix Title]
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+% \section*{Materials and Methods}
 
@@
 
-Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day.
-Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
-
-\paragraph{Acknowledgments}
-The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein. This work was supported in part by the NSA Research Program in Applied Neuroscience.
-% \end{acknowledgments}
-
-%\subparagraph*{References}
-\bibliography{biblist}
-\addcontentsline{toc}{section}{References}
-%\bibliography{Science}
-\bibliographystyle{ieeetr}
-% \bibliographystyle{nature}
-
 \begin{figure}[h!]
-\centering \includegraphics[width=.9\linewidth]{Lhatplot}
+\centering \includegraphics[width=.5\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.
@@
 
+Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
+3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a brain-graph within a day.
+Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
+
+% Do NOT remove this, even if you are not including acknowledgments
+\section*{Acknowledgments}
+
+The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.
+
+% \section*{References}
+% The bibtex filename
+\bibliography{biblist}
+\bibliographystyle{plain}
+
+% \clearpage
+% \section*{Figure Legends}
+% 
+% Figure 1: C.~elegans graph classification simulation results.
+$\hL^{1000}_{F}(g_n)$ is plotted as a function of class-conditional training sample size $n_j$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{BM}$.
+Each dot depicts an estimate for $L_{F}(g_n)$; standard errors are $(L_{F}(g_n)(1-L_{F}(g_n))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$ ; $k_n = 53$ ; $\hL^{1000}_{F}(g_n) = 0.057$; standard error less than 0.01.
+We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
+$L_{F}(g^*) \approx 0$ for this simulation.
+% }
+
+% \section*{Tables}
+%\begin{table}[!ht]
+%\caption{
+%\bf{Table title}}
+%\begin{tabular}{|c|c|c|}
+%table information
+%\end{tabular}
+%\begin{flushleft}Table caption
+%\end{flushleft}
+%\label{tab:label}
+% \end{table}
+
 \end{document}

### 2011-02-17 PLoS_ONE/2011_02_17/supervenience.tex
submitted to plos one, revised version
Signed-off-by: jovo <joshuav@jhu.edu>

--- 2010-12-31 supervenience.tex
@@
 \usepackage{algorithmic}
-\usepackage{times}
+% \usepackage{times}
 \usepackage{fancyhdr}
@@
 \usepackage{color}
+\usepackage[T1]{fontenc}
+\usepackage[scaled]{helvet}
+\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
 \pagestyle{fancy}
@@
 \newtheorem{thm}{Theorem}
-\newtheorem{thex}{Thought experiment}
+\newtheorem{thex}{\emph{Gedanken} experiment}
 \lhead{Vogelstein JT, et al}
@@
 
-The ``mind-brain supervenience'' \tr{conjecture} suggests that all mental properties (e.g.\ consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
-The validity of this \tr{conjecture} has been argued in philosophical terms for over 2,500 years.
-\tr{Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.}
+	The ``mind-brain supervenience'' conjecture suggests that all mental properties (e.g.
+consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
+The validity of this conjecture has been argued in philosophical terms for over 2,500 years.
+Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.
 
-To enable addressing the question, through empirical means, of whether the mind supervenes on the brain, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
-To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
-The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or \emph{connectome}), and $\varepsilon$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate $\varepsilon > 0$, regardless of the relationship between the two.
-%In addition to the theoretical results, we show via simulation that given reasonable assumptions about class conditional probabilities and the amount of data available, the thought experiment can actually be conducted on a simple organism, \emph{Caenorhabditis elegans}, with currently available technology.
+	To address the question of whether the mind supervenes on the brain, we here frame a supervenience hypothesis in rigorous mathematical terms.
+Specifically, we propose a modified version of supervenience (called $\eps$-supervenience) that is amenable to experimental investigation and statistical analysis.
+To illustrate this approach, we perform a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\eps$-supervenience.
+The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or connectome).
+$\eps$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate > 0, regardless of the relationship between the two.
 
-The potential significance of this work can be divided into distinct disciplines.
-To the philosopher, this work demonstrates an example of morphing philosophical conjectures into statistical hypotheses that are amenable to experimental investigations, allowing the philosopher to add empirical support to her rational arguments.
-To the statistician, this work points out the limitations of hypothesis testing in a novel domain, and is suggestive of more general limitations previously (perhaps) under-appreciated.
-To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be considered supervenience investigations, further motivating brain-graph related research paradigms.
+	To the philosopher, this work shows how philosophical conjectures can be transformed into statistical hypotheses that are amenable experimental investigation.
+This allows the philosopher to gain empirical support for her rational arguments.
+To the statistician, this work points out the limitations of hypothesis testing; and suggests that some of these limitations have not previously been fully appreciated.
+To the neuroscientist, this work indicates that much of contemporary research can be framed in terms of investigating supervenience.
+This should provide further motivation for cross-disciplinary research between neuroscientists and statistical graph-theoreticians.
 
@@
 
-The primary contributions of this work are as follows.
-A notion of supervenience amenable to empirical investigation is formally introduced.
+The primary contributions of this work flow from our introduction of a notion of supervenience amenable to empirical investigation.
 This renders the mind-brain dualism debate a hypothesis, rather than an assumption, both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
@@
 Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.
-This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, suggesting avenues for future research.
+This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, and suggests avenues for future research.
 
@@
 \end{quotation}
-This conjecture may be concisely and formally stated.
-Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
+This conjecture may be concisely and formally stated thusly: Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
 % The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
@@
 %Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-Thus, the above supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
+The supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
 : $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
-This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details).
+This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details and some examples).
 
@@
 If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.
-This means that there exists a unique mapping from each brain to a single mind.
-In other words, one can in principle construct a function $g(b): \mB \mapsto \mM$, that is a deterministic mapping from brains to minds.
-It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind (see Appendix \ref{sec:relations} for examples).
+This means that there exists a deterministic function mapping each brain to its supervening mind, $g(\cdot): \mB \mapsto \mM$; therefore, one could in principle construct this function.
+It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.
 Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
@@
 
-Thus, the above argument shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
-Further, statistical supervenience can be thought of as a constraint on the possible models.
-Specifically, let $\mc{P}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{P}_s$ be subset of models for which supervenience holds.
-Then, $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
+The above argument shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
+Statistical supervenience can therefore be thought of as a constraint on the possible distributions on minds and brains.
+Specifically, let $\mc{P}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{P}_s$ be subset of distributions for which supervenience holds.
+Theorem \ref{thm1} implies that $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
 
@@
 
-While the above theorem is of potential theoretical interest, because the arguments rest on knowing $\PP[M,B]$ and $g^*$, which are typically unknown, they are pragmatically useless.
+Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $\PP[M,B]$ and $g^*$, rendering them useless pragmatically.
 However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
-Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to model $\PP[M,B]$.
-Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
+Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $\PP[M,B]$.
+Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
 Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
@@
 
-Given this relaxation, consider the problem of testing for $\eps-$supervenience.
+Given this relaxation, consider the problem of testing for $\eps$-supervenience.
 %First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
@@
 Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-We reject for  values of the test-statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
+We reject for  values of the test   statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
 The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
@@
 % \end{thm}
+Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
 
@@
 
-\section{A thought experiment demonstrating unity power} % (fold)
+\section{A \emph{gedanken} experiment demonstrating consistency and unity power} % (fold)
 \label{sub:uc}
@@
 
-To ensure consistency and therefore unity power, the classifier $g_n(\cdot)$ must be able to converge to the truth, regardless of the true model, $\PP$.
-We therefore make explicit a model for brain, and show that under this very general model, universally consistent classifiers are available.
+To ensure consistency and therefore unity power, the classifier $g_n(\cdot)$ must be able to converge to the truth, regardless of the true distribution, $\PP$.
+We therefore make explicit a model for brains, and show that under this very general model, universally consistent classifiers are available.
 
@@
 We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind/brain property pair is less than $\eps$.
-Furthermore, when we restrict to the setting of Thought Experiment 1, we have shown that $k$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
-Alas, this is a one-sided test, so although power converges to unity, one can never determine whether more data is necessary to get a lower p-value, or the particular $\eps$-supervenience does not hold.
+Furthermore, when we restrict the space of minds and brains to the setting of \emph{gedanken} experiment 1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
+Alas, this is a one-sided test, so although power converges to unity, one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
 
@@
 
-Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans when the vertices represent individual neurons is practically possible for any particular mental property at this time.
+Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans is practically possible for any particular mental property at this time (at least when the vertices represent individual neurons).
 Rather, the claim is that \emph{if} one had sufficient data (a very large number of exchangeable mind/brain property pairs), then \emph{in theory}, some level $\eps$-supervenience hypothesis test could be performed.
-Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$), for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (though, see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
-Whichever classifier is used, it is likely to benefit from a large amount of domain specific knowledge, which the proposed classifier completely neglects.
+Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$) for any particular dataset.
+Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (but see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
+Whichever classifier is used, it is likely to benefit from a large amount of domain-specific knowledge, which the proposed classifier completely neglects.
 
@@
 The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (although perhaps causality, not supervenience, is the true desideratum).
-Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neuropsychological investigations.
+Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neurcognitive investigations.
 
@@
 
-In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has seen an explosion of studies investigating the human connectome, and relating connectomes to cognitive properties.
+In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has driven an explosion of studies investigating the human connectome and relating connectomes to cognitive properties \cite{Sporns10}.
 % \cite{Jones10}.
-These studies can all be thought of as testing various $\eps$-supervenience hypotheses.
-For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenics and controls \cite{ArdekaniSzeszko10}.
+Many of these studies can be framed as $\eps$-supervenience hypothesis tests.
+For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenic individuals and normal (control) individuals \cite{ArdekaniSzeszko10}.
 As the resolution and signal-to-noise ratio of magnetic resonance imaging continue to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
@@
 
-\subsection{Quantum networks} % (fold)
-\label{par:quantum_networks}
+\subsection{Alternative explanations} % (fold)
 
-Recently, several authors have suggested the possibility that brains are better characterized as quantum computers, instead of classical computers \cite{Penrose99,Satinover02}.
-Assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
-This follows because quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
-This means, if the above analysis failed to reject the null at level $\alpha$, the interpretation does not change if one assumes quantum versus classical computations.
+Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{gedanken} experiment in Section \ref{sub:uc}, a natural question to raise is: what are the conceivable alternative hypotheses?
+We consider three such alternatives.
 
-% paragraph quantum_networks (end)
+First, perhaps brains are more accurately characterized as quantum networks over classical networks.
+Several authors have suggested that brains have certain hypercomputational properties that classical computers could not achieve \cite{Penrose99,Satinover02}.
+However, assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
+This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
+This means that if the above analysis failed to reject the null hypothesis at level $\alpha$, it will fail regardless of whether one assumes quantum or classical computations.
 
-\subsection{Stochastic Supervenience} % (fold)
-\label{par:stochastic_supervenience}
+Second, perhaps minds stochastically supervene on brains \cite{Craver09}. While perhaps difficult to imagine, 
+much like a non-deterministic world was difficult to imagine prior to modern physics, it is not inconceivable that mental properties are only stochastically determined by physical ones.
 
-Possible explanations of how it might be the case that $\eps>0$ include stochastic supervenience \cite{Craver09}, and supernatural causal effects.
-Thus, the above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
+A third alternative hypothesis is supernatural causal effects.
+The above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
+
+It therefore seems that failing to reject the null hypothesis that a particular mental property $\eps$-supervenes on a particular brain property could potentially be explained by stochastic or supernatural forces, but not a quantum network brain model.
 
@@
 
-The above \emph{gedanken} experiment did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
-Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model, which determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understanding the neural activity ``code.''  In contrast, if mind-brain supervenience is true, this motivates a search for the neural connectivity ``code,'' an \emph{engram} for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
+The \emph{gedanken} experiment in Section \ref{sub:uc} did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
+Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model that determines the activity (in a potentially stochastic process).
+Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience holds, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
 
@@
 
-This thought experiment, together with (i) the formal definition of $\varepsilon$-supervenience, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
+This \emph{gedanken} experiment, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
 The above discussion suggests that many previously conducted investigations either assume supervenience, or test it.
@@
 
-In this appendix we aim to provide more intuition regarding supervenience, by discussing the limitations and extent of its implications.
+In this appendix we aim to provide more intuition regarding supervenience by discussing the limitations and extent of its implications.
 
@@
 For instance, it might be the case that a brain could change without the mind changing.
-Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
+Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$; or that a single additional synapse is formed between a pair of neruons.
 It is possible (likely?) that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
@@
 Rather, position can change with constant velocity, meaning without acceleration changing.
+Thus, to claim that something supervenes on another is not equivalent to claim that the two are identical.
 
@@
 %One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
-The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $m=m_{(1)}$.
+The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $\mh{m}=m_{(1)}$.
 The $k_n$ nearest neighbor is a straightforward generalization of this approach.
-It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $m=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
+It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
 Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
@@
 Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone' original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone's original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
 When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into finite Euclidean space, in which case Stone's theorem applies.
-Stone's original proof applied to the scenario when $|\mc{M}|$ was infinite, resulting in a regression algorithm as well.
+Stone's original proof applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
+
+Note that the above extension of Stone's original theorem to the graph domain implicitly assumed that vertices were labeled, such that elements of the adjacency matrices could easily be compared across graphs.
+In theory, when vertices are unlabeled, one could first map each graph to a quotient space invariant to isomorphisms, and then proceed as before.
+Unfortunately, solving the graph-matching problem is currently NP-Incomplete (meaning it is not known to be either P or NP) \cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging.
 
@@
 The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons.
-While the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}.
-Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}.
+Although the graph with these neurons as vertices and edges defined by chemical synapses between neurons is likely not identical across individuals, it appears to be reasonably consistent \cite{Durbin87}.
+Furthermore, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
 Thus, one may design an experiment by describing the joint distribution $\PP[M,B]$ via class-conditional distributions $\PP[B|M=m_j]$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
@@
 To generate the data, let $E_{ij}$ be an integer-valued random variable whose value indicates the number of synapses (edges) between neurons (vertices) $i$ and $j$.
-Let the class-conditional random variable $E_{ij} | M=m_0$  be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $0<\eta \ll 1$.
+Let the class-conditional random variable $E_{ij} | M=m_0$  be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $\eta=0.05$.
 Let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
@@
 The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $\PP[M,B]$ used here; however, we demand universal consistency.)
+(Better classifiers can be constructed for the joint distribution $\PP[M,B]$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
 

### 2011-02-17 other/supervenience 20110202 RJV.tex
submitted to plos one, revised version
Signed-off-by: jovo <joshuav@jhu.edu>

--- 2010-12-31 supervenience.tex
@@
 
-To enable addressing the question, through empirical means, of whether the mind supervenes on the brain, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
+To address the questions of whether the mind supervenes on the brain through empirical means, here we frame a supervenience hypothesis in rigorous mathematical terms and propose a modified version of supervenience (called $\varepsilon$-supervenience) that is amenable to empirical investigations and statistical analysis.
 To elucidate this approach, we posit a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\varepsilon$-supervenience.
@@
 To the statistician, this work points out the limitations of hypothesis testing in a novel domain, and is suggestive of more general limitations previously (perhaps) under-appreciated.
-To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be considered supervenience investigations, further motivating brain-graph related research paradigms.
+To the neuroscientist, cognitive scientist and psychologist, this work indicates that much of contemporary research can be framed in terms of investigating supervenience, further motivating brain-graph related research paradigms.
 
@@
 %Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-Thus, the above supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
+The supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
 : $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
@@
 
-While the above theorem is of potential theoretical interest, because the arguments rest on knowing $\PP[M,B]$ and $g^*$, which are typically unknown, they are pragmatically useless.
+Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $\PP[M,B]$ and $g^*$, rendering them useless pragmatically.
 However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
@@
 
-In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has seen an explosion of studies investigating the human connectome, and relating connectomes to cognitive properties.
+In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has driven an explosion of studies investigating the human connectome and relating connectomes to cognitive properties.
 % \cite{Jones10}.
 These studies can all be thought of as testing various $\eps$-supervenience hypotheses.
-For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenics and controls \cite{ArdekaniSzeszko10}.
+For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenic individuals and normal (control) individuals \cite{ArdekaniSzeszko10}.
 As the resolution and signal-to-noise ratio of magnetic resonance imaging continue to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
@@
 Assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
-This follows because quantum networks merely speed up computation for certain classes of problems, they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
-This means, if the above analysis failed to reject the null at level $\alpha$, the interpretation does not change if one assumes quantum versus classical computations.
+This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
+This means that if the above analysis failed to reject the null hypothesis at level $\alpha$, it will fail regardless of whether one assumes quantum or classical computations.
 
@@
 
-Possible explanations of how it might be the case that $\eps>0$ include stochastic supervenience \cite{Craver09}, and supernatural causal effects.
-Thus, the above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
+Possible explanations of how it might be the case that $\eps>0$ include stochastic supervenience \cite{Craver09} and supernatural causal effects.
+Thus, the above analysis could be considered an empirical test for whether we have ``souls'', or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
 
@@
 The above \emph{gedanken} experiment did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
-Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model, which determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understanding the neural activity ``code.''  In contrast, if mind-brain supervenience is true, this motivates a search for the neural connectivity ``code,'' an \emph{engram} for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
+Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model that determines the activity (in a potentially stochastic process).
+Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience is true, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
 
@@
 Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone' original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone's original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
@@
 The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons.
-While the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}.
+Although the graph with these neurons as vertices and edges defined by chemical synapses between neurons is not identical across individuals, it is reasonably consistent \cite{Durbin87}.
 Furthermore, these animals exhibit a rich behavioral repertoire that depends on circuit properties \cite{deBonoMaricq05}.

### 2011-03-04 PLoS_ONE/2011_02_17/superveniencerebuttal.txt
resubmitted to PLOS ONE and posted update on arxiv

Dear Prof Sporns and two helpful reviewers,

Thank you for your insightful comments.  Based on your input, we have significantly revised the text of our manuscript entitled "Are Mental Properties Supervenient on Brain Properties?"  Specifically, the manuscript is now re-organized to emphasize the flow of our thesis.  Moreover, we have clarified the notion of supervenience with an appendix, and expanded upon the relationship of this work to both previous and future work in the discussion.  We hope that you will find that our revisions adequate address your concerns.  Below we provide detailed responses to specific comments:

Reviewer #1:

Thank you for your helpful comments.  We agree that the original presentation of these ideas lacked some organization and clarity.  The main theorem is that supervenience, which is currently known mostly by philosophers, actually makes a very strong claim about statistical classification accuracy.  Only upon realizing this connection can hypothesis tests be created for supervenience.  Given the possibility of such a test, we show that not only do such tests exist, but also that a quite general model of minds and brains admits a universally consistent classifier with power converging to unity.  We believe that this idea is quite deep and interesting.  In particular, the manuscript explains how many previously conducted investigations can be framed as epsilon-supervenience hypothesis tests, unifying previously disparate work.  Moreover, these ideas further motivate the now burgeoning field of connectomics, which seems to operate on the assumption that a certain epsilon-supervenience holds for many mind-brain property pairs.  Thus, it is our belief that this manuscript formally ties together many contemporary and future neurocognitive investigations in previously unrecognized ways.

Reviewer #2:

Thank you for your helpful comments.  We address each separately.

First, regarding the very compact mathematical style, we have expanded to exposition of mathematical details in the main text with the intention that it will be more accessible to many interested philosophers and scientists, while recognizing that a certain level of proficiency with statistical inference and/or pattern recognition will likely be necessary to understand the mathematical details.

Second,  regarding the definition of supervenience, we have expanded the text in the main document and added an appendix (Appendix A) to further expand upon the nature of supervenience relations, as well as compare and contrast it with other relations by providing examples, as you suggested. We have also now provided more description (final paragraph of Appendix B) and a reference to further expound upon the notion of isomorphism that we are employing in this manuscript.

Third, regarding the proof of universal consistency, we have modified the text to demonstrate that our "proof" is merely a realization that the composition of (i) representing a graph by its adjacency matrix  such that it lives in finite dimensional Euclidean space and (ii) a isomorphism-matching operator extends Stone's original proof to this domain.  Therefore, we feel as though Stone's original proof is truly the workhorse here.  Appendix B merely explains how one can extend the proof to this domain.   Part of the beauty of this proof is that the the classifier works regardless of the structure of the brain-graph.

Finally, we fixed all the minor English flaws that you pointed out to us.

Please note that due to the significant restructuring of this manuscript, a red-lined version seemed inappropriate (in fact, upon making a red-lined version, essentially every line was either red or blue, which was effectively useless).  As described above, in addition to re-organizing, Appendix A is wholly novel, and the discussion section has been greatly expanded. We hope the reviewers will be easily able to find these modifications.

### 2011-03-04 PLoS_ONE/2011_02_17/supervenience.tex
resubmitted to PLOS ONE and posted update on arxiv

@@
 \newtheorem{thm}{Theorem}
-\newtheorem{thex}{\emph{Gedanken} experiment}
+\newtheorem{thex}{\emph{Gedankenexperiment}}
 \lhead{Vogelstein JT, et al}
@@
 
-\section{A \emph{gedanken} experiment demonstrating consistency and unity power} % (fold)
+\section{A \emph{Gedankenexperiment} demonstrating consistency and unity power} % (fold)
 \label{sub:uc}
@@
 We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind/brain property pair is less than $\eps$.
-Furthermore, when we restrict the space of minds and brains to the setting of \emph{gedanken} experiment 1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
+Furthermore, when we restrict the space of minds and brains to the setting of \emph{Gedankenexperiment}  1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
 Alas, this is a one-sided test, so although power converges to unity, one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
@@
 
-Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{gedanken} experiment in Section \ref{sub:uc}, a natural question to raise is: what are the conceivable alternative hypotheses?
+Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Section \ref{sub:uc}, a natural question to raise is: what are the conceivable alternative hypotheses?
 We consider three such alternatives.
@@
 
-The \emph{gedanken} experiment in Section \ref{sub:uc} did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
+The \emph{Gedankenexperiment} in Section \ref{sub:uc} did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
 Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model that determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience holds, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
+Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience holds, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise (note that supervenience allows for the particular mengram of a mental property to vary both across individuals and time).
 
@@
 
-This \emph{gedanken} experiment, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
+This \emph{Gedankenexperiment}, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
 The above discussion suggests that many previously conducted investigations either assume supervenience, or test it.

### 2011-05-11 nature_sci_rep/cover_letter.tex
added Nature Scientific Reports submission

% request for removal from telemarketing lists

\documentclass[10pt]{letter}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=6in
\topmargin=-1.2in
\textheight=10.2in

\name{	Joshua T.\ Vogelstein, PhD. \\ 
		Johns Hopkins University \\
		% Ph.D.\ Candidate in the Dept.\ of Neuro. \\
		% Postdoc., 
		Dept. Applied Math.\ \& Stat's \\
		3400 N.\ Charles St.\ \\
		Whitehead 105C \\
		Baltimore, MD 21210 \\
      	c: 443.858.9911 \\
		e: joshuav@jhu.edu
		}

\date{\today}

\begin{document}

\begin{letter}{	%Prof.\ Gerald Edelman, M.D.\, Ph.D.\ \\
               	% Chairman,The Scripps Research Institute \\
               	% 10550 N.\ Torrey Pines Rd.\ \\
               	% La Jolla, CA  92037-1000
}

\opening{Dear Editorial Board:}

On behalf of myself and my coauthors, I submit for your review a copy of our manuscript entitled, ``Are mental properties supervenient on brain properties?'' We hope that you will consider publishing this manuscript in Nature Scientific Reports.

At the core of this paper is a thought experiment in which we mathematically prove that if the graph corresponding to brain connectivity (i.e., the \emph{connectome}) is statistically related to a particular mental property (e.g., propensity for mathematics), one could build a classifier to determine whether any individual's brain exhibits that property with an arbitrarily small misclassification rate. The proof follows directly from the probabilistic theory of pattern recognition and is made possible by our novel exposition of universal consistency for a $k_n$-nearest neighbor classifer on graphs. In addition to these theoretical results, the manuscript describes a simulation that suggests how one could actually build such a classifier for the ``brain'' of \emph{Caenorhabditis elegans} using today's technology.

% Our results have many implications, some philosophical, some statistical, and some pragmatic. First, the philosophical implications: The relationship between the mind and brain has been investigated for millennia, using tools ranging from abstract arguments to microstimulation of individual neurons. However, until now, no one (to our knowledge) has proposed a framework that allows for empirical investigation of this relationship at the scale of the connectome. Such a framework will undoubtedly be of value as the multiple ongoing large-scale efforts to collect a connectome (cf.\ NIH Human Connectome Project) begin paying dividends.
% 
% In addition to the philosophical implications, our work has significant statistical implications. 
% Whereas several groups have recently proposed algorithms for classifying graphs, there exists very little theoretical work on the limitations of these tools. In contrast, our work is founded on a constructive proof of a universally consistent classifier on graphs, the limitations of which are both known and well-understood. Furthermore, the proof is shown to be only one-sided, meaning that we can never gain confidence in the negation of the null hypothesis. Although we only apply this classifier to simulated brain-derived graphs, the methods are applicable to many other kinds of graph data. The recent emergence of networks and graphs as the preferred representations of structured relationships in multiple scientific disciplines suggests that the development of statistical tools for rigorous analysis of network data is a pressing need for the scientific community at large.
% 
% Finally, the results described in this manuscript have pragmatic implications. Studies relating brain structure to brain function are being published with increasing frequency due to the increasing availability of diffusion-weighted magnetic resonance imaging and other connectivity-imaging technologies. However, without sophisticated statistical tools for the analysis of such data, to-date the findings have (mostly) been limited to linear correlations between scalar measures of connectivity and the brain function under investigation. The graph-theoretic approach proposed here allows investigators to utilize the entirety of the information obtained from such measures, rather than using simple descriptive statistics.

We believe that the work described in this manuscript makes significant contributions in several contemporary fields of inquiry, including neuroscience, philosophy, and statistics.  It is for these reasons that we believe it is particularly well suited for Nature Scientific Reports.  
% and is therefore suited for a general-audience journal such as Nature Scientific Reports. %Your interest in the philosophical underpinnings of neuroscience, as well as in inferring causal relationships from neural data and generating large-scale simulations of neural populations suggests to us that this manuscript would of interest to you.

As a scientific reviewer, we suggest Prof. Giorgio Ascoli (ascoli@gmu.edu), as he is a scholar in neuroscience and neuroinformatics, with a solid background in quantitative studies, with interests in whole brain analysis.  Please exclude Prof. Olaf Sporns, as our work and his closely overlap, and he has seen previous versions of this manuscript.

Please do not hesitate to contact us with any questions or comments.

\closing{Respectfully,}

\end{letter}

\end{document}

### 2011-05-11 nature_sci_rep/supervenience.tex
added Nature Scientific Reports submission

--- 2011-03-04 PLoS_ONE/2011_02_17/supervenience.tex
@@
 
-\title{Are mental properties supervenient on brain properties?}
-
-\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
-$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723
-}
-
 \begin{document}
 
-\maketitle
+	\begin{center}
+{\huge	Are mental properties supervenient on brain properties?}
+\end{center}
+
+\vspace{5px}
+
+\begin{center}
+{\large		
+Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
+	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+
+\end{center}
+
+\vspace{5px}
+
+% \maketitle
+% \date{\vspace{-5ex}}
 %\tableofcontents
-\begin{abstract}
+% \begin{abstract}
 
-	The ``mind-brain supervenience'' conjecture suggests that all mental properties (e.g.
-consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
-The validity of this conjecture has been argued in philosophical terms for over 2,500 years.
-Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.
-
-	To address the question of whether the mind supervenes on the brain, we here frame a supervenience hypothesis in rigorous mathematical terms.
+\noindent The``mind-brain supervenience'' conjecture suggests that all mental properties are derived from the physical properties of the brain.
+To address the question of whether the mind supervenes on the brain, we frame a supervenience hypothesis in rigorous statistical terms.
 Specifically, we propose a modified version of supervenience (called $\eps$-supervenience) that is amenable to experimental investigation and statistical analysis.
@@
 $\eps$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate > 0, regardless of the relationship between the two.
+This may provide motivation for cross-disciplinary research between neuroscientists and statisticians.
 
-	To the philosopher, this work shows how philosophical conjectures can be transformed into statistical hypotheses that are amenable experimental investigation.
-This allows the philosopher to gain empirical support for her rational arguments.
-To the statistician, this work points out the limitations of hypothesis testing; and suggests that some of these limitations have not previously been fully appreciated.
-To the neuroscientist, this work indicates that much of contemporary research can be framed in terms of investigating supervenience.
-This should provide further motivation for cross-disciplinary research between neuroscientists and statistical graph-theoreticians.
-
-\end{abstract}
+% \end{abstract}
 
@@
 
-\section{Introduction}
+\newpage
+% \section*{Introduction}
 
-Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
+\noindent Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
 Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
@@
 
-\section{Statistical supervenience} % (fold)
-\label{sec:preliminaries}
+\section*{Results}
 
-% \subsection{Mind Brain Supervenience} % (fold)
-% \label{sub:supervenience}
-
-% subsection supervenience (end)
-% The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
-Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}.
-Therefore, here we define minds, brains, and supervenience.
-% The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
+\subsection*{Statistical supervenience} % (fold)
 
@@
 \end{quotation}
-This conjecture may be concisely and formally stated thusly: Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
-% The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
-In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.
-%Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.
+
+\noindent This conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
+This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).
+To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
+
+Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
 Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-%The set of possible minds $\mM$ is also unrestricted.
-In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences.
-%Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-The supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
-: $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
-This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details and some examples).
+Let $\PP[M,B]$ indicate a joint distribution of minds and brains.
+Statistical supervenience can thusly be defined:
 
-To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
-Let $\PP[M,B]$ indicate a joint distribution of minds and brains.
-%Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$.
-% (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
-% \begin{enumerate}
-% 	\item $0 \leq \PP[B=b,M=m] \leq 1$ for all $(m,b) \in \mM \times \mB$
-% 	\item $\PP[\Omega]=1$ and $\PP[\emptyset]=0$, where $\Omega$ is the whole sample sample, that is, all elements, $(m,b) \in \mc{M} \times \mc{B}$
-% 	\item any countable sequence of pairwise disjoint elements, $(m_1,b_1), (m_2,b_2), \ldots$ satisfies \\
-% 	\mbox{$\PP[(m_1,b_1) \cup (m_2,b_2) \cup \ldots] = \sum_i \PP[(m_i,b_i)]$}.
-% \end{enumerate}
-%Given a distribution, one can then calculate any marginal or conditional distributions.
-For instance, $\PP[B]=\int_{m \in \mM} \PP[M,B]dm$, or $\PP[M|B]=\PP[M,B]/\PP[B]$.
-Having specified a model, statistical supervenience can be defined as follows:
 \begin{defi}
@@
 $\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
-%, $\forall$ $(m,b), (m',b') \in \mc{M} \times \mc{B}$.
 \end{defi}
-\noindent Statistical supervenience is therefore a probabilistic relation on sets (related to, but distinct from correlation; see Appendix \ref{sec:relations} for details).
+\noindent Statistical supervenience is therefore a probabilistic relation on sets (related to, but distinct from correlation; see Appendix 1 for details).
 
- % Consider the following analogy: one desires to understand the relationship between height and weight.
-In such a scenario, height, $h$, could be assumed to be an element from the set of all possible heights, $\mc{H}=(0,\infty)$ cm; and weight, $w$, could be assumed to be an element from the set of all possible weights, $\mc{W}=(0,\infty)$ kgs.
-Then, it may be assumed that an agent's height and weight is sampled identically and independently from some \emph{model}, that is $(h,w)\overset{iid}{\sim}\PP[H,W]=F_{HW}$.
-To gain insight into $F_{HW}$, one might then state hypotheses, collect data, and perform analyses.
-Implicit in the exposition of a model is that both objects, height and weight here, are \emph{random objects}.
-Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.
-% 
-% \subsection{Brain-graphs} % (fold)
-% \label{sub:brain_graphs}
-
-% subsection brain_graphs (end)
-
-% section preliminaries (end)
-
-%\section{Results}
-
-\section{Statistical supervenience is equivalent to perfect classification accuracy} % (fold)
+\subsection*{Statistical supervenience is equivalent to perfect classification accuracy} % (fold)
 \label{sub:theoretical_results}
 
-% subsection theoretical_results (end)
 If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.
 This means that there exists a deterministic function mapping each brain to its supervening mind, $g(\cdot): \mB \mapsto \mM$; therefore, one could in principle construct this function.
-It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.
-Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
-Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as
-% \begin{align}
-	$L_{\PP}(g) = \PP[g(B) \neq M]$, %= \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
-% \end{align}
-% where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.  
-% $L_{\PP}(g)$ therefore effectively 
-which denotes the probability that $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
-% \begin{align}
-	$g^* = \argmin_{g} L_{\PP}(g)$.
-% \end{align}
-% where $\mc{G}$ is the set of all possible classifiers.  
-Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
+The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L_{\PP}(g^*)$.
+The minimum misclassification rate is called Bayes error (see Methods for details).
+% It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.  
 The primary result of casting supervenience as a statistical framework is the following theorem: 
@@
 \end{thm}
-% \begin{thm} 
-% 	\label{thm1} 
-% 	Statistical supervenience implies that there exists a classifier that is \emph{always correct}, that is, $g(b)=m$, for all $(m,b) \in \mc{M} \times \mc{B}$.  
-% \end{thm}
 
@@
 
-\section{A hypothesis test for supervenience} % (fold)
+\subsection*{$\eps$-supervenience admits hypothesis testing} % (fold)
 \label{sub:hypothesis_testing}
@@
 However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
-Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $\PP[M,B]$.
-Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
-Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
-The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
-% Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
-% \begin{align}
-% \EE[L_{\PP}(g_n)] = \EE[\PP_F[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
-% \end{align}
-% Unfortunately, in practice, computing $\EE[L_{\PP}(g_n)]$, requires integrating over all possible training data corpuses of size $n$, and by definition, we only have access to a single training data corpus.
-We therefore define ``hold out'' misclassification performance:
-% \begin{align}
-% \EE[L_{\PP}(g_n)] 
-% &\approx  \hL^{n'}_{\PP}(g_n) = \sum_{\mc{T}_{\mt{n}}} \PP[g_n(B)=M | \mc{T}_{\mt{n}}] \PP[\mc{T}_{\mt{n}}],
-% \end{align}
-$\hL^{n'}_{\PP}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
-where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
-%, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{\mt{n}}$, such that $\hL^{n'}_{\PP}(g_n)$ converges.  
- % $n' \hL^{n'}_{\PP}(g_n)$ is t
-The approximate  number of misclassified minds
-%. Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{\PP}(g_n)$ 
-therefore
-% and 
-has a binomial distribution:  $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{\PP}(g_{\mt{n}}))$.
-%, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect.
-
 Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
@@
 Given this relaxation, consider the problem of testing for $\eps$-supervenience.
-%First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
-% Because we hope to reject the null, in favor of the alternative, l
 Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-We reject for  values of the test   statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
-The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
+Hold-out error, $\hL^{n'}_{\PP}(g_{\mt{n}})$, as defined in Methods, is a test statistic for this hypothesis.
+We reject for values of the test statistic lower than the critical value, that is, we reject if and only if $\hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
+The critical value is available under the least favorable distribution $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',\varepsilon)$.
 Thus, rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
 The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$.
-%Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{P}_\eps = \{\PP | L_{\PP}(g^*)< \eps\}$.
-One could then sort  $\eps$-supervenience subsets,  $\mc{P}_s \subseteq \mc{P}_{\eps} \subseteq \mc{P}_{\eps'} \subseteq \mc{P}$, for any $\eps < \eps'$.
 
-% subsection hypothesis_testing (end)
-
-% \section{Power and consistency} % (fold)
-% \label{ssub:power_and_consistency}
-
-Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis% (in other words, the probability that it will not make a Type II error)
-.
+Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis.
 Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
@@
 A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{\PP}(g_n)] \conv L_{\PP}(g^*)$ as $n\conv \infty$.
-%As the notation suggests, consistency of a classifier is a function of the true model, $\PP$.
-Without any prior knowledge of what the model might be, one desires a \emph{universally consistent} classifier, that is a classifier that is consistent for all $\PP \in \mc{P}$.
-% The above described hypothesis tests suggests the following theorem:
-% 
-% \begin{thm} \label{thm1} 
-% Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$.
-Furthermore, given a consistent classifier the power of the test converges to unity.
-% \end{thm}
 Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
@@
 Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, %.
-This means that we can never be confident that $\MeB$ does \emph{not} hold; rather,
-it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-%Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.
+For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
 Thus, without restrictions on $\PP[M,B]$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-%Given these limitations on even universal consistency, it is still the best one can hope for.
-Therefore, we hope to obtain a universally consistent classifier to test for $\eps$-supervenience.
 
-% subsubsection power_and_consistency (end)
-
-\section{A \emph{Gedankenexperiment} demonstrating consistency and unity power} % (fold)
+\subsection*{A \emph{Gedankenexperiment} demonstrating consistency and unity power} % (fold)
 \label{sub:uc}
-
-% Below, we will assume that brains are characterized by graphs, referred to hereafter as ``brain-graphs.''  This assumption follows naturally from the neuron-doctrine
-% % Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.
-For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.
-It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine''
-% \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex {network} of neurons, largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
-This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
-%Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.
-Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.
-Below, we formalize this concept to enable statistical testing in the graph domain.
-% Below, we will assume that $m$ is binary, that is, $m \in \{0,1\}$ (for example, $m$ might indicate whether the person knows calculus or not).
 
@@
 Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix \ref{sec:knn} for proof).
+A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 2 for proof).
 Therefore, %Theorem \ref{thm1} applies and
@@
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
-Furthermore, Appendix \ref{sec:knn} also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
+Furthermore, Appendix 2 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
 \end{thex}
 
-%(fold)
-% Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $V = \{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neural units, and arcs (or edges) $A=\{A_{ij}\}$, where $A_{ij}$ represents the connectivity from neural unit $V_j$ to $V_i$.
-For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that for all $i,j \in [n]$: (i) there are no self loops, $a_{ii}=0$, (ii) all edges are binary, $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric,  $a_{ij}=a_{ji}$.
-Further assume that the vertices are \emph{labeled}, meaning that there is a known one-to-one mapping from each vertex in any brain to a vertex in any other brain, and that each brain-graph has the same number of vertices.
-Thus, each brain-graph is characterized entirely by its adjacency matrix, $a \in \mc{A}$, where $\mc{A} \subseteq \{0,1\}^{n \times n}$.
-Further, let the mental property under investigation be binary, so $m \in \mc{M} = \{0,1\}$.
-The Bayes classifier is:
-% \begin{align}
-% 	\mh{m} = \argmax_m \PP[b,m] = \argmax_m \PP[b|m] \PP[m] 
-% \end{align}
-% where the likelihood term, $\PP[B|M]$, specifies the probability of observing any particular adjacency matrix for any class, and the prior term, $\PP[M]$, specifies the prior probability of either class.
-% 
-% Under the above model assumptions, the number of possible brain-graph/mental-property pairs is finite, that is, $|\mc{M} \times \mc{B}| = d < \infty$.
-Therefore, one can simply enumerate all possible brain-graphs for each class, such that each class is represented by a multinomial distribution.
-The maximum likelihood estimator (MLE) of the multinomial parameter for each class as guaranteed to converge to the true (but unknown) parameter.
-Formally, one can state $\mh{\PP}_{MLE}[B|M] \conv \PP[B|M]$ as $n \conv \infty$, almost surely.
-Furthermore, the MLE  estimator of the prior also converges to the true (but unknown) prior distribution, $\mh{\PP}_{MLE}[M] \conv \PP[M]$, as $n \conv \infty$, almost surely.
-% 
-% The Bayes plugin classifier is defined by:
-% \begin{align}
-% 	\mh{m} = \argmax_m  \mh{\PP}[b|m] \mh{\PP}[m] 
-% \end{align}
-% where $\mh{\PP}[B|M]$ and $\mh{\PP}[M]$ are plugin estimators.
-Since the MLE estimators defined above are plugin estimators, and consistent, one obtains a consistent classifier by plugging in the MLE parameters.
-This logic holds for any model in which the cardinality of the sample space is finite, that is $|\mc{M}\times \mc{B}| < \infty$.
-In particular, one can relax both the assumed symmetry and hollowness of the adjacency matrix, and this result still holds.
-Furthermore, one can allow for multi-graphs, in which there are a finite number of different kinds of edges (different colored edges), such that $a_{ij} \in \{0,1,\ldots, K\}$, with $K < \infty$, and the set of possible brain-graphs remains finite.
-Finally, one can allow $\mc{M}$ to be any finite space, not just binary.
-Note that in all these cases, the process one uses to enumerate and sort the brain-graph collections is essentially irrelevant, assuming the same process is used for each class.
-Further note that more robust estimators, such as certain M-estimators and the \emph{maximum a posteriori} estimators, also admit universally consistent classifiers, with perhaps faster convergence properties.
-% 
-% The above approach, however, is insufficient in several important cases.
-First, when the edges live in a continuous space, such as $a_{ij} \in \Real$, corresponding to weighted graphs, the MLE plugin is not well defined.
-In such scenarios, the $k_n$ nearest neighbor ($k$-NN) algorithm can be used (see Appendix 1 for a description of the $k$-NN algorithm) \cite{Stone77}.
-Although originally defined to operate on finite dimensional Euclidean space, the same algorithm can be applied here, once one has first embedded the brain-graphs into such a space.
-One possible embedding is to stack the columns of the adjacency matrix, yielding an $n^2$ element vector, instead of an $n \times n$ element matrix.
-Given such an embedding, the $k$-NN classifier may be used without modification, assuming one has defined a suitable distance operator (such as $L_2$ or $L_1$).
-This algorithm may also be used in the above cases where the support of the sample space is finite, and, convergence of the $k$-NN algorithm for those problems will be faster than the MLE plugins, whenever the distance used is ``appropriate'' (where appropriate is assessed by convergence rates under the true (but unknown) model).
-Furthermore, one can allow for multiple kinds of edges, as before, yielding adjacency tensors, with $a_{ij} \in \Real^d$.
-Or, one can allow the mental property to be multidimensional as well, $m \in \Real^l$.
-The $k$-NN algorithm converges in both these generalizations.
-Finally, one could allow the number of vertices to differ across brain samples.
-In such a scenario, one can ``pad'' the adjacency matrices of the smaller brain-graphs, and the result still holds.
-% 
-% The above universal consistency results hold for unlabeled graphs as well, with a minor modification.
-Assuming the brain-graphs are unlabeled, and we have a test brain-graph $b$, one could first align each training brain-graph with the test brain-graph, and then use the above algorithms.
-To align any pair of graphs, one can solve the graph isomorphism problem, $\mh{Q}_i=\argmin_{Q \in \mc{Q}} \norm{Q b_i Q\T - b}$, where $\mc{Q}$ is the set of all permutation matrices (that is, zero matrices with a single unit value in each row and column).
-Given $\mh{Q}_i$, one can compute $\mt{b}_i=\mh{Q}_i b \mh{Q}_i\T$, and then proceed as one would normally.
-Unfortunately, the graph isomorphism problem is known to be NP-incomplete (that is, it is not known to believe in either P or NP-complete) \cite{GareyJohnson79}, and therefore no polynomial time algorithm is available, making this approach quite time-consuming.
-% 
-% 
-% % subsection brain_graphs 
-% 
-% \section{Theoretical implications} % (fold)
-% \label{sub:implications}
-% 
-% Returning to the question of supervenience, consider the following thought experiment.
-Let the physical property under consideration be brain-graphs, and the mental property under investigation be binary (such as, knows calculus or not).
-Now, imagine collecting very large amounts of very accurate exchangeable brain-graph data and the associated mental property values.
-A $k_n$-nearest neighbor algorithm using an isomorphism-matching Frobenius norm is universally consistent.
-Therefore, %Theorem \ref{thm1} applies and
-% the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with edge weights, with or without labeled vertices.
-% 
-% 
-% % subsection implications (end)
-% 
-% 
-% % subsection practical_applications (end)
-%(end)
+\section*{Discussion}
 
-\section{Discussion}
-
-\subsection{Summary} % (fold)
-\label{par:summary}
+% \paragraph{Summary} % (fold)
+% \label{par:summary}
 
@@
 
-\subsection{Practical issues} % (fold)
-\label{par:practical_issues}
+% \paragraph{Practical issues} % (fold)
+% \label{par:practical_issues}
 
@@
 Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$) for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (but see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
+Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (but see Appendix 3 for a simulated example in which convergence is relatively fast).
 Whichever classifier is used, it is likely to benefit from a large amount of domain-specific knowledge, which the proposed classifier completely neglects.
 
-% paragraph practical_issues (end)
-
-% paragraph summary (end)
-
-\subsection{A unified quest} % (fold)
-\label{sub:practical_applications}
+% \paragraph{A unified quest} % (fold)
+% \label{sub:practical_applications}
 
@@
 
-% \subsection{Causality} % (fold)
-% \label{par:causality}
-% 
-% While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality.
-Assuming we have confirmed $\MeB$ for a particular mind/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.
-% 
-% % paragraph causality (end)
-
-\subsection{Human applications} % (fold)
-\label{par:human_applications}
+% \paragraph{Human applications} % (fold)
+% \label{par:human_applications}
 
@@
 
-% paragraph relationship_to_previous_work (end)
+% \paragraph{Alternative explanations} % (fold)
 
-\subsection{Alternative explanations} % (fold)
-
-Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Section \ref{sub:uc}, a natural question to raise is: what are the conceivable alternative hypotheses?
+Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Results, a natural question to raise is: what are the conceivable alternative hypotheses?
 We consider three such alternatives.
@@
 
-% paragraph stochastic_supervenience (end)
-
-\subsection{Dynamics vs. statics} % (fold)
-\label{par:dynamics_vs_statics}
+% \paragraph{Dynamics vs. statics} % (fold)
+% \label{par:dynamics_vs_statics}
 
@@
 
-% paragraph dynamics_vs_statics (end)
-
-\subsection{Concluding thoughts} % (fold)
-\label{par:concluding_thoughts}
-
-% paragraph concluding_thoughts (end)
+% \paragraph{Concluding thoughts} % (fold)
+% \label{par:concluding_thoughts}
 
@@
 
-\clearpage
-\appendix
+\section*{Methods}
+\label{sec:methods}
+% \subsection*{Supervenience} % (fold)
+% \label{sec:preliminaries}
 
-\section{Relations between sets} % (fold)
-\label{sec:relations}
+% \subsection*{Mind Brain Supervenience} % (fold)
+% \label{sub:supervenience}
 
-In this appendix we aim to provide more intuition regarding supervenience by discussing the limitations and extent of its implications.
+% subsection supervenience (end)
+% The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
+Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}.
+Therefore, here we define minds, brains, and supervenience.
+% The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
 
-%For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.
+% \subsection*{Statistical Setting} % (fold)
 
-First, a supervenient relation does not imply an injective relation.
-An injective relation is any relation that preserves distinctness.
-Thus if minds are injective on brains, then $b\neq b' \implies m \neq m'$ (note that the directionality of the implication has been switched relative to supervenience).
-For instance, it might be the case that a brain could change without the mind changing.
-Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$; or that a single additional synapse is formed between a pair of neruons.
-It is possible (likely?) that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
-In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
-This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.
-Minds supervening on brains does not imply that brains supervene on minds.
+Assuming for the moment that the space of all possible minds is finite, that is, $|\mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
+Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
+Define misclassification rate as $L_{\PP}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$.
+The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
+$g^* = \argmin_{g} L_{\PP}(g)$.
+Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
 
-Second,  supervenience does not imply causality. %it may be the case that minds supervene on brains, but that brains do not cause minds.  
-For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.
-Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other.
-This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
-
-Third, supervenience does not imply identity.
-Consider, for example, acceleration and velocity.
-Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).
-Similarly, velocity supervenes on position, as velocity cannot change without position changing.
-Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
-Rather, position can change with constant velocity, meaning without acceleration changing.
-Thus, to claim that something supervenes on another is not equivalent to claim that the two are identical.
-
-\begin{figure}[htbp]
-	\centering
-		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
-	\caption{Possible relations between minds and brains.
-(A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
-(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds.
-(C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
-	\label{fig:rel}
-\end{figure}
-% section relations_between_sets (end)
-
-What supervenience does imply, however, is the following.
-Imagine finding two different minds.
-If $\MeB$, then the brains subvening under those two minds must be different.
-In other words, there cannot be two different minds, either of which could supervene on a single brain.
-Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
-
-Note that statistical supervenience is distinct from statistical correlation.
-\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
-If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_{\PP} \mM$.
-Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
-
-\section{$k_n$ nearest neighbor algorithm} % (fold)
-\label{sec:knn}
-
-Consider the following problem setup.
-We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangably from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
-A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim \PP[M,B]$, and $m$ is unobserved.
-Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$.
-
-The $1$-nearest neighbor ($1$-NN) classifier works as follows.
-Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
-Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and their corresponding mental properties, , $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order.
-%One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
-The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $\mh{m}=m_{(1)}$.
-The $k_n$ nearest neighbor is a straightforward generalization of this approach.
-It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
-Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
-
-Unfortunately, no such algorithm is universally consistent.
-Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training points.
-Then, a collection of such algorithms, $\{g_n\}$,  with $k_n$ increasing with $n$, can be universally consistent under certain constraints.
-In particular, as $n$ increases, $k_n$ must also increase, but not quite as quickly.
-Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone's original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
-Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
-When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into finite Euclidean space, in which case Stone's theorem applies.
-Stone's original proof applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
-
-Note that the above extension of Stone's original theorem to the graph domain implicitly assumed that vertices were labeled, such that elements of the adjacency matrices could easily be compared across graphs.
-In theory, when vertices are unlabeled, one could first map each graph to a quotient space invariant to isomorphisms, and then proceed as before.
-Unfortunately, solving the graph-matching problem is currently NP-Incomplete (meaning it is not known to be either P or NP) \cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging.
-
-% section _k_n_nearest_neighbor_algorithm (end)
-
-\section{Simulation} % (fold)
-\label{sec:sim}
-
-As an example of a feasible experiment, one may consider a species whose nervous system consists of the same (small) number of labeled neurons for each organism.
-{\it Caenorhabditis elegans} is believed to be such a species \cite{Durbin87}.
-The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons.
-Although the graph with these neurons as vertices and edges defined by chemical synapses between neurons is likely not identical across individuals, it appears to be reasonably consistent \cite{Durbin87}.
-Furthermore, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
-Thus, one may design an experiment by describing the joint distribution $\PP[M,B]$ via class-conditional distributions $\PP[B|M=m_j]$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
-Here the mental property corresponds to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
-
-Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
-To generate the data, let $E_{ij}$ be an integer-valued random variable whose value indicates the number of synapses (edges) between neurons (vertices) $i$ and $j$.
-Let the class-conditional random variable $E_{ij} | M=m_0$  be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $\eta=0.05$.
-Let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
-Therefore, the distribution of these edges must differ between the two classes.
-The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$, where the signal parameter $z_{ij}=\eta$ for all edges not in $\mc{E}$, and $z_{ij}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{E}$.
-
-We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm.
-The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $\PP[M,B]$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
-
-\begin{figure}[!ht]
-\centering \includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.  $\hL$  (the misclassification rate estimated upon with $1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
-is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $\PP[M,B]$.
-Each dot depicts an estimate for $L_{\PP}(g_{\mt{n}})$; standard errors are $(L_{\PP}(g_{\mt{n}})(1-L_{\PP}(g_{\mt{n}}))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$; $k_n = 53$; $\hL^{1000}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
-We reject $H_0: L_{\PP}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{\PP}(g^*) \approx 0$ for this simulation.
-}
-\label{fig1}
-\end{figure}
-
-Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a C.
-elegans brain-graph within a day.
-Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
+Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $\PP[M,B]$.
+Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
+Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
+The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
+$\hL^{n'}_{\PP}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
+where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
+The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{\PP}(g_{\mt{n}}))$.
 
@@
 
+\bibliography{/Users/jovo/Research/latex/biblist}
+\bibliographystyle{nature}
+
+% \begin{thebibliography}{10}
+% \expandafter\ifx\csname url\endcsname\relax
+%   \def\url#1{\texttt{#1}}\fi
+% \expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+% \providecommand{\bibinfo}[2]{#2}
+% \providecommand{\eprint}[2][]{\url{#2}}
+% 
+% \bibitem{Plato97}
+% \bibinfo{author}{Plato}.
+% \newblock \emph{\bibinfo{title}{Plato: complete works}}
+%   (\bibinfo{publisher}{Hackett Pub Co}, \bibinfo{year}{1997}).
+% 
+% \bibitem{Descartes1641}
+% \bibinfo{author}{Descartes, R.}
+% \newblock \emph{\bibinfo{title}{Meditationes de prima philosophia}}
+%   (\bibinfo{year}{1641}).
+% 
+% \bibitem{Davidson70}
+% \bibinfo{author}{Davidson, D.}
+% \newblock \emph{\bibinfo{title}{Experience and Theory}}, chap.
+%   \bibinfo{chapter}{Mental Events} (\bibinfo{publisher}{Duckworth},
+%   \bibinfo{year}{1970}).
+% 
+% \bibitem{DGL96}
+% \bibinfo{author}{Devroye, L.}, \bibinfo{author}{Gy{\"o}rfi, L.} \&
+%   \bibinfo{author}{Lugosi, G.}
+% \newblock \emph{\bibinfo{title}{{A Probabilistic Theory of Pattern
+%   Recognition}}} (\bibinfo{publisher}{Springer}, \bibinfo{year}{1996}).
+% 
+% \bibitem{Devroye83}
+% \bibinfo{author}{Devroye, L.}
+% \newblock \bibinfo{title}{{On arbitrarily slow rates of global convergence in
+%   density estimation}}.
+% \newblock \emph{\bibinfo{journal}{Probability Theory and Related Fields}}
+%   \textbf{\bibinfo{volume}{62}}, \bibinfo{pages}{475--483}
+%   (\bibinfo{year}{1983}).
+% 
+% \bibitem{Popper}
+% \bibinfo{author}{Popper, K.}
+% \newblock \bibinfo{title}{{The logic of scientific discovery}}
+%   (\bibinfo{year}{1959}).
+% 
+% \bibitem{NorthGreenspan07}
+% \bibinfo{editor}{Geoffrey~North, R. J.~G.} (ed.)
+%   \emph{\bibinfo{title}{Invertebrate neurobiology}} (\bibinfo{publisher}{CSHL
+%   Press}, \bibinfo{year}{2007}).
+% 
+% \bibitem{Shepherd04}
+% \bibinfo{author}{Shepherd, G.}
+% \newblock \emph{\bibinfo{title}{{The synaptic organization of the brain}}}
+%   (\bibinfo{publisher}{Oxford University Press New York},
+%   \bibinfo{year}{2004}).
+% 
+% \bibitem{Felleman_VanEssen91}
+% \bibinfo{author}{Felleman, D.} \& \bibinfo{author}{Van~Essen, D.}
+% \newblock \bibinfo{title}{{Distributed hierarchical processing in the primate
+%   cerebral cortex}}.
+% \newblock \emph{\bibinfo{journal}{Cerebral cortex}}
+%   \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{1} (\bibinfo{year}{1991}).
+% 
+% \bibitem{Mori05}
+% \bibinfo{author}{S.~Mori, P. v. Z. L. N.-P., S.~Wakana}.
+% \newblock \emph{\bibinfo{title}{MRI Atlas of Human White Matter}}
+%   (\bibinfo{publisher}{Elsevier Science}, \bibinfo{year}{2005}).
+% 
+% \bibitem{Abeles91}
+% \bibinfo{author}{Abeles, M.}
+% \newblock \emph{\bibinfo{title}{Corticonics}} (\bibinfo{publisher}{Cambridge
+%   University Press}, \bibinfo{year}{1991}).
+% 
+% \bibitem{Koch_Davis94}
+% \bibinfo{author}{Koch, C.} \& \bibinfo{author}{Davis, J.}
+% \newblock \emph{\bibinfo{title}{{Large-scale neuronal theories of the brain}}}
+%   (\bibinfo{publisher}{The MIT Press}, \bibinfo{year}{1994}).
+% 
+% \bibitem{Rao_Lewicki02}
+% \bibinfo{author}{Rao, R.}, \bibinfo{author}{Olshausen, B.} \&
+%   \bibinfo{author}{Lewicki, M.}
+% \newblock \emph{\bibinfo{title}{{Probabilistic models of the brain: Perception
+%   and neural function}}} (\bibinfo{publisher}{The MIT Press},
+%   \bibinfo{year}{2002}).
+% 
+% \bibitem{Chow_Dalibard03}
+% \bibinfo{editor}{Chow, C.}, \bibinfo{editor}{Gutkin, B.},
+%   \bibinfo{editor}{Hansel, D.}, \bibinfo{editor}{Meunier, C.} \&
+%   \bibinfo{editor}{Dalibard, J.} (eds.) \emph{\bibinfo{title}{Methods and
+%   Models in Neurophysics}} (\bibinfo{publisher}{Elsevier},
+%   \bibinfo{year}{2003}).
+% 
+% \bibitem{SpornsKotter05}
+% \bibinfo{author}{Sporns, O.}, \bibinfo{author}{Tononi, G.} \&
+%   \bibinfo{author}{Kotter, R.}
+% \newblock \bibinfo{title}{The human connectome: A structural description of the
+%   human brain}.
+% \newblock \emph{\bibinfo{journal}{PLoS Computational Biology}}
+%   \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{e42} (\bibinfo{year}{2005}).
+% 
+% \bibitem{Hagmann05}
+% \bibinfo{author}{Hagmann, P.}
+% \newblock \emph{\bibinfo{title}{{From diffusion MRI to brain connectomics}}}.
+% \newblock Ph.D. thesis, \bibinfo{school}{Institut de traitement des signaux}
+%   (\bibinfo{year}{2005}).
+% 
+% \bibitem{Sporns10}
+% \bibinfo{author}{Sporns, O.}
+% \newblock \emph{\bibinfo{title}{Networks of the Brain}}
+%   (\bibinfo{publisher}{MIT Press}, \bibinfo{year}{2010}).
+% 
+% \bibitem{Basser94}
+% \bibinfo{author}{Basser, P.~J.}, \bibinfo{author}{Mattiello, J.} \&
+%   \bibinfo{author}{LeBihan, D.}
+% \newblock \bibinfo{title}{Mr diffusion tensor spectroscopy and imaging.}
+% \newblock \emph{\bibinfo{journal}{Biophys J}} \textbf{\bibinfo{volume}{66}},
+%   \bibinfo{pages}{259--267} (\bibinfo{year}{1994}).
+% \newblock \urlprefix\url{http://dx.doi.org/10.1016/S0006-3495(94)80775-1}.
+% 
+% \bibitem{ArdekaniSzeszko10}
+% \bibinfo{author}{Ardekani, B.~A.} \emph{et~al.}
+% \newblock \bibinfo{title}{Diffusion tensor imaging reliably differentiates
+%   patients with schizophrenia from healthy volunteers.}
+% \newblock \emph{\bibinfo{journal}{Hum Brain Mapp}}  (\bibinfo{year}{2010}).
+% \newblock \urlprefix\url{http://dx.doi.org/10.1002/hbm.20995}.
+% 
+% \bibitem{TuchWedeen02}
+% \bibinfo{author}{Tuch, D.} \emph{et~al.}
+% \newblock \bibinfo{title}{{High angular resolution diffusion imaging reveals
+%   intravoxel white matter fiber heterogeneity}}.
+% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
+%   \textbf{\bibinfo{volume}{48}}, \bibinfo{pages}{577--582}
+%   (\bibinfo{year}{2002}).
+% 
+% \bibitem{Tuch04}
+% \bibinfo{author}{Tuch, D.}
+% \newblock \bibinfo{title}{{Q-ball imaging}}.
+% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
+%   \textbf{\bibinfo{volume}{52}}, \bibinfo{pages}{1358--1372}
+%   (\bibinfo{year}{2004}).
+% 
+% \bibitem{WedeenWeisskoff05}
+% \bibinfo{author}{Wedeen, V.}, \bibinfo{author}{Hagmann, P.},
+%   \bibinfo{author}{Tseng, W.}, \bibinfo{author}{Reese, T.} \&
+%   \bibinfo{author}{Weisskoff, R.}
+% \newblock \bibinfo{title}{{Mapping complex tissue architecture with diffusion
+%   spectrum magnetic resonance imaging}}.
+% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
+%   \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{1377--1386}
+%   (\bibinfo{year}{2005}).
+% 
+% \bibitem{PalmAmunts10}
+% \bibinfo{author}{Palm, C.} \emph{et~al.}
+% \newblock \bibinfo{title}{{Towards ultra-high resolution fibre tract mapping of
+%   the human brain--registration of polarised light images and reorientation of
+%   fibre vectors}}.
+% \newblock \emph{\bibinfo{journal}{Frontiers in Human Neuroscience}}
+%   \textbf{\bibinfo{volume}{4}} (\bibinfo{year}{2010}).
+% 
+% \bibitem{DenkHorstmann04}
+% \bibinfo{author}{W.~Denk, W.} \& \bibinfo{author}{Horstmann, H.}
+% \newblock \bibinfo{title}{Serial block-face scanning electron microscopy to
+%   reconstruct three-dimensional tissue nanostructure}.
+% \newblock \emph{\bibinfo{journal}{PLOS Biol.}} \textbf{\bibinfo{volume}{2}},
+%   \bibinfo{pages}{e329} (\bibinfo{year}{2004}).
+% 
+% \bibitem{HayworthLichtman06}
+% \bibinfo{author}{Hayworth, K.}, \bibinfo{author}{Kasthuri, N.},
+%   \bibinfo{author}{Schalek, R.} \& \bibinfo{author}{Lichtman, J.}
+% \newblock \bibinfo{title}{{Automating the collection of ultrathin serial
+%   sections for large volume TEM reconstructions}}.
+% \newblock \emph{\bibinfo{journal}{Microscopy and Microanalysis}}
+%   \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{86--87}
+%   (\bibinfo{year}{2006}).
+% 
+% \bibitem{Penrose99}
+% \bibinfo{author}{Penrose, R.} \& \bibinfo{author}{Gardner, M.}
+% \newblock \emph{\bibinfo{title}{{The emperor's new mind: concerning computers,
+%   minds, and the laws of physics}}} (\bibinfo{publisher}{Oxford University
+%   Press, USA}, \bibinfo{year}{1999}).
+% 
+% \bibitem{Satinover02}
+% \bibinfo{author}{Satinover, J.}
+% \newblock \emph{\bibinfo{title}{{The quantum brain: the search for freedom and
+%   the next generation of man}}} (\bibinfo{publisher}{Wiley},
+%   \bibinfo{year}{2002}).
+% 
+% \bibitem{NielsenChuang00}
+% \bibinfo{author}{Nielson, M.~A.} \& \bibinfo{author}{Chuang, I.~L.}
+% \newblock \emph{\bibinfo{title}{Quantum Computation and Quantum Information}}
+%   (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{2000}).
+% 
+% \bibitem{Craver09}
+% \bibinfo{author}{Craver, C.~F.}
+% \newblock \bibinfo{title}{Stochastic supervenience}  (\bibinfo{year}{2009}).
+% 
+% \bibitem{Semon21}
+% \bibinfo{author}{Semon, R.~W.}
+% \newblock \emph{\bibinfo{title}{The Mneme}} (\bibinfo{publisher}{G. Allen \&
+%   Unwin ltd.}, \bibinfo{year}{1921}).
+% 
+% \bibitem{Lashley50}
+% \bibinfo{author}{Lashley, K.}
+% \newblock \bibinfo{title}{{In search of the engram}}.
+% \newblock \emph{\bibinfo{journal}{Symposia of the society for experimental
+%   biology}} \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{30}
+%   (\bibinfo{year}{1950}).
+% 
+% \bibitem{ZhangLinden03}
+% \bibinfo{author}{Zhang, W.} \& \bibinfo{author}{Linden, D.}
+% \newblock \bibinfo{title}{{The other side of the engram: experience-driven
+%   changes in neuronal intrinsic excitability}}.
+% \newblock \emph{\bibinfo{journal}{Nature Reviews Neuroscience}}
+%   \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{885--900}
+%   (\bibinfo{year}{2003}).
+% 
+% \bibitem{ShemaDudai07}
+% \bibinfo{author}{Shema, R.}, \bibinfo{author}{Sacktor, T.} \&
+%   \bibinfo{author}{Dudai, Y.}
+% \newblock \bibinfo{title}{{Rapid erasure of long-term memory associations in
+%   the cortex by an inhibitor of PKM $\{$zeta$\}$}}.
+% \newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{317}},
+%   \bibinfo{pages}{951} (\bibinfo{year}{2007}).
+% 
+% \bibitem{BerryDavis08}
+% \bibinfo{author}{Berry, J.}, \bibinfo{author}{Krause, W.} \&
+%   \bibinfo{author}{Davis, R.}
+% \newblock \bibinfo{title}{{Olfactory memory traces in Drosophila}}.
+% \newblock \emph{\bibinfo{journal}{Progress in brain research}}
+%   \textbf{\bibinfo{volume}{169}}, \bibinfo{pages}{293--304}
+%   (\bibinfo{year}{2008}).
+% 
+% \end{thebibliography}
+
 \section*{Acknowledgments}
@@
 
-% \section{References}
-% \bibliography{biblist}
-\bibliography{/Users/jovo/Research/latex/biblist}
-\bibliographystyle{ieeetr}
+\section*{Author Contributions}
+
+JTV, RJV, and CEP conceived of the manuscript.  JTV and CEP wrote it.  CEP ran the experiment.
+
+\section*{Additional Information}
+
+The authors have no competing financial interests to declare.
 

### 2011-05-11 nature_sci_rep/supp.tex
added Nature Scientific Reports submission

--- 2011-02-17 PLoS_ONE/2011_02_17/supervenience.tex
@@
 \newtheorem{thm}{Theorem}
-\newtheorem{thex}{\emph{Gedanken} experiment}
+\newtheorem{thex}{\emph{Gedankenexperiment}}
 \lhead{Vogelstein JT, et al}
-\rhead{Statistical Supervenience}
-
-\title{Are mental properties supervenient on brain properties?}
-
-\author{Joshua T. Vogelstein$^1$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
-$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723
-}
+\rhead{Statistical Supervenience Supplement}
 
@@
 
-\maketitle
-%\tableofcontents
-\begin{abstract}
+	\begin{center}
+{\huge	Supplementary Materials for: \\ Are mental properties supervenient on brain properties?}
+\end{center}
 
-	The ``mind-brain supervenience'' conjecture suggests that all mental properties (e.g.
-consciousness, intelligence, personality, etc.) are derived from the physical properties of the brain.
-The validity of this conjecture has been argued in philosophical terms for over 2,500 years.
-Alternative conjectures, including various non-physical causes of mental properties, seem rather implausible to many, but proving or disproving these alternatives has remained elusive.
+\vspace{5px}
 
-	To address the question of whether the mind supervenes on the brain, we here frame a supervenience hypothesis in rigorous mathematical terms.
-Specifically, we propose a modified version of supervenience (called $\eps$-supervenience) that is amenable to experimental investigation and statistical analysis.
-To illustrate this approach, we perform a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\eps$-supervenience.
-The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or connectome).
-$\eps$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate > 0, regardless of the relationship between the two.
+\begin{center}
+{\large		
+Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
+	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
 
-	To the philosopher, this work shows how philosophical conjectures can be transformed into statistical hypotheses that are amenable experimental investigation.
-This allows the philosopher to gain empirical support for her rational arguments.
-To the statistician, this work points out the limitations of hypothesis testing; and suggests that some of these limitations have not previously been fully appreciated.
-To the neuroscientist, this work indicates that much of contemporary research can be framed in terms of investigating supervenience.
-This should provide further motivation for cross-disciplinary research between neuroscientists and statistical graph-theoreticians.
+\end{center}
 
-\end{abstract}
-
-% \vspace*{0.5 in}
-
-\section{Introduction}
-
-Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
-Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
-Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience \tr{conjecture}, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}.
-Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation.
-This work is an attempt to bridge the gap between these philosophical conjectures and experimentally testable hypotheses.
-% Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms, and that can be experimentally tested \tr{as a hypothesis}.
-
-The primary contributions of this work flow from our introduction of a notion of supervenience amenable to empirical investigation.
-This renders the mind-brain dualism debate a hypothesis, rather than an assumption, both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
-Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.
-Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists, given sufficient data.
-Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.
-This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, and suggests avenues for future research.
-
- % we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.
-Finally, we discuss both contemporary and future applications of this framework to neurobiological studies in humans.
-
-\section{Statistical supervenience} % (fold)
-\label{sec:preliminaries}
-
-% \subsection{Mind Brain Supervenience} % (fold)
-% \label{sub:supervenience}
-
-% subsection supervenience (end)
-% The intention in this work is to develop greater insight regarding the relationship between minds and brains, using statistical methods, with particular interest in notions of supervenience.
-Modern philosophy of science suggests that adequate scientific explanations must account for the objects of investigation and their relationships \cite{Craver07}.
-Therefore, here we define minds, brains, and supervenience.
-% The mind-brain supervenience conjecture is a relation between the set of mental states and the set of brain states.
-
-Donald Davidson canonized the mind-brain supervenience relation in 1970 with the following quote: \cite{Davidson70}
-\begin{quotation}
-\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
-\end{quotation}
-This conjecture may be concisely and formally stated thusly: Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
-% The set of possible brains $\mB$ is completely unrestricted, meaning that $\mB$ could be an infinite set, with arbitrarily complexity.
-In particular, $b$ might represent the position, momentum, and type of each subatomic particle residing within the skull some agent.
-%Thus, each different $b \in \mB$ could correspond to some difference in the position, momentum, or type of at least one subatomic particle composing the brain.
-Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-%The set of possible minds $\mM$ is also unrestricted.
-In particular, $m$ might represent all an agent's thoughts, beliefs, and preferences.
-%Thus, each different $m \in \mM$ could correspond to a difference in at least one thought, belief, or preference.
-The supervenience conjecture implies that%, the above mind brain supervenience conjecture may be formally stated
-: $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
-This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix \ref{sec:relations} for more details and some examples).
-
-To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
-Let $\PP[M,B]$ indicate a joint distribution of minds and brains.
-%Each distribution $\PP[M,B]$ uniquely specifies the probability of any element $(m,b)$ occurring from the space of all possible elements $\mc{M} \times \mc{B}$.
-% (also called the \emph{sample space}).%  The true distribution could in theory be any possible joint distribution from the model, $\PP[M,B] \in \mc{P}$, meaning that we only assume that it conforms to the \emph{probability axioms}:
-% \begin{enumerate}
-% 	\item $0 \leq \PP[B=b,M=m] \leq 1$ for all $(m,b) \in \mM \times \mB$
-% 	\item $\PP[\Omega]=1$ and $\PP[\emptyset]=0$, where $\Omega$ is the whole sample sample, that is, all elements, $(m,b) \in \mc{M} \times \mc{B}$
-% 	\item any countable sequence of pairwise disjoint elements, $(m_1,b_1), (m_2,b_2), \ldots$ satisfies \\
-% 	\mbox{$\PP[(m_1,b_1) \cup (m_2,b_2) \cup \ldots] = \sum_i \PP[(m_i,b_i)]$}.
-% \end{enumerate}
-%Given a distribution, one can then calculate any marginal or conditional distributions.
-For instance, $\PP[B]=\int_{m \in \mM} \PP[M,B]dm$, or $\PP[M|B]=\PP[M,B]/\PP[B]$.
-Having specified a model, statistical supervenience can be defined as follows:
-\begin{defi}
-\label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
-%, $\forall$ $(m,b), (m',b') \in \mc{M} \times \mc{B}$.
-\end{defi}
-\noindent Statistical supervenience is therefore a probabilistic relation on sets (related to, but distinct from correlation; see Appendix \ref{sec:relations} for details).
-
- % Consider the following analogy: one desires to understand the relationship between height and weight.
-In such a scenario, height, $h$, could be assumed to be an element from the set of all possible heights, $\mc{H}=(0,\infty)$ cm; and weight, $w$, could be assumed to be an element from the set of all possible weights, $\mc{W}=(0,\infty)$ kgs.
-Then, it may be assumed that an agent's height and weight is sampled identically and independently from some \emph{model}, that is $(h,w)\overset{iid}{\sim}\PP[H,W]=F_{HW}$.
-To gain insight into $F_{HW}$, one might then state hypotheses, collect data, and perform analyses.
-Implicit in the exposition of a model is that both objects, height and weight here, are \emph{random objects}.
-Note that this does not imply anything in particular about the distribution (which could be a point mass, for instance); rather, it is a formal definition necessary for statistical analysis and empirical investigation.
-% 
-% \subsection{Brain-graphs} % (fold)
-% \label{sub:brain_graphs}
-
-% subsection brain_graphs (end)
-
-% section preliminaries (end)
-
-%\section{Results}
-
-\section{Statistical supervenience is equivalent to perfect classification accuracy} % (fold)
-\label{sub:theoretical_results}
-
-% subsection theoretical_results (end)
-If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.
-This means that there exists a deterministic function mapping each brain to its supervening mind, $g(\cdot): \mB \mapsto \mM$; therefore, one could in principle construct this function.
-It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.
-Assuming for the moment that the space of all possible minds is finite, that is $| \mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
-Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as
-% \begin{align}
-	$L_{\PP}(g) = \PP[g(B) \neq M]$, %= \frac{1}{|\mB| |\mM|}\iint \II\{g(b)\neq m\} db dm
-% \end{align}
-% where $\II\{\cdot\}$ indicates the indicator function, taking unity value if its argument is true, and zero otherwise.  
-% $L_{\PP}(g)$ therefore effectively 
-which denotes the probability that $g$ misclassifies $b$. The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
-% \begin{align}
-	$g^* = \argmin_{g} L_{\PP}(g)$.
-% \end{align}
-% where $\mc{G}$ is the set of all possible classifiers.  
-Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
-The primary result of casting supervenience as a statistical framework is the following theorem:
-\begin{thm}
-\label{thm1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_{\PP} \mB$, if and only if $L_{\PP}(g^*) = 0$.
-Formally, \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.
-\end{thm}
-% \begin{thm} 
-% 	\label{thm1} 
-% 	Statistical supervenience implies that there exists a classifier that is \emph{always correct}, that is, $g(b)=m$, for all $(m,b) \in \mc{M} \times \mc{B}$.  
-% \end{thm}
-
-\noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.
-This means that one could draw a decision boundary between all equivalence classes of brains, each class corresponding to a different mind, and no mind will reside within two different equivalence classes.
-Thus, the optimal classifier would correctly find these decision boundaries, and therefore have no opportunity to err.
-$\square$
-
-The above argument shows (for the first time to our knowledge) that statistical supervenience and zero Bayes error are equivalent.
-Statistical supervenience can therefore be thought of as a constraint on the possible distributions on minds and brains.
-Specifically, let $\mc{P}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{P}_s$ be subset of distributions for which supervenience holds.
-Theorem \ref{thm1} implies that $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
-
-\section{A hypothesis test for supervenience} % (fold)
-\label{sub:hypothesis_testing}
-
-Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $\PP[M,B]$ and $g^*$, rendering them useless pragmatically.
-However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
-Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $\PP[M,B]$.
-Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
-Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
-The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
-% Therefore, instead of calculating misclassification rate for $g_n$, we compute the expected misclassification rate: % as the expectation of $\PP[g_n(B)=M | \mc{T}_n]$ over all possible training data corpuses:
-% \begin{align}
-% \EE[L_{\PP}(g_n)] = \EE[\PP_F[g_n(B;\tr{\mc{T}_n}) \neq M|\tr{\mc{T}_n}]] = \int \PP[g_n(B)=M | \mc{T}_n] \PP[\mc{T}_n] d\mc{T}_n.
-% \end{align}
-% Unfortunately, in practice, computing $\EE[L_{\PP}(g_n)]$, requires integrating over all possible training data corpuses of size $n$, and by definition, we only have access to a single training data corpus.
-We therefore define ``hold out'' misclassification performance:
-% \begin{align}
-% \EE[L_{\PP}(g_n)] 
-% &\approx  \hL^{n'}_{\PP}(g_n) = \sum_{\mc{T}_{\mt{n}}} \PP[g_n(B)=M | \mc{T}_{\mt{n}}] \PP[\mc{T}_{\mt{n}}],
-% \end{align}
-$\hL^{n'}_{\PP}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
-where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
-%, and the sum is taken over a sufficiently large number of subsets, $\mc{T}_{\mt{n}}$, such that $\hL^{n'}_{\PP}(g_n)$ converges.  
- % $n' \hL^{n'}_{\PP}(g_n)$ is t
-The approximate  number of misclassified minds
-%. Simplifying further by assuming just one hold-out set, then $n' \hL^{n'}_{\PP}(g_n)$ 
-therefore
-% and 
-has a binomial distribution:  $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{\PP}(g_{\mt{n}}))$.
-%, because for any of the $n'$ held-out samples, the classifier could be either correct or incorrect.
-
-Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
-\begin{defi}
-\label{def2}
-Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\MeB$, if and only if $L_{\PP}(g^*) < \varepsilon$.
-\end{defi}
-
-Given this relaxation, consider the problem of testing for $\eps$-supervenience.
-%First, specify a significance level, $\alpha$, such that if the p-value is less than $\alpha$, then the null is rejected.
-% Because we hope to reject the null, in favor of the alternative, l
-Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-We reject for  values of the test   statistic lower than the critical value, that is, we reject if and only if $n' \hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
-The critical value is available under the least favorable distribution $\text{Binomial}(n',\varepsilon)$.
-Thus, rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
-The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$.
-%Similar to the above, one can define the set of $\eps$-supervenience models as the set of models under which $\eps$-supervenience holds, that is: $\mc{P}_\eps = \{\PP | L_{\PP}(g^*)< \eps\}$.
-One could then sort  $\eps$-supervenience subsets,  $\mc{P}_s \subseteq \mc{P}_{\eps} \subseteq \mc{P}_{\eps'} \subseteq \mc{P}$, for any $\eps < \eps'$.
-
-% subsection hypothesis_testing (end)
-
-% \section{Power and consistency} % (fold)
-% \label{ssub:power_and_consistency}
-
-Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis% (in other words, the probability that it will not make a Type II error)
-.
-Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
-A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.
-A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{\PP}(g_n)] \conv L_{\PP}(g^*)$ as $n\conv \infty$.
-%As the notation suggests, consistency of a classifier is a function of the true model, $\PP$.
-Without any prior knowledge of what the model might be, one desires a \emph{universally consistent} classifier, that is a classifier that is consistent for all $\PP \in \mc{P}$.
-% The above described hypothesis tests suggests the following theorem:
-% 
-% \begin{thm} \label{thm1} 
-% Given $\alpha > 0$, we can test $\MeB$ so that rejection implies $\MeB$ holds with probability greater than or equal to $1-\alpha$.
-Furthermore, given a consistent classifier the power of the test converges to unity.
-% \end{thm}
-Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
-
-Unfortunately, the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ depends on the (unknown) distribution $\PP=\PP[M,B]$ \cite{DGL96}.
-Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, %.
-This means that we can never be confident that $\MeB$ does \emph{not} hold; rather,
-it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-%Unfortunately, arbitrarily slow convergence theorems inform us that no matter how much data we collect, we cannot disambiguate between not yet having enough data, and $\MeB$ not holding.
-Thus, without restrictions on $\PP[M,B]$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-%Given these limitations on even universal consistency, it is still the best one can hope for.
-Therefore, we hope to obtain a universally consistent classifier to test for $\eps$-supervenience.
-
-% subsubsection power_and_consistency (end)
-
-\section{A \emph{gedanken} experiment demonstrating consistency and unity power} % (fold)
-\label{sub:uc}
-
-% Below, we will assume that brains are characterized by graphs, referred to hereafter as ``brain-graphs.''  This assumption follows naturally from the neuron-doctrine
-% % Given the above notion of statistical supervenience, one can then design a statistical test for supervenience by carefully considering an appropriate model.
-For mind-brain supervenience, a model of the brain is paramount, and has historically been a bit of a sticky wicket.
-It was not until 1891, that H.~Waldeyer-Hartz first formally proposed the ``neuron doctrine''
-% \cite{Waldeyer-Hartz1891}, which states that the nervous system is a complex {network} of neurons, largely based on Ramon y Cajal's work using the Golgi stain \cite{Finger01}.
-This doctrine has been central to much of the development of neuroscience and artificial intelligence for over 100 years, including the development of neural network theory \cite{Bishop95} and cognitive science \cite{McClellandRumelhart86}.
-%Fundamental to nearly all models of the brain (or components therein) since the neuron doctrine's introduction is the existence of neural units, each connected to (some) other neural units.
-Therefore, it seems appropriate to assume that a reasonable model of a brain is graph, composed of a set of vertices (or nodes), each representing a neural unit, and a set of edges (arcs) connecting the nodes.
-Below, we formalize this concept to enable statistical testing in the graph domain.
-% Below, we will assume that $m$ is binary, that is, $m \in \{0,1\}$ (for example, $m$ might indicate whether the person knows calculus or not).
-
-To ensure consistency and therefore unity power, the classifier $g_n(\cdot)$ must be able to converge to the truth, regardless of the true distribution, $\PP$.
-We therefore make explicit a model for brains, and show that under this very general model, universally consistent classifiers are available.
-
-\begin{thex}
-Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a brain-graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
-Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable.
-Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix \ref{sec:knn} for proof).
-Therefore, %Theorem \ref{thm1} applies and
-the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
-Furthermore, Appendix \ref{sec:knn} also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
-\end{thex}
-
-%(fold)
-% Formally, a brain-graph $G=(V,A)$ is characterized by a set of vertices (or nodes), $V = \{V_i\} = \{V_1,\ldots, V_n\}$, where $n$ is the number of neural units, and arcs (or edges) $A=\{A_{ij}\}$, where $A_{ij}$ represents the connectivity from neural unit $V_j$ to $V_i$.
-For simplicity (to be generalized below), assume that brain-graphs are simple graphs, meaning that for all $i,j \in [n]$: (i) there are no self loops, $a_{ii}=0$, (ii) all edges are binary, $a_{ij} \in \{0,1\}$, and (iii) all edges are symmetric,  $a_{ij}=a_{ji}$.
-Further assume that the vertices are \emph{labeled}, meaning that there is a known one-to-one mapping from each vertex in any brain to a vertex in any other brain, and that each brain-graph has the same number of vertices.
-Thus, each brain-graph is characterized entirely by its adjacency matrix, $a \in \mc{A}$, where $\mc{A} \subseteq \{0,1\}^{n \times n}$.
-Further, let the mental property under investigation be binary, so $m \in \mc{M} = \{0,1\}$.
-The Bayes classifier is:
-% \begin{align}
-% 	\mh{m} = \argmax_m \PP[b,m] = \argmax_m \PP[b|m] \PP[m] 
-% \end{align}
-% where the likelihood term, $\PP[B|M]$, specifies the probability of observing any particular adjacency matrix for any class, and the prior term, $\PP[M]$, specifies the prior probability of either class.
-% 
-% Under the above model assumptions, the number of possible brain-graph/mental-property pairs is finite, that is, $|\mc{M} \times \mc{B}| = d < \infty$.
-Therefore, one can simply enumerate all possible brain-graphs for each class, such that each class is represented by a multinomial distribution.
-The maximum likelihood estimator (MLE) of the multinomial parameter for each class as guaranteed to converge to the true (but unknown) parameter.
-Formally, one can state $\mh{\PP}_{MLE}[B|M] \conv \PP[B|M]$ as $n \conv \infty$, almost surely.
-Furthermore, the MLE  estimator of the prior also converges to the true (but unknown) prior distribution, $\mh{\PP}_{MLE}[M] \conv \PP[M]$, as $n \conv \infty$, almost surely.
-% 
-% The Bayes plugin classifier is defined by:
-% \begin{align}
-% 	\mh{m} = \argmax_m  \mh{\PP}[b|m] \mh{\PP}[m] 
-% \end{align}
-% where $\mh{\PP}[B|M]$ and $\mh{\PP}[M]$ are plugin estimators.
-Since the MLE estimators defined above are plugin estimators, and consistent, one obtains a consistent classifier by plugging in the MLE parameters.
-This logic holds for any model in which the cardinality of the sample space is finite, that is $|\mc{M}\times \mc{B}| < \infty$.
-In particular, one can relax both the assumed symmetry and hollowness of the adjacency matrix, and this result still holds.
-Furthermore, one can allow for multi-graphs, in which there are a finite number of different kinds of edges (different colored edges), such that $a_{ij} \in \{0,1,\ldots, K\}$, with $K < \infty$, and the set of possible brain-graphs remains finite.
-Finally, one can allow $\mc{M}$ to be any finite space, not just binary.
-Note that in all these cases, the process one uses to enumerate and sort the brain-graph collections is essentially irrelevant, assuming the same process is used for each class.
-Further note that more robust estimators, such as certain M-estimators and the \emph{maximum a posteriori} estimators, also admit universally consistent classifiers, with perhaps faster convergence properties.
-% 
-% The above approach, however, is insufficient in several important cases.
-First, when the edges live in a continuous space, such as $a_{ij} \in \Real$, corresponding to weighted graphs, the MLE plugin is not well defined.
-In such scenarios, the $k_n$ nearest neighbor ($k$-NN) algorithm can be used (see Appendix 1 for a description of the $k$-NN algorithm) \cite{Stone77}.
-Although originally defined to operate on finite dimensional Euclidean space, the same algorithm can be applied here, once one has first embedded the brain-graphs into such a space.
-One possible embedding is to stack the columns of the adjacency matrix, yielding an $n^2$ element vector, instead of an $n \times n$ element matrix.
-Given such an embedding, the $k$-NN classifier may be used without modification, assuming one has defined a suitable distance operator (such as $L_2$ or $L_1$).
-This algorithm may also be used in the above cases where the support of the sample space is finite, and, convergence of the $k$-NN algorithm for those problems will be faster than the MLE plugins, whenever the distance used is ``appropriate'' (where appropriate is assessed by convergence rates under the true (but unknown) model).
-Furthermore, one can allow for multiple kinds of edges, as before, yielding adjacency tensors, with $a_{ij} \in \Real^d$.
-Or, one can allow the mental property to be multidimensional as well, $m \in \Real^l$.
-The $k$-NN algorithm converges in both these generalizations.
-Finally, one could allow the number of vertices to differ across brain samples.
-In such a scenario, one can ``pad'' the adjacency matrices of the smaller brain-graphs, and the result still holds.
-% 
-% The above universal consistency results hold for unlabeled graphs as well, with a minor modification.
-Assuming the brain-graphs are unlabeled, and we have a test brain-graph $b$, one could first align each training brain-graph with the test brain-graph, and then use the above algorithms.
-To align any pair of graphs, one can solve the graph isomorphism problem, $\mh{Q}_i=\argmin_{Q \in \mc{Q}} \norm{Q b_i Q\T - b}$, where $\mc{Q}$ is the set of all permutation matrices (that is, zero matrices with a single unit value in each row and column).
-Given $\mh{Q}_i$, one can compute $\mt{b}_i=\mh{Q}_i b \mh{Q}_i\T$, and then proceed as one would normally.
-Unfortunately, the graph isomorphism problem is known to be NP-incomplete (that is, it is not known to believe in either P or NP-complete) \cite{GareyJohnson79}, and therefore no polynomial time algorithm is available, making this approach quite time-consuming.
-% 
-% 
-% % subsection brain_graphs 
-% 
-% \section{Theoretical implications} % (fold)
-% \label{sub:implications}
-% 
-% Returning to the question of supervenience, consider the following thought experiment.
-Let the physical property under consideration be brain-graphs, and the mental property under investigation be binary (such as, knows calculus or not).
-Now, imagine collecting very large amounts of very accurate exchangeable brain-graph data and the associated mental property values.
-A $k_n$-nearest neighbor algorithm using an isomorphism-matching Frobenius norm is universally consistent.
-Therefore, %Theorem \ref{thm1} applies and
-% the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) one will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with edge weights, with or without labeled vertices.
-% 
-% 
-% % subsection implications (end)
-% 
-% 
-% % subsection practical_applications (end)
-%(end)
-
-\section{Discussion}
-
-\subsection{Summary} % (fold)
-\label{par:summary}
-
-We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind/brain property pair is less than $\eps$.
-Furthermore, when we restrict the space of minds and brains to the setting of \emph{gedanken} experiment 1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
-Alas, this is a one-sided test, so although power converges to unity, one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
-
-\subsection{Practical issues} % (fold)
-\label{par:practical_issues}
-
-Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans is practically possible for any particular mental property at this time (at least when the vertices represent individual neurons).
-Rather, the claim is that \emph{if} one had sufficient data (a very large number of exchangeable mind/brain property pairs), then \emph{in theory}, some level $\eps$-supervenience hypothesis test could be performed.
-Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$) for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (but see Appendix \ref{sec:sim} for a simulated example in which convergence is relatively fast).
-Whichever classifier is used, it is likely to benefit from a large amount of domain-specific knowledge, which the proposed classifier completely neglects.
-
-% paragraph practical_issues (end)
-
-% paragraph summary (end)
-
-\subsection{A unified quest} % (fold)
-\label{sub:practical_applications}
-
-A central (perhaps \emph{the} central) quest in much of neuroscience, psychology, and cognitive science is to discover the brain properties that subvene under various mental properties, although questions are rarely cast within a  supervenience formalism.
-Moreover, the particular brain properties that are often believed to subvene under these mental properties are neural circuits, or brain-subgraphs.
-To this end, many investigations in these fields include schematic diagrams showing a particular brain-subgraph subvening under a particular mental phenotype.
-This practice transcends the evolutionary hierarchy of neuroscientific research.
-For instance, in the invertebrate literature, vertices correspond to particular labeled neurons, and edges correspond to synapses \cite{NorthGreenspan07}.
-In the vertebrate literature, vertices often correspond to types of neurons in particular regions, and edges correspond to tendencies of connections \cite{Shepherd04}.
-For primates \cite{Felleman_VanEssen91} and humans \cite{Mori05} vertices frequently represent functionally distinct neuroanatomical regions, and edges represent regional interconnectivity.
-Furthermore, this practice also transcends analytical background, including anatomists \cite{Abeles91}, philosophers \cite{Koch_Davis94}, statisticians \cite{Rao_Lewicki02}, and physicists \cite{Chow_Dalibard03}.
-The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (although perhaps causality, not supervenience, is the true desideratum).
-Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neurcognitive investigations.
-
-% \subsection{Causality} % (fold)
-% \label{par:causality}
-% 
-% While the above thought experiment addresses the question of $\varepsilon$-supervenience, it does not address causality.
-Assuming we have confirmed $\MeB$ for a particular mind/brain property pair with confidence level $\alpha$, then morphing the brain (by altering edges) could be used to determine whether the relation is in fact causal.
-% 
-% % paragraph causality (end)
-
-\subsection{Human applications} % (fold)
-\label{par:human_applications}
-
-In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has driven an explosion of studies investigating the human connectome and relating connectomes to cognitive properties \cite{Sporns10}.
-% \cite{Jones10}.
-Many of these studies can be framed as $\eps$-supervenience hypothesis tests.
-For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenic individuals and normal (control) individuals \cite{ArdekaniSzeszko10}.
-As the resolution and signal-to-noise ratio of magnetic resonance imaging continue to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
-Furthermore, the utilization of other imaging technologies, such as polarized light imaging \cite{PalmAmunts10} and high-throughput electron microscopy \cite{DenkHorstmann04,HayworthLichtman06}, will continue to improve the effective resolution of these inferred connectomes from human brains.
-While determining from a brain scan whether a particular individual knows calculus might be quite distant, many other cognitive and psychological supervenience hypotheses have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be diminishing.
-
-% paragraph relationship_to_previous_work (end)
-
-\subsection{Alternative explanations} % (fold)
-
-Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{gedanken} experiment in Section \ref{sub:uc}, a natural question to raise is: what are the conceivable alternative hypotheses?
-We consider three such alternatives.
-
-First, perhaps brains are more accurately characterized as quantum networks over classical networks.
-Several authors have suggested that brains have certain hypercomputational properties that classical computers could not achieve \cite{Penrose99,Satinover02}.
-However, assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
-This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
-This means that if the above analysis failed to reject the null hypothesis at level $\alpha$, it will fail regardless of whether one assumes quantum or classical computations.
-
-Second, perhaps minds stochastically supervene on brains \cite{Craver09}. While perhaps difficult to imagine, 
-much like a non-deterministic world was difficult to imagine prior to modern physics, it is not inconceivable that mental properties are only stochastically determined by physical ones.
-
-A third alternative hypothesis is supernatural causal effects.
-The above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
-
-It therefore seems that failing to reject the null hypothesis that a particular mental property $\eps$-supervenes on a particular brain property could potentially be explained by stochastic or supernatural forces, but not a quantum network brain model.
-
-% paragraph stochastic_supervenience (end)
-
-\subsection{Dynamics vs. statics} % (fold)
-\label{par:dynamics_vs_statics}
-
-The \emph{gedanken} experiment in Section \ref{sub:uc} did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
-Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model that determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience holds, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise.
-
-% paragraph dynamics_vs_statics (end)
-
-\subsection{Concluding thoughts} % (fold)
-\label{par:concluding_thoughts}
-
-% paragraph concluding_thoughts (end)
-
-This \emph{gedanken} experiment, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
-The above discussion suggests that many previously conducted investigations either assume supervenience, or test it.
-Further, new technologies facilitate testing supervenience of mental properties on brain-graphs more easily.
-
-\clearpage
-\appendix
+\vspace{5px}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h!tbp]
 	\centering
@@
 \begin{figure}[!ht]
-\centering \includegraphics[width=.5\linewidth]{Lhatplot}
+\centering 
+\includegraphics[width=.5\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.  $\hL$  (the misclassification rate estimated upon with $1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
@@
 We reject $H_0: L_{\PP}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{\PP}(g^*) \approx 0$ for this simulation.
-}
+$L_{\PP}(g^*) \approx 0$ for this simulation.}
 \label{fig1}
@@
 
-\clearpage
-
-% subsection simulation (end)
-
-\section*{Acknowledgments}
-
-The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, and S.~Seung.
-
-% \section{References}
-% \bibliography{biblist}
 \bibliography{/Users/jovo/Research/latex/biblist}
-\bibliographystyle{ieeetr}
+\bibliographystyle{nature}
 

### 2011-07-04 nature_sci_rep/submission2/supervenience.tex
responding to referrees

--- 2011-05-11 nature_sci_rep/supervenience.tex
@@
 \newcommand{\hL}{\widehat{L}}
-\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_{\PP} \mB}
-\newcommand{\MsB}{\mM \overset{S}{\sim}_{\PP} \mB}
-\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{\PP} \mB}
+\newcommand{\MeB}{\mM \overset{\varepsilon}{{\sim}}_{F} \mB}
+\newcommand{\MsB}{\mM \overset{S}{\sim}_{F} \mB}
+\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{F} \mB}
 \providecommand{\tr}[1]{\textcolor{black}{#1}}
@@
 
-\noindent Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
+\noindent Donald Davidson canonized the mind-brain supervenience relation in 1970 with the following quote: \cite{Davidson70}
+\begin{quotation}
+\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
+\end{quotation}
+While supervenience can be a very general claim about arbitrary sets of objects, we consider a special case of considerable interest.
+Specifically, this work addresses a novel version of a local supervenience between mental properties and brain properties.
+The determination of supervenience (or lack thereof) of a mental property on a brain property has potentially important implications in a number of fields of inquiry.
+Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which, if falsified, might drastically change modern approaches to learning \cite{Haykin2008}.
+Cognitive neuroscience similarly seems to operate under such assumptions, which if false, could cause a revolution \cite{Fodor1992, Gazzaniga2008}.
+And the question of mind-brain supervenience continues to be debated amongst philosophers \cite{Kim2007}.
+
+This work does not attempt to resolve any mind-brain supervenience debates.
+Rather, we propose a statistical approach for framing mind-brain supervenience questions.
+This approach depends on defining the space of mental and brain properties under investigation and a statistical model characterizing the possible distributions governing their relationship.
+Such definitions transform supervenience from a conjecture or an assumption, into a hypothesis which can be tested.
+
+% Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
 Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
@@
 
-The primary contributions of this work flow from our introduction of a notion of supervenience amenable to empirical investigation.
-This renders the mind-brain dualism debate a hypothesis, rather than an assumption, both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
+% The primary contributions of this work flow from our introduction of a notion of supervenience amenable to empirical investigation.
+This renders supervenience as a hypothesis, rather than a conjecture or an assumption,  both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
 Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.
@@
 
-Donald Davidson canonized the mind-brain supervenience relation in 1970 with the following quote: \cite{Davidson70}
-\begin{quotation}
-\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
-\end{quotation}
-
-\noindent This conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
+\noindent The $\mc{M}=\{m_1, m_2, \ldots\}$ be a set of possible mental properties.
+For example, $m$ might therefore indicate a person's intelligence, psychological state, current thought, gender identity, etc.
+Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be a set of possible brain properties.
+For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the spike train of all neurons in the brain during some time period $t$ to $t'$.
+Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
 This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).
@@
 
-Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
+% Let $b$ correspond to an agent's brain, which is a particular element from the set of all possible brains, $\mB$.
 Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-Let $\PP[M,B]$ indicate a joint distribution of minds and brains.
-Statistical supervenience can thusly be defined:
-
+Let $F_{M,B}$ indicate a joint distribution of minds and brains. Statistical supervenience can thusly be defined:
 \begin{defi}
 \label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_\PP \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
 \end{defi}
@@
 
-If minds statistically supervene on brains, $\MsB$, then two different minds must supervene on two different brains.
-This means that there exists a deterministic function mapping each brain to its supervening mind, $g(\cdot): \mB \mapsto \mM$; therefore, one could in principle construct this function.
-The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L_{\PP}(g^*)$.
+If minds statistically supervene on brains, $\MsB$, then if two minds differ, there must be some brain-based difference to account for the mental difference.
+This means that there exists a deterministic function mapping each brain to its supervening mind, $g: \mB \mapsto \mM$; therefore, one could in principle construct this function.
+The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L_{F}(g^*)$, under distribuion $F$.
 The minimum misclassification rate is called Bayes error (see Methods for details).
@@
 \label{thm1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\mM \overset{S}{\sim}_{\PP} \mB$, if and only if $L_{\PP}(g^*) = 0$.
-Formally, \mbox{$\MsB \Leftrightarrow L_{\PP}(g^*)=0$}.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\mM \overset{S}{\sim}_{F} \mB$, if and only if $L_{F}(g^*) = 0$.
+Formally, \mbox{$\MsB \Leftrightarrow L_{F}(g^*)=0$}.
 \end{thm}
-
 \noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.
@@
 Statistical supervenience can therefore be thought of as a constraint on the possible distributions on minds and brains.
-Specifically, let $\mc{P}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{P}_s$ be subset of distributions for which supervenience holds.
-Theorem \ref{thm1} implies that $\mc{P}_s = \{\PP[M,B] : L_{\PP}(g^*)=0\} \subseteq \mc{P}$.
+Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s$ be subset of distributions for which supervenience holds.
+Theorem \ref{thm1} implies that $\mc{F}_s = \{F_{M,B} : L_{F}(g^*)=0\} \subseteq \mc{F}$.
 
@@
 
-Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $\PP[M,B]$ and $g^*$, rendering them useless pragmatically.
-However, both $\PP[M,B]$ and $g^*$ could be estimated from data.
+Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $F_{M,B}$ and $g^*$, rendering them useless pragmatically.
+However, both $F_{M,B}$ and $g^*$ could be estimated from data.
 Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
@@
 \label{def2}
-Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $\PP=\PP[M,B]$, denoted $\MeB$, if and only if $L_{\PP}(g^*) < \varepsilon$.
+Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
 \end{defi}
-
-Given this relaxation, consider the problem of testing for $\eps$-supervenience.
-Let the null hypothesis be $H_0$: $L_{\PP}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{\PP}(g_n) < \eps$.
-Hold-out error, $\hL^{n'}_{\PP}(g_{\mt{n}})$, as defined in Methods, is a test statistic for this hypothesis.
-We reject for values of the test statistic lower than the critical value, that is, we reject if and only if $\hL^{n'}_{\PP}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
-The critical value is available under the least favorable distribution $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',\varepsilon)$.
+\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience.
+Let the null hypothesis be $H_0$: $L_{F}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{F}(g_n) < \eps$.
+Hold-out error, $\hL^{n'}_{F}(g_{\mt{n}})$, as defined in Methods, is a test statistic for this hypothesis.
+We reject for values of the test statistic lower than the critical value, that is, we reject if and only if $\hL^{n'}_{F}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
+The critical value is available under the least favorable distribution $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',\varepsilon)$.
 Thus, rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
@@
 A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.
-A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{\PP}(g_n)] \conv L_{\PP}(g^*)$ as $n\conv \infty$.
+A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{F}(g_n)] \conv L_{F}(g^*)$ as $n\conv \infty$.
 Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
 
-Unfortunately, the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ depends on the (unknown) distribution $\PP=\PP[M,B]$ \cite{DGL96}.
-Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{\PP}(g_n)$ to $L_{\PP}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
+Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{M,B}$ \cite{DGL96}.
+Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
 For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-Thus, without restrictions on $\PP[M,B]$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+Thus, without restrictions on $F_{M,B}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
 
@@
 
-To ensure consistency and therefore unity power, the classifier $g_n(\cdot)$ must be able to converge to the truth, regardless of the true distribution, $\PP$.
+To ensure consistency and therefore unity power, the classifier $g_n$ must be able to converge to the truth, regardless of the true distribution, $F$.
 We therefore make explicit a model for brains, and show that under this very general model, universally consistent classifiers are available.
@@
 Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a brain-graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
-Further let $\mB$, the observation space, be the collection of all graphs on a finite number of vertices, and let $|\mB|$ be countable.
+Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
 Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
@@
 
+A failure to reject the null therefore has many possible explanations.
+First, the amount of data might be insufficient given the particular distance implemented; collecting more data or utilizing a more informative distance might resolve this difficulty.
+Second, $\eps$-supervenience is defined between a mental property and a brain property.
+Thus, a failure to reject supervenience on a particular brain property does not entail non-supervenience on any other brain property.
+Third, in general, neither mental properties nor brain properties are directly measurable; rather, one typically measures a function of such properties.
+For instance, instead of measuring intelligence, one considers the results of an IQ test as a proxy for intelligence.
+Similarly, instead of measuring neural spike trains, one estimates spike trains from some measurable neural signal, like voltage.
+While one could conceivably ask questions such as: ``is IQ score at some time supervenient on the the output of a voltage measuring device?'', these are less elegant and general than their non-observable analogs, such as: ``is intelligence supervenient on spike trains?''
+
+Thus, given measurement of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the exchangeability assumption, a rejection entails that we are $100(1-a)\%$ confident that the mental property under investigation does not supervene on the brain property under investigation.
+Unfortunately, failure to reject is more ambiguous.
+$\eps-$supervenience tests can therefore be thought of as constraining the space of possible brain properties upon which a mental property supervenes.
+Determining that a mental property does not supervene on \emph{any} brain property is beyond the capacities of this formalism.
+
 % \paragraph{Practical issues} % (fold)
@@
 
-Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans is practically possible for any particular mental property at this time (at least when the vertices represent individual neurons).
+% Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans is practically possible for any particular mental property at this time (at least when the vertices represent individual neurons).
 Rather, the claim is that \emph{if} one had sufficient data (a very large number of exchangeable mind/brain property pairs), then \emph{in theory}, some level $\eps$-supervenience hypothesis test could be performed.
@@
 The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (although perhaps causality, not supervenience, is the true desideratum).
-Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neurcognitive investigations.
+Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neurocognitive investigations.
 
@@
 
-% \paragraph{Alternative explanations} % (fold)
-
-Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Results, a natural question to raise is: what are the conceivable alternative hypotheses?
+% % \paragraph{Alternative explanations} % (fold)
+% 
+% Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Results, a natural question to raise is: what are the conceivable alternative hypotheses?
 We consider three such alternatives.
-
-First, perhaps brains are more accurately characterized as quantum networks over classical networks.
+% 
+% First, perhaps brains are more accurately characterized as quantum networks over classical networks.
 Several authors have suggested that brains have certain hypercomputational properties that classical computers could not achieve \cite{Penrose99,Satinover02}.
 However, assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
-This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{NielsenChuang00}.
+This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{Nielsen2004}.
 This means that if the above analysis failed to reject the null hypothesis at level $\alpha$, it will fail regardless of whether one assumes quantum or classical computations.
-
-Second, perhaps minds stochastically supervene on brains \cite{Craver09}. While perhaps difficult to imagine, 
-much like a non-deterministic world was difficult to imagine prior to modern physics, it is not inconceivable that mental properties are only stochastically determined by physical ones.
-
-A third alternative hypothesis is supernatural causal effects.
+% 
+% Second, perhaps minds stochastically supervene on brains \cite{Craver09}. While perhaps difficult to imagine, 
+% much like a non-deterministic world was difficult to imagine prior to modern physics, it is not inconceivable that mental properties are only stochastically determined by physical ones.
+% 
+% A third alternative hypothesis is supernatural causal effects.
 The above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
-
-It therefore seems that failing to reject the null hypothesis that a particular mental property $\eps$-supervenes on a particular brain property could potentially be explained by stochastic or supernatural forces, but not a quantum network brain model.
+% 
+% It therefore seems that failing to reject the null hypothesis that a particular mental property $\eps$-supervenes on a particular brain property could potentially be explained by stochastic or supernatural forces, but not a quantum network brain model.
 
@@
 Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as $L_{\PP}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$.
-The Bayes optimal classifier $g^*$ minimizes $L_{\PP}(g)$ over all classifiers, that is:
-$g^* = \argmin_{g} L_{\PP}(g)$.
-Thus, the \emph{Bayes error}, or Bayes risk, $L_{\PP}(g^*)$ is the minimum possible misclassification rate.
+Define misclassification rate as $L_{F}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$.
+The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is:
+$g^* = \argmin_{g} L_{F}(g)$.
+Thus, the \emph{Bayes error}, or Bayes risk, $L_{F}(g^*)$ is the minimum possible misclassification rate.
 
-Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $\PP[M,B]$.
+Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $F_{M,B}$.
 Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
@@
 The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
-$\hL^{n'}_{\PP}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
+$\hL^{n'}_{F}(g_{\mt{n}}) = F[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
 where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
-The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{\PP}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{\PP}(g_{\mt{n}}))$.
+The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{F}(g_{\mt{n}}))$.
 
@@
 
-\bibliography{/Users/jovo/Research/latex/biblist}
+\bibliography{/Users/jovo/Research/latex/library}
 \bibliographystyle{nature}

### 2011-07-04 nature_sci_rep/submission2/supp.tex
responding to referrees

--- 2011-05-11 nature_sci_rep/supp.tex
@@
 \emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
-If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_{\PP} \mM$.
+If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_{F} \mM$.
 Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
@@
 Consider the following problem setup.
-We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangably from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} \PP[M,B]$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
+We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangably from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
 A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim \PP[M,B]$, and $m$ is unobserved.
-Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$.
+It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
+Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$, and that  $\mc{B}$ is countable.
 
@@
 Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone's original proof, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone's original proof \cite{Stone77}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
@@
 Furthermore, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
-Thus, one may design an experiment by describing the joint distribution $\PP[M,B]$ via class-conditional distributions $\PP[B|M=m_j]$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
+Thus, one may design an experiment by describing the joint distribution $F_{M,B}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
 Here the mental property corresponds to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
 
-Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $\PP[M,B]$ (Figure \ref{fig1}).
+Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $F_{M,B}$ (Figure \ref{fig1}).
 To generate the data, let $E_{ij}$ be an integer-valued random variable whose value indicates the number of synapses (edges) between neurons (vertices) $i$ and $j$.
@@
 The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $\PP[M,B]$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
+(Better classifiers can be constructed for the joint distribution $F_{M,B}$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
 
@@
 \caption{C.~elegans graph classification simulation results.  $\hL$  (the misclassification rate estimated upon with $1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
-is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $\PP[M,B]$.
-Each dot depicts an estimate for $L_{\PP}(g_{\mt{n}})$; standard errors are $(L_{\PP}(g_{\mt{n}})(1-L_{\PP}(g_{\mt{n}}))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$; $k_n = 53$; $\hL^{1000}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
-We reject $H_0: L_{\PP}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{\PP}(g^*) \approx 0$ for this simulation.}
+is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{M,B}$.
+Each dot depicts an estimate for $L_{F}(g_{\mt{n}})$; standard errors are $(L_{F}(g_{\mt{n}})(1-L_{F}(g_{\mt{n}}))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$; $k_n = 53$; $\hL^{1000}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
+We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
+$L_{F}(g^*) \approx 0$ for this simulation.}
 \label{fig1}

### 2011-07-06 nature_sci_rep/submission2/supervenience.tex
revised due to referee comments

@@
 
-\noindent Donald Davidson canonized the mind-brain supervenience relation in 1970 with the following quote: \cite{Davidson70}
+\noindent The mind-brain supervenience notion was canonized in 1970 with the following quote from Donald Davidson: \cite{Davidson70}
 \begin{quotation}
-\noindent supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
+\noindent [mind-brain] supervenience might be taken to mean that there cannot be two events alike in all physical respects but differing in some mental respect, or that an object cannot alter in some mental respect without altering in some physical respect.
 \end{quotation}
 While supervenience can be a very general claim about arbitrary sets of objects, we consider a special case of considerable interest.
-Specifically, this work addresses a novel version of a local supervenience between mental properties and brain properties.
+Specifically, this work addresses a novel version of a local supervenience between mental properties and brain properties, with particular emphasize on brain-graphs (i.e., connectivity structure).
 The determination of supervenience (or lack thereof) of a mental property on a brain property has potentially important implications in a number of fields of inquiry.
-Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which, if falsified, might drastically change modern approaches to learning \cite{Haykin2008}.
-Cognitive neuroscience similarly seems to operate under such assumptions, which if false, could cause a revolution \cite{Fodor1992, Gazzaniga2008}.
+Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which if falsified, might change modern approaches to learning \cite{Haykin2008}.
+Cognitive neuroscience similarly seems to operate under such assumptions, which if falsified, might result in novel perspectives and theories \cite{Fodor1998, Gazzaniga2008}.
 And the question of mind-brain supervenience continues to be debated amongst philosophers \cite{Kim2007}.
@@
 
-\noindent The $\mc{M}=\{m_1, m_2, \ldots\}$ be a set of possible mental properties.
-For example, $m$ might therefore indicate a person's intelligence, psychological state, current thought, gender identity, etc.
+\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be a set of possible mental properties.
+For example, $m$ might indicate a person's intelligence, psychological state, current thought, gender identity, etc.
 Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be a set of possible brain properties.
@@
 If minds statistically supervene on brains, $\MsB$, then if two minds differ, there must be some brain-based difference to account for the mental difference.
-This means that there exists a deterministic function mapping each brain to its supervening mind, $g: \mB \mapsto \mM$; therefore, one could in principle construct this function.
+This means that there exists a deterministic function mapping each brain to its supervening mind, $g: \mB \mapsto \mM$.
+One could therefore, in principle, construct this function.
 The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L_{F}(g^*)$, under distribuion $F$.
@@
 % It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.  
-The primary result of casting supervenience as a statistical framework is the following theorem: 
+The primary result of casting supervenience in a statistical framework is the following theorem: 
 \begin{thm}
@@
 
-Although the above theorem is of potential theoretical interest, the arguments rely on knowing the typically unknown $F_{M,B}$ and $g^*$, rendering them useless pragmatically.
-However, both $F_{M,B}$ and $g^*$ could be estimated from data.
+Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unknown $F_{M,B}$, rendering it useless pragmatically.
+However, either $F_{M,B}$ or $g^*$ could be estimated from data.
 Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
@@
 Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
-Now, imagine collecting very large amounts of very accurate independent and identically distributed brain-graph data and the associated mental property indicators.
+Now, imagine collecting very large amounts of very accurate exchangeable  brain-graph data and the associated mental property indicators.
 A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 2 for proof).
 Therefore, %Theorem \ref{thm1} applies and
-the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind/brain property pair, if indeed $\varepsilon$-supervenience holds.
+the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds.
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
@@
 
-We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind/brain property pair is less than $\eps$.
-Furthermore, when we restrict the space of minds and brains to the setting of \emph{Gedankenexperiment}  1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind/brain pair properties.
-Alas, this is a one-sided test, so although power converges to unity, one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
+We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind-brain property pair is less than $\eps$.
+Furthermore, when we restrict the space of minds and brains to the setting of \emph{Gedankenexperiment}  1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind-brain pair properties.
 
-A failure to reject the null therefore has many possible explanations.
+Alas, this is a one-sided test, so although power converges to unity, %one can never determine the reason for a failure to reject.
+%one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
+a failure to reject the null has many possible explanations.
 First, the amount of data might be insufficient given the particular distance implemented; collecting more data or utilizing a more informative distance might resolve this difficulty.
@@
 
-% \paragraph{Practical issues} % (fold)
-% \label{par:practical_issues}
-
-% Importantly, we are \emph{not} claiming that actually determining $\eps$-supervenience in humans is practically possible for any particular mental property at this time (at least when the vertices represent individual neurons).
-Rather, the claim is that \emph{if} one had sufficient data (a very large number of exchangeable mind/brain property pairs), then \emph{in theory}, some level $\eps$-supervenience hypothesis test could be performed.
-Even in such a scenario, a universally consistent classifier is just one of many possible kinds of classifiers, and not necessarily the best one (in terms of $\mh{L}$) for any particular dataset.
-Further, even if a universally consistent classifier is used, a more informative and tractable distance on $\mB$ may be desired, as the $k_n$-nearest neighbor classifier under a Frobenius norm may have a rate of convergence so slow and a computational demand so high as to be impractical (but see Appendix 3 for a simulated example in which convergence is relatively fast).
-Whichever classifier is used, it is likely to benefit from a large amount of domain-specific knowledge, which the proposed classifier completely neglects.
-
-% \paragraph{A unified quest} % (fold)
-% \label{sub:practical_applications}
-
-A central (perhaps \emph{the} central) quest in much of neuroscience, psychology, and cognitive science is to discover the brain properties that subvene under various mental properties, although questions are rarely cast within a  supervenience formalism.
-Moreover, the particular brain properties that are often believed to subvene under these mental properties are neural circuits, or brain-subgraphs.
-To this end, many investigations in these fields include schematic diagrams showing a particular brain-subgraph subvening under a particular mental phenotype.
-This practice transcends the evolutionary hierarchy of neuroscientific research.
-For instance, in the invertebrate literature, vertices correspond to particular labeled neurons, and edges correspond to synapses \cite{NorthGreenspan07}.
-In the vertebrate literature, vertices often correspond to types of neurons in particular regions, and edges correspond to tendencies of connections \cite{Shepherd04}.
-For primates \cite{Felleman_VanEssen91} and humans \cite{Mori05} vertices frequently represent functionally distinct neuroanatomical regions, and edges represent regional interconnectivity.
-Furthermore, this practice also transcends analytical background, including anatomists \cite{Abeles91}, philosophers \cite{Koch_Davis94}, statisticians \cite{Rao_Lewicki02}, and physicists \cite{Chow_Dalibard03}.
-The near ubiquity of this practice suggests that a fundamental quest is to determine which brain-subgraphs subvene under which mental properties (although perhaps causality, not supervenience, is the true desideratum).
-Perhaps supervenience is therefore a framework that can fruitfully be applied to myriad and varied neurocognitive investigations.
+Interestingly, much of contemporary research in neuroscience and cognitive science could be cast as mind-brain supervenience investigations.
+Specifically, any investigation that aims to ``explain'' some behavioral or psychological phenomenon by reference to neural (or other brain) activity or structure, could be thought of as mind-brain supervenience investigations.
+Thus, the proposed $\eps$-supervenience framework is perhaps a useful unifying perspective for these fields.
+Perhaps this is especially true in light of the advent of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, a field devoted to estimating whole organism brain-graphs, and relating them to function.
+Testing supervenience of various mental properties on these brain-graphs will likely therefore become increasingly compelling; so the framework developed herein could be fundamental to these investigations.
+Note that these connectomics investigations might lack any dynamics; rather, they effectively assume that structure determines the relevant dynamics.
+Assuming that dynamics are a function of either vertex or edge attributes, whether any particular mental property is supervenient to ``only'' connectivity structure, or also some attributes, is merely another $\eps$-supervenience hypothesis.
+Appendix 3 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
+Similar supervenience tests on larger animals will require either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011} or more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
 
 % \paragraph{Human applications} % (fold)
-% \label{par:human_applications}
-
-In recent years, with the advent of the field of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, neuroimaging has driven an explosion of studies investigating the human connectome and relating connectomes to cognitive properties \cite{Sporns10}.
-% \cite{Jones10}.
-Many of these studies can be framed as $\eps$-supervenience hypothesis tests.
-For instance, a recent study showed that using data from diffusion tensor imaging \cite{Basser94}, one can nearly perfectly differentiate between schizophrenic individuals and normal (control) individuals \cite{ArdekaniSzeszko10}.
-As the resolution and signal-to-noise ratio of magnetic resonance imaging continue to improve, especially with more advanced techniques such as High Angular Diffusion Imaging \cite{TuchWedeen02}, Q-Ball Imaging \cite{Tuch04}, and diffusion spectrum imaging \cite{WedeenWeisskoff05}, similar results could be obtained with other, more subtle cognitive properties.
-Furthermore, the utilization of other imaging technologies, such as polarized light imaging \cite{PalmAmunts10} and high-throughput electron microscopy \cite{DenkHorstmann04,HayworthLichtman06}, will continue to improve the effective resolution of these inferred connectomes from human brains.
-While determining from a brain scan whether a particular individual knows calculus might be quite distant, many other cognitive and psychological supervenience hypotheses have already been tested, and the gap between testing for calculus and testing for schizophrenia seems to be diminishing.
-
-% % \paragraph{Alternative explanations} % (fold)
-% 
-% Although hypothesis testing for a particular $\epsilon$-supervenience appears to be possible based on the \emph{Gedankenexperiment} in Results, a natural question to raise is: what are the conceivable alternative hypotheses?
-We consider three such alternatives.
-% 
-% First, perhaps brains are more accurately characterized as quantum networks over classical networks.
-Several authors have suggested that brains have certain hypercomputational properties that classical computers could not achieve \cite{Penrose99,Satinover02}.
-However, assuming that the computer can be represented as a network, the above results hold regardless of whether computations in the brain are quantum or classical.
-This follows because quantum networks merely speed up computation for certain classes of problems; they cannot, however, solve problems that classical computers cannot \cite{Nielsen2004}.
-This means that if the above analysis failed to reject the null hypothesis at level $\alpha$, it will fail regardless of whether one assumes quantum or classical computations.
-% 
-% Second, perhaps minds stochastically supervene on brains \cite{Craver09}. While perhaps difficult to imagine, 
-% much like a non-deterministic world was difficult to imagine prior to modern physics, it is not inconceivable that mental properties are only stochastically determined by physical ones.
-% 
-% A third alternative hypothesis is supernatural causal effects.
-The above analysis could be considered an empirical test for whether we have souls, or, perhaps whether souls play a causal role in our mental properties over and above the physical role played by the brain, or whether the data we have suggests that the probability that our souls play a measurable causal role over and above the physical is less than $\eps$.
-% 
-% It therefore seems that failing to reject the null hypothesis that a particular mental property $\eps$-supervenes on a particular brain property could potentially be explained by stochastic or supernatural forces, but not a quantum network brain model.
-
 % \paragraph{Dynamics vs. statics} % (fold)
-% \label{par:dynamics_vs_statics}
-
-The \emph{Gedankenexperiment} in Section \ref{sub:uc} did not require simulating any dynamics; rather, the dynamics are necessarily a function of the model parameters (statics).
-Similarly, for the question of mind-brain supervenience in humans, one need not ever observe any activity of the brain, one must merely observe the model that determines the activity (in a potentially stochastic process).
-Thus, this approach to understanding the relationship between mind and brain is distinct from the standard systems neuroscience paradigm, in which the goal is typically to understand the neural activity ``code.''  In contrast, if mind-brain supervenience holds, it motivates a search for the neural \emph{connectivity} ``code,'' a so-called ``engram''  for memories \cite{Semon21, Lashley50, ZhangLinden03, ShemaDudai07, BerryDavis08}, or more generally a \emph{mengram}, the neural signature of any mental property, be it cognitive, psychological, or otherwise (note that supervenience allows for the particular mengram of a mental property to vary both across individuals and time).
-
 % \paragraph{Concluding thoughts} % (fold)
@@
 
-This \emph{Gedankenexperiment}, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
-The above discussion suggests that many previously conducted investigations either assume supervenience, or test it.
+In conclusion, this \emph{Gedankenexperiment}, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
+The above discussion suggests that many previously conducted investigations either assume supervenience, or try to test it.
 Further, new technologies facilitate testing supervenience of mental properties on brain-graphs more easily.
@@
 
-Assuming for the moment that the space of all possible minds is finite, that is, $|\mM| < \infty$, then we call any such function a \emph{classifier} (this assumption will later be relaxed).
+Assuming for the moment that the space of all possible minds is finite, that is, $|\mM| < \infty$, then we call any function, $g: \mc{B} \mapsto \mc{M}$, a \emph{classifier} (this assumption will later be relaxed).
 Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as $L_{F}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$.
+Define misclassification rate as $L_{F}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$ under distribution $F$.
 The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is:
 $g^* = \argmin_{g} L_{F}(g)$.
-Thus, the \emph{Bayes error}, or Bayes risk, $L_{F}(g^*)$ is the minimum possible misclassification rate.
+Thus, the \emph{Bayes error}, or Bayes risk, $L_{F}(g^*)$, is the minimum possible misclassification rate.
 
 Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $F_{M,B}$.
-Generalizing the concept of a classifier $g$ to allow incorporation of training data, consider $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$ which takes as input an observed brain  $b$ and training data $\mc{T}_n$, and produces a classification: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
+An \emph{induced} classifier, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$, takes as input an observed brain $b$ and training data $\mc{T}_n$, and classifies: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
 Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
 The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
-$\hL^{n'}_{F}(g_{\mt{n}}) = F[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
-where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}(\cdot)$).
-The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{F}(g_{\mt{n}}))$.
+$\hL^{n'}_{F}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
+where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}$).
+% The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{F}(g_{\mt{n}}))$.
 
 \clearpage
-
-% subsection simulation (end)
-
 \bibliography{/Users/jovo/Research/latex/library}
 \bibliographystyle{nature}
-
-% \begin{thebibliography}{10}
-% \expandafter\ifx\csname url\endcsname\relax
-%   \def\url#1{\texttt{#1}}\fi
-% \expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-% \providecommand{\bibinfo}[2]{#2}
-% \providecommand{\eprint}[2][]{\url{#2}}
-% 
-% \bibitem{Plato97}
-% \bibinfo{author}{Plato}.
-% \newblock \emph{\bibinfo{title}{Plato: complete works}}
-%   (\bibinfo{publisher}{Hackett Pub Co}, \bibinfo{year}{1997}).
-% 
-% \bibitem{Descartes1641}
-% \bibinfo{author}{Descartes, R.}
-% \newblock \emph{\bibinfo{title}{Meditationes de prima philosophia}}
-%   (\bibinfo{year}{1641}).
-% 
-% \bibitem{Davidson70}
-% \bibinfo{author}{Davidson, D.}
-% \newblock \emph{\bibinfo{title}{Experience and Theory}}, chap.
-%   \bibinfo{chapter}{Mental Events} (\bibinfo{publisher}{Duckworth},
-%   \bibinfo{year}{1970}).
-% 
-% \bibitem{DGL96}
-% \bibinfo{author}{Devroye, L.}, \bibinfo{author}{Gy{\"o}rfi, L.} \&
-%   \bibinfo{author}{Lugosi, G.}
-% \newblock \emph{\bibinfo{title}{{A Probabilistic Theory of Pattern
-%   Recognition}}} (\bibinfo{publisher}{Springer}, \bibinfo{year}{1996}).
-% 
-% \bibitem{Devroye83}
-% \bibinfo{author}{Devroye, L.}
-% \newblock \bibinfo{title}{{On arbitrarily slow rates of global convergence in
-%   density estimation}}.
-% \newblock \emph{\bibinfo{journal}{Probability Theory and Related Fields}}
-%   \textbf{\bibinfo{volume}{62}}, \bibinfo{pages}{475--483}
-%   (\bibinfo{year}{1983}).
-% 
-% \bibitem{Popper}
-% \bibinfo{author}{Popper, K.}
-% \newblock \bibinfo{title}{{The logic of scientific discovery}}
-%   (\bibinfo{year}{1959}).
-% 
-% \bibitem{NorthGreenspan07}
-% \bibinfo{editor}{Geoffrey~North, R. J.~G.} (ed.)
-%   \emph{\bibinfo{title}{Invertebrate neurobiology}} (\bibinfo{publisher}{CSHL
-%   Press}, \bibinfo{year}{2007}).
-% 
-% \bibitem{Shepherd04}
-% \bibinfo{author}{Shepherd, G.}
-% \newblock \emph{\bibinfo{title}{{The synaptic organization of the brain}}}
-%   (\bibinfo{publisher}{Oxford University Press New York},
-%   \bibinfo{year}{2004}).
-% 
-% \bibitem{Felleman_VanEssen91}
-% \bibinfo{author}{Felleman, D.} \& \bibinfo{author}{Van~Essen, D.}
-% \newblock \bibinfo{title}{{Distributed hierarchical processing in the primate
-%   cerebral cortex}}.
-% \newblock \emph{\bibinfo{journal}{Cerebral cortex}}
-%   \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{1} (\bibinfo{year}{1991}).
-% 
-% \bibitem{Mori05}
-% \bibinfo{author}{S.~Mori, P. v. Z. L. N.-P., S.~Wakana}.
-% \newblock \emph{\bibinfo{title}{MRI Atlas of Human White Matter}}
-%   (\bibinfo{publisher}{Elsevier Science}, \bibinfo{year}{2005}).
-% 
-% \bibitem{Abeles91}
-% \bibinfo{author}{Abeles, M.}
-% \newblock \emph{\bibinfo{title}{Corticonics}} (\bibinfo{publisher}{Cambridge
-%   University Press}, \bibinfo{year}{1991}).
-% 
-% \bibitem{Koch_Davis94}
-% \bibinfo{author}{Koch, C.} \& \bibinfo{author}{Davis, J.}
-% \newblock \emph{\bibinfo{title}{{Large-scale neuronal theories of the brain}}}
-%   (\bibinfo{publisher}{The MIT Press}, \bibinfo{year}{1994}).
-% 
-% \bibitem{Rao_Lewicki02}
-% \bibinfo{author}{Rao, R.}, \bibinfo{author}{Olshausen, B.} \&
-%   \bibinfo{author}{Lewicki, M.}
-% \newblock \emph{\bibinfo{title}{{Probabilistic models of the brain: Perception
-%   and neural function}}} (\bibinfo{publisher}{The MIT Press},
-%   \bibinfo{year}{2002}).
-% 
-% \bibitem{Chow_Dalibard03}
-% \bibinfo{editor}{Chow, C.}, \bibinfo{editor}{Gutkin, B.},
-%   \bibinfo{editor}{Hansel, D.}, \bibinfo{editor}{Meunier, C.} \&
-%   \bibinfo{editor}{Dalibard, J.} (eds.) \emph{\bibinfo{title}{Methods and
-%   Models in Neurophysics}} (\bibinfo{publisher}{Elsevier},
-%   \bibinfo{year}{2003}).
-% 
-% \bibitem{SpornsKotter05}
-% \bibinfo{author}{Sporns, O.}, \bibinfo{author}{Tononi, G.} \&
-%   \bibinfo{author}{Kotter, R.}
-% \newblock \bibinfo{title}{The human connectome: A structural description of the
-%   human brain}.
-% \newblock \emph{\bibinfo{journal}{PLoS Computational Biology}}
-%   \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{e42} (\bibinfo{year}{2005}).
-% 
-% \bibitem{Hagmann05}
-% \bibinfo{author}{Hagmann, P.}
-% \newblock \emph{\bibinfo{title}{{From diffusion MRI to brain connectomics}}}.
-% \newblock Ph.D. thesis, \bibinfo{school}{Institut de traitement des signaux}
-%   (\bibinfo{year}{2005}).
-% 
-% \bibitem{Sporns10}
-% \bibinfo{author}{Sporns, O.}
-% \newblock \emph{\bibinfo{title}{Networks of the Brain}}
-%   (\bibinfo{publisher}{MIT Press}, \bibinfo{year}{2010}).
-% 
-% \bibitem{Basser94}
-% \bibinfo{author}{Basser, P.~J.}, \bibinfo{author}{Mattiello, J.} \&
-%   \bibinfo{author}{LeBihan, D.}
-% \newblock \bibinfo{title}{Mr diffusion tensor spectroscopy and imaging.}
-% \newblock \emph{\bibinfo{journal}{Biophys J}} \textbf{\bibinfo{volume}{66}},
-%   \bibinfo{pages}{259--267} (\bibinfo{year}{1994}).
-% \newblock \urlprefix\url{http://dx.doi.org/10.1016/S0006-3495(94)80775-1}.
-% 
-% \bibitem{ArdekaniSzeszko10}
-% \bibinfo{author}{Ardekani, B.~A.} \emph{et~al.}
-% \newblock \bibinfo{title}{Diffusion tensor imaging reliably differentiates
-%   patients with schizophrenia from healthy volunteers.}
-% \newblock \emph{\bibinfo{journal}{Hum Brain Mapp}}  (\bibinfo{year}{2010}).
-% \newblock \urlprefix\url{http://dx.doi.org/10.1002/hbm.20995}.
-% 
-% \bibitem{TuchWedeen02}
-% \bibinfo{author}{Tuch, D.} \emph{et~al.}
-% \newblock \bibinfo{title}{{High angular resolution diffusion imaging reveals
-%   intravoxel white matter fiber heterogeneity}}.
-% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
-%   \textbf{\bibinfo{volume}{48}}, \bibinfo{pages}{577--582}
-%   (\bibinfo{year}{2002}).
-% 
-% \bibitem{Tuch04}
-% \bibinfo{author}{Tuch, D.}
-% \newblock \bibinfo{title}{{Q-ball imaging}}.
-% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
-%   \textbf{\bibinfo{volume}{52}}, \bibinfo{pages}{1358--1372}
-%   (\bibinfo{year}{2004}).
-% 
-% \bibitem{WedeenWeisskoff05}
-% \bibinfo{author}{Wedeen, V.}, \bibinfo{author}{Hagmann, P.},
-%   \bibinfo{author}{Tseng, W.}, \bibinfo{author}{Reese, T.} \&
-%   \bibinfo{author}{Weisskoff, R.}
-% \newblock \bibinfo{title}{{Mapping complex tissue architecture with diffusion
-%   spectrum magnetic resonance imaging}}.
-% \newblock \emph{\bibinfo{journal}{Magnetic Resonance in Medicine}}
-%   \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{1377--1386}
-%   (\bibinfo{year}{2005}).
-% 
-% \bibitem{PalmAmunts10}
-% \bibinfo{author}{Palm, C.} \emph{et~al.}
-% \newblock \bibinfo{title}{{Towards ultra-high resolution fibre tract mapping of
-%   the human brain--registration of polarised light images and reorientation of
-%   fibre vectors}}.
-% \newblock \emph{\bibinfo{journal}{Frontiers in Human Neuroscience}}
-%   \textbf{\bibinfo{volume}{4}} (\bibinfo{year}{2010}).
-% 
-% \bibitem{DenkHorstmann04}
-% \bibinfo{author}{W.~Denk, W.} \& \bibinfo{author}{Horstmann, H.}
-% \newblock \bibinfo{title}{Serial block-face scanning electron microscopy to
-%   reconstruct three-dimensional tissue nanostructure}.
-% \newblock \emph{\bibinfo{journal}{PLOS Biol.}} \textbf{\bibinfo{volume}{2}},
-%   \bibinfo{pages}{e329} (\bibinfo{year}{2004}).
-% 
-% \bibitem{HayworthLichtman06}
-% \bibinfo{author}{Hayworth, K.}, \bibinfo{author}{Kasthuri, N.},
-%   \bibinfo{author}{Schalek, R.} \& \bibinfo{author}{Lichtman, J.}
-% \newblock \bibinfo{title}{{Automating the collection of ultrathin serial
-%   sections for large volume TEM reconstructions}}.
-% \newblock \emph{\bibinfo{journal}{Microscopy and Microanalysis}}
-%   \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{86--87}
-%   (\bibinfo{year}{2006}).
-% 
-% \bibitem{Penrose99}
-% \bibinfo{author}{Penrose, R.} \& \bibinfo{author}{Gardner, M.}
-% \newblock \emph{\bibinfo{title}{{The emperor's new mind: concerning computers,
-%   minds, and the laws of physics}}} (\bibinfo{publisher}{Oxford University
-%   Press, USA}, \bibinfo{year}{1999}).
-% 
-% \bibitem{Satinover02}
-% \bibinfo{author}{Satinover, J.}
-% \newblock \emph{\bibinfo{title}{{The quantum brain: the search for freedom and
-%   the next generation of man}}} (\bibinfo{publisher}{Wiley},
-%   \bibinfo{year}{2002}).
-% 
-% \bibitem{NielsenChuang00}
-% \bibinfo{author}{Nielson, M.~A.} \& \bibinfo{author}{Chuang, I.~L.}
-% \newblock \emph{\bibinfo{title}{Quantum Computation and Quantum Information}}
-%   (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{2000}).
-% 
-% \bibitem{Craver09}
-% \bibinfo{author}{Craver, C.~F.}
-% \newblock \bibinfo{title}{Stochastic supervenience}  (\bibinfo{year}{2009}).
-% 
-% \bibitem{Semon21}
-% \bibinfo{author}{Semon, R.~W.}
-% \newblock \emph{\bibinfo{title}{The Mneme}} (\bibinfo{publisher}{G. Allen \&
-%   Unwin ltd.}, \bibinfo{year}{1921}).
-% 
-% \bibitem{Lashley50}
-% \bibinfo{author}{Lashley, K.}
-% \newblock \bibinfo{title}{{In search of the engram}}.
-% \newblock \emph{\bibinfo{journal}{Symposia of the society for experimental
-%   biology}} \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{30}
-%   (\bibinfo{year}{1950}).
-% 
-% \bibitem{ZhangLinden03}
-% \bibinfo{author}{Zhang, W.} \& \bibinfo{author}{Linden, D.}
-% \newblock \bibinfo{title}{{The other side of the engram: experience-driven
-%   changes in neuronal intrinsic excitability}}.
-% \newblock \emph{\bibinfo{journal}{Nature Reviews Neuroscience}}
-%   \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{885--900}
-%   (\bibinfo{year}{2003}).
-% 
-% \bibitem{ShemaDudai07}
-% \bibinfo{author}{Shema, R.}, \bibinfo{author}{Sacktor, T.} \&
-%   \bibinfo{author}{Dudai, Y.}
-% \newblock \bibinfo{title}{{Rapid erasure of long-term memory associations in
-%   the cortex by an inhibitor of PKM $\{$zeta$\}$}}.
-% \newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{317}},
-%   \bibinfo{pages}{951} (\bibinfo{year}{2007}).
-% 
-% \bibitem{BerryDavis08}
-% \bibinfo{author}{Berry, J.}, \bibinfo{author}{Krause, W.} \&
-%   \bibinfo{author}{Davis, R.}
-% \newblock \bibinfo{title}{{Olfactory memory traces in Drosophila}}.
-% \newblock \emph{\bibinfo{journal}{Progress in brain research}}
-%   \textbf{\bibinfo{volume}{169}}, \bibinfo{pages}{293--304}
-%   (\bibinfo{year}{2008}).
-% 
-% \end{thebibliography}
 

### 2011-07-06 nature_sci_rep/submission2/supp.tex
revised due to referee comments

@@
 Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
-Rather, position can change with constant velocity, meaning without acceleration changing.
+In other words, acceleration cannot change without a corresponding change in position, which is the requirement for supervenience.
+More formally, let $a$ and $a'$ indicate accelerations and $p$ and $p'$ indicate positions, we do not have $\PP[a \neq a' | p \neq p'] = 0$; rather, position can change with constant velocity, meaning without acceleration changing.
 Thus, to claim that something supervenes on another is not equivalent to claim that the two are identical.
@@
 Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+In fact, supervenience may be thought of as a generalization of correlation which can be applied to arbitrary valued random variables (such as mental or brain properties), unlike correlation which is only defined for scalar valued random variables.
 
@@
 Consider the following problem setup.
-We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangably from some unknown joint distribution, $(m_i,b_i)\overset{iid}{\sim} F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
+We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangeably from some unknown joint distribution, $(m_i,b_i)\sim F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
 A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind/brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
+It is further assumed that the test mind-brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
 Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$, and that  $\mc{B}$ is countable.
@@
 Unfortunately, no such algorithm is universally consistent.
-Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training points.
+Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training samples.
 Then, a collection of such algorithms, $\{g_n\}$,  with $k_n$ increasing with $n$, can be universally consistent under certain constraints.
@@
 Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone's original proof \cite{Stone77}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone's original proof \cite{Stone1977}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
-When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into finite Euclidean space, in which case Stone's theorem applies.
-Stone's original proof applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
+When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into a vector space, in which case Stone's theorem applies.
+Stone's original proof also applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
 
@@
 Furthermore, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
-Thus, one may design an experiment by describing the joint distribution $F_{M,B}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $\PP[M=m_1]$.
+Thus, one may design an experiment by describing the joint distribution $F_{M,B}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $F_{M=m_j}$.
 Here the mental property corresponds to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
@@
 \includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.  $\hL$  (the misclassification rate estimated upon with $1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+\caption{C.~elegans graph classification simulation results.  $\hL^{n'}_{F}(g_{\mt{n}})$  (the misclassification rate estimated upon with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
 is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{M,B}$.
-Each dot depicts an estimate for $L_{F}(g_{\mt{n}})$; standard errors are $(L_{F}(g_{\mt{n}})(1-L_{F}(g_{\mt{n}}))/1000)^{1/2}$\tr{; e}.g., $n_j = 180$; $k_n = 53$; $\hL^{1000}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
+Each dot depicts an estimate for $\hL^{n'}_{F}(g_{\mt{n}})$; standard errors are $(\hL^{n'}_{F}(g_{\mt{n}}) (1-\hL^{n'}_{F}(g_{\mt{n}}))/n')^{1/2}$; e.g., $n_j = 180$; $k_n = 53$; $\hL^{n'}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
 We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
@@
 Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{VaziriShank08} combined with neurite tracing algorithms \cite{HelmstaedterDenk08,Mishchenko09,LuLichtman09} allow the collection of a C.
+3D superresolution imaging \cite{Vaziri2008} combined with neurite tracing algorithms \cite{Helmstaedter2008,Mishchenko09,LuLichtman09} allow the collection of a C.
 elegans brain-graph within a day.
-Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{BuckinghamSattelle08}.
+Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{Buckingham2008}.
 

### 2011-07-25 nature_sci_rep/submission2/supervenience.tex
up to (excluding) gedanken is ok now

@@
 \end{quotation}
-While supervenience can be a very general claim about arbitrary sets of objects, we consider a special case of considerable interest.
+We consider a special case of supervenience of considerable interest.
 Specifically, this work addresses a novel version of a local supervenience between mental properties and brain properties, with particular emphasize on brain-graphs (i.e., connectivity structure).
 The determination of supervenience (or lack thereof) of a mental property on a brain property has potentially important implications in a number of fields of inquiry.
-Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which if falsified, might change modern approaches to learning \cite{Haykin2008}.
+Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which if falsified, might change modern approaches to learning \cite{Haykin2008,Ripley2008}.
 Cognitive neuroscience similarly seems to operate under such assumptions, which if falsified, might result in novel perspectives and theories \cite{Fodor1998, Gazzaniga2008}.
@@
 
-This work does not attempt to resolve any mind-brain supervenience debates.
+This work does not attempt to resolve any particular mind-brain supervenience debates.
 Rather, we propose a statistical approach for framing mind-brain supervenience questions.
@@
 
-% Questioning the relationship between the mind (our thoughts, beliefs, preferences, emotions, intelligences, etc.) and the brain (the physical structure inside our skulls) dates back at least as far as 400~BCE, when Plato wrote the dialogues, in which he posited immateriality of the soul \cite{Plato97}.
-Approximately two millennia passed before these ideas reached their canonical form through Descartes's discussion of mind-body dualism \cite{Descartes1641}.
-Then, in the 20th century, Donald Davidson stated and popularized the mind-brain supervenience \tr{conjecture}, which claims that an agent cannot alter in some mental property without altering in some physical property \cite{Davidson70}.
-Contemporary fields of neural network theory and neuroscientific inquiry often assume mind-brain supervenience, or an even stronger assumption about mind-brain causality, but no previously proposed notion of supervenience seems amenable to empirical investigation.
-This work is an attempt to bridge the gap between these philosophical conjectures and experimentally testable hypotheses.
-% Here we define new versions of supervenience that formulate the \tr{conjecture} in rigorous mathematical terms, and that can be experimentally tested \tr{as a hypothesis}.
-
-% The primary contributions of this work flow from our introduction of a notion of supervenience amenable to empirical investigation.
-This renders supervenience as a hypothesis, rather than a conjecture or an assumption,  both expanding the space of questions amenable to hypothesis testing, and placing limits on this space.
-Because hypothesis tests (implicitly sometimes) depend on a model, a very general model of brains and their associated mental properties is proposed.
-Fortunately, this formulation admits universally consistent classifiers, that is, classifiers guaranteed to find the relationship between minds and brains, if one exists, given sufficient data.
-Many previous investigations relating brains and mental properties can therefore be considered $\eps$-supervenience hypothesis tests.
-This paradigm, therefore, generalizes previous approaches, embedding them in a rigorous statistical framework, and suggests avenues for future research.
-
- % we demonstrate through simulation that the proposed universally consistent classifier has reasonable convergence properties on simulated brain-graph data.
-Finally, we discuss both contemporary and future applications of this framework to neurobiological studies in humans.
-
 \section*{Results}
 
-\subsection*{Statistical supervenience} % (fold)
+\subsection*{Statistical supervenience; a definition} % (fold)
 
@@
 Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be a set of possible brain properties.
-For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the spike train of all neurons in the brain during some time period $t$ to $t'$.
+For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$.
 Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
@@
 Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-Let $F_{M,B}$ indicate a joint distribution of minds and brains. Statistical supervenience can thusly be defined:
+Let $F_{MB}$ indicate a joint distribution of minds and brains. Statistical supervenience can thusly be defined:
 \begin{defi}
 \label{def1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
 \end{defi}
-\noindent Statistical supervenience is therefore a probabilistic relation on sets (related to, but distinct from correlation; see Appendix 1 for details).
+\noindent Statistical supervenience is therefore a probabilistic relation on sets (which could be considered a generalization of correlation; see Appendix 1 for details).
 
@@
 One could therefore, in principle, construct this function.
-The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L_{F}(g^*)$, under distribuion $F$.
+When the space of all possible minds is finite; that is, $|\mM| < \infty$, $g$ is called a \emph{classifier}.
+% Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.  
+Define misclassification rate, the probability that $g$ misclassifies $b$ under distribution $F=F_{MB}$,  as
+\begin{align}
+L_{F}(g) = \PP[g(B) \neq M] = \frac{1}{|\mc{M} \times \mc{B}|} \sum_{(m,b) \in \mc{M} \times \mc{B}} \II\{g(b) \neq m\} \PP[B=b, M=m]	
+\end{align}
+where $\II$ denotes the indicator function taking value unity whenever its argument is true, and zero otherwise.
+The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is:
+$g^* = \argmin_{g} L_{F}(g)$.
+The \emph{Bayes error}, or Bayes risk, $L^*=L_{F}(g^*)$, is the minimum possible misclassification rate.
+
+% The optimal such classifier, $g^*$, has the smallest expected misclassification rate, $L^*=L_{F}(g^*)$, under distribuion $F$.
 The minimum misclassification rate is called Bayes error (see Methods for details).
-% It may be the case that subsets of brains form equivalence classes, such that any brain in that subset is mapped to the same mind.  
 The primary result of casting supervenience in a statistical framework is the following theorem: 
@@
 \label{thm1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\mM \overset{S}{\sim}_{F} \mB$, if and only if $L_{F}(g^*) = 0$.
-Formally, \mbox{$\MsB \Leftrightarrow L_{F}(g^*)=0$}.
+$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\mM \overset{S}{\sim}_{F} \mB$, if and only if $L^*= 0$.
+Formally, \mbox{$\MsB \Leftrightarrow L^*=0$}.
 \end{thm}
-\noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.
+\begin{proof}
+Let $\mc{M}_b=\{m: \PP[B=b | M=m] >0\}$.  
+$$L^*=0 \Leftrightarrow |\mc{M}_b|=1 \, \forall \, b \in \mc{B}.$$
+In words, for $L^*$ to equal 0, it must be the case that there exists only a single $m$ for each $b$.
+\end{proof}
+
+% \noindent If minds supervene on brains, then, by the definition of supervenience, there exists a function that maps each brain deterministically to a particular mind.
 This means that one could draw a decision boundary between all equivalence classes of brains, each class corresponding to a different mind, and no mind will reside within two different equivalence classes.
@@
 Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s$ be subset of distributions for which supervenience holds.
-Theorem \ref{thm1} implies that $\mc{F}_s = \{F_{M,B} : L_{F}(g^*)=0\} \subseteq \mc{F}$.
+Theorem \ref{thm1} implies that $\mc{F}_s = \{F_{MB} : L^*=0\} \subseteq \mc{F}$.
+
+\subsection*{The intractability of testing for statistical supervenience} % (fold)
+\label{sub:subsection_name}
+
+% subsection subsection_name (end)
+
+Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unavailable $L^*$, rendering it useless pragmatically.
+However,  $g^*$ could be estimated from data, and its estimate could be utilized to conduct the following statistical test of supervenience:
+% Let $g_n$ be a classifier induced from some training data, $\mc{T}_n$, 
+\begin{align*} 
+	H_0:& L^* > 0 \\
+	H_A:& L^*=0
+\end{align*}
+A statistical test proceeds by calculating a test statistic and a critical value $c_\alpha$.
+We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$.
+The $p$-value, the probability of rejecting a true null hypothesis, is a function of the distribution of the test statistic.
+
+Let $g_n$ be a classifier induced by the data, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$.
+Assume that training data $\mc{T}_n=\{(M_{1},B_{1}), \ldots, (M_{n},B_{n})\}$ are each sampled identically and independently (iid) from the true (but unknown) joint distribution, $F_{MB}$.
+The misclassification rate of a classifier will therefore be a random variable.
+The expected misclassification rate is given by
+\begin{align}
+\EE[L_F(g_n)]=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m].
+\end{align}
+Because calculating the expected misclassification rate requires a sum over all possible mind/brain property pairs, calculating $\EE[L_F(g_n)]$ will often be intractable in practice.
+Instead, expected misclassification rate is approximated by ``hold-out'' error.
+Let $\mc{H}_{n'}=\{(M_{n+1},B_{n+1}), \ldots, (M_{n+n'},B_{n+n'})\}$ be a set of $n'$ hold-out samples, each sampled iid from $F_{MB}$.
+The hold-out approximation to misclassification rate is given by
+\begin{align}
+\hL^{n'}_{F}(g_{n}) = \sum_{(m,b) \in \mc{H}_{n'}}\II \{g_{n}(b; \mc{T}_{n})\neq m] \PP[B=b,M=m].	
+\end{align}
+Under the above assumptions, $\hL^{n'}_F(g_n)$ can be used as a surrogate for $L^*$.
+By definition of $g^*$, we expect $\hL^{n'}_F(g_n) \geq g^*$ for any $g_n$ and all $n$.
+We therefore let $\mh{n}= \lfloor n' \hL^{n'}_{F}(g_n) \rfloor$ be the test statistic, where $\lfloor \cdot \rfloor$ is the floor operator.
+The distribution of $\mh{n}$ is available under the least favorable distribution of the alternate hypothesis.
+Specifically,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', p_{H_A})= \sum_{k \in [\mh{n}]}$Binomial$(k; n'; p_{H_A})$, where $p_{H_A}$ is Bernoulli probability under least favorable distribution of the alternate hypothesis and $[\mh{n}]=\{1,\ldots, \mh{n}\}$.
+In this simple alternate hypothesis, $p_{H_A}=0$.
+Because Binomial$(k; n', 0)=1$ for all $n'$, the $p$-value is always equal to unity.
+Thus, we can \emph{never} reject the above null hypothesis, no matter how much data we obtain.
 
@@
 
-Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unknown $F_{M,B}$, rendering it useless pragmatically.
-However, either $F_{M,B}$ or $g^*$ could be estimated from data.
-Before explicitly considering the problem of testing for statistical supervenience, we define a relaxed notion of supervenience:
+We therefore consider a relaxed notion of supervenience:
 \begin{defi}
 \label{def2}
-Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{M,B}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
+Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
 \end{defi}
-\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience.
-Let the null hypothesis be $H_0$: $L_{F}(g_n) \geq \eps$, and the alternative hypothesis be $H_A$: $L_{F}(g_n) < \eps$.
-Hold-out error, $\hL^{n'}_{F}(g_{\mt{n}})$, as defined in Methods, is a test statistic for this hypothesis.
-We reject for values of the test statistic lower than the critical value, that is, we reject if and only if $\hL^{n'}_{F}(g_{\mt{n}}) <c_{\alpha}(n',\varepsilon)$.
-The critical value is available under the least favorable distribution $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',\varepsilon)$.
-Thus, rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
-The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a statistical test of supervenience, given a specified $\eps$ and $\alpha$.
+\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience:
+\begin{align*}
+	H_0^{\eps}: L^* \geq \eps \\
+	H_A^{\eps}: L^* < \eps
+\end{align*}
+The $p$-value of the above composite alternate hypothesis is given by the binomial distribution under the least favorable distribution, $\mathbb{B}(\mh{n}; n', p_{H_A}=\eps)<1$.
+We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
+The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$.
 
-Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis.
+\subsection*{A \emph{Gedankenexperiment} demonstrating an ideal test} % (fold)
+\label{sub:uc}
+
+Ideally, as $n$ increases, the probability of both a Type I error (rejection of a true null hypothesis) and a Type II error (failure to reject a false null hypothesis) goes to zero.
+
+%For the above described hypothesis of $\eps$-supervenience, both the $p$-value (probability of a type I error)  and power (probability of not making a type II error), will depend on the distribution  of $g_n$.
+We therefore consider the properties of $g_n$, and the implications of those properties with regard to the probability of a Type I or Type II error.
+
+The $p$-value 
+(probability of a type I error)
+of the test statistic for any classifier $g_n$ converges to zero, because $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n))) \conv 0$ as $n,n' \conv \infty$.
+The power of the test (probability of not making a type II error), however, only converges to unity under certain conditions.
+Specifically, let a  \emph{consistent} classifier be any classifier with expected misclassification rate converging to the Bayes optimal limit; that is, $\EE[L_{F}(g_n)] \conv L_{F}(g^*)$ as $n\conv \infty$.
+The power of the test can be written as: $\PP[\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) < \mathbb{B}(\mh{n}; n', \eps)]$.
+As $n,n' \conv \infty$, $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) \conv \mathbb{B}(\mh{n}; n', L^*)$.
+And if $L^* < \eps$, then $\mathbb{B}(\mh{n}; n', L^*) < \mathbb{B}(\mh{n}; n', \eps)$.
+So, the probability of a Type II error also goes to zero as $n,n' \conv \infty$.
+
+% Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis.
 Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
 A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.
-A classifier is consistent if and only if its expected misclassification rate converges to the Bayes optimal limit with sufficient data, that is $\EE[L_{F}(g_n)] \conv L_{F}(g^*)$ as $n\conv \infty$.
+
+Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{MB}$ \cite{DGL96}.
+Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
+For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
+Thus, without restrictions on $F_{MB}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+
 Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
 
-Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{M,B}$ \cite{DGL96}.
-Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-Thus, without restrictions on $F_{M,B}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-
-\subsection*{A \emph{Gedankenexperiment} demonstrating consistency and unity power} % (fold)
-\label{sub:uc}
-
-To ensure consistency and therefore unity power, the classifier $g_n$ must be able to converge to the truth, regardless of the true distribution, $F$.
+% To ensure consistency and therefore unity power, the classifier $g_n$ must be able to converge to the truth, regardless of the true distribution, $F$.
 We therefore make explicit a model for brains, and show that under this very general model, universally consistent classifiers are available.
@@
 \begin{thex}
-Let the physical property under consideration be brain connectivity structure (``connectome''), so $b$ is a brain-graph \tr{(or, network)} with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
+Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
 Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
 Now, imagine collecting very large amounts of very accurate exchangeable  brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 2 for proof).
+A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 2 for proof sketch).
 Therefore, %Theorem \ref{thm1} applies and
@@
 
-Assuming for the moment that the space of all possible minds is finite, that is, $|\mM| < \infty$, then we call any function, $g: \mc{B} \mapsto \mc{M}$, a \emph{classifier} (this assumption will later be relaxed).
-Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.
-Define misclassification rate as $L_{F}(g) = \PP[g(B) \neq M]$ which denotes the probability that $g$ misclassifies $b$ under distribution $F$.
-The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is:
-$g^* = \argmin_{g} L_{F}(g)$.
-Thus, the \emph{Bayes error}, or Bayes risk, $L_{F}(g^*)$, is the minimum possible misclassification rate.
-
-Let $\mc{T}_n=\{(m_1,b_1), (m_2,b_2), \ldots, (m_n,b_n)\}$ be a set of random samples taking their values in $\mc{M} \times \mc{B}$, each independently and identically distributed according to $F_{M,B}$.
-An \emph{induced} classifier, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$, takes as input an observed brain $b$ and training data $\mc{T}_n$, and classifies: $g_n(b;\tr{\mc{T}_n})=\mh{m}$.
-Misclassification rate for this classifier will be a random variable, because the training data $\mc{T}_n$ are random samples.
-The expected misclassification rate for this classifier is therefore approximated by ``hold-out'' error:
-$\hL^{n'}_{F}(g_{\mt{n}}) = \PP[g_{\mt{n}}(B)=M | \mc{T}_{\mt{n}}]$,
-where $\mt{n}=n-n'$, and $n'<n$ is the number of held-out training samples (samples not used to obtain $g_{\mt{n}}$).
-% The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{F}(g_{\mt{n}}) \sim \text{Binomial}(n',L_{F}(g_{\mt{n}}))$.
+% The approximate  number of misclassified minds therefore has a binomial distribution:  $n' \hL^{n'}_{F}(g_{n}) \sim \text{Binomial}(n',L_{F}(g_{n}))$.
 

### 2011-07-26 nature_sci_rep/submission2/supervenience.tex
ready for cep redlines

@@
 \begin{align}
-L_{F}(g) = \PP[g(B) \neq M] = \frac{1}{|\mc{M} \times \mc{B}|} \sum_{(m,b) \in \mc{M} \times \mc{B}} \II\{g(b) \neq m\} \PP[B=b, M=m]	
+L_{F}(g) = \PP[g(B) \neq M] =  \sum_{(m,b) \in \mc{M} \times \mc{B}} \II\{g(b) \neq m\} \PP[B=b, M=m]	
 \end{align}
@@
 
-\subsection*{The intractability of testing for statistical supervenience} % (fold)
+\subsection*{Testing for statistical supervenience} % (fold)
 \label{sub:subsection_name}
@@
 
-Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unavailable $L^*$, rendering it useless pragmatically.
-However,  $g^*$ could be estimated from data, and its estimate could be utilized to conduct the following statistical test of supervenience:
-% Let $g_n$ be a classifier induced from some training data, $\mc{T}_n$, 
-\begin{align*} 
-	H_0:& L^* > 0 \\
-	H_A:& L^*=0
-\end{align*}
-A statistical test proceeds by calculating a test statistic and a critical value $c_\alpha$.
-We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$.
-The $p$-value, the probability of rejecting a true null hypothesis, is a function of the distribution of the test statistic.
+The above theorem implies that if we desire to know whether a particular mental property is supervenient to a particular brain property, we must check whether $L^*=0$.
+Unfortunately, $L^*$ is typically unknown.
+Fortunately, we can approximate $L^*$ using training data.
 
@@
 Assume that training data $\mc{T}_n=\{(M_{1},B_{1}), \ldots, (M_{n},B_{n})\}$ are each sampled identically and independently (iid) from the true (but unknown) joint distribution, $F_{MB}$.
-The misclassification rate of a classifier will therefore be a random variable.
-The expected misclassification rate is given by
+The misclassification rate of such a classifier is given by:
 \begin{align}
-\EE[L_F(g_n)]=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m].
+L_F(g_n)=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
 \end{align}
-Because calculating the expected misclassification rate requires a sum over all possible mind/brain property pairs, calculating $\EE[L_F(g_n)]$ will often be intractable in practice.
-Instead, expected misclassification rate is approximated by ``hold-out'' error.
+%  The expected misclassification rate is given by 
+% \begin{align}
+% \EE[L_F(g_n)]=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
+% \end{align}
+% where the expectation is taken with respect 
+Calculating the misclassification rate is often intractable in practice because it requires a sum over all possible mind/brain property pairs.
+Instead, misclassification rate is approximated by ``hold-out'' error.
 Let $\mc{H}_{n'}=\{(M_{n+1},B_{n+1}), \ldots, (M_{n+n'},B_{n+n'})\}$ be a set of $n'$ hold-out samples, each sampled iid from $F_{MB}$.
@@
 \begin{align}
-\hL^{n'}_{F}(g_{n}) = \sum_{(m,b) \in \mc{H}_{n'}}\II \{g_{n}(b; \mc{T}_{n})\neq m] \PP[B=b,M=m].	
+\hL^{n'}_{F}(g_{n}) = \sum_{(m,b) \in \mc{H}_{n'}}\II \{g_{n}(b; \mc{T}_{n})\neq m], % \PP[B=b,M=m].	
 \end{align}
-Under the above assumptions, $\hL^{n'}_F(g_n)$ can be used as a surrogate for $L^*$.
-By definition of $g^*$, we expect $\hL^{n'}_F(g_n) \geq g^*$ for any $g_n$ and all $n$.
-We therefore let $\mh{n}= \lfloor n' \hL^{n'}_{F}(g_n) \rfloor$ be the test statistic, where $\lfloor \cdot \rfloor$ is the floor operator.
-The distribution of $\mh{n}$ is available under the least favorable distribution of the alternate hypothesis.
-Specifically,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', p_{H_A})= \sum_{k \in [\mh{n}]}$Binomial$(k; n'; p_{H_A})$, where $p_{H_A}$ is Bernoulli probability under least favorable distribution of the alternate hypothesis and $[\mh{n}]=\{1,\ldots, \mh{n}\}$.
-In this simple alternate hypothesis, $p_{H_A}=0$.
-Because Binomial$(k; n', 0)=1$ for all $n'$, the $p$-value is always equal to unity.
-Thus, we can \emph{never} reject the above null hypothesis, no matter how much data we obtain.
+% Under the above assumptions, $\hL^{n'}_F(g_n)$ 
+which can be used as a surrogate for $L^*$. 
+By definition of $g^*$, the expectation of $\hL^{n'}_F(g_n)$ is greater than or equal to $g^*$ for any $g_n$ and all $n$.
+Thus, we can construct a hypothesis test for $L^*$ using the surrogate $\hL^{n'}_F(g_n)$.
 
-\subsection*{$\eps$-supervenience admits hypothesis testing} % (fold)
-\label{sub:hypothesis_testing}
+A statistical test proceeds by calculating a test statistic and a critical value $c_\alpha$.
+We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$.
+The $p$-value, the probability of rejecting a true null hypothesis, is a functional of the distribution of the test statistic.
 
-We therefore consider a relaxed notion of supervenience:
+Ideally, we might consider the following hypothesis test: $H_0: L^*>0$ and $H_A: L^*=0$.
+Unfortunately, because the null hypothesis is not closed, the alternate hypothesis lies on the boundary, so the $p$-value is always equal to unity \cite{Bickel2000}.
+To proceed, therefore, we introduce a relaxed notion of supervenience:
 \begin{defi}
@@
 \end{align*}
-The $p$-value of the above composite alternate hypothesis is given by the binomial distribution under the least favorable distribution, $\mathbb{B}(\mh{n}; n', p_{H_A}=\eps)<1$.
+Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the test statistic.
+The distribution of $\mh{n}$ is available under the least favorable distribution of the alternate hypothesis.
+For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', p_{H_A})= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; p_{H_A})$, where $p_{H_A}$ is Bernoulli probability under least favorable distribution of the alternate hypothesis and $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.
+In this composite alternate hypothesis, $p_{H_A}=\eps$.
 We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
@@
 
-\subsection*{A \emph{Gedankenexperiment} demonstrating an ideal test} % (fold)
+% Because Binomial$(k; n', 0)=1$ for all $n'$, the $p$-value is always equal to unity.    Thus, we can \emph{never} reject the above null hypothesis, no matter how much data we obtain.
+
+% Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unavailable $L^*$, rendering it useless pragmatically.
+However,  $g^*$ could be estimated from data, and its estimate could be utilized to conduct the following statistical test of supervenience:
+% % Let $g_n$ be a classifier induced from some training data, $\mc{T}_n$, 
+% \begin{align*} 
+% 	H_0:& L^* > 0 \\
+% 	H_A:& L^*=0
+% \end{align*}
+
+% \subsection*{$\eps$-supervenience admits hypothesis testing} % (fold)
+% \label{sub:hypothesis_testing}
+% 
+% We therefore consider a relaxed notion of supervenience:
+% 
+% 
+% The $p$-value of the above composite alternate hypothesis is given by the binomial distribution under the least favorable distribution, $\mathbb{B}(\mh{n}; n', p_{H_A}=\eps)<1$.
+
+\subsection*{The power and limits of an ideal test of supervenience} % (fold)
 \label{sub:uc}
 
-Ideally, as $n$ increases, the probability of both a Type I error (rejection of a true null hypothesis) and a Type II error (failure to reject a false null hypothesis) goes to zero.
+Ideally, as $n$ increases, the power of a test (the probability of a rejecting the null when it is false)  goes to unity.
+For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (a point mass at zero), then whenever $\alpha >0$, power goes to unity.
+For the above statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $L^{n'}_F(g_n) \conv L^*$ as $n' \conv \infty$.
+Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) \conv \delta_0$.
+
+% probability of both a Type I error (rejection of a true null hypothesis) and a Type II error (failure to reject a false null hypothesis) goes to zero.
 
@@
 
-The $p$-value 
-(probability of a type I error)
-of the test statistic for any classifier $g_n$ converges to zero, because $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n))) \conv 0$ as $n,n' \conv \infty$.
+% The $p$-value 
+% (probability of a type I error)
+% of the test statistic for any classifier $g_n$ converges to zero, because $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n))) \conv 0$ as $n,n' \conv \infty$.
 The power of the test (probability of not making a type II error), however, only converges to unity under certain conditions.
@@
 
-Below, we show that under a very general mind-brain model, one can construct a consistent classifier whose power approaches unity with sufficient data.
+The above considerations indicate that to obtain a statistical test with unity power, a consistent classifier (one for which $g_n \conv g^*$ as $n \conv \infty$) is required.
+Below, we show that under a very general mind-brain model, a consistent classifier is readily available.
 
@@
 \begin{thex}
-Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or neuroanatomical regions) and edges representing synapses (or white matter tracts).
+Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or collections thereof) and edges representing synapses (or collections thereof).
 Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
 Now, imagine collecting very large amounts of very accurate exchangeable  brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Appendix 2 for proof sketch).
+A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Methods for details).
 Therefore, %Theorem \ref{thm1} applies and
@@
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
-Furthermore, Appendix 2 also extends the proof to deal with other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite.
+Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).
 \end{thex}
@@
 We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind-brain property pair is less than $\eps$.
+This definition admits, for the first time to our knowledge, a viable statistical test for supervenience.
 Furthermore, when we restrict the space of minds and brains to the setting of \emph{Gedankenexperiment}  1, we have shown that $k_n$-NN classifiers are universally consistent, such that one can derive a hypothesis test, with confidence level $\alpha$, that is guaranteed to converge to the Bayes optimal misclassification rate, given sufficient data, no matter the true (but unknown) distribution of mind-brain pair properties.
@@
 
-Thus, given measurement of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the exchangeability assumption, a rejection entails that we are $100(1-a)\%$ confident that the mental property under investigation does not supervene on the brain property under investigation.
+Thus, given measurement of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the exchangeability assumption, a rejection of $H_0^\eps$ entails that we are $100(1-a)\%$ confident that the mental property under investigation does not $\eps$-supervene on the brain property under investigation.
 Unfortunately, failure to reject is more ambiguous.
@@
 Testing supervenience of various mental properties on these brain-graphs will likely therefore become increasingly compelling; so the framework developed herein could be fundamental to these investigations.
-Note that these connectomics investigations might lack any dynamics; rather, they effectively assume that structure determines the relevant dynamics.
+% Note that these connectomics investigations might lack any dynamics; rather, they effectively assume that structure determines the relevant dynamics.
 Assuming that dynamics are a function of either vertex or edge attributes, whether any particular mental property is supervenient to ``only'' connectivity structure, or also some attributes, is merely another $\eps$-supervenience hypothesis.
-Appendix 3 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
+For example, questions about whether connectivity structure alone is sufficient to explain a particular mental property is one possible mind/brain $\eps$-supervenience investigation.
+Appendix 2 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
 Similar supervenience tests on larger animals will require either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011} or more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
@@
 
-In conclusion, this \emph{Gedankenexperiment}, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
+% In conclusion, this \emph{Gedankenexperiment}, together with (i) the formal definition of $\varepsilon$-supervenience as a constraint on distributions, (ii) the brain-graph model, and (iii) the universal consistency proof on graphs, is the first demonstration (to our knowledge) that empirically investigating supervenience is at least theoretically possible.
 The above discussion suggests that many previously conducted investigations either assume supervenience, or try to test it.
@@
 \label{sec:methods}
+
+% Consider the following problem setup.
+We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangeably from some unknown joint distribution, $(m_i,b_i)\sim F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
+A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
+It is further assumed that the test mind-brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
+Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$, and that  $\mc{B}$ is countable.
+
+The $1$-nearest neighbor ($1$-NN) classifier works as follows.
+Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
+Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and their corresponding mental properties, , $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order.
+%One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
+The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $\mh{m}=m_{(1)}$.
+The $k_n$ nearest neighbor is a straightforward generalization of this approach.
+It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
+Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
+
+Let $g_n$ be the $k_n$ nearest neighbor ($k_n$NN) classifier when there are $n$ training samples.  
+A collection of such classifiers $\{g_n\}$,  with $k_n$ increasing with $n$, is called a classifier sequence.  
+A universally consistent classifier sequence is any classifier sequence that is guaranteed to converge to the Bayes optimal classifier regardless of the true distribution from which the data were sampled; that is, a universally consistent classifier sequence satisfies $g_n \conv g^*$ as $n \conv \infty$ for all $F_{MB}$.
+
+% In particular, as $n$ increases, $k_n$ must also increase, but not quite as quickly.  Formally, $k_n$ must satisfy
+The $k_n$NN classifier is consistent if (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$ \cite{Stone1977}.
+In Stone's original proof \cite{Stone1977}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
+When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into a vector space, in which case Stone's theorem applies.
+Stone's original proof also applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
+
+Note that the above extension of Stone's original theorem to the graph domain implicitly assumed that vertices were labeled, such that elements of the adjacency matrices could easily be compared across graphs.
+In theory, when vertices are unlabeled, one could first map each graph to a quotient space invariant to isomorphisms, and then proceed as before.
+Unfortunately,
+there is no known polynomial time complexity algorithm for graph isomorphism
+% solving the graph-matching problem is currently NP-Incomplete (meaning it is not known to be either P or NP) 
+\cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging.
+
 % \subsection*{Supervenience} % (fold)

### 2011-07-26 nature_sci_rep/submission2/supp.tex
ready for cep redlines

@@
 \newcommand{\mM}{\mathcal{M}}
-\newcommand{\PP}{\mathbb{P}}           % probability
+\newcommand{\PP}{F}           % probability
 \newcommand{\EE}{\mathbb{E}}           % expected value
@@
 Thus if minds are injective on brains, then $b\neq b' \implies m \neq m'$ (note that the directionality of the implication has been switched relative to supervenience).
-For instance, it might be the case that a brain could change without the mind changing.
-Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$; or that a single additional synapse is formed between a pair of neruons.
+However, it might be the case that a brain could change without the mind changing.
+Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
 It is possible (likely?) that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
@@
 Third, supervenience does not imply identity.
-Consider, for example, acceleration and velocity.
-Clearly, acceleration supervenes on velocity, as acceleration cannot change without velocity changing (assuming one does not consider gravity as acceleration).
-Similarly, velocity supervenes on position, as velocity cannot change without position changing.
-Therefore, acceleration supervenes on position, by the transitive property of supervenience, but it is not the case that a change in acceleration is equal to a change in position.
-In other words, acceleration cannot change without a corresponding change in position, which is the requirement for supervenience.
-More formally, let $a$ and $a'$ indicate accelerations and $p$ and $p'$ indicate positions, we do not have $\PP[a \neq a' | p \neq p'] = 0$; rather, position can change with constant velocity, meaning without acceleration changing.
-Thus, to claim that something supervenes on another is not equivalent to claim that the two are identical.
+The above example with the two coins demonstrates this, as the two coins are not the same thing, even if one has perfect information about one from the other.
+
+What supervenience does imply, however, is the following.
+Imagine finding two different minds.
+If $\MsB$, then the brains subvening under those two minds must be different.
+In other words, there cannot be two different minds, either of which could supervene on a single brain.
+Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
+
+Note that statistical supervenience is distinct from statistical correlation.
+\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
+If $\rho_{MB}=1$, then both $\MsB$ and $\mB \overset{S}{{\sim}}_{F} \mM$.
+Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+In fact, supervenience may be thought of as a generalization of correlation which can be applied to arbitrary valued random variables (such as mental or brain properties), unlike correlation which is only defined for scalar valued random variables.
 
@@
 % section relations_between_sets (end)
-
-What supervenience does imply, however, is the following.
-Imagine finding two different minds.
-If $\MeB$, then the brains subvening under those two minds must be different.
-In other words, there cannot be two different minds, either of which could supervene on a single brain.
-Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
-
-Note that statistical supervenience is distinct from statistical correlation.
-\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
-If $\rho_{MB}=1$, then both $\MeB$ and $\mB \overset{\varepsilon}{{\sim}}_{F} \mM$.
-Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
-In fact, supervenience may be thought of as a generalization of correlation which can be applied to arbitrary valued random variables (such as mental or brain properties), unlike correlation which is only defined for scalar valued random variables.
-
-\section{$k_n$ nearest neighbor algorithm} % (fold)
-\label{sec:knn}
-
-Consider the following problem setup.
-We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangeably from some unknown joint distribution, $(m_i,b_i)\sim F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
-A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind-brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
-Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$, and that  $\mc{B}$ is countable.
-
-The $1$-nearest neighbor ($1$-NN) classifier works as follows.
-Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
-Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and their corresponding mental properties, , $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order.
-%One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
-The $1$-NN algorithm predicts that the unobserved mind is of the same class as the closest brain's class: $\mh{m}=m_{(1)}$.
-The $k_n$ nearest neighbor is a straightforward generalization of this approach.
-It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
-Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
-
-Unfortunately, no such algorithm is universally consistent.
-Let $g_n$ be the $k_n$ nearest neighbor classifier when there are $n$ training samples.
-Then, a collection of such algorithms, $\{g_n\}$,  with $k_n$ increasing with $n$, can be universally consistent under certain constraints.
-In particular, as $n$ increases, $k_n$ must also increase, but not quite as quickly.
-Formally, $k_n$ must satisfy: (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$.
-In Stone's original proof \cite{Stone1977}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
-Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
-When brain-graphs are represented by their adjacency matrices, one can stack the columns of the adjacency matrices, effectively embedding graphs into a vector space, in which case Stone's theorem applies.
-Stone's original proof also applied to the scenario when $|\mc{M}|$ was infinite, resulting in a universally consistent regression algorithm as well.
-
-Note that the above extension of Stone's original theorem to the graph domain implicitly assumed that vertices were labeled, such that elements of the adjacency matrices could easily be compared across graphs.
-In theory, when vertices are unlabeled, one could first map each graph to a quotient space invariant to isomorphisms, and then proceed as before.
-Unfortunately, solving the graph-matching problem is currently NP-Incomplete (meaning it is not known to be either P or NP) \cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging.
-
-% section _k_n_nearest_neighbor_algorithm (end)
 
@@
 \includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.  $\hL^{n'}_{F}(g_{\mt{n}})$  (the misclassification rate estimated upon with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
-is plotted as a function of class-conditional training sample size $n_j=\mt{n}/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{M,B}$.
-Each dot depicts an estimate for $\hL^{n'}_{F}(g_{\mt{n}})$; standard errors are $(\hL^{n'}_{F}(g_{\mt{n}}) (1-\hL^{n'}_{F}(g_{\mt{n}}))/n')^{1/2}$; e.g., $n_j = 180$; $k_n = 53$; $\hL^{n'}_{F}(g_{\mt{n}}) = 0.057$; standard error less than 0.01.
+\caption{C.~elegans graph classification simulation results.  $\hL^{n'}_{F}(g_{n})$  (the misclassification rate estimated  with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+is plotted as a function of class-conditional training sample size $n_j=n/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{M,B}$.
+Each dot depicts an estimate for $\hL^{n'}_{F}(g_n)$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$; e.g., $n_j = 180$; $k_n = 53$; $\hL^{n'}_{F}(g_{n}) = 0.057$; standard error less than 0.01.
 We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.

### 2011-07-27 nature_sci_rep/submission2/supervenience.tex
finished cep redline comments up to disc and methods

@@
 \newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose
+\newcommand{\from}{\colon}
 
@@
 Specifically, we propose a modified version of supervenience (called $\eps$-supervenience) that is amenable to experimental investigation and statistical analysis.
-To illustrate this approach, we perform a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of $\eps$-supervenience.
+To illustrate this approach, we perform a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a \emph{one}-sided determination of $\eps$-supervenience.
 The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or connectome).
-$\eps$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given misclassification rate > 0, regardless of the relationship between the two.
+$\eps$-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given positive misclassification rate, regardless of the relationship between the two.
 This may provide motivation for cross-disciplinary research between neuroscientists and statisticians.
@@
 
-\noindent The mind-brain supervenience notion was canonized in 1970 with the following quote from Donald Davidson: \cite{Davidson70}
+\noindent The mind-brain supervenience notion was canonized in 1970 with the following quote from Donald Davidson \cite{Davidson70}:
 \begin{quotation}
@@
 We consider a special case of supervenience of considerable interest.
-Specifically, this work addresses a novel version of a local supervenience between mental properties and brain properties, with particular emphasize on brain-graphs (i.e., connectivity structure).
+Specifically, this work addresses a version of a local supervenience between mental properties and brain properties, with particular emphasize on brain-graphs (i.e., connectivity structure).
 The determination of supervenience (or lack thereof) of a mental property on a brain property has potentially important implications in a number of fields of inquiry.
-Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption; which if falsified, might change modern approaches to learning \cite{Haykin2008,Ripley2008}.
+Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption, which if falsified, might change modern approaches to learning \cite{Haykin2008,Ripley2008}.
 Cognitive neuroscience similarly seems to operate under such assumptions, which if falsified, might result in novel perspectives and theories \cite{Fodor1998, Gazzaniga2008}.
@@
 This approach depends on defining the space of mental and brain properties under investigation and a statistical model characterizing the possible distributions governing their relationship.
-Such definitions transform supervenience from a conjecture or an assumption, into a hypothesis which can be tested.
+Such definitions transform supervenience from a conjecture or an assumption into a hypothesis which can be tested.
 
@@
 
-\subsection*{Statistical supervenience; a definition} % (fold)
+\subsection*{Statistical supervenience: a definition} % (fold)
 
-\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be a set of possible mental properties.
-For example, $m$ might indicate a person's intelligence, psychological state, current thought, gender identity, etc.
-Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be a set of possible brain properties.
-For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$.
+\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible mental properties, including all possible thoughts, memories beliefs, etc.
+% respects (in whatever sense Davidson meant \cite{Davidson70}). 
+%For example, $m$ might indicate a person's intelligence, psychological state, current thought, gender identity, etc.   
+Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brain properties,
+% physical respects (which we assume to mean all possible brain properties, 
+including the position and momentum of all subatomic particles within the skull.  
+% For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$. 
 Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
@@
 Similarly, let $m$ correspond to an agent's mind, which is a particular element from the set of all possible minds, $\mM$.
-Let $F_{MB}$ indicate a joint distribution of minds and brains. Statistical supervenience can thusly be defined:
+Let $F_{MB}$ indicate a joint distribution of minds and brains. Statistical supervenience can then be defined as follows:
 \begin{defi}
@@
 
-If minds statistically supervene on brains, $\MsB$, then if two minds differ, there must be some brain-based difference to account for the mental difference.
-This means that there exists a deterministic function mapping each brain to its supervening mind, $g: \mB \mapsto \mM$.
-One could therefore, in principle, construct this function.
-When the space of all possible minds is finite; that is, $|\mM| < \infty$, $g$ is called a \emph{classifier}.
+Given that minds statistically supervene on brains, $\MsB$, then if two minds differ, there must be some brain-based difference to account for the mental difference.
+This means that there must exist a deterministic function mapping each brain to its supervening mind, $g^*$.
+One could therefore, in principle, know this function.
+When the space of all possible minds is finite; that is, $|\mM| < \infty$, any function $g\from \mB \to \mM$ mapping from minds to brains is called a \emph{classifier}.
 % Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.  
@@
 The minimum misclassification rate is called Bayes error (see Methods for details).
-The primary result of casting supervenience in a statistical framework is the following theorem: 
+The primary result of casting supervenience in a statistical framework is the following theorem that we state without proof: 
 \begin{thm}
@@
 $\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\mM \overset{S}{\sim}_{F} \mB$, if and only if $L^*= 0$.
-Formally, \mbox{$\MsB \Leftrightarrow L^*=0$}.
+%Formally, \mbox{$\MsB \Leftrightarrow L^*=0$}.
 \end{thm}
-\begin{proof}
-Let $\mc{M}_b=\{m: \PP[B=b | M=m] >0\}$.  
-$$L^*=0 \Leftrightarrow |\mc{M}_b|=1 \, \forall \, b \in \mc{B}.$$
-In words, for $L^*$ to equal 0, it must be the case that there exists only a single $m$ for each $b$.
-\end{proof}
+% \begin{proof}
+% Let $\mc{M}_b=\{m: \PP[B=b | M=m] >0\}$.  
+% $$L^*=0 \Leftrightarrow |\mc{M}_b|=1 \, \forall \, b\in \mc{B} : \PP[B=b]>0. $$
+% In words, for $L^*$ to equal 0, it must be the case that there exists only a single $m$ for each $b$.
+% \end{proof}
 
@@
 Statistical supervenience can therefore be thought of as a constraint on the possible distributions on minds and brains.
-Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s$ be subset of distributions for which supervenience holds.
-Theorem \ref{thm1} implies that $\mc{F}_s = \{F_{MB} : L^*=0\} \subseteq \mc{F}$.
+Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s = \{F_{MB} \in \mc{F}: L^*=0\}$ be subset of distributions for which supervenience holds.
+Theorem \ref{thm1} implies that $\mc{F}_s  \subsetneqq \mc{F}$.
 
@@
 
-The above theorem implies that if we desire to know whether a particular mental property is supervenient to a particular brain property, we must check whether $L^*=0$.
+The above theorem implies that if we desire to know whether a particular mental property is supervenient on a particular brain property, we must check whether $L^*=0$.
 Unfortunately, $L^*$ is typically unknown.
@@
 
-Let $g_n$ be a classifier induced by the data, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$.
 Assume that training data $\mc{T}_n=\{(M_{1},B_{1}), \ldots, (M_{n},B_{n})\}$ are each sampled identically and independently (iid) from the true (but unknown) joint distribution, $F_{MB}$.
-The misclassification rate of such a classifier is given by:
+Let $g_n$ be a classifier induced by the training data, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$.
+The expected misclassification rate of such a classifier is given by:
 \begin{align}
-L_F(g_n)=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
+\EE[L_F(g_n)]=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
 \end{align}
@@
 % \end{align}
-% where the expectation is taken with respect 
-Calculating the misclassification rate is often intractable in practice because it requires a sum over all possible mind/brain property pairs.
-Instead, misclassification rate is approximated by ``hold-out'' error.
+where the expectation is taken with respect to the training data $\mc{T}_n$.
+Calculating the expected misclassification rate is often intractable in practice because it requires a sum over all possible mind-brain property pairs.
+Instead, expected misclassification rate is approximated by ``hold-out'' error.
 Let $\mc{H}_{n'}=\{(M_{n+1},B_{n+1}), \ldots, (M_{n+n'},B_{n+n'})\}$ be a set of $n'$ hold-out samples, each sampled iid from $F_{MB}$.
-The hold-out approximation to misclassification rate is given by
+The hold-out approximation to misclassification rate is given by:
 \begin{align}
-\hL^{n'}_{F}(g_{n}) = \sum_{(m,b) \in \mc{H}_{n'}}\II \{g_{n}(b; \mc{T}_{n})\neq m], % \PP[B=b,M=m].	
+\EE[L_F(g_n)] \approx \hL^{n'}_{F}(g_{n}) = \sum_{(M_i,B_i) \in \mc{H}_{n'}}\II \{g_{n}(B_i; \mc{T}_{n})\neq M_i\}, % \PP[B=b,M=m].	
 \end{align}
@@
 which can be used as a surrogate for $L^*$. 
-By definition of $g^*$, the expectation of $\hL^{n'}_F(g_n)$ is greater than or equal to $g^*$ for any $g_n$ and all $n$.
+By definition of $g^*$, the expectation of $\hL^{n'}_F(g_n)$ (with respect to both $\mc{T}_n$ and $\mc{H}_{n'}$)  is greater than or equal to $g^*$ for any $g_n$ and all $n$.
 Thus, we can construct a hypothesis test for $L^*$ using the surrogate $\hL^{n'}_F(g_n)$.
@@
 A statistical test proceeds by calculating a test statistic and a critical value $c_\alpha$.
-We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$.
-The $p$-value, the probability of rejecting a true null hypothesis, is a functional of the distribution of the test statistic.
+We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$, the allowable Type I error rate.
+The $p$-value, the probability of rejecting the least favorable true null hypothesis (the hypothesis within the potentially composite null closest to the boundary with the alternate hypothesis), is a functional of the distribution of the test statistic.
 
@@
 Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the test statistic.
-The distribution of $\mh{n}$ is available under the least favorable distribution of the alternate hypothesis.
-For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', p_{H_A})= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; p_{H_A})$, where $p_{H_A}$ is Bernoulli probability under least favorable distribution of the alternate hypothesis and $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.
-In this composite alternate hypothesis, $p_{H_A}=\eps$.
+The distribution of $\mh{n}$ is available under the least favorable null distribution to the alternate hypothesis.
+For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', \eps)= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; \eps)$, where $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.
 We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
@@
 
-Ideally, as $n$ increases, the power of a test (the probability of a rejecting the null when it is false)  goes to unity.
-For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (a point mass at zero), then whenever $\alpha >0$, power goes to unity.
-For the above statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $L^{n'}_F(g_n) \conv L^*$ as $n' \conv \infty$.
-Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) \conv \delta_0$.
-
-% probability of both a Type I error (rejection of a true null hypothesis) and a Type II error (failure to reject a false null hypothesis) goes to zero.
-
-%For the above described hypothesis of $\eps$-supervenience, both the $p$-value (probability of a type I error)  and power (probability of not making a type II error), will depend on the distribution  of $g_n$.
-We therefore consider the properties of $g_n$, and the implications of those properties with regard to the probability of a Type I or Type II error.
-
-% The $p$-value 
-% (probability of a type I error)
-% of the test statistic for any classifier $g_n$ converges to zero, because $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n))) \conv 0$ as $n,n' \conv \infty$.
-The power of the test (probability of not making a type II error), however, only converges to unity under certain conditions.
-Specifically, let a  \emph{consistent} classifier be any classifier with expected misclassification rate converging to the Bayes optimal limit; that is, $\EE[L_{F}(g_n)] \conv L_{F}(g^*)$ as $n\conv \infty$.
-The power of the test can be written as: $\PP[\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) < \mathbb{B}(\mh{n}; n', \eps)]$.
-As $n,n' \conv \infty$, $\mathbb{B}(\mh{n}; n', \hL^{n'}_F(g_n)) \conv \mathbb{B}(\mh{n}; n', L^*)$.
-And if $L^* < \eps$, then $\mathbb{B}(\mh{n}; n', L^*) < \mathbb{B}(\mh{n}; n', \eps)$.
-So, the probability of a Type II error also goes to zero as $n,n' \conv \infty$.
-
-% Importantly, the utility of any statistical test depends both on the p-value, the probability of obtaining a test statistic at least as extreme as the observed value (under the assumed model), and its power, the probability that the test will reject a false null hypothesis.
-Ideally, the power of this test would go to unity, as $n,n' \rightarrow \infty$.
-A sufficient condition for power to approach unity is that $g_n$ is a \emph{consistent} classifier.
+Ideally, our test is \emph{consistent}; that is, as $n$ increases, the power of a test (the probability of a rejecting the null when it is indeed false)  goes to unity.
+For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.
+For the above $\eps$-supervenience statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $\hL^{n'}_F(g_n) \conv L^*$ as $n,n' \conv \infty$.
+Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.
 
@@
 
-The above considerations indicate that to obtain a statistical test with unity power, a consistent classifier (one for which $g_n \conv g^*$ as $n \conv \infty$) is required.
+The above considerations indicate that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required.
 Below, we show that under a very general mind-brain model, a consistent classifier is readily available.
@@
 Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
-Now, imagine collecting very large amounts of very accurate exchangeable  brain-graph data and the associated mental property indicators.
-A $k_n$-nearest neighbor classifier using an isomorphism-matching Frobenius norm is universally consistent (see Methods for details).
+Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators.
+A $k_n$-nearest neighbor classifier using a Frobenius norm is universally consistent (see Methods for details).
 Therefore, %Theorem \ref{thm1} applies and
 the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes.
+This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11unlabeled} for details).
 Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).
@@
 Assuming that dynamics are a function of either vertex or edge attributes, whether any particular mental property is supervenient to ``only'' connectivity structure, or also some attributes, is merely another $\eps$-supervenience hypothesis.
-For example, questions about whether connectivity structure alone is sufficient to explain a particular mental property is one possible mind/brain $\eps$-supervenience investigation.
+For example, questions about whether connectivity structure alone is sufficient to explain a particular mental property is one possible mind-brain $\eps$-supervenience investigation.
 Appendix 2 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.

### 2011-07-27 nature_sci_rep/submission2/supervenience.tex
organized up to including simulation in main text

@@
 \noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible mental properties, including all possible thoughts, memories beliefs, etc.
-% respects (in whatever sense Davidson meant \cite{Davidson70}). 
-%For example, $m$ might indicate a person's intelligence, psychological state, current thought, gender identity, etc.   
 Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brain properties,
-% physical respects (which we assume to mean all possible brain properties, 
 including the position and momentum of all subatomic particles within the skull.  
-% For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$. 
 Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
@@
 
-\subsection*{Testing for statistical supervenience} % (fold)
+\subsection*{The existence of a consistent statistical test for supervenience} % (fold)
 \label{sub:subsection_name}
@@
 We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
-The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$.
 
-% Because Binomial$(k; n', 0)=1$ for all $n'$, the $p$-value is always equal to unity.    Thus, we can \emph{never} reject the above null hypothesis, no matter how much data we obtain.
-
-% Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unavailable $L^*$, rendering it useless pragmatically.
-However,  $g^*$ could be estimated from data, and its estimate could be utilized to conduct the following statistical test of supervenience:
-% % Let $g_n$ be a classifier induced from some training data, $\mc{T}_n$, 
-% \begin{align*} 
-% 	H_0:& L^* > 0 \\
-% 	H_A:& L^*=0
-% \end{align*}
-
-% \subsection*{$\eps$-supervenience admits hypothesis testing} % (fold)
-% \label{sub:hypothesis_testing}
-% 
-% We therefore consider a relaxed notion of supervenience:
-% 
-% 
-% The $p$-value of the above composite alternate hypothesis is given by the binomial distribution under the least favorable distribution, $\mathbb{B}(\mh{n}; n', p_{H_A}=\eps)<1$.
-
-\subsection*{The power and limits of an ideal test of supervenience} % (fold)
-\label{sub:uc}
-
-Ideally, our test is \emph{consistent}; that is, as $n$ increases, the power of a test (the probability of a rejecting the null when it is indeed false)  goes to unity.
+A test is \emph{consistent} whenever its power  (the probability of a rejecting the null when it is indeed false)  goes to unity as $n$ increases.
 For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.
@@
 Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.
+
+The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$.
+Moreover, this test is consistent whenever $g_n$ converges to the Bayes classifier, $g^*$.
 
@@
 
-The above considerations indicate that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required.
-Below, we show that under a very general mind-brain model, a consistent classifier is readily available.
+\subsection*{The availability of a consistent statistical test for supervenience} % (fold)
+\label{sub:uc}
+
+The above considerations indicate the existence of a consistent test for supervenience whenever the classifier used is consistent.  
+% that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required. 
+To actually implement such a test, one must be able to (i) measure mind/brain pairs and (ii) have a consistent classifier $g_n$.
+Unfortunately, we do not know how to measure the entirety of one's brain, much less one's mind.
+Moreover, even if we did, we do not know the joint distribution of mind/brain pairs, $F_{MB}$.
+We therefore must restrict our interest to a mental property/brain property pair.
+A mental property might be a person's intelligence, psychological state, current thought, gender identity, etc.
+A brain property might be the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$.
+Regardless of the details of the specifications of the mental property and the brain property, given such specifications, one can assume a model, $\mc{F}$.
+We would like a classifier, $g_n$, that is guaranteed to be consistent, no matter which of the possible distributions $F_{MB} \in \mc{F}$ is the true distribution.
+A classifier with such a property is called an \emph{universally consistent classifier}.
+Below, we show that under a very general mind-brain model $\mc{F}$, an universally consistent classifier is readily available.
 
@@
 Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or collections thereof) and edges representing synapses (or collections thereof).
-Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
-Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators.
+Further let $\mB$, the brain observation space, be the collection of all graphs on a given finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
+Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators from $F_{MB}$.
 A $k_n$-nearest neighbor classifier using a Frobenius norm is universally consistent (see Methods for details).
@@
 \end{thex}
+
+% \subsection*{The practicality of a consistent statistical test for supervenience} % (fold)
 

### 2011-07-28 nature_sci_rep/submission2/supervenience.tex
finished edits, back to cep

@@
 
-% \subsection*{The practicality of a consistent statistical test for supervenience} % (fold)
+\subsection*{The feasibility of a consistent statistical test for supervenience} % (fold)
+
+The above argument demonstrates the availability of a consistent test under certain restrictions.
+The arbitrary slow convergence theorem, however, demonstrates that convergence rates might be unbearably slow.
+We therefore provide a (hopefully illustrative) example of the feasibility of such a test on synthetic data.
+
+{\it Caenorhabditis elegans} is a species whose nervous system is believed to consist of the same $279$ labeled neurons for each organism \cite{Durbin87}.
+Moreover, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
+These findings motivate the use of C.~elegans for a synthetic data analysis \cite{Tukey}.
+Conducting such an experiment requires specifying a joint distribution $F_{MB}$ over brain-graphs and behaviors.
+The joint distribution decomposes into the product of a class-conditional distribution (likelihood) and a prior, $F_{MB}=F_{B|M}F_M$.
+The prior  specifies the probability of any particular organism exhibiting the behavior.
+The class-conditional distribution specifies the brain-graph distribution given that the organism does (or does not) exhibit the behavior.
+
+Let $A_{uv}$ be the number of chemical synapses between neuron $u$ and neuron $v$ according to \cite{VarshneyChklovskii09}.
+Then, let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
+If odor-evoked behavior is supervenient on this subgraph $\mc{E}$, then the distribution of edges in $\mc{E}$ must differ between the two classes.
+Let each edge in each class be Poisson distributed,  $F_{B_{uv}|M=m_j}$= Poisson$(E_{uv|j})$.
+For class $m_0$, let $E_{uv|j}=A_{uv}+\eta$,  where $\eta=0.05$ is a small noise parameter because it is believed that the C.~elegans connectome is similar across organisms \cite{Durbin87}.
+For class $m_1$, let $E_{uv|1}=A_{ij}+z_{ij}$, where the signal parameter $z_{ij}=\eta$ for all edges not in $\mc{E}$, and $z_{ij}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{E}$.
+
+% Let the mental property correspond to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
+Thus, one can specify $F_{M=m_1}$ and $F_{M=m_0}$, the probability of a C.~elegans exhibiting or not a particular behavior.
+The class-conditional distributions $F_{B|M=m_j}$ are therefore the distributions of brain-graphs conditional on whether the C.~elegans exhibits said property.
+
+% Now, to generate the data, let $E_{uv|j}$ be an integer-valued random variable whose value indicates the number of chemical synapses (edges) between neurons (vertices) $u$ and $v$ for class $j$.
+
+We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm.
+The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
+(Better classifiers can be constructed for the joint distribution $F_{M,B}$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
+
+% Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $F_{M,B}$ (Figure \ref{fig1}).
+
+\begin{figure}[!ht]
+\centering 
+\includegraphics[width=.5\linewidth]{Lhatplot}
+\caption{C.~elegans graph classification simulation results.  The estimated expected misclassification rate $\hL^{n'}_{F}(g_{n})$  (with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+is plotted as a function of class-conditional training sample size $n_j=n/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{MB}$.
+Each dot depicts an estimate for $\hL^{n'}_{F}(g_n)$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$.
+For example, at $n_j = 180$ we have $k_n = 53$, $\hL^{n'}_{F}(g_{n}) = 0.057$, and standard error less than $0.01$.
+We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
+Note that $L_{F}(g^*) \approx 0$ for this simulation.}
+\label{fig1}
+\end{figure}
+
+Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
+3D superresolution imaging \cite{Vaziri2008} combined with neurite tracing algorithms \cite{Helmstaedter2008,Mishchenko09,LuLichtman09} allow the collection of a C.
+elegans brain-graph within a day.
+Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{Buckingham2008}.
 
@@
 
-We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind-brain property pair is less than $\eps$.
+This work makes the following contributions.
+First, we define statistical supervenience.
+This definition makes it apparent that supervenience implies the possibility of perfect classification.
+The determination of supervenience from data requires a statistical test.
+We therefore demonstrate the existence, availability, and feasibility of a consistent test of a relaxed notion of supervenience, $\eps$-supervenience, given specified and measurable mental properties and brain properties.
+In other words, the proposed test is guaranteed to reject the null whenever the null is false, given sufficient data, for any possible distribution governing mental property/brain property pairs.
+
+% We have introduced the notion of $\eps$-supervenience, which states that the Bayes optimal misclassification rate for any mind-brain property pair is less than $\eps$.
 This definition admits, for the first time to our knowledge, a viable statistical test for supervenience.
@@
 Second, $\eps$-supervenience is defined between a mental property and a brain property.
-Thus, a failure to reject supervenience on a particular brain property does not entail non-supervenience on any other brain property.
+Thus, a failure to reject $\eps$-supervenience on a particular brain property does not entail non-$\eps$-supervenience on any other brain property.
 Third, in general, neither mental properties nor brain properties are directly measurable; rather, one typically measures a function of such properties.
-For instance, instead of measuring intelligence, one considers the results of an IQ test as a proxy for intelligence.
-Similarly, instead of measuring neural spike trains, one estimates spike trains from some measurable neural signal, like voltage.
+For instance, instead of measuring intelligence, one may consider the results of an IQ test as a proxy for intelligence.
+
+ % Similarly, instead of measuring neural spike trains, one estimates spike trains from some measurable neural signal, like voltage.
 While one could conceivably ask questions such as: ``is IQ score at some time supervenient on the the output of a voltage measuring device?'', these are less elegant and general than their non-observable analogs, such as: ``is intelligence supervenient on spike trains?''
 
-Thus, given measurement of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the exchangeability assumption, a rejection of $H_0^\eps$ entails that we are $100(1-a)\%$ confident that the mental property under investigation does not $\eps$-supervene on the brain property under investigation.
+Thus, given measurements of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the independent and identically sampled assumption, a rejection of $\eps$-supervenience entails that we are $100(1-a)\%$ confident that the mental property under investigation does not $\eps$-supervene on the brain property under investigation.
 Unfortunately, failure to reject is more ambiguous.
@@
 Perhaps this is especially true in light of the advent of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, a field devoted to estimating whole organism brain-graphs, and relating them to function.
-Testing supervenience of various mental properties on these brain-graphs will likely therefore become increasingly compelling; so the framework developed herein could be fundamental to these investigations.
+Testing supervenience of various mental properties on these brain-graphs will perhaps therefore become increasingly compelling; the framework developed herein could be fundamental to these investigations.
 % Note that these connectomics investigations might lack any dynamics; rather, they effectively assume that structure determines the relevant dynamics.
@@
 For example, questions about whether connectivity structure alone is sufficient to explain a particular mental property is one possible mind-brain $\eps$-supervenience investigation.
-Appendix 2 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
-Similar supervenience tests on larger animals will require either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011} or more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
+The above synthetic data analysis demonstrates the feasibility of $\eps$-supervenience on small brain-graphs.
+% Appendix 2 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
+Similar supervenience tests on larger animals (such as humans) will potentially benefit from either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011}, more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
 
@@
 
-The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, and S.~Seung.
+The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, S.~Seung, and two helpful referees.
 

### 2011-07-28 nature_sci_rep/submission2/supp.tex
finished edits, back to cep

@@
 
-\begin{center}
-{\large		
-Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
-	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+% \begin{center}
+% {\large		
+% Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
+% 	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+% \end{center}
+% 
+% \vspace{5px}
+%
 
-\end{center}
+% \section{Relations between sets} % (fold)
+% \label{sec:relations}
 
-\vspace{5px}
+In this appendix we compare supervenience to several other relations on sets.
 
-\section{Relations between sets} % (fold)
-\label{sec:relations}
-
-In this appendix we aim to provide more intuition regarding supervenience by discussing the limitations and extent of its implications.
+% aim to provide more intuition regarding supervenience by discussing the limitations and extent of its implications.
 
@@
 
-\section{Simulation} % (fold)
-\label{sec:sim}
+% \section{Simulation} % (fold)
+% \label{sec:sim}
 
-As an example of a feasible experiment, one may consider a species whose nervous system consists of the same (small) number of labeled neurons for each organism.
-{\it Caenorhabditis elegans} is believed to be such a species \cite{Durbin87}.
-The hermaphroditic C.~elegans' somatic nervous system consists of 279 interconnected neurons.
-Although the graph with these neurons as vertices and edges defined by chemical synapses between neurons is likely not identical across individuals, it appears to be reasonably consistent \cite{Durbin87}.
-Furthermore, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
-Thus, one may design an experiment by describing the joint distribution $F_{M,B}$ via class-conditional distributions $F_{B|M=m_j}$ for the C.~elegans brain-graph for two mental properties of interest, $m_0$ and $m_1$, along with the prior probability of class membership $F_{M=m_j}$.
-Here the mental property corresponds to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
-
-Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $F_{M,B}$ (Figure \ref{fig1}).
-To generate the data, let $E_{ij}$ be an integer-valued random variable whose value indicates the number of synapses (edges) between neurons (vertices) $i$ and $j$.
-Let the class-conditional random variable $E_{ij} | M=m_0$  be distributed Poisson$(A_{ij}+\eta)$, where $A_{ij}$ is the number of chemical synapses between neuron $i$ and neuron $j$ according to \cite{VarshneyChklovskii09}, with noise parameter $\eta=0.05$.
-Let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
-Therefore, the distribution of these edges must differ between the two classes.
-The class-conditional random variable $E_{ij} | M=m_1$ is distributed Poisson$(A_{ij}+ \tr{z_{ij}})$, where the signal parameter $z_{ij}=\eta$ for all edges not in $\mc{E}$, and $z_{ij}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{E}$.
-
-We consider $k_n$-nearest neighbor classification of labeled multigraphs (directed, with loops) on 279 vertices, under Frobenius norm.
-The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $F_{M,B}$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
-
-\begin{figure}[!ht]
-\centering 
-\includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.  $\hL^{n'}_{F}(g_{n})$  (the misclassification rate estimated  with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
-is plotted as a function of class-conditional training sample size $n_j=n/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{M,B}$.
-Each dot depicts an estimate for $\hL^{n'}_{F}(g_n)$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$; e.g., $n_j = 180$; $k_n = 53$; $\hL^{n'}_{F}(g_{n}) = 0.057$; standard error less than 0.01.
-We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
-$L_{F}(g^*) \approx 0$ for this simulation.}
-\label{fig1}
-\end{figure}
-
-Importantly, conducting this experiment {\it in actu} is not beyond current technological limitations.
-3D superresolution imaging \cite{Vaziri2008} combined with neurite tracing algorithms \cite{Helmstaedter2008,Mishchenko09,LuLichtman09} allow the collection of a C.
-elegans brain-graph within a day.
-Genetic manipulations, laser ablations, and training paradigms can each be used to obtain a non-wild type population for use as $M=m_1$ \cite{deBonoMaricq05}, and the class of each organism ($m_0$ vs.~$m_1$) can also be determined automatically \cite{Buckingham2008}.
-
-\bibliography{/Users/jovo/Research/latex/library}
-\bibliographystyle{nature}
+% \bibliography{/Users/jovo/Research/latex/library}
+% \bibliographystyle{nature}
 

### 2011-07-29 nature_sci_rep/submission2/supervenience.tex
ready for cep redline

@@
 the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds.
-This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11unlabeled} for details).
+This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11_unlabeled} for details).
 Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).
@@
 Moreover, these animals exhibit a rich behavioral repertoire that seemingly depends on circuit properties \cite{deBonoMaricq05}.
-These findings motivate the use of C.~elegans for a synthetic data analysis \cite{Tukey}.
+These findings motivate the use of C.~elegans for a synthetic data analysis \cite{GelmanShalizi11}.
 Conducting such an experiment requires specifying a joint distribution $F_{MB}$ over brain-graphs and behaviors.

### 2011-08-05 nature_sci_rep/submission2/supervenience.tex
finished cep redline corrections

@@
 \newtheorem{thex}{\emph{Gedankenexperiment}}
-\lhead{Vogelstein JT, et al}
+\lhead{Vogelstein JT, et al.}
 \rhead{Statistical Supervenience}
@@
 
-\noindent The mind-brain supervenience notion was canonized in 1970 with the following quote from Donald Davidson \cite{Davidson70}:
+\noindent Questions and assumptions about mind-brain supervenience go back at least as far as Plato's dialogues in circa 400 BCE \cite{Plato97}.
+While there are many different notions of supervenience, we find Davidson's canonical description particularly illustrative \cite{Davidson70}:
+%  The mind-brain supervenience notion was canonized in 1970 with the following quote from Donald Davidson \cite{Davidson70}:
 \begin{quotation}
@@
 \end{quotation}
-We consider a special case of supervenience of considerable interest.
+This philosophical conjecture has potentially widespread implications.  
+% We consider a special case of supervenience of considerable interest.
 Specifically, this work addresses a version of a local supervenience between mental properties and brain properties, with particular emphasize on brain-graphs (i.e., connectivity structure).
-The determination of supervenience (or lack thereof) of a mental property on a brain property has potentially important implications in a number of fields of inquiry.
-Neural network theory and artificial intelligence often implicitly take a generalized notion of local mind-brain supervenience as an assumption, which if falsified, might change modern approaches to learning \cite{Haykin2008,Ripley2008}.
-Cognitive neuroscience similarly seems to operate under such assumptions, which if falsified, might result in novel perspectives and theories \cite{Fodor1998, Gazzaniga2008}.
-And the question of mind-brain supervenience continues to be debated amongst philosophers \cite{Kim2007}.
+Supervenience questions or assumptions are fundamental to
+% The determination of supervenience (or lack thereof) of minds on brains has potentially important implications in 
+% a number of fields of inquiry.  
+For example, neural network theory and artificial intelligence often implicitly assume 
+% take a generalized notion of 
+a local version mind-brain supervenience 
+% as an assumption 
+% , which if falsified, might change modern approaches to learning 
+\cite{Haykin2008,Ripley2008}. Cognitive neuroscience similarly seems to operate under such assumptions
+% , which if falsified, might result in novel perspectives and theories 
+\cite{Fodor1998, Gazzaniga2008}.  Philosophers continue to debate and refine notions of supervenience 
+% And the question of mind-brain supervenience continues to be debated amongst philosophers 
+\cite{Kim2007}.  
+% Moreover, many cognitive neuroscience investigations could be cast as questions
+Yet, to date, relatively scant attention has been paid to what might be empirically learned about supervenience.
 
-This work does not attempt to resolve any particular mind-brain supervenience debates.
+In this work we attempt to bridge the gap between philosophical conjecture and empirical investigations by casting supervenience in a probabilistic framework amenable to hypothesis testing. 
+We then use the probabilistic theory of pattern recognition to determine
+ % This perspective leads to a greater understanding vis-\'a-vis 
+the limits of what one can and cannot learn about supervenience through data analysis.
+
+% This work does not attempt to resolve any particular mind-brain supervenience debates.
 Rather, we propose a statistical approach for framing mind-brain supervenience questions.
@@
 
-\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible mental properties, including all possible thoughts, memories beliefs, etc.
-Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brain properties,
-including the position and momentum of all subatomic particles within the skull.  
+\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible minds and
+let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brains.  $\mc{M}$ includes a mind for each possible collection of thoughts, memories, beliefs, etc.
+$\mc{B}$ includes a brain for each possible position and momentum of all subatomic particles within the skull.  
 Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
@@
 \begin{defi}
-\label{def1} 
+\label{def:1} 
 $\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\mM \overset{S}{\sim}_F \mB$, if and only if $\PP[m \neq m' | b=b']=0$, or equivalently $\PP[m = m' | b = b']=1$.
@@
 
-Given that minds statistically supervene on brains, $\MsB$, then if two minds differ, there must be some brain-based difference to account for the mental difference.
-This means that there must exist a deterministic function mapping each brain to its supervening mind, $g^*$.
+% Given that 
+If minds statistically supervene on brains, 
+% , $\MsB$, 
+then if two minds differ, there must be some brain-based difference to account for the mental difference.
+This means that there must exist a deterministic function $g^*$ mapping each brain to its supervening mind.
 One could therefore, in principle, know this function.
-When the space of all possible minds is finite; that is, $|\mM| < \infty$, any function $g\from \mB \to \mM$ mapping from minds to brains is called a \emph{classifier}.
+When the space of all possible minds is finite---that is, $|\mM| < \infty$---any function $g\from \mB \to \mM$ mapping from minds to brains is called a \emph{classifier}.
 % Let $\mh{m}$ denote the output of a classifier, $g(b)=\mh{m}$.  
 Define misclassification rate, the probability that $g$ misclassifies $b$ under distribution $F=F_{MB}$,  as
-\begin{align}
-L_{F}(g) = \PP[g(B) \neq M] =  \sum_{(m,b) \in \mc{M} \times \mc{B}} \II\{g(b) \neq m\} \PP[B=b, M=m]	
+\begin{align} \label{eq:1}
+L_{F}(g) = \PP[g(B) \neq M] =  \sum_{(m,b) \in \mc{M} \times \mc{B}} \II\{g(b) \neq m\} \PP[B=b, M=m],	
 \end{align}
-where $\II$ denotes the indicator function taking value unity whenever its argument is true, and zero otherwise.
-The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers, that is:
+where $\II\{\cdot\}$ denotes the indicator function taking value unity whenever its argument is true and zero otherwise.  The Bayes optimal classifier $g^*$ minimizes $L_{F}(g)$ over all classifiers:	
 $g^* = \argmin_{g} L_{F}(g)$.
@@
 The minimum misclassification rate is called Bayes error (see Methods for details).
-The primary result of casting supervenience in a statistical framework is the following theorem that we state without proof: 
+The primary result of casting supervenience in a statistical framework is the below theorem, which 
+follows immediately from Definition \ref{def:1} and Eq. \eqref{eq:1}: 
 \begin{thm}
-\label{thm1} 
-$\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\mM \overset{S}{\sim}_{F} \mB$, if and only if $L^*= 0$.
+\label{thm:1} 
+% $\mM$ is said to \textit{statistically supervene} on $\mB$ for distribution $F=F_{MB}$, denoted 
+$\mM \overset{S}{\sim}_{F} \mB \Leftrightarrow L^*= 0$.
  %Formally, \mbox{$\MsB \Leftrightarrow L^*=0$}.  
@@
 Statistical supervenience can therefore be thought of as a constraint on the possible distributions on minds and brains.
-Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s = \{F_{MB} \in \mc{F}: L^*=0\}$ be subset of distributions for which supervenience holds.
-Theorem \ref{thm1} implies that $\mc{F}_s  \subsetneqq \mc{F}$.
+Specifically, let $\mc{F}$ indicate the set of all possible joint distributions on minds and brains, and let $\mc{F}_s = \{F_{MB} \in \mc{F}: L^*=0\}$ be the subset of distributions for which supervenience holds.
+Theorem \ref{thm:1} implies that $\mc{F}_s  \subsetneqq \mc{F}$.
+Mind-brain supervenience is therefore an extremely restrictive assumption about the possible relationships between minds and brains.
+It seems that such a restrictive assumption begs for empirical evaluation, vis-\'a-vis, for instance, a hypothesis test.
 
-\subsection*{The existence of a consistent statistical test for supervenience} % (fold)
+\subsection*{The non-existence of a viable statistical test for supervenience} % (fold)
 \label{sub:subsection_name}
@@
 
-The above theorem implies that if we desire to know whether a particular mental property is supervenient on a particular brain property, we must check whether $L^*=0$.
+The above theorem implies that if we desire to know whether  minds supervene on brains, we can check whether $L^*=0$.
 Unfortunately, $L^*$ is typically unknown.
@@
 
-Assume that training data $\mc{T}_n=\{(M_{1},B_{1}), \ldots, (M_{n},B_{n})\}$ are each sampled identically and independently (iid) from the true (but unknown) joint distribution, $F_{MB}$.
+Assume that training data $\mc{T}_n=\{(M_{1},B_{1}), \ldots, (M_{n},B_{n})\}$ are each sampled identically and independently (iid) from the true (but unknown) joint distribution $F=F_{MB}$.
 Let $g_n$ be a classifier induced by the training data, $g_n:\mB \times (\mc{M} \times \mc{B})^n \mapsto \mM$.
-The expected misclassification rate of such a classifier is given by:
+The  misclassification rate of such a classifier is given by
 \begin{align}
-\EE[L_F(g_n)]=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
+L_F(g_n)=\sum_{(m,b)  \in \mc{M}\times \mc{B}} \II\{g_n(b; \mc{T}_n) \neq m\} \PP[B=b,M=m],
 \end{align}
@@
 % \end{align}
-where the expectation is taken with respect to the training data $\mc{T}_n$.
-Calculating the expected misclassification rate is often intractable in practice because it requires a sum over all possible mind-brain property pairs.
-Instead, expected misclassification rate is approximated by ``hold-out'' error.
+which is a random variable due to the dependence on a randomly sampled training set $\mc{T}_n$.  
+% where the expectation is taken with respect to the training data $\mc{T}_n$. 
+Calculating the expected misclassification rate $\EE[L_F(g_n)]$ is often intractable in practice because it requires a sum over all possible training sets.
+Instead, expected misclassification rate can be approximated by ``hold-out'' error.
 Let $\mc{H}_{n'}=\{(M_{n+1},B_{n+1}), \ldots, (M_{n+n'},B_{n+n'})\}$ be a set of $n'$ hold-out samples, each sampled iid from $F_{MB}$.
-The hold-out approximation to misclassification rate is given by:
+The hold-out approximation to the misclassification rate is given by
 \begin{align}
-\EE[L_F(g_n)] \approx \hL^{n'}_{F}(g_{n}) = \sum_{(M_i,B_i) \in \mc{H}_{n'}}\II \{g_{n}(B_i; \mc{T}_{n})\neq M_i\}, % \PP[B=b,M=m].	
+\hL^{n'}_{F}(g_{n}) = \sum_{(M_i,B_i) \in \mc{H}_{n'}}\II \{g_{n}(B_i; \mc{T}_{n})\neq M_i\} \approx \EE[L_F(g_n)] \geq L^*. % \PP[B=b,M=m].	
 \end{align}
 % Under the above assumptions, $\hL^{n'}_F(g_n)$ 
-which can be used as a surrogate for $L^*$. 
-By definition of $g^*$, the expectation of $\hL^{n'}_F(g_n)$ (with respect to both $\mc{T}_n$ and $\mc{H}_{n'}$)  is greater than or equal to $g^*$ for any $g_n$ and all $n$.
+% which can be used as a surrogate for $L^*$. 
+By definition of $g^*$, the expectation of $\hL^{n'}_F(g_n)$ (with respect to both $\mc{T}_n$ and $\mc{H}_{n'}$)  is greater than or equal to $L^*$ for any $g_n$ and all $n$.
 Thus, we can construct a hypothesis test for $L^*$ using the surrogate $\hL^{n'}_F(g_n)$.
 
-A statistical test proceeds by calculating a test statistic and a critical value $c_\alpha$.
-We reject if the test statistic is more extreme than the critical value, or equivalently, if the $p$-value is less than $\alpha$, the allowable Type I error rate.
-The $p$-value, the probability of rejecting the least favorable true null hypothesis (the hypothesis within the potentially composite null closest to the boundary with the alternate hypothesis), is a functional of the distribution of the test statistic.
+A statistical test proceeds by 
+specifying the allowable Type I error rate $\alpha>0$ and then
+calculating a 
+test statistic.
+The $p$-value---the probability of rejecting the least favorable null hypothesis (the simple hypothesis within the potentially composite null which is closest to the boundary with the alternative hypothesis)---is the probability of observing a result at least as extreme as the observed.
+In other words, the $p$-value is the cumulative distribution function of the test statistic evaluated at the observed test statistic with parameter given by the least favorable null distribution.
+% a functional of the distribution of the test statistic.  
+% and a critical value $c_\alpha$.  
+We reject if 
+% the test statistic is more extreme than the critical value, or equivalently, if 
+the $p$-value 
+is less than $\alpha$.
+A test is \emph{consistent} whenever its power  (the probability of rejecting the null when it is indeed false)  goes to unity as $n\conv \infty$ .
+For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.
 
-Ideally, we might consider the following hypothesis test: $H_0: L^*>0$ and $H_A: L^*=0$.
-Unfortunately, because the null hypothesis is not closed, the alternate hypothesis lies on the boundary, so the $p$-value is always equal to unity \cite{Bickel2000}.
+% Ideally, 
+Based on the above considerations,
+we might consider the following hypothesis test: $H_0: L^*>0$ and $H_A: L^*=0$; rejecting the null indicates that $\MsB$. Unfortunately, 
+% because the null hypothesis is not closed, 
+the alternative hypothesis lies on the boundary, so the $p$-value is always equal to unity \cite{Bickel2000}.  From this, Theorem \ref{thm:2} follows immediately:
+\begin{thm} \label{thm:2}
+	There does not exist a viable test of $\MsB$
+\end{thm}
+
+In other words, we can \emph{never} reject $L^*>0$ in favor of supervenience, no matter how much data we obtain.
+Supervenience therefore does not satisfy Popper's falsifiability criterion \cite{Popper1959}.
+
+\subsection*{The existence of a consistent statistical test for $\eps$-supervenience} % (fold)
+
 To proceed, therefore, we introduce a relaxed notion of supervenience: 
+% Let $\MeB$ denote that $\mc{M}$ is $\eps$-supervenient on $\mc{B}$ for distribution $F=F_{MB}$, for $\eps>0$.  We define $\MeB$ as follows:
 \begin{defi}
-\label{def2}
-Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
+\label{def:2}
+$\mM$ is said to $\eps$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\MeB$, if and only if $L^*< \eps$ for some $\eps>0$.
+% $$\MeB \Leftrightarrow L^*< \eps.$$
+% Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$ if and only if $L_F(g^*)< \eps$. 
+%, denoted $\MeB \Leftrightarrow L_{F}(g^*) < \varepsilon$.
 \end{defi}
-\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience:
+\noindent 
+% For brevity, let $\MeB$ denote that $\mc{M}$ is $\eps$-supervenient on $\mc{B}$ 
+Given this relaxation, consider the problem of testing for $\eps$-supervenience:
 \begin{align*}
 	H_0^{\eps}: L^* \geq \eps \\
-	H_A^{\eps}: L^* < \eps
+	H_A^{\eps}: L^* < \eps.
 \end{align*}
-Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the test statistic.
-The distribution of $\mh{n}$ is available under the least favorable null distribution to the alternate hypothesis.
-For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', \eps)= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; \eps)$, where $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.
-We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
+Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the \emph{test statistic}. 
+The distribution of $\mh{n}$ is available under the least favorable null distribution. 
+% to the alternative hypothesis.  
+% The $p$-value is given by the least favorable null distribution to the alternative hypothesis.
+For the above hypothesis test,  
+% for this hypothesis test 
+the $p$-value is therefore the binomial cumulative distribution function with parameter $\eps$; that is, $p$-value $=\mathbb{B}(\mh{n}; n', \eps)= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; \eps)$, where
+% Let 
+% and
+$[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.  We reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.
 
-A test is \emph{consistent} whenever its power  (the probability of a rejecting the null when it is indeed false)  goes to unity as $n$ increases.
-For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.
-For the above $\eps$-supervenience statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $\hL^{n'}_F(g_n) \conv L^*$ as $n,n' \conv \infty$.
-Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.
-
+ For the above $\eps$-supervenience statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $\hL^{n'}_F(g_n) \conv L^*$ as $n,n' \conv \infty$.  Thus, if $L^* < \eps$, 
+power goes to unity.
+% the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.
 The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$.
-Moreover, this test is consistent whenever $g_n$ converges to the Bayes classifier, $g^*$.
+Moreover, this test is consistent whenever $g_n$ converges to the Bayes classifier $g^*$.
 
@@
 Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient for any number of reasons, including that we simply have not yet collected enough data.
-Thus, without restrictions on $F_{MB}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient
+% for any number of reasons, including that 
+because we simply have not yet collected enough data. 
+% Thus, without restrictions on $F_{MB}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
 
-\subsection*{The availability of a consistent statistical test for supervenience} % (fold)
+\subsection*{The availability of a consistent statistical test for $\eps$-supervenience} % (fold)
 \label{sub:uc}
 
-The above considerations indicate the existence of a consistent test for supervenience whenever the classifier used is consistent.  
+The above considerations indicate the existence of a consistent test for $\eps$-supervenience whenever the classifier used is consistent.  
 % that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required. 
@@
 Unfortunately, we do not know how to measure the entirety of one's brain, much less one's mind.
-Moreover, even if we did, we do not know the joint distribution of mind/brain pairs, $F_{MB}$.
-We therefore must restrict our interest to a mental property/brain property pair.
-A mental property might be a person's intelligence, psychological state, current thought, gender identity, etc.
+% Moreover, even if we did, we do not know the joint distribution of mind/brain pairs, $F_{MB}$.  
+We therefore must restrict our interest to a mind/brain \emph{property} pair.  
+A mind (mental) property might be a person's intelligence, psychological state, current thought, gender identity, etc.
 A brain property might be the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$.
@@
 We would like a classifier, $g_n$, that is guaranteed to be consistent, no matter which of the possible distributions $F_{MB} \in \mc{F}$ is the true distribution.
-A classifier with such a property is called an \emph{universally consistent classifier}.
-Below, we show that under a very general mind-brain model $\mc{F}$, an universally consistent classifier is readily available.
+A classifier with such a property is called a \emph{universally consistent classifier}.
+Below, under a very general mind-brain model $\mc{F}$, we construct a universally consistent classifier. 
+% is readily available.
 
@@
 
-\begin{thex}
+\begin{thex} \label{exp:1}
 Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or collections thereof) and edges representing synapses (or collections thereof).
-Further let $\mB$, the brain observation space, be the collection of all graphs on a given finite number of vertices, and $\mc{M}$, the mental property observation space, be finite.
-Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators from $F_{MB}$.
+Further let $\mB$, the brain observation space, be the collection of all graphs on a given finite number of vertices, and let $\mc{M}$, the mental property observation space, be finite.
+Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and associated mental property indicators from $F_{MB}$.
 A $k_n$-nearest neighbor classifier using a Frobenius norm is universally consistent (see Methods for details).
-Therefore, %Theorem \ref{thm1} applies and
-the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds.
+% Therefore, %Theorem \ref{thm:1} applies and 
+The existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds.
 This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11_unlabeled} for details).
-Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).
+Furthermore, the proof holds for other matrix norms (which might speed up convergence and hence reduce the required $n$), and the regression scenario where $|\mM|$ is infinite (again, see Methods for details).
 \end{thex}
+Thus, under the conditions stated in the above \emph{Gedankenexperiment}, we have:
+\begin{thm} \label{thm:3}
+	$\MeB \implies \beta \conv 1$ as $n,n'\conv \infty$, but there is no universal $n,n'$ that guarantes $\beta=1$.
+\end{thm}
 
-\subsection*{The feasibility of a consistent statistical test for supervenience} % (fold)
+\subsection*{The feasibility of a consistent statistical test for $\eps$-supervenience} % (fold)
 
-The above argument demonstrates the availability of a consistent test under certain restrictions.
+Theorem \ref{thm:3} demonstrates the availability of a consistent test under certain restrictions.
 The arbitrary slow convergence theorem, however, demonstrates that convergence rates might be unbearably slow.
-We therefore provide a (hopefully illustrative) example of the feasibility of such a test on synthetic data.
+We therefore provide an illustrative example of the feasibility of such a test on synthetic data.
 
@@
 Let $A_{uv}$ be the number of chemical synapses between neuron $u$ and neuron $v$ according to \cite{VarshneyChklovskii09}.
-Then, let $\mc{E}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
-If odor-evoked behavior is supervenient on this subgraph $\mc{E}$, then the distribution of edges in $\mc{E}$ must differ between the two classes.
-Let each edge in each class be Poisson distributed,  $F_{B_{uv}|M=m_j}$= Poisson$(E_{uv|j})$.
-For class $m_0$, let $E_{uv|j}=A_{uv}+\eta$,  where $\eta=0.05$ is a small noise parameter because it is believed that the C.~elegans connectome is similar across organisms \cite{Durbin87}.
-For class $m_1$, let $E_{uv|1}=A_{ij}+z_{ij}$, where the signal parameter $z_{ij}=\eta$ for all edges not in $\mc{E}$, and $z_{ij}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{E}$.
-
+Then, let $\mc{S}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
+If odor-evoked behavior is supervenient on this signal subgraph $\mc{S}$, then the distribution of edges in $\mc{S}$ must differ between the two classes of odor evoked behavior \cite{VP11_sigsub}.
+Let $E_{uv|j}$ denote the expected number of edges from vertex $v$ to vertex $u$ in class $j$.
+For class $m_0$, let $E_{uv|1}=A_{uv}+\eta$,  where $\eta=0.05$ is a small noise parameter  (it is believed that the C.~elegans connectome is similar across organisms \cite{Durbin87}).
+For class $m_1$, let $E_{uv|1}=A_{uv}+z_{uv}$, where the signal parameter $z_{uv}=\eta$ for all edges not in $\mc{S}$, and $z_{uv}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{S}$.
+For both classes, let each edge be Poisson distributed, $F_{A_{uv}|M=m_j}$= Poisson$(E_{uv|j})$.
 % Let the mental property correspond to the C.~elegans exhibiting (or not exhibiting) a particular behavior (e.g., response to an odor).
@@
 The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $F_{M,B}$ used here; however, we demand universal consistency.)  Figure \ref{fig1} shows that for this simulation, rejecting $\eps=0.1$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
+(Better classifiers can be constructed for the joint distribution $F_{MB}$ used here; however, we demand universal consistency.)  Figure \ref{fig:1} shows that for this simulation, rejecting $(\eps=0.1)$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
 
-% Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $F_{M,B}$ (Figure \ref{fig1}).
+% Simulations suggest that one may build a classifier, practically and with a manageable training sample size $n$, that demonstrates $\varepsilon$-supervenience with reasonable choices for $\varepsilon$ and $\alpha$, and a plausible joint distribution $F_{MB}$ (Figure \ref{fig:1}).
 
@@
 \includegraphics[width=.5\linewidth]{Lhatplot}
-\caption{C.~elegans graph classification simulation results.  The estimated expected misclassification rate $\hL^{n'}_{F}(g_{n})$  (with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+\caption{C.~elegans graph classification simulation results.  The estimated hold-out misclassification rate $\hL^{n'}_{F}(g_{n})$  (with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
 is plotted as a function of class-conditional training sample size $n_j=n/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{MB}$.
-Each dot depicts an estimate for $\hL^{n'}_{F}(g_n)$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$.
-For example, at $n_j = 180$ we have $k_n = 53$, $\hL^{n'}_{F}(g_{n}) = 0.057$, and standard error less than $0.01$.
-We reject $H_0: L_{F}(g^*) \geq 0.10$ at $\alpha=0.01$.
-Note that $L_{F}(g^*) \approx 0$ for this simulation.}
-\label{fig1}
+Each dot depicts $\hL^{n'}_{F}(g_n)$ for some $n$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$.
+For example, at $n_j = 150$ we have $k_n = \lfloor\sqrt{n}\rfloor=11$ (where $\lfloor\cdot\rfloor$ indicates the floor operator), $\hL^{n'}_{F}(g_{n}) = 0.069$, and standard error less than $0.01$.
+We reject $H_0^{0.1}: L^* \geq 0.1$ at $\alpha=0.01$.
+Note that $L^* \approx 0$ for this simulation.}
+\label{fig:1}
 \end{figure}
@@
 This work makes the following contributions.
-First, we define statistical supervenience.
-This definition makes it apparent that supervenience implies the possibility of perfect classification.
-The determination of supervenience from data requires a statistical test.
-We therefore demonstrate the existence, availability, and feasibility of a consistent test of a relaxed notion of supervenience, $\eps$-supervenience, given specified and measurable mental properties and brain properties.
+First, we define statistical supervenience based on Davidson's canonical statement (Definition \ref{def:1}).
+This definition makes it apparent that supervenience implies the possibility of perfect classification (Theorem \ref{thm:1}).
+We then prove that there is no viable test against supervenience, so one can \emph{never} reject a null hypothesis in favor of supervenience, regardless of the amount of data (Theorem \ref{thm:2}).
+This motivates the introduction of a relaxed notion called $\eps$-supervenience (Definition \ref{def:2}), against which consistent statistical tests are readily available.
+Under a very general brain-graph/mental property model (\emph{Gedankenexperiment} \ref{exp:1}),  a consistent statistical test against $\eps$-supervenience is always available no matter the true distribution $F_{MB}$ (Theorem \ref{thm:3}).
 In other words, the proposed test is guaranteed to reject the null whenever the null is false, given sufficient data, for any possible distribution governing mental property/brain property pairs.
+
+Alas, arbitrary slow convergence theorems demonstrate that there is no universal $n,n'$ for which convergence is guaranteed.
+Thus, a failure to reject is ambiguous: even if the data satisfy the above assumptions, the failure to reject may be due to either (i) an insufficient amount of data or (ii) $\mc{M}$ may not be $\eps$-supervenient on $\mc{B}$.
+Moreover, the data will not, in general, satisfy the above assumptions.
+In addition to dependence (because each human does not exist in a vacuum), the mental property measurements will often be ``noisy'' (for example, accurately diagnosing psychiatric disorders is a sticky wicket \cite{Kessler2005}).
+% has a number of possible explanations.
+First, the amount of data might be insufficient; collecting more data or utilizing a more informative distance might resolve this difficulty.
+Second, the class labels (mental property indicators) may not be faithful representations of the mental property of interest (mental property observations may be ``noisy'' indicators).
+% 
+% Second, it may be that the mental property under investigation is supervenient to a different brain property.  
+% Third,  Fourth, the mental property may really not be supervenient to any brain properties (although testing this more global notion of supervenience is not possible, as one can never be sure that one as measured all brain properties).
+Nonetheless, synthetic data analysis suggests that under somewhat realistic assumptions, convergence obtains with an amount of data one might conceivably collect (Figure \ref{fig:1} and ensuing discussion).
 
@@
 
-Alas, this is a one-sided test, so although power converges to unity, %one can never determine the reason for a failure to reject.
+ %one can never determine the reason for a failure to reject.
 %one can never determine whether (i) more data is necessary to get a lower p-value, or (ii) that the particular $\eps$-supervenience does not hold.
-a failure to reject the null has many possible explanations.
-First, the amount of data might be insufficient given the particular distance implemented; collecting more data or utilizing a more informative distance might resolve this difficulty.
-Second, $\eps$-supervenience is defined between a mental property and a brain property.
+
+% Second, $\eps$-supervenience is defined between a mental property and a brain property.
 Thus, a failure to reject $\eps$-supervenience on a particular brain property does not entail non-$\eps$-supervenience on any other brain property.
@@
 
-Thus, given measurements of mental and brain properties that we believe reflect the properties of interest, and a sufficient amount data satisfying the independent and identically sampled assumption, a rejection of $\eps$-supervenience entails that we are $100(1-a)\%$ confident that the mental property under investigation does not $\eps$-supervene on the brain property under investigation.
+Thus, given measurements of mental and brain properties that we believe reflect the properties of interest, and given a sufficient amount of data satisfying the independent and identically sampled assumption, a rejection of $H_0^{\eps}: L^*\geq \eps$ in favor of $\MeB$ entails that we are $100(1-a)\%$ confident that the mental property under investigation is $\eps$-supervenient on the brain property under investigation.
 Unfortunately, failure to reject is more ambiguous.
-$\eps-$supervenience tests can therefore be thought of as constraining the space of possible brain properties upon which a mental property supervenes.
-Determining that a mental property does not supervene on \emph{any} brain property is beyond the capacities of this formalism.
+% $\eps-$supervenience tests can therefore be thought of as constraining the space of possible brain properties upon which a mental property supervenes.  
+% Determining that a mental property does not supervene on \emph{any} brain property is beyond the capacities of this formalism.
 
 Interestingly, much of contemporary research in neuroscience and cognitive science could be cast as mind-brain supervenience investigations.
-Specifically, any investigation that aims to ``explain'' some behavioral or psychological phenomenon by reference to neural (or other brain) activity or structure, could be thought of as mind-brain supervenience investigations.
-Thus, the proposed $\eps$-supervenience framework is perhaps a useful unifying perspective for these fields.
-Perhaps this is especially true in light of the advent of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, a field devoted to estimating whole organism brain-graphs, and relating them to function.
+Specifically, searches for ``engrams'' of  memory traces \cite{Lashley50} or ``neural correlates''  of various behaviors or mental properties (for example, consciousness \cite{Koch2010}), may be more aptly called searches for the ``neural supervenia'' of such properties.
+% any investigation that aims to ``explain'' some behavioral or psychological phenomenon by reference to neural (or other brain) activity or structure could be thought of as mind-brain supervenience investigations.
+% Thus, the proposed $\eps$-supervenience framework is perhaps a useful unifying perspective.
+ % for these fields.  
+% The above Gedankenexperiment is an 
+Letting the brain property be a brain-graph is perhaps especially pertinent in light of the advent of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, a field devoted to estimating whole organism brain-graphs and relating them to function.
 Testing supervenience of various mental properties on these brain-graphs will perhaps therefore become increasingly compelling; the framework developed herein could be fundamental to these investigations.
@@
 % Consider the following problem setup.
-We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangeably from some unknown joint distribution, $(m_i,b_i)\sim F_{M,B}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
+We have a collection of training data, $\mc{T}_n =\{(m_i,b_i)\}_{i=1}^n$, each sampled exchangeably from some unknown joint distribution, $(m_i,b_i)\sim F_{MB}$, where $m_i$ and $b_i$ are the observed mental and brain properties of experiment $i$, respectively.
 A new brain, $b$, called the ``test brain'', is then observed, and one desires to find the most likely class of the new brain, $m$.
-It is further assumed that the test mind-brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{M,B}$, and $m$ is unobserved.
+It is further assumed that the test mind-brain pair is sampled from the same distribution as the training data, $(m,b)\sim F_{MB}$, and $m$ is unobserved.
 Further assume that $m$ can take one of a finite number of possible values, that is, $|\mc{M}|<\infty$, and that  $\mc{B}$ is countable.
@@
 The $1$-nearest neighbor ($1$-NN) classifier works as follows.
-Compute the distance between the test brain and all the training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
-Then, sort them, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and their corresponding mental properties, , $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order.
+Compute the distance between the test brain  $b$ and all $n$ training brains, $d_i=d(b,b_i)$ for all $i \in [n]$, where $[n]=1,2,\ldots, n$.
+Then, sort these distances, $d_{(1)} < d_{(2)} < \ldots < d_{(n)}$, and consider their corresponding minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where parenthetical indices indicate rank order among $\{d_i\}_{i\in[n]}$.
 %One can then also obtain a rank order for the training minds, $m_{(1)}, m_{(2)}, \ldots, m_{(n)}$, where $m_{(i)}$ is the class of the $i^{th}$ closest training brain to $b$.
@@
 The $k_n$ nearest neighbor is a straightforward generalization of this approach.
-It says that the test mind is in the same class as which ever class is the plurality class of the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
-Given a particular choice of $k_n$ (the number of nearest neighbors to consider), and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one then has a relatively simple and intuitive algorithm.
+It says that the test mind is in the same class as whichever class is the plurality class among the $k_n$ nearest neighbors, $\mh{m}=\argmax_{m'}\II\{\sum_{i=1}^{k_n} m_{(i)}=m'\}$.
+Given a particular choice of $k_n$ (the number of nearest neighbors to consider) and a choice of $d(\cdot,\cdot)$ (the distance metric used to compare the test datum and training data), one has a relatively simple and intuitive algorithm.
 
@@
 A collection of such classifiers $\{g_n\}$,  with $k_n$ increasing with $n$, is called a classifier sequence.  
-A universally consistent classifier sequence is any classifier sequence that is guaranteed to converge to the Bayes optimal classifier regardless of the true distribution from which the data were sampled; that is, a universally consistent classifier sequence satisfies $g_n \conv g^*$ as $n \conv \infty$ for all $F_{MB}$.
+A universally consistent classifier sequence is any classifier sequence that is guaranteed to converge to the Bayes optimal classifier regardless of the true distribution from which the data were sampled; that is, a universally consistent classifier sequence satisfies $L_F(g_n) \conv L_F(g^*)$ as $n \conv \infty$ for all $F_{MB}$.
+In the main text, we refer to the whole sequence as a classifier.
 
@@
 The $k_n$NN classifier is consistent if (i) $k_n \conv \infty$ as $n \conv \infty$ and (ii) $k_n/n \conv 0$ as $n\conv\infty$ \cite{Stone1977}.
-In Stone's original proof \cite{Stone1977}, $b$ was assumed to be a $d$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^d (b_j-b_j')^2$, where $j$ indexes elements of the $d$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
+In Stone's original proof \cite{Stone1977}, $b$ was assumed to be a $q$-dimensional vector, and the $L_2$ norm ($d(b,b')=\sum_{j=1}^q (b_j-b_j')^2$, where $j$ indexes elements of the $q$-dimensional vector) was shown to satisfy the constraints on a distance metric for this collection of classifiers to be universally consistent.
 Later, others extended these results to apply to any $L_p$ norm \cite{DGL96}.
@@
 % solving the graph-matching problem is currently NP-Incomplete (meaning it is not known to be either P or NP) 
-\cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging.
+\cite{GareyJohnson79}, so in practice, dealing with unlabeled vertices will likely be computationally challenging \cite{VP11_unlabeled}.
 

### 2011-08-05 nature_sci_rep/submission2/supp.tex
finished cep redline corrections

@@
 
-In this appendix we compare supervenience to several other relations on sets.
+In this appendix we compare supervenience to several other relations on sets (see Figure \ref{fig:rel}).
 
@@
 
-Third, supervenience does not imply identity.
-The above example with the two coins demonstrates this, as the two coins are not the same thing, even if one has perfect information about one from the other.
+Third, supervenience does not imply identity.  The above example with the two coins demonstrates this, as the two coins are not the same thing, even if one has perfect information about the other.
 
 What supervenience does imply, however, is the following.
-Imagine finding two different minds.
-If $\MsB$, then the brains subvening under those two minds must be different.
-In other words, there cannot be two different minds, either of which could supervene on a single brain.
+Imagine finding two unequal minds.
+If $\MsB$, then the brains on which those two minds supervene must be different.
+In other words, there cannot be two unequal minds, either of which could supervene on a single brain.
 Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
@@
 Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
-In fact, supervenience may be thought of as a generalization of correlation which can be applied to arbitrary valued random variables (such as mental or brain properties), unlike correlation which is only defined for scalar valued random variables.
+In fact, supervenience may be thought of as a generalization of correlation which incorporates directionality, can be applied to arbitrary valued random variables (such as mental or brain properties), can depend on any moment of a distribution (not just the first two).
 
@@
 (A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
-(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds.
+(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds, but not a bijective one.
 (C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}

### 2011-08-05 nature_sci_rep/submission2/supervenience.tex
updated refs
finished redline cep comments of aug 5 2011

@@
 % , which if falsified, might result in novel perspectives and theories 
-\cite{Fodor1998, Gazzaniga2008}.  Philosophers continue to debate and refine notions of supervenience 
+\cite{Gazzaniga2008}.  Philosophers continue to debate and refine notions of supervenience 
 % And the question of mind-brain supervenience continues to be debated amongst philosophers 
@@
 \begin{thm} \label{thm:2}
-	There does not exist a viable test of $\MsB$
+	There does not exist a viable test of $\MsB$.
 \end{thm}
@@
 
-\subsection*{The existence of a consistent statistical test for $\eps$-supervenience} % (fold)
+\subsection*{Conditions for a consistent statistical test for $\eps$-supervenience} % (fold)
 
@@
 
-Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{MB}$ \cite{DGL96}.
-Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
-For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient
-% for any number of reasons, including that 
-because we simply have not yet collected enough data. 
-% Thus, without restrictions on $F_{MB}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
-
-\subsection*{The availability of a consistent statistical test for $\eps$-supervenience} % (fold)
+\subsection*{The existence and construction of a consistent statistical test for $\eps$-supervenience} % (fold)
 \label{sub:uc}
@@
 Regardless of the details of the specifications of the mental property and the brain property, given such specifications, one can assume a model, $\mc{F}$.
-We would like a classifier, $g_n$, that is guaranteed to be consistent, no matter which of the possible distributions $F_{MB} \in \mc{F}$ is the true distribution.
+We desire a classifier $g_n$ that is guaranteed to be consistent, no matter which of the possible distributions $F_{MB} \in \mc{F}$ is the true distribution.
 A classifier with such a property is called a \emph{universally consistent classifier}.
@@
 \end{thex}
-Thus, under the conditions stated in the above \emph{Gedankenexperiment}, we have:
+Thus, under the conditions stated in the above \emph{Gedankenexperiment}, universal consistency yields:
 \begin{thm} \label{thm:3}
-	$\MeB \implies \beta \conv 1$ as $n,n'\conv \infty$, but there is no universal $n,n'$ that guarantes $\beta=1$.
+	$\MeB \implies \beta \conv 1$ as $n,n'\conv \infty$.
+\end{thm}
+
+Unfortunately, the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ depends on the (unknown) distribution $F=F_{MB}$ \cite{DGL96}.
+Furthermore, arbitrarily slow convergence theorems regarding the rate of convergence of $L_{F}(g_n)$ to $L_{F}(g^*)$ demonstrate that there is no universal $n,n'$ which will guarantee that the test has power greater than any specified target $\beta > \alpha$ \cite{Devroye83}.
+For this reason, the test outlined above can provide only a one-sided conclusion: if we reject we can be $100(1-\alpha)$\% confident that $\MeB$ holds, but we can never be confident in its negation; rather, it may be the case that the evidence in favor of $\MeB$ is insufficient
+% for any number of reasons, including that 
+because we simply have not yet collected enough data. 
+% Thus, without restrictions on $F_{MB}$, arbitrarily slow convergence theorems imply that our theorem of $\varepsilon$-supervenience does not \tr{strictly} satisfy Popper's {\it falsifiability} requirement \cite{Popper}.
+This leads immediately to the following theorem:
+\begin{thm} \label{thm:4}
+There is no universal $n,n'$ that guarantees $\beta \geq \beta_{min}>\alpha$.
 \end{thm}
@@
 Theorem \ref{thm:3} demonstrates the availability of a consistent test under certain restrictions.
-The arbitrary slow convergence theorem, however, demonstrates that convergence rates might be unbearably slow.
+Theorem \ref{thm:4}, however, demonstrates that convergence rates might be unbearably slow.
 We therefore provide an illustrative example of the feasibility of such a test on synthetic data.
@@
 Let $E_{uv|j}$ denote the expected number of edges from vertex $v$ to vertex $u$ in class $j$.
-For class $m_0$, let $E_{uv|1}=A_{uv}+\eta$,  where $\eta=0.05$ is a small noise parameter  (it is believed that the C.~elegans connectome is similar across organisms \cite{Durbin87}).
+For class $m_0$, let $E_{uv|0}=A_{uv}+\eta$,  where $\eta=0.05$ is a small noise parameter  (it is believed that the C.~elegans connectome is similar across organisms \cite{Durbin87}).
 For class $m_1$, let $E_{uv|1}=A_{uv}+z_{uv}$, where the signal parameter $z_{uv}=\eta$ for all edges not in $\mc{S}$, and $z_{uv}$ is uniformly sampled from $[-5,5]$ for all edges within $\mc{S}$.
@@
 The $k_n$-nearest neighbor classifier used here satisfies $k_n \rightarrow \infty$ as $n \rightarrow \infty$ and $k_n/n \rightarrow 0$ as $n \rightarrow \infty$, ensuring universal consistency.
-(Better classifiers can be constructed for the joint distribution $F_{MB}$ used here; however, we demand universal consistency.)  Figure \ref{fig:1} shows that for this simulation, rejecting $(\eps=0.1)$-supervenience at $\alpha=0.01$ only requires a few hundred training samples.
+(Better classifiers can be constructed for the joint distribution $F_{MB}$ used here; however, we demand universal consistency.)  Figure \ref{fig:1} shows that for this simulation, rejecting $(\eps=0.1)$-supervenience at $\alpha=0.01$ requires only a few hundred training samples.
 
@@
 Each dot depicts $\hL^{n'}_{F}(g_n)$ for some $n$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$.
-For example, at $n_j = 150$ we have $k_n = \lfloor\sqrt{n}\rfloor=11$ (where $\lfloor\cdot\rfloor$ indicates the floor operator), $\hL^{n'}_{F}(g_{n}) = 0.069$, and standard error less than $0.01$.
+For example, at $n_j = 180$ we have $k_n = \lfloor\sqrt{8 n}\rfloor=53$ (where $\lfloor\cdot\rfloor$ indicates the floor operator), $\hL^{n'}_{F}(g_{n}) = 0.057$, and standard error less than $0.01$.
 We reject $H_0^{0.1}: L^* \geq 0.1$ at $\alpha=0.01$.
@@
 
-Alas, arbitrary slow convergence theorems demonstrate that there is no universal $n,n'$ for which convergence is guaranteed.
+Alas, arbitrary slow convergence theorems demonstrate that there is no universal $n,n'$ for which convergence is guaranteed (Theorem \ref{thm:4}).
 Thus, a failure to reject is ambiguous: even if the data satisfy the above assumptions, the failure to reject may be due to either (i) an insufficient amount of data or (ii) $\mc{M}$ may not be $\eps$-supervenient on $\mc{B}$.

### 2011-08-05 nature_sci_rep/submission2/supervenience.tex
finished redline cep comments of aug 5 2011
defined beta_min

@@
 In other words, we can \emph{never} reject $L^*>0$ in favor of supervenience, no matter how much data we obtain.
-Supervenience therefore does not satisfy Popper's falsifiability criterion \cite{Popper1959}.
 
@@
 \begin{thm} \label{thm:4}
-There is no universal $n,n'$ that guarantees $\beta \geq \beta_{min}>\alpha$.
+For any target power $\beta_{min} > \alpha$, there is no universal $n,n'$ that guarantees $\beta \geq \beta_{min}$.
 \end{thm}
+
+% Moreover, switching the null and alternative hypotheses---yielding $H_0^{\eps'}: L^*<\eps$ and $H_A^{\eps'}: L^* \geq \eps$---does not resolve the matter.
+Therefore, even $\eps$-supervenience does not satisfy Popper's falsifiability criterion \cite{Popper1959}.
 

### 2011-08-10 arxiv/v1/supervenience.tex
added implications paragraph, perhaps to keep

--- 2010-01-05 manuscript/PNAS/PNAS_submission.tex
@@
 \newtheorem{thex}{Thought experiment}
+\lhead{Vogelstein JT, et al}
+\rhead{Neurocognitive Graph Theory}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2008}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+\title{Are mental properties supervenient on brain properties?}
+
+\author{Joshua T. Vogelstein, R. Jacob Vogelstein, Carey E. Priebe\\{Johns Hopkins University}}
 
@@
 
-\begin{acknowledgments}
+\paragraph{Acknowledgments}
 The authors would like to acknowledge helpful discussions with J Lande and B Vogelstein.
-\end{acknowledgments}
+% \end{acknowledgments}
 
-%% PNAS does not support submission of supporting .tex files such as BibTeX.
-%% Instead all references must be included in the article .tex document. 
-%% If you currently use BibTeX, your bibliography is formed because the 
-%% command \verb+\bibliography{}+ brings the <filename>.bbl file into your
-%% .tex document. To conform to PNAS requirements, copy the reference listings
-%% from your .bbl file and add them to the article .tex file, using the
-%% bibliography environment described above.
-
-%%  Contact pnas@nas.edu if you need assistance with your
-%%  bibliography.
-
-% Sample bibliography item in PNAS format:
-%% \bibitem{in-text reference} comma-separated author names up to 5,
-%% for more than 5 authors use first author last name et al. (year published)
-%% article title  {\it Journal Name} volume #: start page-end page.
-%% ie,
-% \bibitem{Neuhaus} Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) 
-% A short C-terminal sequence is necessary and sufficient for the
-% targeting of chitinases to the plant vacuole. 
-% {\it Proc Natl Acad Sci USA} 88:10362-10366.
-
-%% Enter the largest bibliography number in the facing curly brackets
-%% following \begin{thebibliography}
-
-\begin{thebibliography}{26}
-
-\bibitem{Plato97}
-Plato.
-\newblock {\em Plato: complete works}.
-\newblock Hackett Pub Co,  (1997).
-
-\bibitem{Descartes1641}
-Descartes, R.
-\newblock {\em Meditationes de prima philosophia}.
-\newblock  (1641).
-
-\bibitem{Davidson70}
-Davidson, D.
-\newblock {\em Experience and Theory}, chapter Mental Events.
-\newblock Duckworth (1970).
-
-\bibitem{SpornsKotter05}
-Sporns, O., Tononi, G., and Kotter, R.
-\newblock {\em PLoS Computational Biology}{ \bf 1}(4), e42 (2005).
-
-\bibitem{LichtmanSanes08}
-Lichtman, J.~W., Livet, J., and Sanes, J.~R.
-\newblock {\em Nat Rev Neurosci}{ \bf 9}(6), 417--422 Jun  (2008).
-
-\bibitem{Seung09}
-Seung, H.
-\newblock {\em Neuron}{ \bf 62}(1), 17--29 (2009).
-
-\bibitem{Kim95}
-Kim, J.
-\newblock {\em Phiilosophy of Mind}.
-\newblock Westview Press, second edition,  (2005).
-
-\bibitem{DGL96}
-Devroye, L., Gy{\"o}rfi, L., and Lugosi, G.
-\newblock {\em {A Probabilistic Theory of Pattern Recognition}}.
-\newblock Springer,  (1996).
-
-\bibitem{Devroye83}
-Devroye, L.
-\newblock {\em Probability Theory and Related Fields}{ \bf 62}(4), 475--483
-  (1983).
-
-\bibitem{Popper}
-Popper, K.
-\newblock { \bf } (1959).
-
-\bibitem{WhiteBrenner86}
-White, J., Southgate, E., Thomson, J.~N., and Brenner, S.
-\newblock {\em Philosophical Transactions of Royal Society London. Series B,
-  Biological Sciences}{ \bf 314}(1165), 1--340 (1986).
-
-\bibitem{DenkHorstmann04}
-W.~Denk, W. and Horstmann, H.
-\newblock {\em PLOS Biol.}{ \bf 2}, e329 (2004).
-
-\bibitem{BriggmanDenk06}
-Briggman, K. and Denk, W.
-\newblock {\em Current opinion in neurobiology}{ \bf 16}(5), 562--570 (2006).
-
-\bibitem{MackeBorst08}
-Macke, J.~H., Maack, N., Gupta, R., Denk, W., Schölkopf, B., and Borst, A.
-\newblock {\em J Neurosci Methods}{ \bf 167}(2), 349--357 Jan  (2008).
-
-\bibitem{Mishchenko09}
-Mishchenko, Y.
-\newblock {\em J Neurosci Methods}{ \bf 176}(2), 276--289 Jan  (2009).
-
-\bibitem{LuLichtman09}
-Lu, J., Fiala, J.~C., and Lichtman, J.~W.
-\newblock {\em PLoS ONE}{ \bf 4}(5), e5655 05  (2009).
-
-\bibitem{HornJohnson90}
-Horn, R. and Johnson, C.
-\newblock {\em {Matrix analysis}}.
-\newblock Cambridge Univ Pr,  (1990).
-
-\bibitem{ConroyLouis97}
-Conroy, J.~M., Kratzer, S.~G., and Podrazik, L.~J.
-\newblock In {\em Society for Industrial and Applied Mathematics},  (1997).
-
-\bibitem{ZaslavskiyVert08}
-Zaslavskiy, M., Bach, F., and Vert, J.
-\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence}{
-  \bf 31}, 2227--2242 (2008).
-
-\bibitem{Durbin87}
-Durbin, R.~M.
-\newblock {\em Studies on the Development and Organisation of the Nervous
-  System of Caenorhabditis elegans}.
-\newblock PhD thesis, University of Cambridge,  (1987).
-
-\bibitem{deBonoMaricq05}
-de~Bono, M. and Maricq, A.~V.
-\newblock {\em Annu Rev Neurosci}{ \bf 28}, 451--501 (2005).
-
-\bibitem{VarshneyChklovskii09}
-Varshney, L., Chen, B., Paniagua, E., Hall, D., and Chklovskii, D.
-\newblock {\em ArXiv}{ \bf } (2009).
-
-\bibitem{ChalasaniBargmann07}
-Chalasani, S.~H., Chronis, N., Tsunozaki, M., Gray, J.~M., Ramot, D., Goodman,
-  M.~B., and Bargmann, C.~I.
-\newblock {\em Nature}{ \bf 450}(7166), 63--70 Nov  (2007).
-
-\bibitem{VaziriShank08}
-Vaziri, A., Tang, J., Shroff, H., and Shank, C.~V.
-\newblock {\em Proc Natl Acad Sci U S A}{ \bf 105}(51), 20221--20226 Dec
-  (2008).
-
-\bibitem{HelmstaedterDenk08}
-Helmstaedter, M., Briggman, K.~L., and Denk, W.
-\newblock {\em Curr Opin Neurobiol}{ \bf 18}(6), 633--641 Dec  (2008).
-
-\bibitem{BuckinghamSattelle08}
-Buckingham, S.~D. and Sattelle, D.~B.
-\newblock {\em Invert Neurosci}{ \bf 8}(3), 121--131 Sep  (2008).
-
-\end{thebibliography}
-
-\end{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Adding Figure and Table References
-%% Be sure to add figures and tables after \end{article}
-%% and before \end{document}
-
-%% For figures, put the caption below the illustration.
-%%
-%% \begin{figure}
-%% \caption{Almost Sharp Front}\label{afoto}
-%% \end{figure}
+%\subparagraph*{References}
+\bibliography{/Users/joshyv/Research/misc/biblist}
+\addcontentsline{toc}{section}{References}
+%\bibliography{Science}
+\bibliographystyle{ieeetr}
+% \bibliographystyle{nature}
 
 \begin{figure}[h!]
-% \centering 
-\includegraphics[width=.5\linewidth]{Lhatplot}
+\centering \includegraphics[width=.9\linewidth]{Lhatplot}
 \caption{C.~elegans graph classification simulation results.

### 2011-08-10 arxiv/v2/super.tex
added implications paragraph, perhaps to keep

--- 2011-08-05 nature_sci_rep/submission2/supervenience.tex
@@
-%\input{/Users/joshyv/Research/misc/latex_paper.tex}
 \documentclass{article}
@@
 \usepackage{amsfonts}
-\usepackage{amssymb}
 \usepackage{amsthm}
@@
 \usepackage{algorithmic}
-% \usepackage{times}
+\usepackage{times}
 \usepackage{fancyhdr}
-\usepackage{graphicx}
-\usepackage{verbatim}
-\usepackage{color}
-\usepackage[T1]{fontenc}
-\usepackage[scaled]{helvet}
-\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
 \pagestyle{fancy}
@@
 
-\usepackage{hyperref}
-% \usepackage{ulem}
-% \usepackage{color}
+\newcommand{\jv}{Joshua Vogelstein}
+\newcommand{\jhu}{Johns Hopkins University}
+%\newcommand{\hp}{https://jshare.johnshopkins.edu/jvogels3/public_html/}
+\newcommand{\ema}{joshuav@jhu.edu}
 
-% \newcommand{\loo}{$L^{(1)}_{h; \mD_n}$}
-\newcommand{\conv}{\rightarrow}
-% \newcommand{\Real}{\mathbb{R}}
-% \providecommand{\tr}[1]{\textcolor{red}{#1}}
+\newcommand{\Vr}{V_{reset}}
+\newcommand{\Vl}{V_{leat}}
+\newcommand{\eqdef}{\overset{\triangle}{=}}
+\newcommand{\D}[2]{\frac{\partial #1}{\partial #2}}
+\newcommand{\DD}[2]{\frac{\partial ^2 #1}{\partial #2 ^2}}
+\newcommand{\DDD}[3]{\frac{\partial ^2 #1}{\partial #2 \partial #3}}
+\newcommand{\Di}[2]{\frac{\partial ^i #1}{\partial #2 ^i}}
+\newcommand{\grad}{\nabla}
+\newcommand{\Hess}{\nabla\nabla}
 
-\newcommand{\mB}{\mathcal{B}}
-\newcommand{\mD}{\mathcal{D}}
-\newcommand{\mM}{\mathcal{M}}
-\newcommand{\PP}{\mathbb{P}}           % probability
+\providecommand{\tg}[1]{\textcolor{green}{#1}}
+\providecommand{\tb}[1]{\textcolor{blue}{#1}}
+\providecommand{\tr}[1]{\textcolor{red}{#1}}
+\providecommand{\tk}[1]{\textcolor{black}{#1}}
+\providecommand{\twhite}[1]{\textcolor{white}{#1}}
+\providecommand{\ve}[1]{\boldsymbol{#1}}
+\providecommand{\ma}[1]{\boldsymbol{#1}}
+%\providecommand{\ve}[1]{\boldsymbol{#1}}
+%\newcommand{\norm}[1]{\left|\left|#1\right|\right|}
+\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
+\providecommand{\deter}[1]{\lvert #1 \rvert}
+\providecommand{\abs}[1]{\left \lvert #1 \right \rvert}
+\newcommand{\trans}[1]{{#1}^{\ensuremath{\mathsf{T}}}}           % transpose
+\newcommand{\transpose}[1]{{#1}^{\ensuremath{\mathsf{T}}}}           % transpose
+\newcommand{\argmax}{\operatornamewithlimits{argmax}}
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
+\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose
+
+\newcommand{\FF}{\mathbb{F}}         
+\newcommand{\PP}{\mathbb{P}}         
+\newcommand{\GG}{\mathbb{G}}         
 \newcommand{\EE}{\mathbb{E}}           % expected value
-\newcommand{\II}{\mathbb{I}}           % expected value
-\newcommand{\Real}{\mathbb{R}}           % expected value
+\newcommand{\II}{\mathbb{I}}           % indicator function
+\newcommand{\QQ}{\mathbb{Q}}           
+\newcommand{\SSS}{\mathbb{S}}           
+\newcommand{\YY}{\mathbb{Y}}
+\newcommand{\Aa}{\mathbb{A}}
+\newcommand{\VV}{\mathbb{V}}
+\newcommand{\LL}{\mathbb{L}}
+
+\newcommand{\iid}{\overset{iid}{\sim}}
+
+% \DeclareMathOperator*{\argmax}{argmax}
+% \DeclareMathOperator*{\argmin}{argmin}
+% \DeclareMathOperator{\find}{find}
+
+% \newtheorem{thm}{Theorem}
+\newcommand{\thma}{\begin{thm}}
+\newcommand{\thmb}{\end{thm}}
+
+\newtheorem{defi}{Definition}
+\newtheorem{thm}{Theorem}
+
+% \newtheorem{defi}{Definition}
+% \newcommand{\defa}{\begin{defi}}
+% \newcommand{\defb}{\end{defi}}
+
+% \newcommand{\eq}{\begin{equation}}
+% \newcommand{\en}{\end{equation}}
+% \newcommand{\eqa}{\begin{equation}}
+% \newcommand{\eqb}{\end{equation}}
+
+\newcommand{\enuma}{\begin{enumerate}}
+\newcommand{\enumb}{\end{enumerate}}
+\newcommand{\ena}{\begin{enumerate}}
+\newcommand{\enb}{\end{enumerate}}
+
+\newcommand{\itema}{\begin{itemize}}
+\newcommand{\itemb}{\end{itemize}}
+\newcommand{\ita}{\begin{itemize}}
+\newcommand{\itb}{\end{itemize}}
+
+% \newcommand{\aligna}{\begin{align}}
+\newcommand{\alignb}{\end{align}}
+
+\newcommand{\proofa}{\begin{proof}}
+\newcommand{\proofb}{\end{proof}}
+
+\newcommand{\bla}{\begin{block}}
+\newcommand{\blb}{\end{block}}
+
+% \newcommand{\seqa}{\begin{equation*}}
+\newcommand{\seqb}{\end{equation*}}
+
+\newcommand{\bth}{\ve{\theta}}
+\newcommand{\bhth}{\mh{\ve{\theta}}}
+\newcommand{\thetn}{\ve{\theta}}
+\newcommand{\thet}{\thetn}
+\newcommand{\theth}{\widehat{\ve{\theta}}}
+\newcommand{\theto}{\ve{\theta}'}
+\newcommand{\wht}{\widehat{\thet}}
+\newcommand{\wtt}{\widetilde{\thet}}
+\newcommand{\vth}{\ve{\thet}}
+\newcommand{\vTh}{\ve{\Theta}}
+\newcommand{\hvth}{\widehat{\ve{\thet}}}
+\newcommand{\bTh}{\ve{\Theta}}
+\newcommand{\hbth}{\widehat{\thet}}
+
+\newcommand{\p}{P_{\bth}}
+\newcommand{\pold}{P_{\bth'}}
+\newcommand{\pk}{P_{\widehat{\ve{\theta}}^{(k)}}}
+\newcommand{\pT}{P_{\thetn_{Tr}}} %\thetn_T
+\newcommand{\pO}{P_{\thetn_o}} %\thetn_o
+\newcommand{\Q}{Q(\thetn,\theto)}
+\newcommand{\m}{m^{\ast}}
+\newcommand{\q}{q(\ve{H}_t)}
+\newcommand{\Ca}{[\text{Ca}^{2+}]}
+
+\newcommand{\Lik}{\mathcal{L}}
+\newcommand{\Cae}{[\widehat{\text{Ca}}^{2+}]}
+\newcommand{\Cav}{\ve{C}}%[\ve{\text{Ca}}^{2+}]}
+\newcommand{\sml}{\sqrt{\ma{\lambda}}}
+\newcommand{\ml}{\ma{\lambda}}
+\newcommand{\nw}{\widehat{n}}
+\newcommand{\nv}{\vec{n}}
+\newcommand{\Ae}{\widehat{A}}
+\newcommand{\te}{\widehat{\tau}}
+\newcommand{\maxn}{\max_{\ve{n}: n_t \geq 0}}
+\newcommand{\V}{\text{Var}}
+
+\providecommand{\mc}[1]{\mathcal{#1}}
+\providecommand{\mb}[1]{\boldsymbol{#1}}
+\providecommand{\mbb}[1]{\mathbb{#1}}
+\providecommand{\mv}[1]{\vec{#1}}
+\providecommand{\mh}[1]{\hat{#1}}
+\providecommand{\mt}[1]{\widetilde{#1}}
+\providecommand{\mhc}[1]{\hat{\mathcal{#1}}}
+\providecommand{\mhb}[1]{\hat{\boldsymbol{#1}}}
+\providecommand{\mvb}[1]{\vec{\boldsymbol{#1}}}
+\providecommand{\mtb}[1]{\widetilde{\boldsymbol{#1}}}
+
+\newcommand{\mP}{\mathbb{P}}
 
@@
 
-\providecommand{\mc}[1]{\mathcal{#1}}
-\providecommand{\mb}[1]{\boldsymbol{#1}}
-\providecommand{\mbb}[1]{\mathbb{#1}}
-\providecommand{\mv}[1]{\vec{#1}}
-\providecommand{\mh}[1]{\widehat{#1}}
-\providecommand{\mt}[1]{\widetilde{#1}}
-\providecommand{\mhc}[1]{\hat{\mathcal{#1}}}
-\providecommand{\mhb}[1]{\hat{\boldsymbol{#1}}}
-\providecommand{\mvb}[1]{\vec{\boldsymbol{#1}}}
-\providecommand{\mtb}[1]{\widetilde{\boldsymbol{#1}}}
+\newcommand{\Del}{\Delta}
+\newcommand{\Sig}{\Sigma}
+\newcommand{\Lam}{\Lambda}
+\newcommand{\Gam}{\Gamma}
 
-\newcommand{\argmax}{\operatornamewithlimits{argmax}}
-\newcommand{\argmin}{\operatornamewithlimits{argmin}}
+\newcommand{\dvs}{\dot{\bs}_t}
+\newcommand{\dvw}{\dot{\bw}_t}
+\newcommand{\dvx}{\dot{\bx}_t}
+\newcommand{\dvy}{\dot{\by}_t}
 
-% \newcommand{\mN}{\mathcal{N}}
+\newcommand{\ft}{f_{\ve{\thet}}}
+\newcommand{\gt}{g_{\ve{\thet}}}
+\newcommand{\hht}{h_{\thetn}}
 
+\newcommand{\Real}{\mathbb{R}}
+
+\newcommand{\wconv}{\overset{i.p.}{\rightarrow}}
+\newcommand{\sconv}{\overset{i.p.}{\rightarrow}}
+\newcommand{\conv}{\rightarrow}
+
+\newcommand{\mcE}{\mathcal{E}}
+\newcommand{\mcT}{\mathcal{T}}
+\newcommand{\mcG}{\mathcal{G}}
+\newcommand{\mcM}{\mathcal{M}}
+\newcommand{\mcL}{\mathcal{L}}
+\newcommand{\hatmcE}{\widehat{\mcE}}
+\newcommand{\hatp}{\widehat{p}}
+\newcommand{\hatP}{\widehat{P}}
+\newcommand{\hatQ}{\widehat{Q}}
+\newcommand{\hatL}{\widehat{L}}
+\newcommand{\mhP}{\widehat{\PP}}
+\newcommand{\tildeA}{\widetilde{A}}
+
+\newcommand{\defa}{\begin{defi}}
+\newcommand{\defb}{\end{defi}}
+\newcommand{\defeq}{\overset{\triangle}{=}}
+
+\newcommand{\qap}{\texttt{QAP} }
+\newcommand{\qapa}{\texttt{QAP}$_1$ }
+\newcommand{\qapn}{\texttt{QAP$_n$} }
+\newcommand{\qapb}{\texttt{QAP$_{100}$} }
+
+\newtheorem{Rem}{Remark}%[section]
+\newtheorem{Alg}{Algorithm}%[section]
+\newtheorem{Thm}{Theorem}[section]
+%\theoremstyle{marginbreak}
+\newtheorem{Lem}{Lemma}[section]
+%\newtheorem{Lem}[Cor]{Lemma}
+%\theoremstyle{change}
+%\theorembodyfont{\itshape}
+\newtheorem{Def}{Definition}[section]
+
+% \newcommand{\ss}{\mc{S}}
+% \newcommand{\mhc}{\mh{\mc{S}}_n}
+
+% \newcommand{\Real}{\mathbb{R}}
+% \newcommand{\bTh}{\mb{\Theta}}
+% \newcommand{\ra}{\rightarrow}
+% \newcommand{\la}{\leftarrow}
+% \newcommand{\bx}{\mb{x}}
+% \newcommand{\bv}{\mb{v}}
+% \newcommand{\ba}{\mb{a}}
+% \newcommand{\bd}{\mb{d}}
+% \newcommand{\balpha}{\mb{\alpha}}
+% \newcommand{\conv}{\rightarrow}  
+% \newcommand{\xx}{\langle \bx_u^\la, \bx_v^\ra\rangle}  
+% \DeclareMathOperator*{\argmax}{arg \hspace{1pt} max \hspace{2pt}}
+% \DeclareMathOperator*{\argmin}{arg \hspace{1pt} min \hspace{2pt}}
+% \newcommand{\norm}[1]{\left \lVert#1 \right  \rVert}
+%
+
+\usepackage{amssymb}
+\usepackage{hyperref}
+
+\newcommand{\mB}{\mathcal{B}}
+\newcommand{\mM}{\mathcal{M}}
 \newcommand{\hL}{\widehat{L}}
@@
 \newcommand{\MsB}{\mM \overset{S}{\sim}_{F} \mB}
-\newcommand{\MnoteB}{\mM \overset{\varepsilon}{{\not\sim}}_{F} \mB}
-\providecommand{\tr}[1]{\textcolor{black}{#1}}
-\providecommand{\norm}[1]{\left \lVert#1 \right  \rVert}
-\newcommand{\T}{^{\ensuremath{\mathsf{T}}}}           % transpose
 \newcommand{\from}{\colon}
 
-\newtheorem{defi}{Definition}
-\newtheorem{thm}{Theorem}
 \newtheorem{thex}{\emph{Gedankenexperiment}}
@@
 
-	\begin{center}
-{\huge	Are mental properties supervenient on brain properties?}
-\end{center}
+\title{Are mental properties supervenient on brain properties?}
+\author{Joshua T.~Vogelstein\\ Department of Applied Mathematics \& Statistics,  Johns Hopkins University
+        \and 
+		R.~Jacob Vogelstein \\ National Security Technology Department, Johns Hopkins University Applied Physics Laboratory
+		\and 
+		Carey E.~Priebe \\ Department of Applied Mathematics \& Statistics,  Johns Hopkins University}
 
-\vspace{5px}
+\date{}
+\maketitle
 
-\begin{center}
-{\large		
-Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
-	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
-
-\end{center}
-
-\vspace{5px}
+% 	\begin{center}
+% {\huge	Are mental properties supervenient on brain properties?}
+% \end{center}
+% 
+% \vspace{5px}
+% 
+% \begin{center}
+% {\large		
+% Joshua T. Vogelstein$^{1*}$, R. Jacob Vogelstein$^2$, Carey E. Priebe$^1$\\
+% 	$^1$Department of Applied Mathematics \& Statistics, \\ Johns Hopkins University, Baltimore, MD, 21218,\\ $^2$National Security Technology Department, \\ Johns Hopkins University Applied Physics Laboratory, Laurel, MD 20723}
+% 	
+% \end{center}
+% 
+% \vspace{5px}
 
@@
 
+% \noindent 
+\begin{abstract}
 \noindent The``mind-brain supervenience'' conjecture suggests that all mental properties are derived from the physical properties of the brain.
@@
 This may provide motivation for cross-disciplinary research between neuroscientists and statisticians.
+\end{abstract}
 
@@
 
-\newpage
-% \section*{Introduction}
+% \newpage
+\section*{Introduction}
 
@@
 Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.
-This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).
+This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix for more details and some examples).
 To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
@@
 \end{defi}
-\noindent Statistical supervenience is therefore a probabilistic relation on sets (which could be considered a generalization of correlation; see Appendix 1 for details).
+\noindent Statistical supervenience is therefore a probabilistic relation on sets (which could be considered a generalization of correlation; see Appendix for details).
 
@@
 
+\newpage
+\appendix
+\section*{Appendix}
+
+In this appendix we compare supervenience to several other relations on sets (see Figure \ref{fig:rel}).
+
+% aim to provide more intuition regarding supervenience by discussing the limitations and extent of its implications.
+
+%For instance, it may be the case that minds supervene on brains, but one cannot form an \emph{injective} relation from brains to minds.
+
+First, a supervenient relation does not imply an injective relation.
+An injective relation is any relation that preserves distinctness.
+Thus if minds are injective on brains, then $b\neq b' \implies m \neq m'$ (note that the directionality of the implication has been switched relative to supervenience).
+However, it might be the case that a brain could change without the mind changing.
+Consider the case that a single subatomic particle shifts its position by a Plank length, changing brain state from $b$ to $b'$.
+It is possible (likely?) that the mental state supervening on brain state $b$ remains $m$, even after $b$ changes to $b'$.
+In such a scenario, the mind might still supervene on the brain, but the relation from brains to minds is not injective.
+This argument also shows that supervenience is not necessarily a \emph{symmetric} relation.
+Minds supervening on brains does not imply that brains supervene on minds.
+
+Second,  supervenience does not imply causality. %it may be the case that minds supervene on brains, but that brains do not cause minds.  
+For instance, consider an analogy where $M$ and $B$ correspond to two coins being flipped, each possibly landing on heads or tails.
+Further assume that every time one lands on heads so does the other, and every time one lands on tails, so do the other.
+This implies that $M$ supervenes on $B$, but assumes nothing about whether $M$ causes $B$, or $B$ causes $M$, or some exogenous force causes both.
+
+Third, supervenience does not imply identity.  The above example with the two coins demonstrates this, as the two coins are not the same thing, even if one has perfect information about the other.
+
+What supervenience does imply, however, is the following.
+Imagine finding two unequal minds.
+If $\MsB$, then the brains on which those two minds supervene must be different.
+In other words, there cannot be two unequal minds, either of which could supervene on a single brain.
+Figure \ref{fig:rel} shows several possible relations between the sets of minds and brains.
+
+Note that statistical supervenience is distinct from statistical correlation.
+\emph{Statistical correlation} between brain states and mental states is defined as $\rho_{MB}=\EE[(B-\mu_B)(M-\mu_M)]/(\sig_B \sig_M)$, where $\mu_X$ and $\sig_X$ are the mean and variance of $X$, and $\EE[X]$ is the expected value of $X$.
+If $\rho_{MB}=1$, then both $\MsB$ and $\mB \overset{S}{{\sim}}_{F} \mM$.
+Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
+In fact, supervenience may be thought of as a generalization of correlation which incorporates directionality, can be applied to arbitrary valued random variables (such as mental or brain properties), can depend on any moment of a distribution (not just the first two).
+
+\begin{figure}[h!tbp]
+	\centering
+		\includegraphics[width=1\linewidth]{supervenience_relations.pdf}
+	\caption{Possible relations between minds and brains.
+(A) Minds supervene on brains, and it so happens that there is a bijective relation from brains to minds.
+(B) Minds supervene on brains, and it so happens that there is a surjective (a.k.a., onto) relation from brains to minds, but not a bijective one.
+(C) Minds are \emph{not} supervenient on brains, because two different minds supervene on the same brain.}
+	\label{fig:rel}
+\end{figure}
+% section relations_between_sets (end)
+
 \clearpage

### 2011-08-10 arxiv/v2/super_abstract.tex
added implications paragraph, perhaps to keep

The "mind-brain supervenience" conjecture suggests that all mental properties are derived from the physical properties of the brain. To address the question of whether the mind supervenes on the brain, we frame a supervenience hypothesis in rigorous statistical terms. Specifically, we propose a modified version of supervenience (called epsilon-supervenience) that is amenable to experimental investigation and statistical analysis. To illustrate this approach, we perform a thought experiment that illustrates how the probabilistic theory of pattern recognition can be used to make a one-sided determination of epsilon-supervenience. The physical property of the brain employed in this analysis is the graph describing brain connectivity (i.e., the brain-graph or connectome). epsilon-supervenience allows us to determine whether a particular mental property can be inferred from one's connectome to within any given positive misclassification rate, regardless of the relationship between the two. This may provide motivation for cross-disciplinary research between neuroscientists and statisticians.

### 2011-08-10 nature_sci_rep/submission2/supervenience.tex
explained that super need not be crazy hard or eps crazy small

@@
 We then use the probabilistic theory of pattern recognition to determine
- % This perspective leads to a greater understanding vis-\'a-vis 
 the limits of what one can and cannot learn about supervenience through data analysis.  
+The implications of this work are varied.
+It provides a probabilistic framework for converting philosophical conjectures into statistical hypotheses that are amenable experimental investigation, which  allows the philosopher to gain empirical support for her rational arguments.
+This leads to the construction of the first explicit proof (to our knowledge) of a universally consistent classifier on graphs, and the first demonstration of the tractability of answering supervenience questions.
+Supervenience therefore seems to perhaps be a useful (but under utilized) concept for neuroscientific investigations.
+This should provide further motivation for cross-disciplinary efforts across three fields---philosophy, statistics, and neuroscience---with shared goals but mostly disjoint jargon and methods of analysis.
 
@@
 Specifically, searches for ``engrams'' of  memory traces \cite{Lashley50} or ``neural correlates''  of various behaviors or mental properties (for example, consciousness \cite{Koch2010}), may be more aptly called searches for the ``neural supervenia'' of such properties.
-% any investigation that aims to ``explain'' some behavioral or psychological phenomenon by reference to neural (or other brain) activity or structure could be thought of as mind-brain supervenience investigations.
-% Thus, the proposed $\eps$-supervenience framework is perhaps a useful unifying perspective.
- % for these fields.  
-% The above Gedankenexperiment is an 
 Letting the brain property be a brain-graph is perhaps especially pertinent in light of the advent of ``connectomics'' \cite{SpornsKotter05,Hagmann05}, a field devoted to estimating whole organism brain-graphs and relating them to function.
 Testing supervenience of various mental properties on these brain-graphs will perhaps therefore become increasingly compelling; the framework developed herein could be fundamental to these investigations.
-% Note that these connectomics investigations might lack any dynamics; rather, they effectively assume that structure determines the relevant dynamics.
-Assuming that dynamics are a function of either vertex or edge attributes, whether any particular mental property is supervenient to ``only'' connectivity structure, or also some attributes, is merely another $\eps$-supervenience hypothesis.
 For example, questions about whether connectivity structure alone is sufficient to explain a particular mental property is one possible mind-brain $\eps$-supervenience investigation.
 The above synthetic data analysis demonstrates the feasibility of $\eps$-supervenience on small brain-graphs.
-% Appendix 2 presents an illustrative simulated example using the only currently known connectome---that of a Caenorhabditis elegans---demonstrating that $\eps$-supervenience can be rejected at some critical value $c_\alpha$ with only a reasonably small number of samples.
-Similar supervenience tests on larger animals (such as humans) will potentially benefit from either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011}, more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
+Note that $\eps$-supervenience tests need not investigate seemingly intractable problems, like consciousness.
+For example, it already seems quite clear that visual perception supervenes on visual cortical activity \cite{??}.
+Moreover, an inability to reject $\eps$-supervenience for small $\eps$ is also potentially meaningful.
+For example, perhaps auditory localization precision supervenes on a rate code only to some $\eps > c$, the rest supervening on a spike timing code \cite{??}.
+Similar supervenience tests on increasingly complex mental properties will potentially benefit from either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011}, more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
 

### 2011-08-10 nature_sci_rep/submission2/supervenience.tex
moved a paranthesis

@@
 \end{defi}
-\noindent Statistical supervenience is therefore a probabilistic relation on sets (which could be considered a generalization of correlation; see Appendix 1 for details).
+\noindent Statistical supervenience is therefore a probabilistic relation on sets which could be considered a generalization of correlation (see Appendix 1 for details).
 

### 2011-08-11 nature_sci_rep/submission2/cover_letter.tex
ready to resubmit?

The philosophical notion of supervenience, colloquially, that there cannot be an A-difference without a corresponding B-difference, has received much attention in the philosophy literature; but relatively scant attention from the statistics and science communities.  We believe this is an unfortunate and unnecessary consequence of scientific reductionism.  In this work, we very briefly present the philosophical notion of supervenience canonized by Davidson (Experience and Theory, 1970).  We then rapidly launch into a statistical understanding

### 2011-08-11 nature_sci_rep/submission2/supervenience.tex
ready to resubmit?

@@
 \end{quotation}
+Colloquially, supervenience means ``there cannot be a mind-difference without a physical-difference.''
 This philosophical conjecture has potentially widespread implications.  
@@
 The implications of this work are varied.
-It provides a probabilistic framework for converting philosophical conjectures into statistical hypotheses that are amenable experimental investigation, which  allows the philosopher to gain empirical support for her rational arguments.
+It provides a probabilistic framework for converting philosophical conjectures into statistical hypotheses that are amenable to experimental investigation, which  allows the philosopher to gain empirical support for her rational arguments.
 This leads to the construction of the first explicit proof (to our knowledge) of a universally consistent classifier on graphs, and the first demonstration of the tractability of answering supervenience questions.
-Supervenience therefore seems to perhaps be a useful (but under utilized) concept for neuroscientific investigations.
-This should provide further motivation for cross-disciplinary efforts across three fields---philosophy, statistics, and neuroscience---with shared goals but mostly disjoint jargon and methods of analysis.
+Supervenience therefore seems to perhaps be a useful but under-utilized concept for neuroscientific investigations.
+This work should provide further motivation for cross-disciplinary efforts across three fields---philosophy, statistics, and neuroscience---with shared goals but mostly disjoint jargon and methods of analysis.
 
@@
 This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).
-To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.
+To facilitate both statistical analysis and empirical investigation, we convert this local supervenience relation from a logical to a probabilistic relation.
 
@@
 
-% Given that 
-If minds statistically supervene on brains, 
-% , $\MsB$, 
-then if two minds differ, there must be some brain-based difference to account for the mental difference.
+If minds statistically supervene on brains, then if two minds differ, there must be some brain-based difference to account for the mental difference.
 This means that there must exist a deterministic function $g^*$ mapping each brain to its supervening mind.
@@
 Note that $\eps$-supervenience tests need not investigate seemingly intractable problems, like consciousness.
-For example, it already seems quite clear that visual perception supervenes on visual cortical activity \cite{??}.
+For example, aspects of visual perception appear to supervene on visual cortical activity (for example, binocular rivalry \cite{Tong2006}).
 Moreover, an inability to reject $\eps$-supervenience for small $\eps$ is also potentially meaningful.
-For example, perhaps auditory localization precision supervenes on a rate code only to some $\eps > c$, the rest supervening on a spike timing code \cite{??}.
+For example, perhaps auditory localization precision supervenes on a rate code only to some $\eps > c$, the rest supervening on a spike timing code \cite{Chase2006}.
 Similar supervenience tests on increasingly complex mental properties will potentially benefit from either higher-throughput imaging modalities \cite{HayworthLichtman06, Bock2011}, more coarse brain-graphs \cite{PalmAmunts10,Johansen-Berg2009}, or both.
@@
 
-The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, S.~Seung, and two helpful referees.
+The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, S.~Seung, and K.~Kording.
 

### 2011-08-11 nature_sci_rep/submission2/supp.tex
finished redline cep comments of aug 5 2011
ready to resubmit?

@@
 \newtheorem{thex}{\emph{Gedankenexperiment}}
-\lhead{Vogelstein JT, et al}
+\lhead{Vogelstein JT, et al.}
 \rhead{Statistical Supervenience Supplement}
@@
 Thus, perfect correlation implies supervenience, but supervenience does not imply correlation.
-In fact, supervenience may be thought of as a generalization of correlation which incorporates directionality, can be applied to arbitrary valued random variables (such as mental or brain properties), can depend on any moment of a distribution (not just the first two).
+In fact, supervenience may be thought of as a generalization of correlation which incorporates directionality, can be applied to arbitrary valued random variables (such as mental or brain properties), and can depend on any moment of a distribution (not just the first two).
 

### 2011-09-16 nature_sci_rep/submission2/supervenience.tex
made decision letter response less light
waiting for proofs now

@@
 For example, neural network theory and artificial intelligence often implicitly assume 
-% take a generalized notion of 
-a local version mind-brain supervenience 
-% as an assumption 
-% , which if falsified, might change modern approaches to learning 
-\cite{Haykin2008,Ripley2008}. Cognitive neuroscience similarly seems to operate under such assumptions
+a local version mind-brain supervenience \cite{Haykin2008,Ripley2008}. Cognitive neuroscience similarly seems to operate under such assumptions
 % , which if falsified, might result in novel perspectives and theories 
@@
 
-Let $A_{uv}$ be the number of chemical synapses between neuron $u$ and neuron $v$ according to \cite{VarshneyChklovskii09}.
+Let $A_{uv}$ be the number of chemical synapses between neuron $u$ and neuron $v$ according to \cite{Varshney2011}.
 Then, let $\mc{S}$ be the set of edges deemed responsible for odor-evoked behavior according to \cite{ChalasaniBargmann07}.
@@
 \bibliography{/Users/jovo/Research/latex/library}
-\bibliographystyle{nature}
+\bibliographystyle{naturemag}
 
@@
 
-The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, S.~Seung, and K.~Kording.
+This work was partially supported by the Research Program in Applied Neuroscience. The authors would like to acknowledge helpful discussions with J.~Lande, B.~Vogelstein, S.~Seung, and K.~Kording.
 
@@
 
+\newpage
+\section*{Figure Legends} % (fold)
+\label{sec:section_name}
+
+% section section_name (end)
+\paragraph{Figure 1} % (fold)
+\label{par:figure_1}
+
+% paragraph figure_1 (end)
+
+C.~elegans graph classification simulation results.  The estimated hold-out misclassification rate $\hL^{n'}_{F}(g_{n})$  (with $n'=1000$ testing samples) %$\hL^{1000}_{\PP}(g_{\mt{n}})$ 
+is plotted as a function of class-conditional training sample size $n_j=n/2$, suggesting that for $\varepsilon=0.1$ we can determine that $\MeB$ holds with $99\%$ confidence with just a few hundred training samples generated from $F_{MB}$.
+Each dot depicts $\hL^{n'}_{F}(g_n)$ for some $n$; standard errors are $(\hL^{n'}_{F}(g_{n}) (1-\hL^{n'}_{F}(g_{n}))/n')^{1/2}$.
+For example, at $n_j = 180$ we have $k_n = \lfloor\sqrt{8 n}\rfloor=53$ (where $\lfloor\cdot\rfloor$ indicates the floor operator), $\hL^{n'}_{F}(g_{n}) = 0.057$, and standard error less than $0.01$.
+We reject $H_0^{0.1}: L^* \geq 0.1$ at $\alpha=0.01$.
+Note that $L^* \approx 0$ for this simulation.
+
 \end{document}
